### 2013-11-14 ThanoRepro/ThanoRepro.tex
first commit, full ThanoRepro

\documentclass{article}
\usepackage{amsmath}
\usepackage{caption}
\usepackage{placeins}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{natbib}
\bibpunct{(}{)}{,}{a}{}{;} 
\usepackage{url}
\usepackage{nth}
% for the d in integrals
\newcommand{\dd}{\; \mathrm{d}}

\usepackage[top=2in, bottom=1.5in, left=1in, right=1in]{geometry}

% matrix numbering
\usepackage{trivfloat}
\trivfloat{matrix}
\floatstyle{plaintop}
    \restylefloat{matrix}
\AtBeginDocument{\numberwithin{matrix}{section}}
% end preamble
%-------------------------------------------------------

\begin{document}

\title{Renewal and stability in populations structured by remaining
years of life}
\author{Tim Riffe \\ University of California, Berkeley}
\maketitle

\begin{abstract}
We transform data classified by chronological age into data classified by
remaining years of life (thanatological age). A model for population renewal and
the corresponding projection matrix are presented for populations structured by
thanatological age. Period results are derived using all available data from the
HMD and HFD. We compare the intrinsic growth rate, $r$, as derived from the
classic Lotka equation versus that derived on the basis of thanatological age.
We also compare some transient indicators between the two
models. Empirical results suggest that $r$ from the thanatological model tends to be less erratic than
Lotka's $r$, and the trajectory to stability tends to be faster with less
oscillation.
\end{abstract}

*This is a work in progress. All findings are preliminary at this time, so
please don't cite without permission of the author.
\vspace{2em}

All demographic forces vary over age in known and regular ways. Information on
such forces, along with the size and age structure of a population allows the
demographer to make predictions about the future size and structure of a population. In this
paper, we explore some formal demographic consequences of a particular
redefinition of age. Instead of counting age as the time passed between birth and the present (or
some other moment), consider age as the amount of time left from the present
until death. Individuals in this case move in the same direction along
imaginary life lines, but the reference point for age is now at the end of the
life line instead of at the beginning. Of course, such information cannot be
requested of survey respondents or register data, because one's time at death is
typically unknown. Instead we approximate this perspective using a similar
strategy to that used by \citet{miller2001increasing} or
\citet{lee2002approach}, which uses lifetables and aggregated
population counts.

Temporal perspectives and rescalings are abundant in demography already,
but questions are left unanswered. Likely some demographic phenomena are best
described as a function of time since birth, and others of time until death,
while still others can be a function of both to some degree. This author
supposes that fertility and reproduction can be a function of both age
perspectives, and that the question of how to partition forces and events over
time is a difficult one. Thanatological reproduction sits at the opposite
extreme of chronological reproduction, the Lotka-Euler renewal
model. By specifying a model of thanatological renewal, we both expand the
toolbox available to approach complex temporal structure and we provide a set of broad
findings to help understand the implications of temporal structure.

\section*{From chronological to thanatological age}

Our point of departure is a little formula that gives the probability of
dying at some age $n$ years in the future, given survival to some
earlier age, $a$. For age 0 this is simply the $d(a)$ column of the lifetable,
standardized to sum to 1. For higher ages, we simply condition on survival:

\begin{equation}
\label{eq:vaupel1}
f(n | a) = \mu(a+n)\frac{l(a+n)}{l(a)}
\end{equation}

A convenient discretization of \eqref{eq:vaupel1} is to simply work with the
$d_x$ column of the lifetable

\begin{equation}
d_{x+n, x} = \frac{d_{x+n}}{\sum_{i=x}^\omega d_i }
\end{equation}
where $\omega$ is the highest age.\footnote{In later formulas we just use $l_x$
in the denominator to reduce clutter, but the above sum over $d_x$ will eliminate
rounding error in case one is working from a published life table.} One
produces in this way a probability density function over future death times for each age, and the $d_{x+n, x}$- weighted average of $n$ for a given age $x$ will simply produce the familiar remaining life expectancy column of the lifetable, $e_x$. Since $1 = \sum_{n=0}^{\omega - x}d_{x+n, x}$ for each $x$, we can use this to
decompose age-classified demographic data. For example, we can break down a
population pyramid into remaining years categories, like in
Figure~\ref{fig:USdecomp}, which shows this exercise for 2010 US data. The fun starts when we proceed to regroup
the data by $n$ instead of collapsing back to $x$. Now instead of remaining lifetime 
heterogeneity within each age group we have age heterogeneity within each
discrete remaining years of life. Figure~\ref{fig:USrecomp} shows this
\textit{re}composition for the same 2010 US data.

\begin{figure}
	\caption{2010 US population}
	\begin{center}
	\begin{subfigure}{.45\textwidth}
		\caption{Chronological age-structure decomposed by thanatological age.}
		\label{fig:USdecomp}
		\includegraphics[scale=.6]{Figures/US2010Age.pdf}
	\end{subfigure}
	~
	\begin{subfigure}{.45\textwidth}
		\caption{Thanatological age-structure decomposed by chronological age.}
		\label{fig:USrecomp}
		\includegraphics[scale=.6]{Figures/US2010Thano.pdf}
	\end{subfigure}
	\\
	\small{*Population and mortality data from the HMD.}
	\end{center}
\end{figure}

If $P_x$ is an age-classified discrete population count, we can derive $P_y$, a
remaining-years-classified population count, as follows:

\begin{align}
\label{eq:transform}
P_y =& \sum_{x=0}^\omega P_x \frac{d_{x+y}}{l_x}
\intertext{and of course}
P =& \sum P_y = \sum P_a
\end{align}

We refer to \eqref{eq:transform} as the thanatological age
transformation\footnote{I have on good authority that Ken Wachter
coined the term thanatological age.}, and this complements the notion of
chronological age. It bears repeating that thanatological age is merely probabilistic, since it
refers to the future according to particular assumptions about the survival
curve. As such, Figure~\ref{fig:USrecomp} is projective in nature, a
forward-looking glance at a population given the age-structure and lifetable of a particular moment, whereas
Figure~\ref{fig:USdecomp} is reflective in nature, since a population's
age-structure is mostly the fruit of past fertility, but also migration and to a
lesser degree attrition. In a different paper we explore population structure by remaining years of life
for its own sake. Note that any age-classified count can be
reclassified in this way, as long as you have a lifetable that plausibly
represents the population whose data you wish to reclassify. This means we can
derive thanatological fertility rates, $F_y$, by applying \eqref{eq:transform}
to age-classified birth counts, $B_x$, to get $B_y$ and again to
exposure-to-risk, $E_x$ (here we take exposures from all ages), to get $E_y$ and then dividing:

\begin{equation}
F_y = \frac{B_y}{E_y}
\end{equation}

We are interested to use thanatological fertility rates, $F_y$ in
models of population renewal, rather than to explicitly study the nature of these rates.
Nonetheless it is best not to throw such rates into a model blindly, so we
offer a schematic overview of some of their characteristics.
Figure~\ref{fig:FySpaghetti} represents the full variety in
female thanatological period fertility rates that can be found for all years
of data that overlap in the HFD and HMD.\footnote{There are as of this writing
1834 population-years of overlap between the HMD and HFD, including a wide
variety of fertility and mortality combinations.} Figure~\ref{fig:FxSpaghetti} gives ASFR for the same
populations and years as a more familiar reference, but note that both scales are different! With no indication of particular populations or time series, one already concludes that thanatological fertility rates have a characteristic shape and are not random, erratic or informationless: there is a pattern to fertility rates by remaining years of life, and it varies over time and between populations within some range of normality. Thanatological fertility rates have a wider distribution than do chronological rates. Note that this spaghetti plot includes both fertility booms and busts, as
well as some mortality crises (1918, WWII). Some patterns to note, but which we
do not separate graphically here, are that the left tail (a gauge of how
orphan-prone a population is) has tended to fall over time. All populations have
shown a rightward shift in both the mean and modal female thanatological age at
birth (TAB) over time (that's unambiguously a good thing). Several populations
now show modal TABs of over 60 years. Thanatological total fertility rates track
standard PTFR rather closely, but tend to be somewhat higher.

\begin{figure}[h!]
	\caption{Chronological and thantological fertility rates, all 1600
	country-year combinations present in both the HFD and HMD.* Note different x
	and y scales.}
	\label{fig:Fxcompare}
	\begin{center}
	\makebox[\textwidth]{
	\begin{subfigure}{.45\textwidth}
		\caption{Chronological period fertility rates.}
		\label{fig:FxSpaghetti}
		\includegraphics[scale=.45]{Figures/FxSpaghettiDraft.png}
	\end{subfigure}
	~
	\begin{subfigure}{.45\textwidth}
		\caption{Thanatological period fertility rates.}
		\label{fig:FySpaghetti}
		\includegraphics[scale=.45]{Figures/FySpaghettiDraft.png}
	\end{subfigure}
	}
	\\
	\end{center}
	\begin{tiny}
	*AUT, $1951-2010$; BGR, $1947-2009$; BLR, $1964-2009$; CAN, $1921-2007$; 
	CHE, $1932-2011$; CZE, $1950-2011$; DEUTE, $1956-2010$; DEUTNP, $1990-2010$; 
	DEUTW, $1956-2010$; ESP, $1922-2006$; EST, $1959-2010$; FIN, $1939-2009$; 
	FRATNP, $1946-2010$; GBR\_NIR, $1974-2009$; GBR\_NP, $1974-2009$; GBR\_SCO,
	$1945-2009$; GBRTENW, $1938-2009$; HUN, $1950-2009$; IRL, $1955-2009$; JPN, $1947-2009$; 
	LTU, $1959-2010$; NLD, $1950-2009$; NOR, $1967-2009$; PRT, $1940-2009$; 
	RUS, $1959-2010$; SVK, $1950-2009$; SVN, $1983-2009$; SWE, $1891-2010$; 
	TWN, $1976-2010$; UKR, $1959-2006$; USA, $1933-2010$
	\end{tiny}
\end{figure}

\section*{The thanatological renewal model}
For the population model that follows, we are most interested in using the kind
of fertility rates shown in Figure~\ref{fig:FySpaghetti}, and we refer to a
unisex population. Assuming constant vital rates, the births for the present year are given by

\begin{align}
B(t) =& \int _0^\infty P(y,t)F(y) \dd y = \int _0^\infty P(a,t)F(a) \dd a
\intertext{where $P(a), P(y)$ are population counts and $F(y), F(a)$ are exact
specific fertility probabilities (rates), and $d(a)$ is the continuous
lifetable death distribution with radix of 1, so $d(a)/l(0)$. The
thanatological integral can be broken down back in terms of chronological age.} =&
\int_{y=0}^\infty
\int_{a=0}^\infty F(y)\frac{P(a,t)d_(a+y)}{l(a)} \dd a \dd y 
\intertext{We can relate the present population to past
births with $P(a,t) = B(t-a)l(a)$.} 
\label{eq:ergostep1}
=& \int_{y=0}^\infty \int_{a=0}^\infty F(y) B(t-a)d(a+y)\dd a \dd y
\intertext{And eventually-- a bit quicker than is the case for chronological
age-- strong ergodicity will assert itself, and $B(t)$ will be related to
$B(t-a)$ according to a constant factor $e^{ra}$, where $r$ is the familiar
intrinsic rate of growth.}
\label{eq:ergostep2}
=& \int_{y=0}^\infty \int_{a=0}^\infty F(y) B(t)e^{-ra}d(a+y)\dd a \dd y
\intertext{Divide out $B(t)$ to get back a familiar-looking renewal equation.}
\label{eq:thanoren}
1 =& \int_{y=0}^\infty \int_{a=0}^\infty F(y) d(a+y)e^{-ra}\dd a \dd y
\end{align}
Now compare this to Lotka's chronological formulation
\begin{equation}
\label{eq:lotka}
1 = \int_{a=0}^\infty F(a)l(a)e^{-ra}\dd a
\end{equation}
and note that these are really quite similar, since $\int _{y=0}^\infty
d(a+y)\dd a = l(a)$. For intuition, notice that $l(a)$ is here split
up into pieces of $d(a)$, and imagine a 2D surface of these, where one axis is chronological age and the
other axis is thanatological age. For the chronological case \eqref{eq:lotka},
we multiply chronological age-specific fertility rates over one margin, and for
the thanatological case over the other margin. There is a strong parallel here
with the case of two-sex age-specific fertility rates. As in the case of
divergence between male and female single-sex models, there will always be divergence between
single-sex models under chronological versus thanatological age. Even though the
modeled population stocks are in a way commensurable, the rates used, $F(a)$
versus $F(y)$ are calculated on the basis of differently distributed
denominators $E(a)$ versus $E(y)$.

There are gaps in the above line of development, since the jump from
\eqref{eq:ergostep1} to \eqref{eq:ergostep2} (strong ergodicity) is unproven. A
further conjecture is that the jump will also hold for the case of weak
ergodicity. The mechanisms at play unfold in the same way as those so
intuitively described by \citet{arthur1982ergodic}, and said proof may apply
here without further modification. A proof of
the uniqueness of the solution to \eqref{eq:thanoren} is given in Appendix~\ref{app:A}.

Once one finds $r$ from \eqref{eq:thanoren}, other familiar stable population
parameters can be calculated. For instance, we may calculate the mean
thanatological generation time, $T^y$, as:
\begin{equation}
\label{eq:Ty}
 T^y =  \frac{\int _{y=0}^\infty \int _{a=y}^\infty y e^{-ra} d(a) F(y) \dd a
\dd y}{\int _{y=0}^\infty \int _{a=y}^\infty e^{-ra} d(a) F(y) \dd a \dd y}
\end{equation}
The net reproduction rate, $R_0$ is related by, e.g.,
\begin{equation}
\label{eq:R0fromTy}
R_0 = e^{r T^y}
\end{equation}
The birth rate, $b$, is given by
\begin{equation}
\label{eq:eybrate}
b = \frac{1}{\int _{y=0}^\infty \int _{a=y}^\infty e^{-ra} d(a) \dd a
\dd y}
\end{equation}
The stable age structure, $c$, where $c_y$ is the
proportion of the stable population with remaining years to live $y$, is given
by
\begin{equation}
\label{eq:cy}
c(y) = b \int _{a=y}^\infty e^{-ra} d(a) \dd a
\end{equation}
Other stable population quantities may be estimated by similarly translating the
various common definitions (e.g., in the glossary of \citet{coale1972growth}) to
the present perspective. We will focus on the main model rather than on these.

What should one imagine under the model of thanatological population renewal? A
useful mnemonic bases itself on Figures~\ref{fig:USdecomp} and
\ref{fig:USrecomp}. In the age-structured model, new generations
appear at the bottom of the pyramid, and move up one rung per year. All ages are subject to decrement. In the thanatological \textit{leaf}, each birth cohort increments to
the population over the whole range of thanatological age according to $d(a)$,
as seen in \eqref{eq:thanoren}. Birth cohorts become the layers seen in
Figure~\ref{fig:USrecomp}. Each horizontal step is a death cohort, and these
move one step down the pyramid each year without any decrement (indeed
incrementing) until reaching the very bottom. In short, the locations of
increment and decrement, and the direction of movement are all switched.

\section*{The thanatological projection matrix}
These descriptions can be made more explicit by hashing out the projection
matrix that corresponds to the thanatological renewal model. As with the
age-structured Leslie matrix, the thanatological projection matrix,
$\textbf{Y}$, is square and of dimension $n \times n$, where $n$ is the number
of remaining-years classifications into which the population is divided. 
The matrix contains elements for survival and elements for fertility. Unlike
Leslie matrices, $\textbf{Y}$ is not sparse, but is populated primarily with non-zero entries.

Of interest is that mortality occurs in only the population class with zero
remaining years of life. Thanatological age 1 in year $t$ moves to 0 in year
$t + 1$. Thus, instead of in the subdiagonal, we place survival in the
superdiagonal. All survival values are 1, since there is no decrement, and the
upper-left corner of this superdiagonal contains a 0, for full decrement.
We illustrate using a 6$\times$6 matrix. The survival component of $\textbf{Y}$ is organized as in
Matrix~\ref{matrix:Ysurv}.

\begin{matrix}[h!]
\centering
\caption{Survival component of unisex thanatological projection matrix,
$\textbf{Y}$}
\label{matrix:Ysurv}
$\bordermatrix{{e_y } & 0_t & 1_t & 2_t & 3_t & 4_t & 5_t\cr 
                0_{t+1} & 0    &  1   & 0    & 0    & 0    & 0   \cr
                1_{t+1} & 0    &  0   & 1    & 0    & 0    & 0   \cr 
                2_{t+1} & 0    &  0   & 0    & 1    & 0    & 0   \cr 
                3_{t+1} & 0    &  0   & 0    & 0    & 1    & 0   \cr 
                4_{t+1} & 0    &  0   & 0    & 0    & 0    & 1   \cr
                5_{t+1} & 0    &  0   & 0    & 0    & 0    & 0   }$
\end{matrix}

 Fertility inputs to the matrix are derived from single-sex thanatological
 fertility and the lifetable $d_a$ distribution, where $a$ indexes age, but is
 equal to $y$, remaining years of life for persons aged 0. Fertility
 in a thanatologically structured population occurs in all but the highest
 remaining years classes. For our example, say that fertility is observed in
 classes 0-4, while the final class has no fertility, where $F_y$ indicates the
 fertility probability for class $y$ in the year $t$ entering population (in the 
 matrix columns). Each $F_y$ is then distributed according to
 $d_a$ with no further translation, since the $d_a$ lifetable column already
 refers to age 0. Thus, the fertility entry in row $m$ and column $n$ of
 $\textbf{Y}$ will be $F_n \cdot d_m$. We assume that those dying over the course of year
 $t$ (the first column) are exposed to fertility for half of the
 year,\footnote{One might be tempted to not allow for fertility at all for
 females dying in year $t$, but recall that fertility is measured in the moment of
 birth, and not conception.}
 and so discount the fertility entry accordingly. Further, infant mortality, 
 $F_y \cdot d_0$, located in the first row, must also be discounted, since part
 of the mortality will occur in the same year $t$ and the rest in year $t + 1$. 
 The first row of fertility must be further discounted by a factor, $\lambda$,
 in order to account for the fact that infant mortality is higher in the lower Lexis 
 triangle than in the upper: of those infants who die in the first year of life, a proportion equal to
 $\lambda$ do not make it to December \nth{31} of the calendar year in which
 they were born.\footnote{$\lambda$ can be derived directly from death counts
 data classified by Lexis triangles. In the US, $\lambda$ has behaved similarly
 for males and females, falling steadily from around $0.9$ in 1969 to $0.86$
 around 1990, since which time it has steadily risen to around $0.87$. That is
 to say, $\lambda$ has varied, but not drastically. These numbers are just meant to give a
feel for the ranges that $\lambda$ can be expected to receive. If the demographer
 does not have information to derive $\lambda$ directly, ad hoc or semidirect
 methods may be used to assign a reasonable proportion. } The
 fertility component of $\textbf{Y}$ is then composed as in Matrix~\ref{matrix:Yfert}.

\begin{matrix}[h!]
\centering
\caption{Fertility component of unisex thanatological projection matrix,
$\textbf{Y}$}
\label{matrix:Yfert}
$\bordermatrix{
  {e_y } \vspace{.6em}&                0_t  & 1_t  & 2_t  & 3_t  & 4_t  & 5_t\cr 
   0_{t+1} \vspace{.6em}& (1-\lambda) \tfrac{F_0d_0}{2} & (1-\lambda) F_1d_0 & (1-\lambda)
   F_2d_0 & (1-\lambda) F_3d_0 & (1-\lambda) F_4d_0 & 0 \cr 
   1_{t+1} \vspace{.6em}& \tfrac{F_0d_1}{2} & F_1d_1 & F_2d_1 & F_3d_1 & F_4d_1
   & 0   \cr 2_{t+1} \vspace{.6em}& \tfrac{F_0d_2}{2} & F_1d_2 & F_2d_2 & F_3d_2 & F_4d_2
   & 0   \cr 3_{t+1} \vspace{.6em}& \tfrac{F_0d_3}{2} & F_1d_3 & F_2d_3 & F_3d_3 & F_4d_3
   & 0   \cr 4_{t+1} \vspace{.6em}& \tfrac{F_0d_4}{2} & F_1d_4 & F_2d_4 & F_3d_4 & F_4d_4
   & 0   \cr 5_{t+1} \vspace{.6em}& \tfrac{F_0d_5}{2} & F_1d_5 & F_2d_5 & F_3d_5 & F_4d_5
   & 0   }$
\end{matrix}

The survival and fertility components of $\textbf{Y}$ add together elementwise,
thus the full 6$\times$6 matrix is composed as in Matrix~\ref{matrix:Y}.

\begin{matrix}[h!]
\centering
\caption{A full unisex thanatological projection matrix, $\textbf{Y}$} 
\label{matrix:Y}
$\textbf{Y} = \bordermatrix{
  {e_y } \vspace{.6em} & 0_t  & 1_t  & 2_t  & 3_t  & 4_t  & 5_t\cr 
  0_{t+1} \vspace{.6em}&  (1-\lambda) \tfrac{F_0d_0}{2} & (1-\lambda) F_1d_0 + 1 &
  (1-\lambda) F_2d_0 & (1-\lambda) F_3d_0 & (1-\lambda) F_4d_0 & 0 \cr 
    1_{t+1} \vspace{.6em}& \tfrac{F_0d_1}{2} & F_1d_1 & F_2d_1 + 1 & F_3d_1 & F_4d_1 & 0 \cr 
    2_{t+1} \vspace{.6em}& \tfrac{F_0d_2}{2} & F_1d_2 & F_2d_2 & F_3d_2 + 1 & F_4d_2 & 0 \cr 
   3_{t+1} \vspace{.6em}& \tfrac{F_0d_3}{2} & F_1d_3 & F_2d_3 & F_3d_3 & F_4d_3 + 1 & 0 \cr 
   4_{t+1} \vspace{.6em}& \tfrac{F_0d_4}{2} & F_1d_4 & F_2d_4 & F_3d_4 & F_4d_4 & 1 \cr 
   5_{t+1} \vspace{.6em}& \tfrac{F_0d_5}{2} & F_1d_5 & F_2d_5 & F_3d_5 & F_4d_5 & 0 }$
\end{matrix}

Thanatological age-classes will ideally terminate at the highest value permitted
by data. For the data used here, there are 111 total age classes, which
translate to 111 total remaining-years classes (0-110+). In practice $\textbf{Y}$ becomes
a 111$\times$111 matrix, with most entries non-zero. Construction may appear
tedious for this reason. However, note that the bulk of fertility entries can
be derived as the outer (tensor) product $d_a \otimes f_y$, leaving only the 
first row and first column mortality discounting followed by the addition of the
survival superdiagonal. In most statistical programming languages constructing $\textbf{Y}$ entails only
a couple more lines of code than constructing a Leslie matrix.

As with Leslie matrices, the above projection matrix may be manipulated using
standard matrix techniques. Where $\textbf{p}$ is our population vector, we
project by multiplying $\textbf{Y}$ from the left: $\textbf{p}(t + 1) =
\textbf{Y}\textbf{p}(t)$. From $\textbf{Y}$, we can extract such information as 
the intrinsic growth rate, $r$ (natural log of the largest
real eigenvalue), the stable thanatological age-structure (the real part of the
eigenvector that corresponds to the largest real eigenvalue), or the pace of
convergence to stability (the ratio of the \nth{1} to the \nth{2} eigenvalues).
\footnote{See \citet[p.86-87]{caswell2001matrix}.}

\section*{Some empirical findings}
The renewal equation and the discrete projection matrix can be put
to work with data. At this time we are only ready to report some early
results, and we do not yet have explanations some of the patterns that we
report. Of our 1834 population-years, we optimize $r$ from both
\eqref{eq:thanoren} and \eqref{eq:lotka}\footnote{It is possible to optimize $r$ using a variety of approximations, or by using a generic optimizer. We have used the method proposed by \citet{coale1957new} for Lotka's $r$ and a modified version of the same for
the thanatological $r$. Details and / or implementation code available on
request.}. In this sample, both versions of $r$ are only plausibly equal in a
single instance. Usually thantological $r$ is greater than Lotka's $r$ (1373
cases). When Lotka's $r$ is positive (693 cases), thanatological $r$ is greater
just over of 50\% of the time (356), but when Lotka's $r$ is negative,
thanatological $r$ is the greater of the two around 90\% of the time (1017
cases). These two approximations of $r$ are of opposite sign 138 times. We
provide a comparison of the $r$ distributions in Figure~\ref{fig:rDist}. Mean
locations for each distribution are indicated with vertical dashed lines; thano.
-0.0010; chrono. -0.0027. The distribution of thanatological $r$ is more compact
than, with the ratio of variances (thano./chrono.) of about 0.75. Usually the
two theoretical values of $r$ move in the same direction, but thanatological $r$
is over time the less erratic of the two, and it usually paints a less dire picture when both are negative.

\begin{figure}[h!]
	\caption{Distribution of $r$, chronological (Lotka) and thanatological*.}
	\begin{center}
		\label{fig:rDist}
		\includegraphics[scale=.7]{Figures/rDist.pdf}
	\end{center}
	\begin{tiny}
     * Data from HMD and HFD. Countries and years listed in
     Figure~\ref{fig:Fxcompare}.
	\end{tiny}
\end{figure}

From the projection matrix, the ratio of the largest to the second largest
eigenvalue, the damping ratio, is an indicator of the \texit{springiness} of the
stable population structure, $c_y$ or $c_a$, with respect to disturbances from the stable state
as determined by vital rate distributions. A higher damping ratio means that the
population structure oscillates back to its stable state faster, i.e., that
oscillations decrease in size more rapidly. The thanatological damping ratio was
greater than the Leslie damping ratio for all 1834 population-years included in
our empirical analysis. Leslie damping ratios ranged from 1.01258 to 1.0518,
while thanatological damping ratios ranged from 1.0455 to 1.0868.

\vspace{2em}

\appendix
\section{Unique solution for thanatological $r$}
\label{app:A}
The solution for the intrinsic growth rate, $r$, is unique for the case of the
thanatological renewal model, \eqref{eq:thanoren}, and can be proven so in
essentially the same fashion as those in existence for the Lotka-Euler model,
\eqref{eq:lotka}. This little proof follows that given in
\textit{pressat1972demographic}. Define a convenience function, $I(r)$, for the
integrand of \eqref{eq:thanoren} for a given $r$ and fixed $F(y)$ and $d(a)$:

\begin{equation}
I(r) = \int_{y=0}^\infty \int_{a=0}^\infty F(y) d(a+y)e^{-ra}\dd a \dd y
\end{equation}
Since the death distribution function, $d(a)$, and fertility
function, $F(y)$, are continuous and non-negative, $\lim_{r \to +\infty} I(r)
= 0$ and $\lim_{r \to -\infty} I(r)= \infty$. If $r_2 > r_1$, then $I(r_1) >
I(r_2)$. $I()$ is therefore a continuous and monotonically decreasing function
of $r$ with boundaries that include the value 1 of \eqref{eq:thanoren}, and
 necessarily only obtain this value once.

[\ldots paper to be updated and explanded\ldots]
\nocite{HMD,HFD}
% --------------------------------------------------
% bibliography
\bibliographystyle{plainnat}
  \bibliography{References}

\end{document}

### 2013-11-14 ThanoRepro/R/DataPrep.R

@@ -1,0 +1,225 @@
+source("/home/triffe/workspace/ThanoRepro/R/Functions.R")
+HMDpath <- "/home/triffe/DATA/HMD"
+HFDpath <- "/home/triffe/DATA/HFD"
+#list.files(HFDpath)
+
+Bx <- read.table(file.path(HFDpath,"birthsRR.txt"), 
+        skip = 2, as.is = TRUE, header = TRUE)
+others <- c("IRL","UKR","ESP","BLR")
+XXX <- others[1]
+BxOTHER <- do.call(rbind, lapply(others, function(XXX, HFDpath){
+                    Dat <- read.table(file.path(HFDpath,"OTHER",paste0(XXX,"birthsRR.txt")),skip=2,as.is=TRUE,header=TRUE)
+                    cbind(Code = XXX, Dat)
+                },HFDpath=HFDpath))
+Bx <- rbind(Bx, BxOTHER)
+
+Bx$Age <- as.integer(gsub(pattern = "\\+", replacement = "",(gsub(pattern = "\\-", replacement = "", Bx$Age))))
+colnames(Bx)[4] <-"Births"
+
+Ex <- read.table(file.path(HFDpath,"exposRR.txt"), 
+        skip = 2, as.is = TRUE, header = TRUE)
+
+ExOTHER <- do.call(rbind, lapply(others, function(XXX, HFDpath){
+                    Dat <- read.table(file.path(HFDpath,"OTHER",paste0(XXX,"exposRR.txt")),skip=2,as.is=TRUE,header=TRUE)
+                    cbind(Code = XXX, Dat)
+                },HFDpath=HFDpath))
+Ex <- rbind(Ex, ExOTHER)
+
+HFDcountries <- unique(Ex$Code)
+
+# pad out HFD data to make conformable with HMD data (ages 0 - 110)
+HFDpad <- function(DATAyr, colname = "Total"){
+    young       <- DATAyr[1:12, ]
+    young$Age   <- 0:11
+    young[,colname] <- 0
+    
+    old         <- rbind(young[1:11, ], DATAyr)
+    old$Age     <- 56:(55+nrow(old))
+    old[,colname]   <- 0
+    
+    fyr <- rbind(young, DATAyr, old)  
+}
+# takes a min
[…]

### 2013-11-14 ThanoRepro/R/Functions.R

@@ -1,0 +1,109 @@
+Mna0 <- compiler::cmpfun(function(M){
+            M[is.na(M)]  <- 0
+            M[is.nan(M)] <- 0
+            M
+        })
+
+Minf0 <- compiler::cmpfun(function(M){
+            M[is.infinite(M)]  <- 0
+            M
+        })
+MinfNA <- compiler::cmpfun(function(M){
+            M[is.infinite(M)]  <- NA
+            M
+        })
+
+Thano <- compiler::cmpfun(function(Px, dx){
+            ay      <- 0:110
+            dx      <- Mna0(dx)
+            # this is the refinement:
+            dx      <- (dx + c(dx[2:length(dx)],0)) / 2
+            
+            dx      <- c(dx, dx * 0) / sum(dx) # pad out with 0s
+            EDx     <- matrix(dx[col(matrix(nrow = 111, ncol = 111)) + ay], 
+                    nrow = 111, ncol = 111, dimnames = list(Ex = ay, Age = ay))
+            
+            (EDx + cbind(EDx[,2:ncol(EDx)],0)) / 2
+            
+            t(Px * Minf0(Mna0(EDx / rowSums(EDx))))
+        })
+
+# original
+ThanoSimple <- compiler::cmpfun(function(Px, dx){
+            ay      <- 0:110
+            dx      <- Mna0(dx)
+            dx      <- c(dx, dx * 0) / sum(dx) # pad out with 0s
+            EDx     <- matrix(dx[col(matrix(nrow = 111, ncol = 111)) + ay], 
+                    nrow = 111, ncol = 111, dimnames = list(Ex = ay, Age = ay))
+            t(Px * Minf0(Mna0(EDx / rowSums(EDx))))
+        })
+
+FindSmoothMode <- function(x, fy, delta = .05){
+    NA0ind <- fy == 0 | is.na(fy)
+    xy <- smooth.spline(x=x[!NA0ind],y = log(fy[!NA0ind]))
+    xy <- predict(xy, seq(min(x),max(x), by = delta))
+    xy$x[which.max(xy$y[xy$x<90])]
+}
+
+ThanoMin <- compiler::cmpfun(function(r, da, Fy, .a = .5:110.5){
+            # get the overlapped / staggered dx structure
+            N    <- length(Fy)
+            daM  <- matrix(0, ncol = N, nrow = N)
+            dai  <- da
+            for (i in 1:N){
[…]

### 2013-11-14 ThanoRepro/R/HFDspaghetti.R

@@ -1,0 +1,162 @@
+source("/home/triffe/workspace/ThanoRepro/R/Functions.R")
+# read in data and select females
+Data <- local(get(load("/home/triffe/workspace/ThanoRepro/Data/DataAll.Rdata")))
+Data <- Data[Data$Sex == "f", ]
+
+
+
+# wow, the max age-specific rate is .29! WTF?
+#max(Data$Fx)
+#Data$Fx[Data$Fx==0] <- NA
+plot(NULL, type="n",xlim = c(12,55),ylim=c(0,.3))
+lines(Data$Age, Data$Fx, col = "#00000010")
+
+
+# ----------------------------------------------
+
+Data$Fx[Data$Fx == 0] <- NA
+Data$Fy[Data$Fy == 0] <- NA
+
+par(mfrow=c(1,2))
+# -----------------------------------------------------
+# all data
+plot(NULL, type="n",xlim = c(0,95),ylim=c(0,.15))
+lines(Data$Age, Data$Fy, col = "#00000005")
+
+plot(NULL, type="n",xlim = c(12,55),ylim=c(0,.3))
+lines(Data$Age, Data$Fx, col = "#FF000005")
+
+# -----------------------------------------------------
+# pre war
+plot(NULL, type="n",xlim = c(12,55),ylim=c(0,.3))
+lines(with(Data, Age[Year > 1920 & Year < 1939]), with(Data, Fx[Year > 1920 & Year < 1939]), col = "#00000050")
+
+plot(NULL, type="n",xlim = c(0,95),ylim=c(0,.15))
+lines(with(Data, Age[Year > 1920 & Year < 1939]), with(Data, Fy[Year > 1920 & Year < 1939]), col = "#00000050")
+
+# -----------------------------------------------------
+# post war to 1970
+plot(NULL, type="n",xlim = c(12,55),ylim=c(0,.3))
+lines(with(Data, Age[Year > 1945 & Year < 1970]), with(Data, Fx[Year > 1945 & Year < 1970]), col = "#00000010")
+
+plot(NULL, type="n",xlim = c(0,95),ylim=c(0,.15))
+lines(with(Data, Age[Year > 1945 & Year < 1970]), with(Data, Fy[Year > 1945 & Year < 1970]), col = "#00000010")
+
[…]

### 2013-11-14 ThanoRepro/R/MatrixStuff.R

@@ -1,0 +1,63 @@
+library(popbio)
+# 
+source("/home/triffe/workspace/ThanoRepro/R/Functions.R")
+
+Data <- local(get(load("/home/triffe/workspace/ThanoRepro/Data/DataAll.Rdata")))
+Data <- Data[Data$Sex == "f", ]
+
+lambdaT <- local(get(load("/home/triffe/workspace/ThanoRepro/Data/lambda.Rdata")))
+lambdaL <- local(get(load("/home/triffe/workspace/ThanoRepro/Data/lambdaLeslie.Rdata")))
+# made in commented-out code below:
+#AllMatrices <- local(get(load("/home/triffe/workspace/ThanoRepro/Data/AllMatrices.Rdata")))
+#print(object.size(AllMatrices),units = "Mb")
+#AllMatrices <- lapply(split(Data, with(Data, paste(Code,Year))), function(Dat, lambdaL, lambdaT){
+#           yr <- unique(Dat$Year)
+#           Cd <- unique(Dat$Code)
+#           N  <- nrow(Dat)
+#           px <- Mna0(Dat$lx[2:N] / Dat$lx[1:(N-1)])
+#           lL <- lambdaL$lambda.f[lambdaL$Code == Cd & lambdaL$Year == yr]
+#           lT <- lambdaT$lambda.f[lambdaL$Code == Cd & lambdaL$Year == yr]
+#         
+#           list(Y = ThanoProjMatrix(Mna0(Dat$Fyf), Dat$dx, lT),
+#               L = Leslie(Mna0(Dat$Fxf), px, lL))
+#        },lambdaL = lambdaL, lambdaT = lambdaT)
+# best to save this out, as it's quite big and take a while to calculate:
+#save(AllMatrices, file="/home/triffe/workspace/ThanoRepro/Data/AllMatrices.Rdata")
+IDs <- names(AllMatrices)
+
+# calculate Damping Ratios --- takes a while
+# this is a recursive apply, because we have a nested list
+Damping <- rapply(AllMatrices, function(X){
+            eigen.analysis(X)$damping.ratio
+        })
+Damping <- matrix(Damping,ncol=2,byrow=TRUE,dimnames = list(IDs, c("DThano","DLeslie")))
+# 100% of sample shows higher Damping ratio for thanatological matrix than for Leslie
+sum(Damping[,1] > Damping[,2])
+
[…]

### 2013-11-14 ThanoRepro/R/Oscillation.R

@@ -1,0 +1,51 @@
+
+# seems like oscillation comparisons should be done on standard populations.
+
+# could start with Px, get Py, and then derive Cohen's measure, but this adds an initial smoothing
+# to Py. Instead I'll start with identical but randomly generated starting populations.
+AllMatrices <- local(get(load("/home/triffe/workspace/ThanoRepro/Data/AllMatrices.Rdata")))
+
+CohenA <- compiler::cmpfun(function(P0, LorY, N=500){
+    Pops <- matrix(nrow=length(P0), ncol = N)
+    Pops[,1] <- P0
+    for (i in 2:N){
+        Pops[,i] <- c(LorY %*% Pops[, i - 1])
+    }
+    # make sum to 1
+    Pops <- t(t(Pops) / colSums(Pops)) 
+    sum(abs(Pops - Pops[,N]))/2
+})
+
+P0 <- runif(111)
+CohenA(P0, AllMatrices[[1]][[2]])
+
+library(parallel)
+P0mat <- matrix(runif(111*1e3),ncol=1e3)
+
+
+apply(P0mat,2,function(P0, AY){
+       CohenA(P0, AY[["L"]]) / CohenA(P0, AY[["Y"]])
+        }, AY = AllMatrices[[1]])
+
+extremeCase1 <- c(1,rep(0,110))
+CohenA(extremeCase1, AllMatrices[[1]][["L"]]) / CohenA(extremeCase1, AllMatrices[[1]][["Y"]])
+CohenA(extremeCase1, AllMatrices[[1]][["L"]]) / CohenA(rev(extremeCase1), AllMatrices[[1]][["Y"]])
+
+extremeCase2 <- c(rep(0,15),1,rep(0,95))
+CohenA(extremeCase2, AllMatrices[[1]][["L"]]) / CohenA(extremeCase2, AllMatrices[[1]][["Y"]])
+CohenA(extremeCase2, AllMatrices[[1]][["L"]]) / CohenA(rev(extremeCase2), AllMatrices[[1]][["Y"]])
+
+extremeCase3 <- rep(1,111)
+CohenA(extremeCase3, AllMatrices[[1]][["L"]]) / CohenA(extremeCase3, AllMatrices[[1]][["Y"]])
+
+
+Y <- AllMatrices[[1]][["Y"]]
+L <- AllMatrices[[1]][["L"]]
+plot(Y %*% (Y %*% extremeCase3))
+plot(L %*% (L %*% extremeCase3))
+
+
+
+
+
+

### 2013-11-14 ThanoRepro/R/PlotFunctions.R

@@ -1,0 +1,154 @@
+
+makeVectorAgeGroups <- function(vec, ages = 0:110, N = 2){
+    x.new  <- ages - ages %% N
+    c(tapply(vec, x.new, sum))
+}
+# draw rect based on x,y of cetroid, plus width and height
+makeRect <- function(x, y, w, h, ...){
+    w2 <- w / 2
+    h2 <- h / 2
+    rect(x - w2, y - h2, x + w2, y + h2, ...)
+}
+
+PyramidOutline2 <- function(males, females, 
+        scale = sum(c(males, females)), 
+        x = 0, y = 0, ...){
+    N       <- length(males)
+    Total   <- sum(c(males, females), na.rm = TRUE)
+    widths  <- rep(1, N)
+    age     <- c(0,cumsum(widths)[-N])
+    u.age   <- age[N] + widths[N]
+    
+    males   <- scale * males / Total
+    females <- scale * females / Total
+    
+    polygon(x = c(0, rep(females, each = 2) + 0,0, rev(c(-0, rep(-males, each = 2) - 0, -0))) + x, 
+            y =  c(rep(c(age, u.age), each = 2), rev(c(rep(c(age, u.age), each = 2)))) + y, ...)
+}
+
+# Males, Females should be the full thano x chrono population matrix, as returned by Thano()
+# Pyramid for age-structure. FALSE for thanatological age structure instead
+# N interval width for remaining-years heterogeneity (age pyramid) or age heterogeneity (thano leaf)
+# BrewerPal the colo ramp from color brewer package.
+# revcol you can decide direction of ramp explicitly
+# x,y for positioning of pyramid
+# ycex, for vertical scaling.
+# ---------------------
+# *only does pyramids / leaves scaled to N  <- 1
+PyramidOrLeafWithHeterogeneity <- function(Males, Females, Pyramid = TRUE,
+        N = 10, BrewerPal = "YlGnBu", revcol = FALSE,x=0,y=0){
+    # define color functions
+    if (revcol){
+        colR        <- grDevices::colorRampPalette(rev(RColorBrewer::brewer.pal(9, BrewerPal)), space = "Lab")
+    } else {
+        colR        <- grDevices::colorRampPalette(RColorBrewer::brewer.pal(9, BrewerPal), space = "Lab")
+    }
+    # if we're plotting a thanatological leaf
+    if (!Pyramid){
[…]

### 2013-11-14 ThanoRepro/R/Pyramids.R

@@ -1,0 +1,60 @@
+
+Data <- local(get(load("/home/triffe/workspace/ThanoRepro/Data/DataAll.Rdata")))
+source("/home/triffe/workspace/ThanoRepro/R/Functions.R")
+source("/home/triffe/workspace/ThanoRepro/R/PlotFunctions.R")
+
+# Produces Figures 1a and 1b... and some other fun stuff.
+
+
+US2010 <- GetThanoMatrices(.Code = "USA", .Year = 2010, Data = Data)
+
+percent.labs <- c("1.0%","0.8%","0.6%","0.4%","0.2%","","0.2%","0.4%","0.6%","0.8%","1.0%")
+
+pdf("/home/triffe/workspace/ThanoRepro/Figures/US2010Age.pdf",width=4.8,height=4.5)
+par(xaxs = "i", yaxs = "i", mai = c(.4,.4,.4,.8), xpd = TRUE)
+plot(NULL, type = 'n', xlim = c(-.01,.01),ylim = c(0,111), axes = FALSE, xlab = "", ylab = "",
+        panel.first = list(rect(-.01,0,.01,111,col = gray(.97),border = NA),
+                           segments(-.01,seq(0,110,by=10),.01,seq(0,110,by=10),col="white",lwd=.8),
+                           segments(-.01,seq(0,110,by=10),-.0102,seq(0,110,by=10),lwd=.8),
+                           segments(seq(-.01,.01,by=.002),0,seq(-.01,.01,by=.002),111,col="white",lwd=.8),
+                           segments(seq(-.01,.01,by=.002),0,seq(-.01,.01,by=.002),-1,lwd=.8),
+                           text(seq(-.01,.01,by=.002),-1,percent.labs,pos=1,cex=.7),
+                           text(-.0102,seq(0,110,by=10),seq(0,110,by=10),pos=2,cex=.7)
+                           ))
+PyramidOrLeafWithHeterogeneity(Males = US2010$Males, Females = US2010$Females, revcol = TRUE)
+colorStrip(BrewerPal = "YlGnBu",w=.0015,h=80,lwd=.8,border=gray(.7),rising = TRUE,revcol=FALSE)
+text(x=.01,95,"Remaining\nyears\ngroups",pos=4,cex=.8)
+text(-.0125,117,"Age",pos=4,cex=.8)
+dev.off()
+
+
[…]

### 2013-11-14 ThanoRepro/R/UniqueProofTests.R

@@ -1,0 +1,11 @@
+source("/home/triffe/workspace/ThanoRepro/R/Functions.R")
+Data <- local(get(load("/home/triffe/workspace/ThanoRepro/Data/DataAll.Rdata")))
+Data <- Data[Data$Sex == "f", ]
+
+# 
+Dat <- Data[Data$Year == 2000 & Data$Code == "SWE",]
+Mna0(Minf0(rowSums(Thano(Dat$Births, Dat$dx)) / 
+                        rowSums(Thano(Dat$Exposure, Dat$dx))))   -
+
+plot(sapply(seq(-.015,-.005,length.out=100),ThanoMin,da=Dat$dx, Fy=Dat$Fyf), type = 'l')
+

### 2013-11-14 ThanoRepro/R/rCalc.R

@@ -1,0 +1,163 @@
+source("/home/triffe/workspace/ThanoRepro/R/Functions.R")
+library(data.table)
+# read in data and select females
+Data <- local(get(load("/home/triffe/workspace/ThanoRepro/Data/DataAll.Rdata")))
+Data <- Data[Data$Sex == "f", ]
+Data$Fy[is.na(Data$Fy)] <- 0
+Data$Fyf[is.nan(Data$Fyf)] <- 0
+DATA <- data.table(Data)
+
+# functions to optimize r quickish
+#  thanatological age 
+rThanoCoale <- compiler::cmpfun(function(Fy, da, .a = .5:110.5, maxit = 1e2, tol = 1e-15){  
+    
+    # Based on Coale (1957), modified.
+    N    <- length(Fy)
+    dxM  <- matrix(0, ncol = N, nrow = N)
+    dxi  <- da
+    for (i in 1:N){
+        dxM[i, 1:length(dxi)  ] <- dxi 
+        dxi <- dxi[2:length(dxi) ]
+    }     
+    R0      <- sum(dxM * Fy)
+    T.guess <- sum(.a * dxM * Fy) / R0 # assuming r = 0
+    r.i     <- log(R0) / T.guess
+    # be careful to discount Fy by SRB appropriately for males / females
+    # prior to specification
+    for (i in 1:maxit){ # 15 is more than enough!
+        #cat(r2,i,"\n")
+        deltai <- 1 - sum(rowSums(t(t(dxM) * exp(-r.i * .a))) * Fy)
+        # the mean generation time self-corrects 
+        # according to the error produced by the Lotka equation
+        r.i <- r.i - (deltai / (T.guess - (deltai / r.i))) 
+        if (abs(deltai) <= tol){
+            break
+        }
+    }
+    return(r.i)  
+})
+# chronological age
+Rmomentn <- compiler::cmpfun(function(fx,Lx,x,n=0){
+            sum((x^n)*fx*Lx)
+        })
+rLotkaCoale <- compiler::cmpfun(function(fx,Lx,x=.5:110.5, maxit = 1e2, tol = 1e-15){
[…]

### 2013-11-15 ThanoRepro/ThanoRepro.tex
.

@@
 \label{eq:transform}
-P_y =& \sum_{x=0}^\omega P_x \frac{d_{x+y}}{l_x}
+P_y =& \sum_{x=0}^\omega P_x \frac{d_{x+y}+d_{x+y+1}}{l_x+l_{x+1}}
 \intertext{and of course}
@@
 transformation\footnote{I have on good authority that Ken Wachter
-coined the term thanatological age.}, and this complements the notion of
-chronological age. It bears repeating that thanatological age is merely probabilistic, since it
+coined the term \textit{thanatological} age.}, and this complements the notion
+of chronological age. It bears repeating that thanatological age is merely probabilistic, since it
 refers to the future according to particular assumptions about the survival
@@
 
+In general, such fertility rates will not be palatable for purposes of
+projection, unless the demographer believes that their empirical regularity is
+somehow stronger than typical ASFR --- an argument we do not make. There may be
+alternative ways of defining thanatological fertility rates, for instance
+\textit{after} stability, which would make this model component more conformable
+with the Lotka renewal model. First let us define the basic model.
+
 \section*{The thanatological renewal model}
@@
 of fertility rates shown in Figure~\ref{fig:FySpaghetti}, and we refer to a
-unisex population. Assuming constant vital rates, the births for the present year are given by
-
+unisex population. Assuming constant vital rates, the births for the present
+year are given by:
 \begin{align}
@@
 There are gaps in the above line of development, since the jump from
-\eqref{eq:ergostep1} to \eqref{eq:ergostep2} (strong ergodicity) is unproven. A
-further conjecture is that the jump will also hold for the case of weak
-ergodicity. The mechanisms at play unfold in the same way as those so
-intuitively described by \citet{arthur1982ergodic}, and said proof may apply
-here without further modification. A proof of
-the uniqueness of the solution to \eqref{eq:thanoren} is given in Appendix~\ref{app:A}.
+\eqref{eq:ergostep1} to \eqref{eq:ergostep2} (strong ergodicity) is unproven,
+although it is rather intuitive, given the much greater density of connections
+within the model; persons from nearly any thanatologcal age can produce
+offspring that can have any other thanatological age. In this sense, the
+smoothing mechanism must be much stronger than that for chronoloical age, at
+least in most cases. A further conjecture is that the jump will also hold for
+the case of weak ergodicity.
+The mechanisms at play unfold in the same way as those so intuitively described by \citet{arthur1982ergodic}, and said proof may apply here without further modification.
+A proof of the uniqueness of the solution to \eqref{eq:thanoren} is given in Appendix~\ref{app:A}.
 
@@
 
-What should one imagine under the model of thanatological population renewal? A
-useful mnemonic bases itself on Figures~\ref{fig:USdecomp} and
+The thanatological renewal model is coherent, but the idea may seem strange.
+What should one imagine under the model of thanatological population renewal? 
+A useful mnemonic bases itself on Figures~\ref{fig:USdecomp} and
 \ref{fig:USrecomp}. In the age-structured model, new generations
-appear at the bottom of the pyramid, and move up one rung per year. All ages are subject to decrement. In the thanatological \textit{leaf}, each birth cohort increments to
+appear at the bottom of the pyramid, and move up one rung per year. All ages 
+are subject to decrement. In the thanatological \textit{leaf}, each birth cohort increments to
 the population over the whole range of thanatological age according to $d(a)$,
@@
 
+* this is a work in progress, and by the time the EPC comes around I'll
+have shifted the focus away on comparing stable results estimated from data
+(the last couple figures), and more toward the reversibility of chronological
+and thanatological stable age structures, as well as a comparative
+eigen-analysis of the Leslie and thanatological projection matrices.
+
 \vspace{2em}
@@
 
-[\ldots paper to be updated and explanded\ldots]
 \nocite{HMD,HFD}

### 2013-11-15 ThanoRepro/R/DepRatio.R

@@ -1,0 +1,41 @@
+
+# note, Robert's neat dependency ratio parlor trick:
+
+# P65+ / P18-64 = alpha
+# how does alpha change with r?
+
+Plotka <- function(Lx, r, a){
+    sum(Lx * exp(-r*a))
+}
+xbarLotka <- function(Lx, r, a){
+    sum(a * Lx * exp(-r*a)) /  sum(Lx * exp(-r*a))
+}
+Data <- local(get(load("/home/triffe/git/ThanoRepro/ThanoRepro/Data/DataAll.Rdata")))
+
+
+Lx <- Data$Lx[with(Data, Code == "USA" & Year == 2000 & Sex == "m")]
+
+age <- 0:110
+
+r <- 0
+
+Diffa <- function(Lx, r){
+    xbarLotka(Lx[age >= 18 & age < 65], r, a=c(18.5:64.5)) -
+            xbarLotka(Lx[age >= 65], r, a=c(65.5:110.5))
+}
+alphaLotka <- function(Lx,r){
+    Plotka(Lx[age >= 65], r, a=c(65.5:110.5)) /
+            Plotka(Lx[age >= 18 & age < 65], r, a=c(18.5:64.5))
+    
+}
+r <- 0
+DRLotka(Lx,r)
+
+r <- .01
+(alphaLotka(Lx,.01) - alphaLotka(Lx,0)) / alphaLotka(Lx,0) 
+(alphaLotka(Lx,-.01) - alphaLotka(Lx,0)) / alphaLotka(Lx,0) 
+Diffa(Lx,0.0) / 100
+Diffa(Lx,.01) / 100
+Diffa(Lx,-.01) / 100
+
+

### 2014-04-22 ThanoRepro/ThanoRepro.tex
changes before final sending to discussant for PAA

@@
 \begin{abstract}
-We transform data classified by chronological age into data classified by
-remaining years of life (thanatological age). A model for population renewal and
+Aggregate data classified by chronological age can be transformed into data
+classified by remaining years of life (thanatological age). A model for population renewal and
 the corresponding projection matrix are presented for populations structured by
@@
 demographer to make predictions about the future size and structure of a population. In this
-paper, we explore some formal demographic consequences of a particular
+paper, I explore some formal demographic consequences of a particular
 redefinition of age. Instead of counting age as the time passed between birth and the present (or
@@
 life line instead of at the beginning. Of course, such information cannot be
-requested of survey respondents or register data, because one's time at death is
-typically unknown. Instead we approximate this perspective using a similar
-strategy to that used by \citet{miller2001increasing} or
+requested of survey respondents or register data, because one's future time of
+death is typically unknown. Instead we approximate this perspective using a
+strategy similar to that used by \citet{miller2001increasing} or
 \citet{lee2002approach}, which uses lifetables and aggregated
-population counts.
+population counts to redistribute counts within age-classes over future death
+times.
 
-Temporal perspectives and rescalings are abundant in demography already,
-but questions are left unanswered. Likely some demographic phenomena are best
+Temporal perspectives and rescalings of the lifecourse are abundant in
+demography already. It is my view that these efforts have enriched the
+discipline by producing new insights and questions much more than they have obscured it by
+producing conundrums. Likely some demographic phenomena are best
 described as a function of time since birth, and others of time until death,
-while still others can be a function of both to some degree. This author
-supposes that fertility and reproduction can be a function of both age
-perspectives, and that the question of how to partition forces and events over
-time is a difficult one. Thanatological reproduction sits at the opposite
-extreme of chronological reproduction, the Lotka-Euler renewal
-model. By specifying a model of thanatological renewal, we both expand the
-toolbox available to approach complex temporal structure and we provide a set of broad
-findings to help understand the implications of temporal structure.
+while still others can be a function of both to some degree. Fertility and
+reproduction are plausibly a function of both age perspectives, and the question
+of how to partition such forces and events over time is a new conundrum that I
+do not tackle here. The model presented here should therefore not be interpreted
+as a suggestion for a better way to model or project population, but as an alternative way to tell the
+story of population renewal within the confines of a philosophical bubble.
+Thanatological and chronological renewal models, such as the Lotka-Euler renewal model, both approximate the movement of population stocks over time, but from opposite vantage points.
+Specifying a model of thanatological renewal expands the toolbox available to approach complex temporal population structure.
+I present a set of broad findings to help understand the implications of temporal structure.
 
@@
 Our point of departure is a little formula that gives the probability of
-dying at some age $n$ years in the future, given survival to some
-earlier age, $a$. For age 0 this is simply the $d(a)$ column of the lifetable,
-standardized to sum to 1. For higher ages, we simply condition on survival:
+dying at some age $y$ years in the future, given survival to some
+earlier age, $a$. For age 0 this probability is simply the $d(a)$
+column of the lifetable, standardized to sum to 1. For higher ages, we simply condition on survival:
 
@@
 \label{eq:vaupel1}
-f(n | a) = \mu(a+n)\frac{l(a+n)}{l(a)}
+d(y | a) = \mu(a+y)\frac{l(a+y)}{l(a)}
 \end{equation}
@@
 A convenient discretization of \eqref{eq:vaupel1} is to simply work with the
-$d_x$ column of the lifetable
+$d_a$ column of the lifetable (here single age):
 
 \begin{equation}
-d_{x+n, x} = \frac{d_{x+n}}{\sum_{i=x}^\omega d_i }
+d_{y, a} =  \frac{d_{a+y}+d_{a+y+1}}{l_a + l_{a+1}} \quad\quad \text{,}
 \end{equation}
-where $\omega$ is the highest age.\footnote{In later formulas we just use $l_x$
-in the denominator to reduce clutter, but the above sum over $d_x$ will eliminate
-rounding error in case one is working from a published life table.} One
-produces in this way a probability density function over future death times for each age, and the $d_{x+n, x}$- weighted average of $n$ for a given age $x$ will simply produce the familiar remaining life expectancy column of the lifetable, $e_x$.
-Since $1 = \sum_{n=0}^{\omega - x}d_{x+n, x}$ for each $x$, we can use this to
-decompose age-classified demographic data. For example, we can break down a
-population pyramid into remaining years categories, like in
-Figure~\ref{fig:USdecomp}, which shows this exercise for 2010 US data. The fun starts when we proceed to regroup
-the data by $n$ instead of collapsing back to $x$. Now instead of remaining lifetime 
-heterogeneity within each age group we have age heterogeneity within each
-discrete remaining years of life. Figure~\ref{fig:USrecomp} shows this
-\textit{re}composition for the same 2010 US data.
+
+\noindent where $\omega$ is the highest age.\footnote{In later formulas we just
+use $l_a$ in the denominator to reduce clutter, but the above sum over $d_a$
+will eliminate rounding error in case one is working from a published life table.} One
+produces in this way a probability density function over future death times for
+each age, and the $d_{y,a}$- weighted average of $y$ for a given age $a$ returns
+the familiar remaining life expectancy column of the lifetable, $e_a$. Since $1
+= \sum_{n=0}^{\omega - a}d_{a+y, a}$ for each $a$, we can use this to decompose
+age-classified demographic data. For example, we can break down a population
+pyramid into remaining years categories, like in Figure~\ref{fig:USdecomp},
+which shows this exercise for 2010 US data. The fun starts when we proceed to
+regroup cross-classified data by $y$ instead of collapsing back to $a$. Now
+instead of remaining lifetime heterogeneity within each age group we have age heterogeneity within each discrete remaining years of life.
+Figure~\ref{fig:USrecomp} shows this \textit{re}composition for the same 2010 US data.
 
@@
 
-If $P_x$ is an age-classified discrete population count, we can derive $P_y$, a
+If $P_a$ is an age-classified discrete population count, we can derive $P_y$, a
 remaining-years-classified population count, as follows:
@@
 \label{eq:transform}
-P_y =& \sum_{x=0}^\omega P_x \frac{d_{x+y}+d_{x+y+1}}{l_x+l_{x+1}}
+P_y =& \sum_{a=0}^\omega P_a \frac{d_{a+y}+d_{a+y+1}}{l_a+l_{a+1}}
 \intertext{and of course}
-P =& \sum P_y = \sum P_a
+\sum P_y =& \sum P_a
 \end{align}
@@
 coined the term \textit{thanatological} age.}, and this complements the notion
-of chronological age. It bears repeating that thanatological age is merely probabilistic, since it
-refers to the future according to particular assumptions about the survival
-curve. As such, Figure~\ref{fig:USrecomp} is projective in nature, a
-forward-looking glance at a population given the age-structure and lifetable of a particular moment, whereas
-Figure~\ref{fig:USdecomp} is reflective in nature, since a population's
-age-structure is mostly the fruit of past fertility, but also migration and to a
-lesser degree attrition. In a different paper we explore population structure by remaining years of life
-for its own sake. Note that any age-classified count can be
-reclassified in this way, as long as you have a lifetable that plausibly
-represents the population whose data you wish to reclassify. This means we can
-derive thanatological fertility rates, $F_y$, by applying \eqref{eq:transform}
-to age-classified birth counts, $B_x$, to get $B_y$ and again to
-exposure-to-risk, $E_x$ (here we take exposures from all ages), to get $E_y$ and then dividing:
+of chronological age. It bears repeating that thanatological age is
+probabilistic until after death, since it refers to the future according to
+particular assumptions about mortality. As such, Figure~\ref{fig:USrecomp} is
+projective in nature, a forward-looking glance at a population attrition given the age-structure and
+lifetable of a particular moment, whereas Figure~\ref{fig:USdecomp} is reflective in nature, since a population's
+chronological age-structure is mostly the fruit of past fertility, but also
+migration and to a lesser degree attrition. I explore population structure by remaining years of life
+for its own sake in a different paper.\footnote{Poster Session 6 (Friday
+mid-day))} Note that any count classified by chronological age can be
+reclassified in this way, as long as you have a lifetable that plausibly represents the population whose data you wish to reclassify.
+This means we can derive thanatological fertility rates, $F_y$, by applying \eqref{eq:transform}
+to age-classified birth counts, $B_a$, to get $B_y$ and again to
+exposure-to-risk, $E_a$ (here we take exposures from all ages), to get $E_y$ and then dividing:
 
 \begin{equation}
-F_y = \frac{B_y}{E_y}
+f^\star_y = \frac{B_y}{E_y}
 \end{equation}
@@
 We are interested to use thanatological fertility rates, $F_y$ in
-models of population renewal, rather than to explicitly study the nature of these rates.
-Nonetheless it is best not to throw such rates into a model blindly, so we
-offer a schematic overview of some of their characteristics.
+models of population renewal,\footnote{This is probably premature, but helps to
+understand the nature and provenance of model components.} rather than to
+explicitly study the nature of these rates. Nonetheless it is best not to throw such rates into a model blindly, so we offer a schematic overview of some of their characteristics.
 Figure~\ref{fig:FySpaghetti} represents the full variety in
@@
 variety of fertility and mortality combinations.} Figure~\ref{fig:FxSpaghetti} gives ASFR for the same
-populations and years as a more familiar reference, but note that both scales are different!
-With no indication of particular populations or time series, one already concludes that thanatological fertility rates have a characteristic shape and are not random, erratic or informationless: there is a pattern to fertility rates by remaining years of life, and it varies over time and between populations within some range of normality.
-Thanatological fertility rates have a wider distribution than do chronological rates.
-Note that this spaghetti plot includes both fertility booms and busts, as
-well as some mortality crises (1918, WWII). Some patterns to note, but which we
-do not separate graphically here, are that the left tail (a gauge of how
-orphan-prone a population is) has tended to fall over time. All populations have
-shown a rightward shift in both the mean and modal female thanatological age at
-birth (TAB) over time (that's unambiguously a good thing). Several populations
-now show modal TABs of over 60 years. Thanatological total fertility rates track
-standard PTFR rather closely, but tend to be somewhat higher.
+populations and years as a more familiar reference, but note that both scales
+are different! With no indication of particular populations or time series, one
+already concludes that period thanatological fertility rates have a
+characteristic shape and are not random, erratic or informationless: There is a
+pattern to fertility rates by remaining years of life, and it varies over time
+and between populations within some range of normality. Thanatological fertility
+rates have a wider distribution than do chronological rates. Note that this
+spaghetti plot includes both fertility booms and busts, as well as some
+mortality crises (1918, WWII). Some patterns to note, but which we do not
+separate graphically here, are that the left tail (a gauge of how orphan-prone a
+population is) has tended to fall over time. All populations have shown a
+rightward shift in both the mean and modal mother's thanatological age at birth
+(TAB) over time (that's unambiguously a good thing). Several populations now
+show modal TABs of over 60 years. Thanatological total fertility rates (TTFR)
+track standard period total fertility rates (PTFR) rather closely, but tend to
+be somewhat higher (not shown).
 
@@
 somehow stronger than typical ASFR --- an argument we do not make. There may be
-alternative ways of defining thanatological fertility rates, for instance
-\textit{after} stability, which would make this model component more conformable
-with the Lotka renewal model. First let us define the basic model.
+alternative ways of defining thanatological fertility rates, for instance by
+transforming from a stable population, which would make the resulting model more
+conformable with the Lotka renewal model. First let us define the basic model.
 
+\FloatBarrier
 \section*{The thanatological renewal model}
@@
 of fertility rates shown in Figure~\ref{fig:FySpaghetti}, and we refer to a
-unisex population. Assuming constant vital rates, the births for the present
-year are given by:
+unisex population based on female vital rates. Assuming constant vital rates,
+the births for the present year are given by
 \begin{align}
-B(t) =& \int _0^\infty P(y,t)F(y) \dd y = \int _0^\infty P(a,t)F(a) \dd a
-\intertext{where $P(a), P(y)$ are population counts and $F(y), F(a)$ are exact
-specific fertility probabilities (rates), and $d(a)$ is the continuous
-lifetable death distribution with radix of 1, so $d(a)/l(0)$. The
-thanatological integral can be broken down back in terms of chronological age.} =&
-\int_{y=0}^\infty
-\int_{a=0}^\infty F(y)\frac{P(a,t)d_(a+y)}{l(a)} \dd a \dd y 
-\intertext{We can relate the present population to past
-births with $P(a,t) = B(t-a)l(a)$.} 
+B(t) =& \int _0^\infty P(y,t)f^\star(y) \dd y = \int _0^\infty P(a,t)f(a) \dd a
+\quad\text{,} \intertext{where $a$ indexes chronological age, $y$ indexes
+thanatological age, $P(a), P(y)$ are population counts, and $f^\star(y), f(a)$ are
+exact specific fertility probabilities (rates), which here are best thought of
+as the fertility rates of females born to females. The thanatological integral can
+be broken down back in terms of chronological age:} =& \int_{y=0}^\infty
+\int_{a=0}^\infty f^\star(y)\frac{P(a,t)d(a+y)}{l(a)} \dd a \dd y \quad\text{,}
+\intertext{where $d(a)$ is the continuous lifetable death distribution with
+radix of 1, so $d(a)/l(0)$. We can relate the present population to past births
+with $P(a,t) = B(t-a)l(a)$:}
 \label{eq:ergostep1}
-=& \int_{y=0}^\infty \int_{a=0}^\infty F(y) B(t-a)d(a+y)\dd a \dd y
-\intertext{And eventually-- a bit quicker than is the case for chronological
-age-- strong ergodicity will assert itself, and $B(t)$ will be related to
+=& \int_{y=0}^\infty \int_{a=0}^\infty f^\star(y) B(t-a)d(a+y)\dd a \dd y
+\quad\text{.} \intertext{Eventually-- usually a bit quicker than is the case for
+chronological age-- strong ergodicity will assert itself, and $B(t)$ will be related to
 $B(t-a)$ according to a constant factor $e^{ra}$, where $r$ is the familiar
-intrinsic rate of growth.}
+intrinsic rate of growth:}
 \label{eq:ergostep2}
-=& \int_{y=0}^\infty \int_{a=0}^\infty F(y) B(t)e^{-ra}d(a+y)\dd a \dd y
-\intertext{Divide out $B(t)$ to get back a familiar-looking renewal equation.}
+=& \int_{y=0}^\infty \int_{a=0}^\infty f^\star(y) B(t)e^{-ra}d(a+y)\dd a \dd y
+\quad\text{.}
+\intertext{Divide out $B(t)$ to get back a familiar-looking renewal equation:}
 \label{eq:thanoren}
-1 =& \int_{y=0}^\infty \int_{a=0}^\infty F(y) d(a+y)e^{-ra}\dd a \dd y
+1 =& \int_{y=0}^\infty \int_{a=0}^\infty f^\star(y) d(a+y)e^{-ra}\dd a \dd y
+\quad\text{.}
 \end{align}
@@
 \label{eq:lotka}
-1 = \int_{a=0}^\infty F(a)l(a)e^{-ra}\dd a
+1 = \int_{a=0}^\infty f(a)l(a)e^{-ra}\dd a \quad\text{,}
 \end{equation}
@@
 divergence between male and female single-sex models, there will always be divergence between
-single-sex models under chronological versus thanatological age. Even though the
-modeled population stocks are in a way commensurable, the rates used, $F(a)$
-versus $F(y)$ are calculated on the basis of differently distributed
-denominators $E(a)$ versus $E(y)$.
+single-sex models under chronological versus thanatological age, unless the
+starting population is stable. Even though the modeled population stocks are in
+a way commensurable, the rates used here, $f(a)$ versus $f^\star(y)$ are
+calculated on the basis of differently distributed denominators $E(a)$ versus $E(y)$.
 
@@
 although it is rather intuitive, given the much greater density of connections
-within the model; persons from nearly any thanatologcal age can produce
-offspring that can have any other thanatological age. In this sense, the
-smoothing mechanism must be much stronger than that for chronoloical age, at
-least in most cases. A further conjecture is that the jump will also hold for
-the case of weak ergodicity.
+within the model; persons from nearly any thanatologcal age\footnote{I say
+\textit{nearly} because the very highest thanatological ages tend to be
+composed of pre-menarchical females, so these can be thought of as structural
+zeros. I am uncertain if it is a record, but it is certainly an extreme case
+useful for illustration: Jeanne Calment, who lived to 122, had a daughter at age
+22, which means she was about thanatological age 100
+(\url{http://en.wikipedia.org/wiki/Jeanne_Calment}). If the lifetable used to
+transform to thanatological closes out at age 110+, as is the case with HMD data, then fertility rates for thantological ages 101-110+ would be structural
+zeros.} can produce offspring that can have any other thanatological age. In
+this sense, the smoothing mechanism at play as the population passes through
+time must be much stronger than that for chronological age, at least in most
+cases.
+A further conjecture is that the jump will also hold for the case of weak ergodicity.
 The mechanisms at play unfold in the same way as those so intuitively described by \citet{arthur1982ergodic}, and said proof may apply here without further modification.
@@
 parameters can be calculated. For instance, we may calculate the mean
-thanatological generation time, $T^y$, as:
+thanatological generation time, $T^y$:
 \begin{equation}
 \label{eq:Ty}
- T^y =  \frac{\int _{y=0}^\infty \int _{a=y}^\infty y e^{-ra} d(a) F(y) \dd a
-\dd y}{\int _{y=0}^\infty \int _{a=y}^\infty e^{-ra} d(a) F(y) \dd a \dd y}
+ T^y =  \frac{\int _{y=0}^\infty \int _{a=y}^\infty y e^{-ra} d(a) f^\star(y) \dd a
+\dd y}{\int _{y=0}^\infty \int _{a=y}^\infty e^{-ra} d(a) f^\star(y) \dd a \dd
+y} \quad \text{.}
 \end{equation}
-The net reproduction rate, $R_0$ is related by, e.g.,
+Literally, this is the mean remaining life expectancy of new mothers in the
+stable population. The net reproduction rate, $R_0$ is related by, e.g.,
 \begin{equation}
@@
 b = \frac{1}{\int _{y=0}^\infty \int _{a=y}^\infty e^{-ra} d(a) \dd a
-\dd y}
+\quad\text{.} \dd y}
 \end{equation}
-The stable age structure, $c$, where $c_y$ is the
-proportion of the stable population with remaining years to live $y$, is given
-by
+The stable thanatological age structure, $c_y$, is the
+proportion of the stable population with remaining years to live $y$:
 \begin{equation}
 \label{eq:cy}
-c(y) = b \int _{a=y}^\infty e^{-ra} d(a) \dd a
+c(y) = b \int _{a=y}^\infty e^{-ra} d(a) \dd a \quad\text{.}
 \end{equation}
-Other stable population quantities may be estimated by similarly translating the
-various common definitions (e.g., in the glossary of \citet{coale1972growth}) to
-the present perspective. We will focus on the main model rather than on these.
+We already knew that when $r = 0$ and $y = a$, that $c(a)=c(y)$
+\citep{vaupel2009life}, and now we can arrive at the same conclusion directly
+via the chronological and thanatological renewal equations. For small non-zero
+magnitude values of $r$ and survival patterns typical of humans, it is also the
+case that $c(y|r)\approx c(a|-r)$. The relationship is asymptotically exact as
+$r$ approaches zero. Other findings regarding stable age structure will be
+presented in a future version of this paper. Other stable population quantities
+may be estimated by similarly translating the various common definitions (e.g.,
+in the glossary of \citet{coale1972growth}) to the present perspective. We will
+focus on the main model rather than on these.
 
-The thanatological renewal model is coherent, but the idea may seem strange.
-What should one imagine under the model of thanatological population renewal? 
-A useful mnemonic bases itself on Figures~\ref{fig:USdecomp} and
-\ref{fig:USrecomp}. In the age-structured model, new generations
-appear at the bottom of the pyramid, and move up one rung per year. All ages 
-are subject to decrement. In the thanatological \textit{leaf}, each birth cohort increments to
-the population over the whole range of thanatological age according to $d(a)$,
-as seen in \eqref{eq:thanoren}. Birth cohorts become the layers seen in
-Figure~\ref{fig:USrecomp}. Each horizontal step is a death cohort, and these
-move one step down the pyramid each year without any decrement (indeed
-incrementing) until reaching the very bottom. In short, the locations of
-increment and decrement, and the direction of movement are all switched.
+The thanatological renewal model is coherent, but these ideas may take some time
+to gestate before accepting that it is also a valid description of how
+populations renew. What should one imagine under the model of thanatological
+population renewal? A useful mnemonic bases itself on Figures~\ref{fig:USdecomp}
+and \ref{fig:USrecomp}. In the chronological age-structured model, new
+generations appear at the bottom of the pyramid, and move up one rung per year.
+All age-classes are subject to attrition, which is spread out over ages and not
+readily visible in the pyramid. In the thanatological \textit{leaf}, each birth cohort increments to the population over the whole range of thanatological age
+according to $d(a)$, as seen in \eqref{eq:thanoren}, becoming the
+shaded layers seen in Figure~\ref{fig:USrecomp}. Each horizontal step is a death
+cohort, and these move one step down the pyramid each year without any decrement
+(indeed incrementing due to births) until reaching the very bottom. In short,
+the locations of increment and decrement, and the direction of movement are all
+switched. The chronological and thanatological renewal models are almost
+perfectly opposite descriptions of the same two processes.
 
 \section*{The thanatological projection matrix}
-These descriptions can be made more explicit by hashing out the projection
-matrix that corresponds to the thanatological renewal model. As with the
-age-structured Leslie matrix, the thanatological projection matrix,
-$\textbf{Y}$, is square and of dimension $n \times n$, where $n$ is the number
-of remaining-years classifications into which the population is divided. 
+These descriptions can be made more explicit, and in some ways more tractable,
+by hashing out the projection matrix that corresponds to the thanatological renewal model. As with the
+chronological age-structured Leslie matrix, the thanatological projection
+matrix, $\textbf{Y}$, is square and of dimension $y \times y$, where $y$ is the number
+of remaining-years classifications on which the population is structured. 
 The matrix contains elements for survival and elements for fertility. Unlike
-Leslie matrices, $\textbf{Y}$ is not sparse, but is populated primarily with non-zero entries.
+Leslie matrices, $\textbf{Y}$ is not sparse, but is instead populated primarily
+with non-zero entries. This is an odd notion, to be explained in following.
 
@@
 $t + 1$. Thus, instead of in the subdiagonal, we place survival in the
-superdiagonal. All survival values are 1, since there is no decrement, and the
-upper-left corner of this superdiagonal contains a 0, for full decrement.
-We illustrate using a 6$\times$6 matrix. The survival component of $\textbf{Y}$ is organized as in
-Matrix~\ref{matrix:Ysurv}.
+superdiagonal. All survival values are 1, since there is no decrement until
+after living through age 0, and the upper-left corner of this superdiagonal
+contains a 0, for full decrement. In the following, I illustrate using a
+6$\times$6 matrix. The survival component of $\textbf{Y}$ is organized as in
+Matrix~\ref{matrix:Ysurv}. To imagine a lifecycle graph, each age decrements by
+one through each lower age in successive order until being absorbed by death
+after completing thanatological age zero. This chain unfolds in the opposite
+index-direction of survival in the chronological age system.
 
@@
  Fertility inputs to the matrix are derived from single-sex thanatological
- fertility and the lifetable $d_a$ distribution, where $a$ indexes age, but is
- equal to $y$, remaining years of life for persons aged 0. Fertility
+ fertility and the lifetable $d_a$ distribution, where $a$ indexes age and is
+ equated to $y$ remaining years of life for members just born. Fertility
  in a thanatologically structured population occurs in all but the highest
- remaining years classes. For our example, say that fertility is observed in
- classes 0-4, while the final class has no fertility, where $F_y$ indicates the
- fertility probability for class $y$ in the year $t$ entering population (in the 
- matrix columns). Each $F_y$ is then distributed according to
- $d_a$ with no further translation, since the $d_a$ lifetable column already
- refers to age 0. Thus, the fertility entry in row $m$ and column $n$ of
- $\textbf{Y}$ will be $F_n \cdot d_m$. We assume that those dying over the course of year
+ remaining years classes, which can be thought of as structural zeros for
+ simplicity, as they only contain pre-menarchical females. For our
+ $6\times 6$ example, say that fertility is observed in classes 0-4, while the
+ final class has no fertility, where $f^\star_y$ indicates the fertility
+ probability for class $y$ in the year $t$ entering population (in the matrix
+ columns). Each $f^\star_y$ is then distributed out according to $d_a$ with no
+ further translation, since the $d_a$ lifetable column already refers to age 0.
+Thus, the fertility entry in row $m$ and column $n$ of $\textbf{Y}$ will be $f^\star_n \cdot d_m$.
+We assume that those dying over the course of year
  $t$ (the first column) are exposed to fertility for half of the
@@
  and so discount the fertility entry accordingly. Further, infant mortality, 
- $F_y \cdot d_0$, located in the first row, must also be discounted, since part
+ $f^\star_y \cdot d_0$, located in the first row, must also be discounted, since part
  of the mortality will occur in the same year $t$ and the rest in year $t + 1$. 
@@
   {e_y } \vspace{.6em}&                0_t  & 1_t  & 2_t  & 3_t  & 4_t  & 5_t\cr 
-   0_{t+1} \vspace{.6em}& (1-\lambda) \tfrac{F_0d_0}{2} & (1-\lambda) F_1d_0 & (1-\lambda)
-   F_2d_0 & (1-\lambda) F_3d_0 & (1-\lambda) F_4d_0 & 0 \cr 
-   1_{t+1} \vspace{.6em}& \tfrac{F_0d_1}{2} & F_1d_1 & F_2d_1 & F_3d_1 & F_4d_1
-   & 0   \cr 2_{t+1} \vspace{.6em}& \tfrac{F_0d_2}{2} & F_1d_2 & F_2d_2 & F_3d_2 & F_4d_2
-   & 0   \cr 3_{t+1} \vspace{.6em}& \tfrac{F_0d_3}{2} & F_1d_3 & F_2d_3 & F_3d_3 & F_4d_3
-   & 0   \cr 4_{t+1} \vspace{.6em}& \tfrac{F_0d_4}{2} & F_1d_4 & F_2d_4 & F_3d_4 & F_4d_4
-   & 0   \cr 5_{t+1} \vspace{.6em}& \tfrac{F_0d_5}{2} & F_1d_5 & F_2d_5 & F_3d_5 & F_4d_5
+   0_{t+1} \vspace{.6em}& (1-\lambda) \tfrac{f^\star_0d_0}{2} & (1-\lambda) f^\star_1d_0 & (1-\lambda)
+   f^\star_2d_0 & (1-\lambda) f^\star_3d_0 & (1-\lambda) f^\star_4d_0 & 0 \cr 
+   1_{t+1} \vspace{.6em}& \tfrac{f^\star_0d_1}{2} & f^\star_1d_1 & f^\star_2d_1 & f^\star_3d_1 & f^\star_4d_1
+   & 0   \cr 2_{t+1} \vspace{.6em}& \tfrac{f^\star_0d_2}{2} & f^\star_1d_2 & f^\star_2d_2 & f^\star_3d_2 & f^\star_4d_2
+   & 0   \cr 3_{t+1} \vspace{.6em}& \tfrac{f^\star_0d_3}{2} & f^\star_1d_3 & f^\star_2d_3 & f^\star_3d_3 & f^\star_4d_3
+   & 0   \cr 4_{t+1} \vspace{.6em}& \tfrac{f^\star_0d_4}{2} & f^\star_1d_4 & f^\star_2d_4 & f^\star_3d_4 & f^\star_4d_4
+   & 0   \cr 5_{t+1} \vspace{.6em}& \tfrac{f^\star_0d_5}{2} & f^\star_1d_5 & f^\star_2d_5 & f^\star_3d_5 & f^\star_4d_5
    & 0   }$
 \end{matrix}
+\FloatBarrier
 
@@
   {e_y } \vspace{.6em} & 0_t  & 1_t  & 2_t  & 3_t  & 4_t  & 5_t\cr 
-  0_{t+1} \vspace{.6em}&  (1-\lambda) \tfrac{F_0d_0}{2} & (1-\lambda) F_1d_0 + 1 &
-  (1-\lambda) F_2d_0 & (1-\lambda) F_3d_0 & (1-\lambda) F_4d_0 & 0 \cr 
-    1_{t+1} \vspace{.6em}& \tfrac{F_0d_1}{2} & F_1d_1 & F_2d_1 + 1 & F_3d_1 & F_4d_1 & 0 \cr 
-    2_{t+1} \vspace{.6em}& \tfrac{F_0d_2}{2} & F_1d_2 & F_2d_2 & F_3d_2 + 1 & F_4d_2 & 0 \cr 
-   3_{t+1} \vspace{.6em}& \tfrac{F_0d_3}{2} & F_1d_3 & F_2d_3 & F_3d_3 & F_4d_3 + 1 & 0 \cr 
-   4_{t+1} \vspace{.6em}& \tfrac{F_0d_4}{2} & F_1d_4 & F_2d_4 & F_3d_4 & F_4d_4 & 1 \cr 
-   5_{t+1} \vspace{.6em}& \tfrac{F_0d_5}{2} & F_1d_5 & F_2d_5 & F_3d_5 & F_4d_5 & 0 }$
+  0_{t+1} \vspace{.6em}&  (1-\lambda) \tfrac{f^\star_0d_0}{2} & (1-\lambda) f^\star_1d_0 + 1 &
+  (1-\lambda) f^\star_2d_0 & (1-\lambda) f^\star_3d_0 & (1-\lambda) f^\star_4d_0 & 0 \cr 
+    1_{t+1} \vspace{.6em}& \tfrac{f^\star_0d_1}{2} & f^\star_1d_1 & f^\star_2d_1 + 1 & f^\star_3d_1 & f^\star_4d_1 & 0 \cr 
+    2_{t+1} \vspace{.6em}& \tfrac{f^\star_0d_2}{2} & f^\star_1d_2 & f^\star_2d_2 & f^\star_3d_2 + 1 & f^\star_4d_2 & 0 \cr 
+   3_{t+1} \vspace{.6em}& \tfrac{f^\star_0d_3}{2} & f^\star_1d_3 & f^\star_2d_3 & f^\star_3d_3 & f^\star_4d_3 + 1 & 0 \cr 
+   4_{t+1} \vspace{.6em}& \tfrac{f^\star_0d_4}{2} & f^\star_1d_4 & f^\star_2d_4 & f^\star_3d_4 & f^\star_4d_4 & 1 \cr 
+   5_{t+1} \vspace{.6em}& \tfrac{f^\star_0d_5}{2} & f^\star_1d_5 & f^\star_2d_5 & f^\star_3d_5 & f^\star_4d_5 & 0 }$
 \end{matrix}
@@
 
-\section*{Some empirical findings}
-The renewal equation and the discrete projection matrix can be put
-to work with data. At this time we are only ready to report some early
+\section*{Some preliminary empirical findings}
+The thanatological renewal equation and the discrete projection matrix can be
+put to work with data. At this time we are only ready to report some early
 results, and we do not yet have explanations some of the patterns that we
@@
 -0.0010; chrono. -0.0027. The distribution of thanatological $r$ is more compact
-than, with the ratio of variances (thano./chrono.) of about 0.75. Usually the
-two theoretical values of $r$ move in the same direction, but thanatological $r$
+than for chronological $r$, with the ratio of variances (thano./chrono.) of
+about 0.75. Usually the two theoretical values of $r$ move in the same direction, but thanatological $r$
 is over time the less erratic of the two, and it usually paints a less dire picture when both are negative.
@@
 
+The differences described here are not due primarily to rounding errors in the
+data, but rather to the way in which $f^\star$ has been calculated. If the
+births and exposure used to calculate $f^\star$ come from a population that is not stable
+then $f^\star$ may be subject to some compositional ruses. Specifically, in this
+case we have transformed fertility rates from empirical data with a particular
+chronological age structure. If the chronological age structure was far from its
+stable distribution, then it is hard to imagine that $f^\star$ would remain
+fixed over time. Some simple contrived examples will be presented in a future
+version of this paper in order to eliminate this problem.
+
 From the projection matrix, the ratio of the largest to the second largest
@@
 our empirical analysis. Leslie damping ratios ranged from 1.01258 to 1.0518,
-while thanatological damping ratios ranged from 1.0455 to 1.0868.
+while thanatological damping ratios ranged from 1.0455 to 1.0868. Again, this
+was for the raw data, and I am not sure whether this consistent finding is a
+natural property of the model or an artifact of the unstable starting state of
+the data used to derive thanatological fertility rates. Intuitively, the
+thanatological smoothing process should be an order of magnitude stronger than
+the chronological smoothing process because each thanatological age class
+(node) -- except for the oldest-old -- is connected directly to each other
+thanatological age class in both directions. Again, I think contrived examples
+will be necessary to shed more light on this finding.
 
@@
 \textit{pressat1972demographic}. Define a convenience function, $I(r)$, for the
-integrand of \eqref{eq:thanoren} for a given $r$ and fixed $F(y)$ and $d(a)$:
+integrand of \eqref{eq:thanoren} for a given $r$ and fixed $f^\star(y)$ and $d(a)$:
 
 \begin{equation}
-I(r) = \int_{y=0}^\infty \int_{a=0}^\infty F(y) d(a+y)e^{-ra}\dd a \dd y
+I(r) = \int_{y=0}^\infty \int_{a=0}^\infty f^\star(y) d(a+y)e^{-ra}\dd a \dd y
 \end{equation}
 Since the death distribution function, $d(a)$, and fertility
-function, $F(y)$, are continuous and non-negative, $\lim_{r \to +\infty} I(r)
+function, $f^\star(y)$, are continuous and non-negative, $\lim_{r \to +\infty} I(r)
 = 0$ and $\lim_{r \to -\infty} I(r)= \infty$. If $r_2 > r_1$, then $I(r_1) >

### 2014-04-22 ThanoRepro/R/DepRatio.R

@@ -1,8 +1,29 @@
-
 # note, Robert's neat dependency ratio parlor trick:
 
 # P65+ / P18-64 = alpha
 # how does alpha change with r?
+
+# change in alpha with r = (xbar18-64 - xbar65+) / 100
+# (or something like that)
+
+
+# example data, HMD, I forget which country, year, sex.... legit tho
+Lx <- c(0.99273, 0.99194, 0.99147, 0.99114, 0.99089, 0.99068, 0.99048, 
+        0.99031, 0.99015, 0.98999, 0.98982, 0.98963, 0.98942, 0.98918, 
+        0.98888, 0.98846, 0.98786, 0.98705, 0.98601, 0.98478, 0.98347, 
+        0.9821, 0.98074, 0.97941, 0.97809, 0.9768, 0.97556, 0.97431, 
+        0.97304, 0.97176, 0.97047, 0.96916, 0.96779, 0.96634, 0.96476, 
+        0.9631, 0.96135, 0.95946, 0.9574, 0.9552, 0.95283, 0.95027, 0.94753, 
+        0.94453, 0.94127, 0.93777, 0.93401, 0.92996, 0.92553, 0.9208, 
+        0.91584, 0.91063, 0.9051, 0.899, 0.89265, 0.88585, 0.87828, 0.87005, 
+        0.86113, 0.85155, 0.8412, 0.83022, 0.8183, 0.80528, 0.79125, 
+        0.77627, 0.76047, 0.74349, 0.72533, 0.70625, 0.68594, 0.66435, 
+        0.64136, 0.61696, 0.59165, 0.56534, 0.53797, 0.50965, 0.4803, 
+        0.44982, 0.41786, 0.38513, 0.35184, 0.31817, 0.2848, 0.2517, 
+        0.21943, 0.18854, 0.1596, 0.13276, 0.10839, 0.08691, 0.06812, 
+        0.05206, 0.03879, 0.02823, 0.02006, 0.01387, 0.00932, 0.00608, 
+        0.00385, 0.00237, 0.00141, 0.00081, 0.00046, 0.00025, 0.00013, 
+        7e-05, 3e-05, 2e-05, 1e-05)
 
 Plotka <- function(Lx, r, a){
     sum(Lx * exp(-r*a))
@@ -10,32 +31,55 @@ Plotka <- function(Lx, r, a){
 xbarLotka <- function(Lx, r, a){
     sum(a * Lx * exp(-r*a)) /  sum(Lx * exp(-r*a))
 }
[…]

### 2014-04-22 ThanoRepro/R/HFDspaghetti.R

@@ -1,6 +1,6 @@
-source("/home/triffe/workspace/ThanoRepro/R/Functions.R")
+source("/home/triffe/git/ThanoRepro/ThanoRepro/R/Functions.R")
 # read in data and select females
-Data <- local(get(load("/home/triffe/workspace/ThanoRepro/Data/DataAll.Rdata")))
+Data <- local(get(load("/home/triffe/git/ThanoRepro/ThanoRepro/Data/DataAll.Rdata")))
 Data <- Data[Data$Sex == "f", ]
 
 
@@ -71,8 +71,14 @@ lines(with(Data, Age[Year >= 2000]), with(Data, Fy[Year >= 2000]), col = "#00000
 # 
 library(data.table)
 DT <- as.data.table(Data)
-DT[, FxSt := Fx / sum(Fx, na.rm = TRUE), by = list(Code, Year)]
-DT[, FySt := Fy / sum(Fy, na.rm = TRUE), by = list(Code, Year)]
+
+st <- function(x){
+    x/sum(x,na.rm=TRUE)
+}
+
+#update.packages()
+DT[, FxSt := st(Fx), by = list(Code, Year)]
+DT[, FySt := st(Fy), by = list(Code, Year)]
 Data <- as.data.frame(DT)
 
 par(mfrow=c(2,2))
@@ -91,8 +97,8 @@ lines(Data$Age, Data$FxSt, col = "#00000005")
 
 graphics.off()
 # Fy 
-png("/home/triffe/workspace/ThanoRepro/Figures/FySpaghettiDraft.png")
-#pdf("/home/triffe/workspace/ThanoRepro/Figures/FySpaghetti.pdf") # high res vector
+png("/home/triffe/git/ThanoRepro/ThanoRepro/Figures/FySpaghettiDraft.png")
+#pdf("/home/triffe/git/ThanoRepro/ThanoRepro/Figures/FySpaghetti.pdf") # high res vector
 par(xaxs = "i", yaxs = "i", mai=c(.5,.5,.5,.5))
 plot(NULL, type = "n",xlim = c(0,95),ylim=c(0,.13), axes = FALSE, xlab = "", ylab = "",
         panel.first = list(rect(0,0,95,.15,col = gray(.95),border = NA),
[…]

### 2014-04-23 ThanoRepro/ThanoRepro.tex
included lifecycle graph, more text edits

@@
 \usepackage{subcaption}
+\usepackage{tikz}
+%\usepackage[active,tightpage]{preview}
 \usepackage{natbib}
@@
 life line instead of at the beginning. Of course, such information cannot be
-requested of survey respondents or register data, because one's future time of
-death is typically unknown. Instead we approximate this perspective using a
-strategy similar to that used by \citet{miller2001increasing} or
-\citet{lee2002approach}, which uses lifetables and aggregated
-population counts to redistribute counts within age-classes over future death
-times.
+requested of survey respondents,
+because one's future time of death is typically unknown. It would be
+possible to organize lifecourse data by remaining years of life in a
+very long running linked population register, but such registers are few and far
+between, and the analysis would be very retrospective in nature.
+Instead we approximate this perspective using a strategy similar to that used by
+\citet{miller2001increasing} or \citet{lee2002approach}, which uses lifetables
+and aggregated population counts to redistribute counts within age-classes over
+future death times. This allows us to make statements about the remaining
+lifelines within a current stock of population, under certain assumptions.
 
@@
 story of population renewal within the confines of a philosophical bubble.
-Thanatological and chronological renewal models, such as the Lotka-Euler renewal model, both approximate the movement of population stocks over time, but from opposite vantage points.
-Specifying a model of thanatological renewal expands the toolbox available to approach complex temporal population structure.
-I present a set of broad findings to help understand the implications of temporal structure.
+Thanatological (counting down towards death) and chronological (counting up
+from birth) renewal models, such as the Lotka-Euler renewal model, both
+approximate the movement of population stocks over time, but from opposite
+vantage points. Specifying a model of thanatological renewal expands the toolbox
+available to approach complex temporal population structure. I present a set of
+preliminary findings to help understand the implications of temporal structure.
 
@@
 
-Our point of departure is a little formula that gives the probability of
-dying at some age $y$ years in the future, given survival to some
-earlier age, $a$. For age 0 this probability is simply the $d(a)$
-column of the lifetable, standardized to sum to 1. For higher ages, we simply condition on survival:
+Before hashing out the thanatological renewal model, I present the
+underlying data transformation --- a little formula that
+gives the probability of dying at some age $y$ years in the future, given survival to some earlier age, $a$.
+For age 0 this probability is simply the $d(a)$ column of the lifetable, standardized to sum to 1.
+For higher ages, we simply condition on survival:
 
@@
 A convenient discretization of \eqref{eq:vaupel1} is to simply work with the
-$d_a$ column of the lifetable (here single age):
+$d_a$ column of the lifetable (here with single ages):
 
@@
 
-If $P_a$ is an age-classified discrete population count, we can derive $P_y$, a
-remaining-years-classified population count, as follows:
+If $P_a$ is an (chronological) age-classified discrete population count, we can
+derive $P_y$, a thanatological age-classified population count, as follows:
 
@@
 \section*{The thanatological renewal model}
-For the population model that follows, we are most interested in using the kind
+For the renewal model that follows, we are most interested in using the kind
 of fertility rates shown in Figure~\ref{fig:FySpaghetti}, and we refer to a
-unisex population based on female vital rates. Assuming constant vital rates,
-the births for the present year are given by
+unisex population based on female vital rates. In the following, I
+explain describe the model in four different ways: as a continuous
+model, as a matrix model, with a lifecylcle graph, and intuitively by
+appealing to how population pyramids and leaves (described earlier) change over
+time. Assuming constant vital rates, the births for the present year are given by
 \begin{align}
@@
 \end{equation}
-Literally, this is the mean remaining life expectancy of new mothers in the
+Literally, this is the mean of the remaining lifespans of new mothers in the
 stable population. The net reproduction rate, $R_0$ is related by, e.g.,
@@
 \citep{vaupel2009life}, and now we can arrive at the same conclusion directly
-via the chronological and thanatological renewal equations. For small non-zero
-magnitude values of $r$ and survival patterns typical of humans, it is also the
+by equating the chronological and thanatological renewal equations. For
+small non-zero magnitude values of $r$ and survival patterns typical of humans, it is also the
 case that $c(y|r)\approx c(a|-r)$. The relationship is asymptotically exact as
 $r$ approaches zero. Other findings regarding stable age structure will be
-presented in a future version of this paper. Other stable population quantities
-may be estimated by similarly translating the various common definitions (e.g.,
+presented in a future version of this paper. Further stable population
+quantities may be estimated by similarly translating the various common definitions (e.g.,
 in the glossary of \citet{coale1972growth}) to the present perspective. We will
@@
 
+Figure~\ref{fig:graph} depicts the projection matrix $\textbf{Y}$ in the form of
+a lifecycle graph. In this graph, each thanatological age is a node. Population
+members are born into any of the thanatological age classes, and then follow
+clockwise around the graph on the thick blue paths in order until finally being
+absorbed into death (the node labeled ``X'', not included in the projection matrix). The light gray paths represent fertility. Thse originate from any age class that has a fertility rate greater
+than zero, which for this toy matrix are all age classes except the \nth{5}
+(highest). The y$^{th}$ age class is subject to the fertility rate $f^\star
+(y)$, and these births distribute out to all other age classes according to the
+lifetable deaths distribution, $d(a)$. Hence, the rate instensity of a fertility
+path is determined by the fertility rate of the origin age, and the probability of dying at the destination age. A fraction of the fertility destined for infant mortality equal
+to $lambda$ leads straight to the absorbing state of death. Fertility paths from
+the \nth{4} thanatological age class have been highlighted. These correspond
+to the fifth column of the example matrix $\textbf{Y}$, plus an additional path
+direct to death for the fraction of infant mortality prior to December \nth{31}
+in year $t$. Paths that loop back into the same age class simply indicate those
+offspring that are destined to die in the same year as their mother. It is
+perhaps clearer to see from the density of this graph and by recalling that
+thanatological fertility distributions are relatively wide (see
+Figure~\ref{fig:FySpaghetti}) that there is a high degree of mixing of
+thanatological age within populations via reproduction -- a strong tendency
+toward stability.
+
+\begin{figure}
+\centering
+\caption{The thanatological lifecycle graph}
+\label{fig:graph}
+\begin{tikzpicture}[->,shorten >=1pt,auto,node distance=4.3cm,
+  thick,main node/.style={circle,fill=blue!20,draw,font=\sffamily\Large\bfseries}]
+  \definecolor{gray}{rgb}{0.8,0.8,0.8}
+  \definecolor{lightgreen}{rgb}{0,0.8,0.8}
+  \node[main node] (0) {0};
+  \node[main node] (5) [above right of=0] {5};
+  \node[main node] (4) [right of=5] {4};
+  \node[main node] (3) [below right of=4] {3};
+  \node[main node] (2) [below left of=3] {2};
+  \node[main node] (1) [left of=2] {1};
+  \node[main node] (X) [left of = 5] {X};
+
+  \path[every node/.style={font=\sffamily\small},draw=gray,thin]
+    (3) edge [bend left] node {} (5)
+        edge [bend left] node {} (4)
+        edge [bend left] node {} (2)
+        edge [bend left] node {} (1)
+        edge [bend left] node {} (0)  
+        edge [bend left] node {} (X)
+        edge [loop right] node {} (3) 
+    (2) edge [bend left] node {} (5)
+        edge [bend left] node {} (4)
+        edge [bend left] node {} (3)
+        edge [bend left] node {} (1)
+        edge [bend left] node {} (0)     
+        edge [loop below] node {} (2) 
+        edge [bend left] node {} (X)
+    (1) edge [bend left] node {} (5)
+        edge [bend left] node {} (4)
+        edge [bend left] node {} (3)
+        edge [bend left] node {} (2)
+        edge [bend left] node {} (0)    
+        edge [loop below] node {} (1) 
+        edge [bend left] node {} (X)
+    (0) edge [bend left] node {} (5)
+        edge [bend left] node {} (4)
+        edge [bend left] node {} (3)
+        edge [bend left] node {} (2)
+        edge [bend left] node {} (1)
+        edge [loop left] node {} (0) 
+        edge [bend left] node {} (X);
+
+    \path[every node/.style={font=\sffamily\small}]
+    (4) edge [bend right] node {$f^\star (4)d(5)$} (5)
+        edge [bend left] node {$f^\star (4)d(3)$} (3)
+        edge [bend left] node {$f^\star (4)d(2)$} (2)
+        edge [bend left] node {$f^\star (4)d(1)$} (1)
+        edge [bend left] node {$(1-\lambda)f^\star (4)d(0)$} (0)
+        edge [loop above] node {$f^\star (4)d(4)$} (4)
+        edge [bend left] node {$\lambda f^\star (4)d(0)$} (X);
+
+    \path[every node/.style={font=\sffamily\small}, ultra thick, color =
+    lightgreen ] (5) edge node {} (4)
+    (4) edge node {} (3)
+    (3) edge node {} (2)
+    (2) edge node {} (1)
+    (1) edge node {} (0)
+    (0) edge node {} (X);
+
+\end{tikzpicture}
+\end{figure}
+
 Thanatological age-classes will ideally terminate at the highest value permitted
@@
 translate to 111 total remaining-years classes (0-110+). In practice $\textbf{Y}$ becomes
-a 111$\times$111 matrix, with most entries non-zero. Construction may appear
-tedious for this reason. However, note that the bulk of fertility entries can
-be derived as the outer (tensor) product $d_a \otimes f_y$, leaving only the 
-first row and first column mortality discounting followed by the addition of the
+a 111$\times$111 matrix, with most entries non-zero (nearly complete
+connectivity, such as in Figure~\ref{fig:graph}). Construction may appear tedious for
+this reason.
+However, note that the bulk of fertility entries can be derived as the outer (tensor) product $d_a \otimes f_y$, leaving only the first row and first column mortality discounting followed by the addition of the
 survival superdiagonal. In most statistical programming languages constructing $\textbf{Y}$ entails only

### 2014-06-05 ThanoRepro/ThanoRepro.tex
incorporated Brouard refs.

@@
 some other moment), consider age as the amount of time left from the present
-until death. Individuals in this case move in the same direction along
-imaginary life lines, but the reference point for age is now at the end of the
+until death\footnote{This quantity is referred to as \textit{residual} life in
+the reliability literaure.}.
+Individuals in this case move in the same direction along imaginary life lines, but the reference point for age is now at the end of the
 life line instead of at the beginning. Of course, such information cannot be
@@
 Instead we approximate this perspective using a strategy similar to that used by
-\citet{miller2001increasing} or \citet{lee2002approach}, which uses lifetables
-and aggregated population counts to redistribute counts within age-classes over
+\citet{brouard1986structure}, \citet{miller2001increasing}, or
+\citet{lee2002approach}, which uses lifetables and aggregated population counts to redistribute counts within age-classes over
 future death times. This allows us to make statements about the remaining
@@
 underlying data transformation --- a little formula that
-gives the probability of dying at some age $y$ years in the future, given survival to some earlier age, $a$.
-For age 0 this probability is simply the $d(a)$ column of the lifetable, standardized to sum to 1.
-For higher ages, we simply condition on survival:
+gives the probability of dying at some age $y$ years in the future, given
+survival to some earlier age, $a$. For age 0 this probability is simply the $d(a)$ column of the
+lifetable, standardized to sum to 1. For higher ages, we simply condition on survival:\footnote{In the reliability literature, this
+our concept of $d(y|a)$ is denoted by \begin{equation}
+X - t | X > t
+\end{equation}
+where $X$ is the total lifespan and $t$ is the current age attained.}
 
@@
 coined the term \textit{thanatological} age.}, and this complements the notion
-of chronological age. It bears repeating that thanatological age is
-probabilistic until after death, since it refers to the future according to
+of chronological age. \citep{brouard1989mouvements}, and later
+\citep{vaupel2009life}, showed that in stationary populations the chronological
+and thanatological age structures are identical.
+It bears repeating that thanatological age is probabilistic until after death, since it refers to the future according to
 particular assumptions about mortality. As such, Figure~\ref{fig:USrecomp} is

### 2014-06-05 ThanoRepro/R/Pyramids.R
made pyramids/leaves render better in different pdf viewers. instead of many horizontal bars, unify color regions with single polygons

@@ -1,37 +1,62 @@
+setwd("/home/triffe/git/ThanoRepro/ThanoRepro")
+source("R/Functions.R")
+source("R/PlotFunctions.R")
+devtools::load_all("/home/triffe/git/Leaves/PlosOne/R/RiffeetalFunctions")
 
-Data <- local(get(load("/home/triffe/workspace/ThanoRepro/Data/DataAll.Rdata")))
-source("/home/triffe/workspace/ThanoRepro/R/Functions.R")
-source("/home/triffe/workspace/ThanoRepro/R/PlotFunctions.R")
+plotPyr <- function(XXX,x,y,Thano=TRUE){
+    N <- ncol(XXX[["Cm"]])
+    mgrab <- ifelse(Thano,"Tm","Cm")
+    fgrab <- ifelse(Thano, "Tf", "Cf")
+    cgrab <- ifelse(Thano, "Tcol", "Ccol")
+    for (i in N:1){
+        m <- rep(XXX[[mgrab]][, i], each = 2)
+        f <- rep(XXX[[fgrab]][, i], each = 2)
+        polygon(c(m, rev(f)) + x, XXX[["y"]] + y, border = NA, col = XXX[[cgrab]][i])
+    }
+    
+    PyramidOutline(abs(XXX[[mgrab]][, N]), XXX[[fgrab]][, N], 
+            x = x, y = y, scale = 1, border = gray(.1), lwd = .2)
+}
 
-# Produces Figures 1a and 1b... and some other fun stuff.
+# OK data object in order
+# transform to plot-specific data
+prepare4plot <- function(XXX, colsFun1, colsFun2){
+    Mf          <- XXX$Females
+    Mm          <- XXX$Males
+    TOT         <- sum(Mf) + sum(Mm)
+    
+    Het1f       <- apply(Mf ,2, aggN, N = 10) / TOT             # age with ey het
+    Het1m       <- apply(Mm ,2, aggN, N = 10) / TOT             # age with ey het
+    Het2f       <- apply(Mf ,1, aggN, N = 10) / TOT             # age with ey het
+    Het2m       <- apply(Mm ,1, aggN, N = 10) / TOT 
+    
+# cumulative sum for cleaner plotting
+    cumHet1f    <- t(apply(Het1f, 2, cumsum))
+    cumHet2f    <- t(apply(Het2f, 2, cumsum))
[…]

### 2014-08-26 ThanoRepro/R/DataPrep.R
changed working dir to new os

@@ -1,6 +1,6 @@
-source("/home/triffe/workspace/ThanoRepro/R/Functions.R")
-HMDpath <- "/home/triffe/DATA/HMD"
-HFDpath <- "/home/triffe/DATA/HFD"
+source("/home/tim/git/ThanoRepro/ThanoRepro/R/Functions.R")
+HMDpath <- "/home/tim/DATA/HMD"
+HFDpath <- "/home/tim/DATA/HFD"
 #list.files(HFDpath)
 
 Bx <- read.table(file.path(HFDpath,"birthsRR.txt"), 
@@ -166,14 +166,14 @@ Data <- as.data.frame(DATA)
 rownames(Data) <- NULL
 #-----------------------------------------
 # save it out for wider use
-save(Data,file = "/home/triffe/workspace/ThanoRepro/Data/DataAll.Rdata")
+save(Data,file = "/home/tim/git/ThanoRepro/ThanoRepro/Data/DataAll.Rdata")
 
 #unique(Data$Code)
 
 # ------------------------------------------------------------
 # now we need to read in deaths by Lexis triangle to calculate age 0 lambda for
 # thanatological projection matrices.
-dlpath <- "/home/triffe/DATA/HMD/deaths/Deaths_lexis"
+dlpath <- "/home/tim/DATA/HMD/deaths/Deaths_lexis"
 
 lambda <- do.call(rbind, lapply(Allcountries, function(XXX, dlpath){
                     Dat <- read.table(file.path(dlpath,paste0(XXX,".Deaths_lexis.txt")),
@@ -200,7 +200,7 @@ lambda <- do.call(rbind, lapply(Allcountries, function(XXX, dlpath){
                             lambda.f = out[["Female"]], 
                             stringsAsFactors = FALSE)
                 }, dlpath = dlpath))
-save(lambda, file = "/home/triffe/workspace/ThanoRepro/Data/lambda.Rdata")
+save(lambda, file = "/home/tim/git/ThanoRepro/ThanoRepro/Data/lambda.Rdata")
 
 # something similar for use in Leslie matrices:
 
@@ -221,5 +221,5 @@ lambdaLeslie <- do.call(rbind,lapply(Allcountries, function(XXX, HMDpath){
                     lambda.f <- 1 - DLf / Bf
                     data.frame(Code = XXX,Year = yrs, lambda.m,lambda.f,stringsAsFactors = FALSE)
[…]

### 2014-08-26 ThanoRepro/R/HFDspaghetti.R

@@ -1,6 +1,6 @@
-source("/home/triffe/git/ThanoRepro/ThanoRepro/R/Functions.R")
+source("/home/tim/git/ThanoRepro/ThanoRepro/R/Functions.R")
 # read in data and select females
-Data <- local(get(load("/home/triffe/git/ThanoRepro/ThanoRepro/Data/DataAll.Rdata")))
+Data <- local(get(load("/home/tim/git/ThanoRepro/ThanoRepro/Data/DataAll.Rdata")))
 Data <- Data[Data$Sex == "f", ]
 
 
@@ -97,8 +97,8 @@ lines(Data$Age, Data$FxSt, col = "#00000005")
 
 graphics.off()
 # Fy 
-png("/home/triffe/git/ThanoRepro/ThanoRepro/Figures/FySpaghettiDraft.png")
-#pdf("/home/triffe/git/ThanoRepro/ThanoRepro/Figures/FySpaghetti.pdf") # high res vector
+png("/home/tim/git/ThanoRepro/ThanoRepro/Figures/FySpaghettiDraft.png")
+#pdf("/home/tim/git/ThanoRepro/ThanoRepro/Figures/FySpaghetti.pdf") # high res vector
 par(xaxs = "i", yaxs = "i", mai=c(.5,.5,.5,.5))
 plot(NULL, type = "n",xlim = c(0,95),ylim=c(0,.13), axes = FALSE, xlab = "", ylab = "",
         panel.first = list(rect(0,0,95,.15,col = gray(.95),border = NA),
@@ -113,8 +113,8 @@ text(mean(c(0,95)),-.01,"Thanatological age (years left)",xpd=TRUE)
 dev.off()
 
 # Fx 
-png("/home/triffe/git/ThanoRepro/ThanoRepro/Figures/FxSpaghettiDraft.png")
-#pdf("/home/triffe/git/ThanoRepro/ThanoRepro/Figures/FxSpaghetti.pdf") # high res vector
+png("/home/tim/git/ThanoRepro/ThanoRepro/Figures/FxSpaghettiDraft.png")
+#pdf("/home/tim/git/ThanoRepro/ThanoRepro/Figures/FxSpaghetti.pdf") # high res vector
 par(xaxs = "i", yaxs = "i", mai=c(.5,.5,.5,.5))
 plot(NULL, type = "n",xlim = c(12,55),ylim=c(0,.3), axes = FALSE, xlab = "", ylab = "",
         panel.first = list(rect(0,0,95,.3,col = gray(.95),border = NA),

### 2014-08-26 ThanoRepro/R/MatrixStuff.R

@@ -1,5 +1,5 @@
 library(popbio)
-setwd("/home/triffe/git/ThanoRepro/ThanoRepro")
+setwd("/home/tim/git/ThanoRepro/ThanoRepro")
 # 
 source("R/Functions.R")
 

### 2014-08-26 ThanoRepro/R/Oscillation.R

@@ -3,7 +3,7 @@
 
 # could start with Px, get Py, and then derive Cohen's measure, but this adds an initial smoothing
 # to Py. Instead I'll start with identical but randomly generated starting populations.
-AllMatrices <- local(get(load("/home/triffe/workspace/ThanoRepro/Data/AllMatrices.Rdata")))
+AllMatrices <- local(get(load("/home/tim/git/ThanoRepro/ThanoRepro/Data/AllMatrices.Rdata")))
 
 CohenA <- compiler::cmpfun(function(P0, LorY, N=500){
     Pops <- matrix(nrow=length(P0), ncol = N)

### 2014-08-26 ThanoRepro/R/Pyramids.R

@@ -1,4 +1,4 @@
-setwd("/home/triffe/git/ThanoRepro/ThanoRepro")
+setwd("/home/tim/git/ThanoRepro/ThanoRepro")
 source("R/Functions.R")
 source("R/PlotFunctions.R")
 devtools::load_all("/home/triffe/git/Leaves/PlosOne/R/RiffeetalFunctions")

### 2014-08-26 ThanoRepro/R/UniqueProofTests.R

@@ -1,5 +1,5 @@
-source("/home/triffe/workspace/ThanoRepro/R/Functions.R")
-Data <- local(get(load("/home/triffe/workspace/ThanoRepro/Data/DataAll.Rdata")))
+source("/home/tim/git/ThanoRepro/ThanoRepro/R/Functions.R")
+Data <- local(get(load("/home/tim/git/ThanoRepro/ThanoRepro/Data/DataAll.Rdata")))
 Data <- Data[Data$Sex == "f", ]
 
 # 

### 2014-08-26 ThanoRepro/R/rCalc.R

@@ -1,7 +1,8 @@
-source("/home/triffe/workspace/ThanoRepro/R/Functions.R")
+source("/home/tim/git/ThanoRepro/ThanoRepro/R/Functions.R")
+#install.packages("data.table")
 library(data.table)
 # read in data and select females
-Data <- local(get(load("/home/triffe/workspace/ThanoRepro/Data/DataAll.Rdata")))
+Data <- local(get(load("/home/tim/git/ThanoRepro/ThanoRepro/Data/DataAll.Rdata")))
 Data <- Data[Data$Sex == "f", ]
 Data$Fy[is.na(Data$Fy)] <- 0
 Data$Fyf[is.nan(Data$Fyf)] <- 0
@@ -65,7 +66,7 @@ rTM <- acast(rT,Year~Code,value.var = "V1")
 rLM <- acast(rL,Year~Code,value.var = "V1")
 rDiff <- rLM - rTM
 
-pdf("/home/triffe/workspace/ThanoRepro/Figures/rDist.pdf")
+pdf("/home/tim/git/ThanoRepro/ThanoRepro/Figures/rDist.pdf")
 par(xaxs="i",yaxs="i",mai=c(1,.5,.5,.5))
 plot(density(rL$V1),xlim=c(-.04,.03),ylim=c(0,55),col= "blue", main = "",axes = FALSE, ylab = "",
         panel.first = list(rect(-.04,0,.03,55,col=gray(.96), border = NA),
@@ -92,13 +93,14 @@ abline(h=0)
 matplot(1891:2011, rLM, type = 'l', lty = 1, col = "#0000FF40")
 abline(h=0)
 
+graphics.off()
 for (i in 1:ncol(rTM)){
 plot(NULL, type = "n",xlim=c(1891,2011),ylim=c(-.04,.02),main=colnames(rLM)[i])
 abline(h=0)
 lines(1891:2011, rLM[,i], col = "blue")
 lines(1891:2011, rTM[,i], col = "red")
 legend("bottomleft",lty=1,col=c("red","blue"),legend=c("Thano","Lotka"))
-Sys.sleep(3)
+Sys.sleep(1)
 }
 
 matplot(1891:2011, rLM - rTM, type = 'l', lty = 1, col = "#0000FF40")
@@ -161,3 +163,14 @@ sum(t(dM)*exp(-r*.5:110.5)*t(Fyx))
 
[…]

### 2014-08-27 ThanoRepro/R/Functions.R
code to demonstrate equality of r if the starting population is already stable.

@@ -1,32 +1,81 @@
+#' @title Minf0 utility function to replace \code{Inf} with 0
+#' 
+#' @description This function is not the most rigorous computation practice, but it reduces headache when working with real data that may have 0 in the denominator.
+#'  
+#' @param M any vector of data
+#' 
+#' @importFrom compiler cmpfun
+#' 
+#' @return M the same vector or array, with 0s imputed.
+#' 
+#' @export 
+#'
+
+Minf0 <- compiler::cmpfun(function(M){
+            M[is.infinite(M)]  <- 0
+            M
+        })
+
+#' @title Mna0 utility function to replace \code{NA} or \code{NaN} with 0
+#' 
+#' @description This function is not the most rigorous computation practice, but it reduces headache when working with real data that may have 0 in the denominator.
+#'  
+#' @param M any vector of data
+#' 
+#' @importFrom compiler cmpfun
+#' 
+#' @return M the same vector or array, with 0s imputed.
+#' 
+#' @export 
+#'
 Mna0 <- compiler::cmpfun(function(M){
             M[is.na(M)]  <- 0
             M[is.nan(M)] <- 0
             M
         })
 
-Minf0 <- compiler::cmpfun(function(M){
-            M[is.infinite(M)]  <- 0
-            M
-        })
-MinfNA <- compiler::cmpfun(function(M){
-            M[is.infinite(M)]  <- NA
-            M
-        })
 
-Thano <- compiler::cmpfun(function(Px, dx){
-            ay      <- 0:110
-            dx      <- Mna0(dx)
-            # this is the refinement:
-            dx      <- (dx + c(dx[2:length(dx)],0)) / 2
+
+#' @title Thano redistribute age an age-classified vector into an age-remaining years cross-classified matrix
+#' 
+#' @description A discrete approximation of the thanatological redistribution formula. If \code{length(Px) != length(dx)}, we pad out the shorter of the two with 0s.
+#'  
+#' @param Px any single-year age-classified vector of data. Age classes must align with those of \code{dx} starting at 0.
+#' @param dx single-age density column of the lifetable. Age classes must align with those of \code{Px} starting at 0.
+#' 
+#' @importFrom compiler cmpfun
+#' 
+#' @export 
+#' 
+Thano <- compiler::cmpfun(function(Px, dx, stagger = TRUE){
+            Np <- length(Px)
+            Nd <- length(dx)
+            if (Np != Nd){
[…]

### 2014-08-27 ThanoRepro/R/rEqual.R

@@ -1,0 +1,142 @@
+source("/home/tim/git/ThanoRepro/ThanoRepro/R/Functions.R")
+#  thanatological age 
+
+rThanoCoale <- compiler::cmpfun(function(Fy, da, .a = .5:110.5, maxit = 1e2, tol = 1e-15){  
+            
+            # Based on Coale (1957), modified.
+            N    <- length(Fy)
+            dxM  <- matrix(0, ncol = N, nrow = N)
+            dxi  <- da
+            for (i in 1:N){
+                dxM[i, 1:length(dxi)  ] <- dxi 
+                dxi <- dxi[2:length(dxi) ]
+            }     
+            R0      <- sum(dxM * Fy)
+            T.guess <- sum(.a * dxM * Fy) / R0 # assuming r = 0
+            r.i     <- log(R0) / T.guess
+            # be careful to discount Fy by SRB appropriately for males / females
+            # prior to specification
+            for (i in 1:maxit){ # 15 is more than enough!
+                #cat(r2,i,"\n")
+                deltai <- 1 - sum(rowSums(t(t(dxM) * exp(-r.i * .a))) * Fy)
+                # the mean generation time self-corrects 
+                # according to the error produced by the Lotka equation
+                r.i <- r.i - (deltai / (T.guess - (deltai / r.i))) 
+                if (abs(deltai) <= tol){
+                    break
+                }
+            }
+            return(r.i)  
+        })
+# chronological age
+Rmomentn <- compiler::cmpfun(function(fx,Lx,x,n=0){
+            sum((x^n)*fx*Lx)
+        })
+rLotkaCoale <- compiler::cmpfun(function(fx,Lx,x=.5:110.5, maxit = 1e2, tol = 1e-15){
+            # from Coale, Ansley J. (1957) A New Method for Calculating Lotka's r- the Intrinsic Rate of Growth in a Stable Population.
+            # Population Studies, Vol. 11 no. 1, pp 92-94
+            R0 <- Rmomentn(fx,Lx,x,0)
+            # first assuming a mean generation time of 29
+            ri <- log(R0)/29
+            for (i in 1:maxit){ # 10 is more than enough!
+                deltai <- sum(exp(-ri*x)*fx*Lx)-1
+                # the mean generation time self-corrects 
+                # according to the error produced by the Lotka equation
+                ri <- ri+(deltai/(29-(deltai/ri)))
[…]

### 2014-08-27 ThanoRepro/ThanoRepro.tex
abstract

@@
 \begin{abstract}
-Aggregate data classified by chronological age can be transformed into data
-classified by remaining years of life (thanatological age). A model for population renewal and
-the corresponding projection matrix are presented for populations structured by
-thanatological age. Period results are derived using all available data from the
-HMD and HFD. We compare the intrinsic growth rate, $r$, as derived from the
-classic Lotka equation versus that derived on the basis of thanatological age.
-We also compare some transient indicators between the two
-models. Empirical results suggest that $r$ from the thanatological model tends to be less erratic than
-Lotka's $r$, and the trajectory to stability tends to be faster with less
-oscillation.
+A unisex model of population renewal for populations structured by
+thanatological age is presented in both continuous and discrete form. Various
+stable and transient properties populations are compared and related when viewed
+by chronological versus thanatological age. Results from the two age perspectives are conformable if the starting
+population is stable, but are otherwise divergent.
 \end{abstract}

### 2014-08-27 ThanoRepro/ThanoRepro.tex
.

@@
 \usepackage[top=2in, bottom=1.5in, left=1in, right=1in]{geometry}
-
+\usepackage{setspace}
 % matrix numbering
@@
 
+\onehalfspacing
 All demographic forces vary over age in known and regular ways. Information on
@@
 some other moment), consider age as the amount of time left from the present
-until death\footnote{This quantity is referred to as \textit{residual} life in
-the reliability literaure.}.
-Individuals in this case move in the same direction along imaginary life lines, but the reference point for age is now at the end of the
-life line instead of at the beginning. Of course, such information cannot be
-requested of survey respondents,
-because one's future time of death is typically unknown. It would be
-possible to organize lifecourse data by remaining years of life in a
-very long running linked population register, but such registers are few and far
-between, and the analysis would be very retrospective in nature.
-Instead we approximate this perspective using a strategy similar to that used by
-\citet{brouard1986structure}, \citet{miller2001increasing}, or
-\citet{lee2002approach}, which uses lifetables and aggregated population counts to redistribute counts within age-classes over
-future death times. This allows us to make statements about the remaining
-lifelines within a current stock of population, under certain assumptions.
+until death\footnote{This quantity is referred to as \textit{residual} lifetime
+in the reliability literaure.}. Individuals in this case move in the same
+direction along imaginary life lines, but the reference point for age is now at
+the end of the life line instead of at the beginning. For individuals, the
+timing of events in life is mirrored exactly between the two definitions of age,
+but aggregate patterns differ.
+
+Direct information on the timing of demographic transitions with respect to
+death is only available from long-running panel or linked historical data
+studies that are few and far between. Some surveys also ask respondents in various ways about
+their perceived future lifetime. Perceived remaining lifetime may underly those
+demographic transitions governed to some degree by volition, but this is not
+treated here. The present inquiry is related to projected
+remaining lifetime, using lifetable methods, which is an approximation of
+actual remaining lifetime.
+The lifetable methods applied here follow those of \citet{brouard1986structure}, \citet{miller2001increasing}, which redistribute age-classified counts over future
+death times according to a fixed mortality schedule. This allows us to make
+statements about the remaining lifelines within a current stock of population, under certain assumptions.
 

### 2014-08-27 ThanoRepro/ThanoRepro.tex
.

@@
 From the projection matrix, the ratio of the largest to the second largest
-eigenvalue, the damping ratio, is an indicator of the \texit{springiness} of the
-stable population structure, $c_y$ or $c_a$, with respect to disturbances from the stable state
+eigenvalue, the damping ratio, is an indicator of the \textit{springiness} of
+the stable population structure, $c_y$ or $c_a$, with respect to disturbances from the stable state
 as determined by vital rate distributions. A higher damping ratio means that the

### 2014-08-28 ThanoRepro/ThanoRepro.tex
edits to first half of text

@@
 Direct information on the timing of demographic transitions with respect to
-death is only available from long-running panel or linked historical data
-studies that are few and far between. Some surveys also ask respondents in various ways about
-their perceived future lifetime. Perceived remaining lifetime may underly those
-demographic transitions governed to some degree by volition, but this is not
-treated here. The present inquiry is related to projected
-remaining lifetime, using lifetable methods, which is an approximation of
-actual remaining lifetime.
-The lifetable methods applied here follow those of \citet{brouard1986structure}, \citet{miller2001increasing}, which redistribute age-classified counts over future
-death times according to a fixed mortality schedule. This allows us to make
-statements about the remaining lifelines within a current stock of population, under certain assumptions.
+death is only available from long-running panel or linked historical data that are few and far between. Some surveys also ask respondents in various ways about
+their perceived future lifetime, which may underly those
+demographic transitions governed to some degree by volition. The present inquiry
+draws only from lifetable-projected remaining lifetime, which
+is intended to approximate actual remaining lifetime.
+The lifetable methods applied here follow those of \citet{brouard1986structure}
+or \citet{miller2001increasing}, which redistribute age-classified counts over
+future death times according to a fixed mortality schedule. This allows us to make statements about the remaining lifelines within a current stock of population, under certain assumptions.
 
@@
 do not tackle here. The model presented here should therefore not be interpreted
-as a suggestion for a better way to model or project population, but as an alternative way to tell the
-story of population renewal within the confines of a philosophical bubble.
+as a suggestion for a better way to model or project population, but as an
+alternative way to tell the story of population renewal.
+
 Thanatological (counting down towards death) and chronological (counting up
@@
 
-\section*{From chronological to thanatological age}
+\section*{A lifetable approximation of thanatological age}
 
-Before hashing out the thanatological renewal model, I present the
-underlying data transformation --- a little formula that
-gives the probability of dying at some age $y$ years in the future, given
-survival to some earlier age, $a$. For age 0 this probability is simply the $d(a)$ column of the
-lifetable, standardized to sum to 1. For higher ages, we simply condition on survival:\footnote{In the reliability literature, this
-our concept of $d(y|a)$ is denoted by \begin{equation}
-X - t | X > t
+A lifetable approach to transforming chronologically classified data into
+thanatologically classified data is valid when the given lifetable accurately
+describes the future mortality of age-classified population. To make
+a projective statement about a current stock of population, one does well to
+factor in future changes in mortality, and to allow the trajectory of changes
+to vary between ages/cohorts. In the case of a stable population model,
+mortality is assumed invariant over time, which means that the same survival
+pattern underlies each chronological age class, and we therefore draw from a
+single lifetable. Under these conditions, data classified by chronological age
+can be decomposed and redistributed into thanatological age classes with a
+single mortality pattern. The basic operation is to multiply the probability of
+surviving from exact chronological age $a$ to exact age $a+y$,
+$\frac{l(a+y)}{l(a)}$, and then dying at exact age $a+y$, $\mu(a+y)$. The
+index $y$ in this case means $y$ years in the future. For age 0 this probability
+is simply the $d(a)$ column of the lifetable, standardized to sum to 1.
+For higher ages, simply condition on survival:\footnote{In the reliability literature, our concept of $d(y|a)$ is denoted by \begin{equation} X - t | X > t
 \end{equation}
-where $X$ is the total lifespan and $t$ is the current age attained.}
+where $X$ is the total lifespan and $t$ is the current chronological age
+attained.}
 
@@
 \begin{equation}
-d_{y, a} =  \frac{d_{a+y}+d_{a+y+1}}{l_a + l_{a+1}} \quad\quad \text{,}
+d_{y, a} =  \frac{d_{a+y}+d_{a+y+1}}{L_a + L_{a+1}} \quad\quad \text{,}
 \end{equation}
 
-\noindent where $\omega$ is the highest age.\footnote{In later formulas we just
-use $l_a$ in the denominator to reduce clutter, but the above sum over $d_a$
-will eliminate rounding error in case one is working from a published life table.} One
-produces in this way a probability density function over future death times for
-each age, and the $d_{y,a}$- weighted average of $y$ for a given age $a$ returns
-the familiar remaining life expectancy column of the lifetable, $e_a$. Since $1
-= \sum_{n=0}^{\omega - a}d_{a+y, a}$ for each $a$, we can use this to decompose
-age-classified demographic data. For example, we can break down a population
-pyramid into remaining years categories, like in Figure~\ref{fig:USdecomp},
-which shows this exercise for 2010 US data. The fun starts when we proceed to
-regroup cross-classified data by $y$ instead of collapsing back to $a$. Now
-instead of remaining lifetime heterogeneity within each age group we have age heterogeneity within each discrete remaining years of life.
-Figure~\ref{fig:USrecomp} shows this \textit{re}composition for the same 2010 US data.
+\noindent where $\omega$ is the highest age. One
+produces in this way a probability density function over future death times
+for each chronological age, $a$, and the $d(y|a)$- weighted average of $y$ for
+a given $a$ returns the familiar remaining life expectancy column of the lifetable,
+$e(a)$. Since $1 = \int_0^\infty d(y|a) \dd y$ for each $a$,
+this density is useful for decomposing data classified by chronological age. For
+example, we can break down a population pyramid into thanatological age classes.
+Figure~\ref{fig:USdecomp} shows this exercise for 2010 US data, assuming the
+2010 period mortality schedule remains constant. When looking from the
+chronological age perspective, this decomposition reveals thanatological
+heterogeneity, but when viewed from the thanatological perspective
+(Figure~\ref{fig:USdecomp}) it reveals chronological age heterogeneity and the
+overall thanatological age profile, the future death flow of the current
+population.
+Figure~\ref{fig:USrecomp} shows this \textit{re}composition for the same 2010
+US data.
 
 \begin{figure}
-	\caption{2010 US population}
+	\caption{2010 US population structure}
 	\begin{center}
 	\begin{subfigure}{.45\textwidth}
-		\caption{Chronological age-structure decomposed by thanatological age.}
+		\caption{Chronological age-structure (years lived) decomposed by
+		thanatological age (years left).}
 		\label{fig:USdecomp}
@@
 	\begin{subfigure}{.45\textwidth}
-		\caption{Thanatological age-structure decomposed by chronological age.}
+		\caption{Thanatological age-structure (years left) decomposed by chronological
+		age (years lived).}
 		\label{fig:USrecomp}
@@
 
-If $P_a$ is an (chronological) age-classified discrete population count, we can
+If $P_a$ is an chronological age-classified discrete population count, we can
 derive $P_y$, a thanatological age-classified population count, as follows:
@@
 \label{eq:transform}
-P_y =& \sum_{a=0}^\omega P_a \frac{d_{a+y}+d_{a+y+1}}{l_a+l_{a+1}}
+P_y =& \sum_{a=0}^\omega P_a \frac{d_{a+y}+d_{a+y+1}}{L_a+L_{a+1}}
 \intertext{and of course}
@@
 
-We refer to \eqref{eq:transform} as the thanatological age
-transformation\footnote{I have on good authority that Ken Wachter
-coined the term \textit{thanatological} age.}, and this complements the notion
-of chronological age. \citep{brouard1989mouvements}, and later
-\citep{vaupel2009life}, showed that in stationary populations the chronological
+\citet{brouard1989mouvements}, and later
+\citet{vaupel2009life}, showed that in stationary populations the chronological
 and thanatological age structures are identical.
-It bears repeating that thanatological age is probabilistic until after death, since it refers to the future according to
-particular assumptions about mortality. As such, Figure~\ref{fig:USrecomp} is
-projective in nature, a forward-looking glance at a population attrition given the age-structure and
-lifetable of a particular moment, whereas Figure~\ref{fig:USdecomp} is reflective in nature, since a population's
-chronological age-structure is mostly the fruit of past fertility, but also
+It bears repeating that lifetable-derived thanatological age structure is
+probabilistic until after death, since it refers to the future according to particular assumptions about mortality.
+As such, Figure~\ref{fig:USrecomp} is projective in nature, a forward-looking glance at a population attrition given the age-structure and
+lifetable of a particular moment, whereas Figure~\ref{fig:USdecomp} is
+reflective in nature, since a population's chronological age-structure is mostly the fruit of past fertility, but also
 migration and to a lesser degree attrition. I explore population structure by remaining years of life

### 2014-08-29 ThanoRepro/ThanoRepro.tex
added new references, updating more manuscript to account for Jim's new manuscript

@@
 a given $a$ returns the familiar remaining life expectancy column of the lifetable,
-$e(a)$. Since $1 = \int_0^\infty d(y|a) \dd y$ for each $a$,
+$e(a)$. If $P(a)$ is an chronological age-classified population count,
+we can derive $P(y)$, a thanatological age-classified population count, as
+follows:
+
+\begin{align}
+\label{eq:transform}
+P(y) =& \int_0^\infty P(a) d(y | a) \dd a
+\end{align}
+
+Since $1 = \int_0^\infty d(y|a) \dd y$ for each $a$,
 this density is useful for decomposing data classified by chronological age. For
@@
 
-\begin{figure}
+\begin{figure}[ht!]
 	\caption{2010 US population structure}
@@
 
-If $P_a$ is an chronological age-classified discrete population count, we can
-derive $P_y$, a thanatological age-classified population count, as follows:
+\citet{brouard1986structure} applied this method to highlight regional
+differences in French age structure, at times combining this projective method
+with historical stocks. \citet{brouard1989mouvements}, and later
+\citet{vaupel2009life}, showed that in stationary populations the chronological
+and thanatological age structures are identical, and this provides a
+limited basis for comparison between the two perspectives. \citet{wachter2014essential} hints at, and
+\citet{vaupel2014stable} expands on how to make chronological and
+thanatological age structure commensurate under a constant growth rate, $r$.
 
-\begin{align}
-\label{eq:transform}
-P_y =& \sum_{a=0}^\omega P_a \frac{d_{a+y}+d_{a+y+1}}{L_a+L_{a+1}}
-\intertext{and of course}
-\sum P_y =& \sum P_a
-\end{align}
-
-\citet{brouard1989mouvements}, and later
-\citet{vaupel2009life}, showed that in stationary populations the chronological
-and thanatological age structures are identical.
-It bears repeating that lifetable-derived thanatological age structure is
-probabilistic until after death, since it refers to the future according to particular assumptions about mortality.
-As such, Figure~\ref{fig:USrecomp} is projective in nature, a forward-looking glance at a population attrition given the age-structure and
-lifetable of a particular moment, whereas Figure~\ref{fig:USdecomp} is
-reflective in nature, since a population's chronological age-structure is mostly the fruit of past fertility, but also
-migration and to a lesser degree attrition. I explore population structure by remaining years of life
-for its own sake in a different paper.\footnote{Poster Session 6 (Friday
-mid-day))} Note that any count classified by chronological age can be
+This result provides a standard on which to compare chronological and thanatological age structure.
+It bears repeating that lifetable-derived thanatological age structure is probabilistic until after death, since it refers to the future according to an assumed mortality pattern.
+As such, Figure~\ref{fig:USrecomp} is projective in nature, a forward-looking
+glance at population attrition given the population stock and lifetable of a
+particular moment, whereas Figure~\ref{fig:USdecomp} is reflective in nature, since a population's chronological age-structure is mostly the fruit of past fertility, but also migration and to a lesser degree attrition.
+Note that any count classified by chronological age can be
 reclassified in this way, as long as you have a lifetable that plausibly represents the population whose data you wish to reclassify.

### 2014-08-30 ThanoRepro/ThanoRepro.tex
minor text edits, adding in R proofs before i work out the final one on paper

@@
 limited basis for comparison between the two perspectives. \citet{wachter2014essential} hints at, and
-\citet{vaupel2014stable} expands on how to make chronological and
+\citet{vaupel2014stable} elaborates on how to make chronological and
 thanatological age structure commensurate under a constant growth rate, $r$.
+These results will also follow from the following specification of a
+thanatological renewal model.
 
-This result provides a standard on which to compare chronological and thanatological age structure.
 It bears repeating that lifetable-derived thanatological age structure is probabilistic until after death, since it refers to the future according to an assumed mortality pattern.

### 2014-08-30 ThanoRepro/R/Functions.R

@@ -156,3 +156,22 @@ mx2dxHMD <- compiler::cmpfun(function(mx){
             dx                  <- lx * qx                                                                                # Eq 66 MPv5
             Mna0(dx)
         })
+da2fya <- function(da, stagger = FALSE){
+    N       <- length(da)
+    ay      <- 1:N - 1
+    
+    da      <- Mna0(da)   # remove NAs if any       
+    da      <- c(da, da * 0) / sum(da) # pad out with 0s
+    fya     <- matrix(da[col(matrix(nrow = N, 
+                                    ncol = N)) + ay], 
+            nrow = N, 
+            ncol = N, 
+            dimnames = list(Ex = ay, 
+                    Age = ay)
+    )
+    if (stagger){
+        fya <- (fya + cbind(fya[, 2:ncol(fya)], 0)) / 2
+    }
+    fya <- Minf0(Mna0(fya / rowSums(fya)))
+    fya
+}

### 2014-08-30 ThanoRepro/R/rEqual.R

@@ -113,7 +113,7 @@ rThanoCoale(.fy,da,.a=a) # 4th decimal place... is it a rounding error or not?
 # further make age strictly integer...
 la2 <- with(Data, lx[Year == 2000 & Code == "SWE"])
 # 2) force d(a) to entail l(a) and vice versa.
-da2 <- -diff(c(la,0)) # extra step to force perfect consistency
+da2 <- -diff(c(la2,0)) # extra step to force perfect consistency
                       # between d(a) and l(a)
 
 # 3) do away with mid-interval age. Use integer.
@@ -140,3 +140,84 @@ rThanoCoale(.fy2,da2,.a=a2)
 # was due to rounding, in order to get a better approximation.
 
 # easier to prove on paper if we know this is true...
+
+# one thing to verify:
+#f^\star(y) is the c(a)d(y|a)-weighted average of f(a)...
+
+sum(.fy2)
+sum(fa)
+
+dya <- da2fya(da2,FALSE)
+dxM  <- matrix(0, ncol = 111, nrow = 111)
+dxi  <- da2
+for (i in 1:111){
+    dxM[i, 1:length(dxi)  ] <- dxi 
+    dxi <- dxi[2:length(dxi) ]
+} 
+# these are identical, save for some machine error:
+any(abs(dya * la2 - dxM) < 1e-12)
+# demonstration:
+all(colSums(fa * ca2 * dya) / colSums(ca2 * dya) == .fy2)
+
+# plug into thano renewal eqn:
+sum(
+t(exp(-a2*rL2) * t(dya * la2)) * 
+        colSums(fa * ca2 * dya) / colSums(ca2 * dya)
+)
+wya <- ca2 * dya
+# or define it as weights:
+sum(
+        t(exp(-a2*rL2) * t(dya * la2)) * 
+                colSums(fa * wya) / colSums(wya)
+)
+
+# compare again w Lotka:
+sum(exp(-a2*rL2)*la2*fa)
+
+# note:
+all(la2 == colSums(dya * la2))    # 1) d(y|a) * l(a) is just d(a) ...
+all(la2 == rev(cumsum(rev(da2)))) # 2) d(a) and l(a) are commensurable here
+
+# this is because d(y|a) is rescaled by dividing out l(a), 
+# so we can just multiply it right back in
+
+
+# this gives c(a):
+all(la2 * exp(-a2 * rL2) / sum(la2 * exp(-a2 * rL2)) == ca2)
+
+# plug it in:
+sum(
[…]

### 2014-08-31 ThanoRepro/R/rEqual.R
starting proof appendix

@@ -155,7 +155,7 @@ for (i in 1:111){
     dxi <- dxi[2:length(dxi) ]
 } 
 # these are identical, save for some machine error:
-any(abs(dya * la2 - dxM) < 1e-12)
+all(abs(dya * la2 - dxM) < 1e-12)
 # demonstration:
 all(colSums(fa * ca2 * dya) / colSums(ca2 * dya) == .fy2)
 
@@ -164,12 +164,12 @@ sum(
 t(exp(-a2*rL2) * t(dya * la2)) * 
         colSums(fa * ca2 * dya) / colSums(ca2 * dya)
 )
-wya <- ca2 * dya
-# or define it as weights:
-sum(
-        t(exp(-a2*rL2) * t(dya * la2)) * 
-                colSums(fa * wya) / colSums(wya)
-)
+#wya <- ca2 * dya
+## or define it as weights:
+#sum(
+#        t(exp(-a2*rL2) * t(dya * la2)) * 
+#                colSums(fa * wya) / colSums(wya)
+#)
 
 # compare again w Lotka:
 sum(exp(-a2*rL2)*la2*fa)
@@ -187,14 +187,14 @@ all(la2 * exp(-a2 * rL2) / sum(la2 * exp(-a2 * rL2)) == ca2)
 
 # plug it in:
 sum(
-        t(exp(-a2*rL2) * t(dya * la2)) * 
+        colSums(exp(-a2*rL2) * t(dya * la2)) * 
                 colSums(fa * dya * la2 * exp(-a2 * rL2) / sum(la2 * exp(-a2 * rL2))) / 
                 colSums(dya * la2 * exp(-a2 * rL2) / sum(la2 * exp(-a2 * rL2)))
 )
 
 # back to dxM, which is straight dx, unscaled:
 sum(
-        t(exp(-a2*rL2) * t(dxM)) * 
+        colSums(exp(-a2*rL2) * dxM) * 
                 colSums(fa * dxM * exp(-a2 * rL2) / sum(la2 * exp(-a2 * rL2))) / 
                 colSums(dxM * exp(-a2 * rL2) / sum(la2 * exp(-a2 * rL2)))
 )
@@ -202,22 +202,53 @@ sum(
 growth <- exp(-a2*rL2) 
 # this is still a thanatological setup:
 sum(
-        t(growth * t(dxM)) * 
+        t(growth * dxM) * 
                 colSums(fa * dxM * growth / sum(la2 * growth)) / 
[…]

### 2014-09-02 ThanoRepro/ThanoRepro.tex
starting proof appendix
mostly finished appendix B, included pressat in bib

@@
 \documentclass{article}
+\usepackage{appendix}
 \usepackage{amsmath}
@@
 
-* this is a work in progress, and by the time the EPC comes around I'll
-have shifted the focus away on comparing stable results estimated from data
-(the last couple figures), and more toward the reversibility of chronological
-and thanatological stable age structures, as well as a comparative
-eigen-analysis of the Leslie and thanatological projection matrices.
-
 \vspace{2em}
 
-\appendix
+\begin{appendices}
 \section{Unique solution for thanatological $r$}
@@
 \eqref{eq:lotka}. This little proof follows that given in
-\textit{pressat1972demographic}. Define a convenience function, $I(r)$, for the
+\citet{pressat1973analyse}. Define a convenience function, $I(r)$, for the
 integrand of \eqref{eq:thanoren} for a given $r$ and fixed $f^\star(y)$ and $d(a)$:
@@
 
+\section{Equivalence in stability}
+\label{app:B}
+This appendix contains a proof that the thanatological renewal equation implies
+the same intrinsic growth rate, $r$, as Lotka's chronological renewal model if
+the starting population is stable. This is identical to claiming that the right
+side of the rewal equation is equal given the same $r$. Since there is a unique
+real solution, demonstrated in the previous appendix, then it is sufficient to
+show that the two integral equations are equal under these conditions. The basic
+relationship is:
+\begin{align}
+\int _{a=0}^\infty l(a)e^{-ra}f(a) \dd a &= \int _{y=0}^\infty \int
+_{a=0}^\infty d(a+y)e^{-ra}f^\star (y) \dd a \dd y
+\intertext{when}
+f^\star(y) &= \frac{b^\star(y)}{c^\star(y)} \\
+b^\star (y) &= \int_{a=0}^\infty b(a) \mu(a+y))\frac{l(a+y)}{l(a)} \dd a \\
+c^\star (y) &= \int_{a=0}^\infty c(a) \mu(a+y))\frac{l(a+y)}{l(a)} \dd a
+\intertext{Note that:}
+c(a) &= \frac{l(a)r^{-ra}}{\int l(a) e^{-ra} \dd a} \\
+b(a) &= f(a)c(a)
+\end{align}
+Replacing $f^\star (y)$ with the full expanded expression and
+plugging into the thanatological renewal equation gives:
+\begin{align}
+=& \int _{y=0}^\infty \int
+_{a=0}^\infty d(a+y)e^{-ra} \frac{\frac{\int_{a=0}^\infty
+f(a)l(a)e^{-ra}\mu(a+y)\frac{l(a+y)}{l(a)} \dd a}{\int l(a) e^{-ra} \dd
+a}}{\frac{\int_{a=0}^\infty l(a)e^{-ra} \mu(a+y)\frac{l(a+y)}{l(a)} \dd a}{\int l(a) e^{-ra} \dd a}} \dd a \dd y
+\intertext{The denominator of $c(a)$ cancels out:}
+=& \int _{y=0}^\infty \int
+_{a=0}^\infty d(a+y)e^{-ra} \frac{
+\int_{a=0}^\infty
+f(a)l(a)e^{-ra}\mu(a+y)\frac{l(a+y)}{l(a)} \dd a}{
+\int_{a=0}^\infty l(a)e^{-ra} \mu(a+y)\frac{l(a+y)}{l(a)} \dd a} \dd a \dd y
+\intertext{Some $l(a)$'s also cancel out:}
+=& \int _{y=0}^\infty \int
+_{a=0}^\infty d(a+y)e^{-ra} \frac{
+\int_{a=0}^\infty
+f(a)e^{-ra}\mu(a+y)l(a+y) \dd a}{
+\int_{a=0}^\infty e^{-ra} \mu(a+y)l(a+y) \dd a} \dd a \dd y
+\intertext{Now $d(a+y) = \mu (a+y)l(a+y)$, so we can cancel the inner integral
+with the denominator:}
+=& \int _{y=0}^\infty 
+\int_{a=0}^\infty
+f(a)e^{-ra}\mu(a+y)l(a+y) \dd a \dd y
+\intertext{Note also that $l(a) = \int _{x=a}^\infty d(x) \dd x$ (present
+livings are future deaths), which brings us back to the chronological
+formulation:} =& \int_{a=0}^\infty e^{-ra}l(a)f(a)
+\dd a
+\end{align}
+\end{appendices}
 \nocite{HMD,HFD}

### 2014-09-03 ThanoRepro/ThanoRepro.tex
changed integrand to t

@@
 \end{align}
-Replacing $f^\star (y)$ with the full expanded expression and
-plugging into the thanatological renewal equation gives:
+Replacing
+$f^\star (y)$ with the full expanded expression and plugging into the thanatological renewal equation gives:
 \begin{align}
 =& \int _{y=0}^\infty \int
-_{a=0}^\infty d(a+y)e^{-ra} \frac{\frac{\int_{a=0}^\infty
-f(a)l(a)e^{-ra}\mu(a+y)\frac{l(a+y)}{l(a)} \dd a}{\int l(a) e^{-ra} \dd
-a}}{\frac{\int_{a=0}^\infty l(a)e^{-ra} \mu(a+y)\frac{l(a+y)}{l(a)} \dd a}{\int l(a) e^{-ra} \dd a}} \dd a \dd y
+_{a=0}^\infty d(a+y)e^{-ra} \frac{\frac{\int_{t=0}^\infty
+f(t)l(t)e^{-rt}\mu(a+y)\frac{l(t+y)}{l(t)} \dd t}{\int l(t) e^{-rt} \dd
+t}}{\frac{\int_{t=0}^\infty l(t)e^{-rt} \mu(t+y)\frac{l(t+y)}{l(t)} \dd t}{\int
+l(t) e^{-rt} \dd t}} \dd a \dd y
 \intertext{The denominator of $c(a)$ cancels out:}
@@
 _{a=0}^\infty d(a+y)e^{-ra} \frac{
-\int_{a=0}^\infty
-f(a)l(a)e^{-ra}\mu(a+y)\frac{l(a+y)}{l(a)} \dd a}{
-\int_{a=0}^\infty l(a)e^{-ra} \mu(a+y)\frac{l(a+y)}{l(a)} \dd a} \dd a \dd y
+\int_{t=0}^\infty
+f(t)l(t)e^{-rt}\mu(t+y)\frac{l(t+y)}{l(t)} \dd t}{
+\int_{t=0}^\infty l(t)e^{-rt} \mu(t+y)\frac{l(t+y)}{l(t)} \dd t} \dd a \dd y
 \intertext{Some $l(a)$'s also cancel out:}
@@
 _{a=0}^\infty d(a+y)e^{-ra} \frac{
-\int_{a=0}^\infty
-f(a)e^{-ra}\mu(a+y)l(a+y) \dd a}{
-\int_{a=0}^\infty e^{-ra} \mu(a+y)l(a+y) \dd a} \dd a \dd y
+\int_{t=0}^\infty
+f(t)e^{-rt}\mu(t+y)l(t+y) \dd t}{
+\int_{t=0}^\infty e^{-rt} \mu(t+y)l(t+y) \dd t} \dd a \dd y
 \intertext{Now $d(a+y) = \mu (a+y)l(a+y)$, so we can cancel the inner integral

### 2014-09-04 ThanoRepro/ThanoRepro.tex
lots of edits. new appendix for optimization method. more work on matrix stuff, but still not perfectly equal...

@@
 Now compare this to Lotka's chronological formulation
-\begin{equation}
+\begin{align}
 \label{eq:lotka}
-1 = \int_{a=0}^\infty f(a)l(a)e^{-ra}\dd a \quad\text{,}
-\end{equation}
+1 =& \int_{a=0}^\infty f(a)l(a)e^{-ra}\dd a \quad\text{,}
+\intertext{which can be expressed as well as:}
+  =& \int_{a=0}^\infty \int_{y=0}^\infty f(a)d(a+y)e^{-ra}\dd y \dd a
+\end{align}
 and note that these are really quite similar, since $\int _{y=0}^\infty
@@
 other axis is thanatological age. For the chronological case \eqref{eq:lotka},
-we multiply chronological age-specific fertility rates over one margin, and for
-the thanatological case over the other margin. There is a strong parallel here
-with the case of two-sex age-specific fertility rates. As in the case of
+we multiply chronological age-specific fertility rates, $f(a)$ over one margin,
+and for the thanatological case we multiply $f^\star(y)$ over the other margin.
+There is a strong parallel here with the case of two-sex age-specific fertility rates. As in the case of
 divergence between male and female single-sex models, there will always be divergence between
 single-sex models under chronological versus thanatological age, unless the
-starting population is stable. Even though the modeled population stocks are in
-a way commensurable, the rates used here, $f(a)$ versus $f^\star(y)$ are
-calculated on the basis of differently distributed denominators $E(a)$ versus $E(y)$.
+starting population is stable. A proof of equivalence in the case of
+$f^\star(y)$ being calculated from a stable starting population is provided
+in Appendix~\ref{app:B}. Even though the modeled population stocks are in a way
+commensurable, the rates used here, $f(a)$ versus $f^\star(y)$ are calculated on the basis of differently distributed denominators $E(a)$ versus $E(y)$.
 
@@
 parameters can be calculated. For instance, we may calculate the mean
-thanatological generation time, $T^y$:
+thanatological generation time, $T^\star$:
 \begin{equation}
 \label{eq:Ty}
- T^y =  \frac{\int _{y=0}^\infty \int _{a=y}^\infty y e^{-ra} d(a) f^\star(y) \dd a
-\dd y}{\int _{y=0}^\infty \int _{a=y}^\infty e^{-ra} d(a) f^\star(y) \dd a \dd
-y} \quad \text{.}
+ T^\star =  \frac{\int _{y=0}^\infty \int _{a=0}^\infty y e^{-ra} d(a+y)
+ f^\star(y) \dd a \dd y}{\int _{y=0}^\infty \int _{a=0}^\infty e^{-ra} d(a+y)
+ f^\star(y) \dd a \dd y} \quad \text{.}
 \end{equation}
@@
 \label{eq:R0fromTy}
-R_0 = e^{r T^y}
+R_0 = e^{r T^\star}
 \end{equation}
 The birth rate, $b$, is given by
-\begin{equation}
+\begin{align}
 \label{eq:eybrate}
-b = \frac{1}{\int _{y=0}^\infty \int _{a=y}^\infty e^{-ra} d(a) \dd a
-\quad\text{.} \dd y}
-\end{equation}
-The stable thanatological age structure, $c_y$, is the
-proportion of the stable population with remaining years to live $y$:
-\begin{equation}
+b =& \frac{1}{\int _{y=0}^\infty \int _{a=0}^\infty e^{-ra} d(a+y) \dd a \dd
+y}\quad\text{,}
+\intertext{the denominator of which is the inside of
+equation~\eqref{eq:thanoren} less $f^\star(y)$. This simplifies to} =&
+\frac{1}{ \int _{a=0}^\infty e^{-ra} l(a) \dd a } \quad \text{,}
+\intertext{which is exactly the same as that valid for
+chronological age in the Lotka-Euler setup, except that $r$ here comes from the
+thanatological renewal model. If the stable age structure, $c^\star(y)$, is
+known, then we may take the sum of hypothetical births over thanatological age:}
+=& \int _{y=0}^\infty f^\star(y) c^\star(y) \dd y \quad{.}
+\end{align}
+
+The stable thanatological age structure, $c^\star(y)$, is the proportion of the stable population with remaining years to live $y$:
+\begin{align}
 \label{eq:cy}
-c(y) = b \int _{a=y}^\infty e^{-ra} d(a) \dd a \quad\text{.}
-\end{equation}
-We already knew that when $r = 0$ and $y = a$, that $c(a)=c(y)$
-\citep{vaupel2009life}, and now we can arrive at the same conclusion directly
-by equating the chronological and thanatological renewal equations. For
+c^\star(y) =& b \int _{a=0}^\infty e^{-ra} d(a+y) \dd a 
+\intertext{or, given the stable chronological age structure and a lifetable:}
+           =& \int _{a=0}^\infty c(a)d(y|a) \dd a\\ \notag \\
+           =& \frac{\int _{a=0}^\infty d(a+y)e^{-ra}\dd a}{\int
+           _{a=0}^\infty l(a)e^{-ra}\dd a}  \quad\text{.}
+\end{align}
+
+It has previously been shown that when $r = 0$ and $y = a$, that
+$c(a)=c^\star(y)$ \citep{brouard1989mouvements, vaupel2009life}, and now we can arrive at the same
+conclusion directly by equating the chronological and thanatological renewal equations. For
 small non-zero magnitude values of $r$ and survival patterns typical of humans, it is also the
-case that $c(y|r)\approx c(a|-r)$. The relationship is asymptotically exact as
-$r$ approaches zero. Other findings regarding stable age structure will be
-presented in a future version of this paper. Further stable population
-quantities may be estimated by similarly translating the various common definitions (e.g.,
+case that $c^\star(y|r)\approx c(a|-r)$. This could make a useful heuristic,
+and the relationship is asymptotically exact as $r$ approaches zero, but the
+exact relationship is to be found in equation~\eqref{} Further
+stable population quantities may be estimated by similarly translating the various common definitions (e.g.,
 in the glossary of \citet{coale1972growth}) to the present perspective. We will
@@
 
-The thanatological renewal model is coherent, but these ideas may take some time
-to gestate before accepting that it is also a valid description of how
+The thanatological renewal model is valid, but this perspective may take some
+time to gestate before accepting that it is also a sound description of how
 populations renew. What should one imagine under the model of thanatological
@@
 (indeed incrementing due to births) until reaching the very bottom. In short,
-the locations of increment and decrement, and the direction of movement are all
-switched. The chronological and thanatological renewal models are almost
-perfectly opposite descriptions of the same two processes.
+the locations of increment and decrement, and the direction of movement (when
+so visualized) are all switched. The chronological and thanatological renewal
+models are almost perfectly opposite descriptions of the same process.
 
@@
 \label{app:A}
-The solution for the intrinsic growth rate, $r$, is unique for the case of the
-thanatological renewal model, \eqref{eq:thanoren}, and can be proven so in
+This appendix contains a brief proof that the real solution for the
+intrinsic growth rate, $r$, is unique for the case of the thanatological renewal model, \eqref{eq:thanoren}, and can be proven so in
 essentially the same fashion as those in existence for the Lotka-Euler model,
-\eqref{eq:lotka}. This little proof follows that given in
+\eqref{eq:lotka}. This proof follows that given in
 \citet{pressat1973analyse}. Define a convenience function, $I(r)$, for the
@@
 of $r$ with boundaries that include the value 1 of \eqref{eq:thanoren}, and
- necessarily only obtain this value once.
+ necessarily only obtain this value once. As with the Lotka-Euler equation, there will be more complex
+ conjugate solutions for $r$ in the thanatological model,
+ and these are not explored in this paper.
 
@@
 This appendix contains a proof that the thanatological renewal equation implies
-the same intrinsic growth rate, $r$, as Lotka's chronological renewal model if
+the same intrinsic growth rate, $r$, as the Lotka-Euler model if
 the starting population is stable. This is identical to claiming that the right
@@
 b(a) &= f(a)c(a)
-\end{align}
-Replacing
-$f^\star (y)$ with the full expanded expression and plugging into the thanatological renewal equation gives:
+\end{align}Replacing $f^\star (y)$ with the full expanded
+expression and plugging into the thanatological renewal equation gives:
 \begin{align}
@@
 f(a)e^{-ra}\mu(a+y)l(a+y) \dd a \dd y
-\intertext{Note also that $l(a) = \int _{x=a}^\infty d(x) \dd x$ (present
+\intertext{Note also that $l(a) = \int _{=a}^\infty d(x) \dd x$ (present
 livings are future deaths), which brings us back to the chronological
@@
 \end{align}
+
+\section{An iterative method to find the thanatological $r$}
+\citet{coale1957new} proposed a fast-converging iterative approach to estimate
+the intrinsic growth rate for the Lotka-Euler equation. For the thanatological
+renewal model, a similar approach may be taken, with some slight
+modifications to Coale's original.
+The following steps can be followed to estimate $r$ from
+Equation~\eqref{eq:thanoren}:
+
+\begin{enumerate}
+  \item Derive a first rough estimate of the mean remaining years of life at
+  reproduction, $\widehat{T^\star}$, akin to Lotka's mean generation time, $T$.
+  To start, a good-enough guess is to just assume $r=0$:
+\begin{equation}
+\widehat{T^\star} = \frac{\int _{y=0}^\infty \int _{a=0}^\infty y d (a+y)
+f^\star(y) \dd a \dd y}{\int _{y=0}^\infty \int _{a=0}^\infty d(a+y) f^\star(y)
+\dd a \dd y}
+\end{equation}
+  \item A first rough guess at the net reproduction rate, $R_0$ is given by
+ \begin{equation}
+  R_0 = \int _{y=0}^\infty \int _{a=0}^\infty d(a+y) f^\star(y) \dd a
+\dd y
+\end{equation}
+  \item A first rough estimate of $r$, $r^0$, is given by
+   \begin{equation}
+   r^0 = \frac{ln(R_0)}{\widehat{T^\star}}
+   \end{equation}
+  \item Plug $r^0$ into Equation~\eqref{eq:thanoren} to calculate a
+  residual, $\delta^0$.
+  \item Use $\delta^0$ and $\widehat{T^\star}$ to calibrate the estimate of $r$
+  using
+  \begin{equation}
+  r^{1} = r^0 + \frac{\delta^0}{\widehat{T^\star} - \frac{\delta^0}{r^0}}
+  \end{equation}
+  \item Repeat step (4) to to derive a new $\delta^i$, then step (5) to refine
+  $r^i$, until converging on a stable $r$ after some 20 or so iterations,
+  depending on the degree of precision desired ($\widehat{T^\star}$ is not updated
+  in this process).
+\end{enumerate}
+
+The above procedure is more computationally efficient than minimizing the
+absolute residual of Equation~\eqref{eq:thanoren} using a generic
+optimizer. Alternatively, one could use the method of cumulants to estimate $r$,
+bbut this would entail less precision than the above (unless one goes beyond,
+say, five cumulants), and this has not yet been worked out at this 
 \end{appendices}
 \nocite{HMD,HFD}
-% --------------------------------------------------
+-------------------------
 % bibliography

### 2014-09-04 ThanoRepro/R/Functions.R

@@ -175,3 +175,55 @@ da2fya <- function(da, stagger = FALSE){
     fya <- Minf0(Mna0(fya / rowSums(fya)))
     fya
 }
+
+getca <- function(La,r,a=.5:110.5){
+    La * exp(a * -r) / sum(La * exp(a * -r))
+}
+rThanoCoale <- compiler::cmpfun(function(Fy, da, .a = .5:110.5, maxit = 1e2, tol = 1e-15){  
+            
+            # Based on Coale (1957), modified.
+            N    <- length(Fy)
+            dxM  <- matrix(0, ncol = N, nrow = N)
+            dxi  <- da
+            for (i in 1:N){
+                dxM[i, 1:length(dxi)  ] <- dxi 
+                dxi <- dxi[2:length(dxi) ]
+            }     
+            R0      <- sum(dxM * Fy)
+            T.guess <- sum(.a * dxM * Fy) / R0 # assuming r = 0
+            r.i     <- log(R0) / T.guess
+            # be careful to discount Fy by SRB appropriately for males / females
+            # prior to specification
+            for (i in 1:maxit){ # 15 is more than enough!
+                #cat(r2,i,"\n")
+                deltai <- 1 - sum(sum(colSums(dxM * exp(-r.i * .a)) * Fy) * Fy)
+                # the mean generation time self-corrects 
+                # according to the error produced by the Lotka equation
+                r.i <- r.i - (deltai / (T.guess - (deltai / r.i))) 
+                if (abs(deltai) <= tol){
+                    break
+                }
+            }
+            return(r.i)  
+        })
+# chronological age
+Rmomentn <- compiler::cmpfun(function(fx,Lx,x,n=0){
+            sum((x^n)*fx*Lx)
+        })
+rLotkaCoale <- compiler::cmpfun(function(fx,Lx,x=.5:110.5, maxit = 1e2, tol = 1e-15){
+            # from Coale, Ansley J. (1957) A New Method for Calculating Lotka's r- the Intrinsic Rate of Growth in a Stable Population.
+            # Population Studies, Vol. 11 no. 1, pp 92-94
+            R0 <- Rmomentn(fx,Lx,x,0)
+            # first assuming a mean generation time of 29
+            ri <- log(R0)/29
+            for (i in 1:maxit){ # 10 is more than enough!
+                deltai <- sum(exp(-ri*x)*fx*Lx)-1
+                # the mean generation time self-corrects 
+                # according to the error produced by the Lotka equation
[…]

### 2014-09-04 ThanoRepro/R/rEqual.R

@@ -1,54 +1,6 @@
 source("/home/tim/git/ThanoRepro/ThanoRepro/R/Functions.R")
 #  thanatological age 
 
-rThanoCoale <- compiler::cmpfun(function(Fy, da, .a = .5:110.5, maxit = 1e2, tol = 1e-15){  
-            
-            # Based on Coale (1957), modified.
-            N    <- length(Fy)
-            dxM  <- matrix(0, ncol = N, nrow = N)
-            dxi  <- da
-            for (i in 1:N){
-                dxM[i, 1:length(dxi)  ] <- dxi 
-                dxi <- dxi[2:length(dxi) ]
-            }     
-            R0      <- sum(dxM * Fy)
-            T.guess <- sum(.a * dxM * Fy) / R0 # assuming r = 0
-            r.i     <- log(R0) / T.guess
-            # be careful to discount Fy by SRB appropriately for males / females
-            # prior to specification
-            for (i in 1:maxit){ # 15 is more than enough!
-                #cat(r2,i,"\n")
-                deltai <- 1 - sum(rowSums(t(t(dxM) * exp(-r.i * .a))) * Fy)
-                # the mean generation time self-corrects 
-                # according to the error produced by the Lotka equation
-                r.i <- r.i - (deltai / (T.guess - (deltai / r.i))) 
-                if (abs(deltai) <= tol){
-                    break
-                }
-            }
-            return(r.i)  
-        })
-# chronological age
-Rmomentn <- compiler::cmpfun(function(fx,Lx,x,n=0){
-            sum((x^n)*fx*Lx)
-        })
-rLotkaCoale <- compiler::cmpfun(function(fx,Lx,x=.5:110.5, maxit = 1e2, tol = 1e-15){
-            # from Coale, Ansley J. (1957) A New Method for Calculating Lotka's r- the Intrinsic Rate of Growth in a Stable Population.
-            # Population Studies, Vol. 11 no. 1, pp 92-94
-            R0 <- Rmomentn(fx,Lx,x,0)
-            # first assuming a mean generation time of 29
-            ri <- log(R0)/29
-            for (i in 1:maxit){ # 10 is more than enough!
-                deltai <- sum(exp(-ri*x)*fx*Lx)-1
-                # the mean generation time self-corrects 
-                # according to the error produced by the Lotka equation
-                ri <- ri+(deltai/(29-(deltai/ri)))
-                if (abs(deltai) <= tol){
[…]

### 2014-09-04 ThanoRepro/R/rMatrixEqual.R

@@ -1,0 +1,44 @@
+source("/home/tim/git/ThanoRepro/ThanoRepro/R/Functions.R")
+
+# I finally got the discrete (pseudo continuous) equation to match
+# perfectly, so now it seems relevant to try again with the matrix.
+# this will require making some compromises on what seems to be best 
+# practice for discretizing actual projection matrices so that things line
+# up exactly. Let's see how far we can take this.
+
+Data                        <- local(get(load("/home/tim/git/ThanoRepro/ThanoRepro/Data/DataAll.Rdata")))
+Data                        <- Data[Data$Sex == "f", ]
+Data$Fy[is.na(Data$Fy)]     <- 0
+Data$Fyf[is.nan(Data$Fyf)]  <- 0
+fa <- with(Data, Fxf[Year == 2000 & Code == "SWE"])
+# 1) use l(a) instead of L(a)
+# further make age strictly integer...
+la2 <- with(Data, lx[Year == 2000 & Code == "SWE"])
+# 2) force d(a) to entail l(a) and vice versa.
+da2 <- -diff(c(la2,0))
+a2      <- 0:110
+(rL2    <- rLotkaCoale(fa,la2,x=a2))
+# 5) recalc stable age and births using these 'clean' inputs
+ca2     <- getca(la2,rL2,a2)
+ba2     <- fa*ca2 
+# 6) and convert these to thanatological equivalents *without*
+# staggering the d(a) vector.
+.cy2    <- rowSums(Thano(ca2,da2,stagger=FALSE))
+.by2    <- rowSums(Thano(ba2,da2,stagger=FALSE))
+# these is the 'clean' stable thanatological fertility schedule
+.fy2    <- .by2 / .cy2
+# ok, now make simplified Leslie:
+
+L <- rbind(fa, cbind(diag(la2[2:111] / la2[1:110]), 0))
+Y       <- outer(da2, .fy2, "*") + rbind(cbind(0, diag(110)), 0)
+
+# 
+log(Re(eigen(L)$values[1]))
+log(Re(eigen(Y)$values[1]))
+
+# 5th decimal. Can it be the close-out?
+
+# same first eigenvalue(roughly), but different first eigenvectors...
+# these are the stable age structures
+plot(Re(eigen(L)$vectors[, 1]),type='l')
+lines(abs(Re(eigen(Y)$vectors[, 1])),col="green

### 2014-09-09 ThanoRepro/ThanoRepro.tex
got perfect matrix match. further edits in text, formula corrections

@@
 Leslie matrices, $\textbf{Y}$ is not sparse, but is instead populated primarily
-with non-zero entries. This is an odd notion, to be explained in following.
+with positive entries greater than zero.
 
@@
 remaining years of life. Thanatological age 1 in year $t$ moves to 0 in year
-$t + 1$. Thus, instead of in the subdiagonal, we place survival in the
-superdiagonal. All survival values are 1, since there is no decrement until
+$t + 1$ and dies in the course of that year. Thus, rather than in the
+subdiagonal, we place survival in the superdiagonal. All survival values are 1, since there is no decrement until
 after living through age 0, and the upper-left corner of this superdiagonal
-contains a 0, for full decrement. In the following, I illustrate using a
-6$\times$6 matrix. The survival component of $\textbf{Y}$ is organized as in
+contains a 0, for full decrement. In the following, I illustrate
+using a 6$\times$6 matrix. The survival component of $\textbf{Y}$ is organized as in
 Matrix~\ref{matrix:Ysurv}. To imagine a lifecycle graph, each age decrements by
@@
  Fertility inputs to the matrix are derived from single-sex thanatological
- fertility and the lifetable $d_a$ distribution, where $a$ indexes age and is
- equated to $y$ remaining years of life for members just born. Fertility
- in a thanatologically structured population occurs in all but the highest
- remaining years classes, which can be thought of as structural zeros for
- simplicity, as they only contain pre-menarchical females. For our
- $6\times 6$ example, say that fertility is observed in classes 0-4, while the
- final class has no fertility, where $f^\star_y$ indicates the fertility
+ fertility and the lifetable $d(a)$ distribution, where $a$ indexes
+ chronological age and is equated to $y$ remaining years of life for members
+ just born. Fertility in a thanatologically structured human population occurs
+ in all but the very highest remaining years classes, which can be thought of as
+ structural zeros for simplicity, as they only contain pre-menarchical females
+ that will have very long lives.
+ For our $6\times 6$ example, say that fertility is observed in classes 0-4, while the
+ final class has no fertility, where $f^\star(y)$ indicates the fertility
  probability for class $y$ in the year $t$ entering population (in the matrix
- columns). Each $f^\star_y$ is then distributed out according to $d_a$ with no
- further translation, since the $d_a$ lifetable column already refers to age 0.
-Thus, the fertility entry in row $m$ and column $n$ of $\textbf{Y}$ will be $f^\star_n \cdot d_m$.
-We assume that those dying over the course of year
- $t$ (the first column) are exposed to fertility for half of the
- year,\footnote{One might be tempted to not allow for fertility at all for
- females dying in year $t$, but recall that fertility is measured in the moment of
- birth, and not conception.}
- and so discount the fertility entry accordingly. Further, infant mortality, 
- $f^\star_y \cdot d_0$, located in the first row, must also be discounted, since part
+ columns). Each $f^\star(y)$ is then distributed out according to $d(a)$ with no
+ further translation, since the $d(a)$ lifetable column already refers to
+ the lifespan distribution of newborns.
+ Thus, the fertility entry in row $m$ and column $n$ of $\textbf{Y}$ will be
+ $f^\star_n \cdot d_m$. One can assume that those dying over the course of year
+ $t$ (the first column) are exposed to fertility for half of the year,\footnote{One might be tempted to not allow for fertility at all for females dying in year $t$, but recall that fertility is measured in the moment of
+ birth, and not conception.} and so discount the fertility entry
+ accordingly.
+ Further, infant mortality, $f^\star_y \cdot d_0$, located in the first row, must also be discounted, since part
  of the mortality will occur in the same year $t$ and the rest in year $t + 1$. 
@@
 Thanatological age-classes will ideally terminate at the highest value permitted
-by data. For the data used here, there are 111 total age classes, which
+by data. For the data used in this paper, there are 111 total age classes, which
 translate to 111 total remaining-years classes (0-110+). In practice $\textbf{Y}$ becomes
@@
 f^\star(y) &= \frac{b^\star(y)}{c^\star(y)} \\
-b^\star (y) &= \int_{a=0}^\infty b(a) \mu(a+y))\frac{l(a+y)}{l(a)} \dd a \\
-c^\star (y) &= \int_{a=0}^\infty c(a) \mu(a+y))\frac{l(a+y)}{l(a)} \dd a
+b^\star (y) &= \int_{a=0}^\infty b(a) \mu(a+y)\frac{l(a+y)}{l(a)} \dd a \\
+c^\star (y) &= \int_{a=0}^\infty c(a) \mu(a+y)\frac{l(a+y)}{l(a)} \dd a
 \intertext{Note that:}
-c(a) &= \frac{l(a)r^{-ra}}{\int l(a) e^{-ra} \dd a} \\
+c(a) &= \frac{l(a)e^{-ra}}{\int l(a) e^{-ra} \dd a} \\
 b(a) &= f(a)c(a)
@@
 bbut this would entail less precision than the above (unless one goes beyond,
-say, five cumulants), and this has not yet been worked out at this 
+say, five cumulants), and this has not yet been worked out at this time.
+
 \end{appendices}

### 2014-09-09 ThanoRepro/R/Functions.R

@@ -210,12 +210,12 @@ rThanoCoale <- compiler::cmpfun(function(Fy, da, .a = .5:110.5, maxit = 1e2, tol
 Rmomentn <- compiler::cmpfun(function(fx,Lx,x,n=0){
             sum((x^n)*fx*Lx)
         })
-rLotkaCoale <- compiler::cmpfun(function(fx,Lx,x=.5:110.5, maxit = 1e2, tol = 1e-15){
+rLotkaCoale <- compiler::cmpfun(function(fx,Lx,x=.5:110.5, T.guess = 29, maxit = 1e2, tol = 1e-15){
             # from Coale, Ansley J. (1957) A New Method for Calculating Lotka's r- the Intrinsic Rate of Growth in a Stable Population.
             # Population Studies, Vol. 11 no. 1, pp 92-94
             R0 <- Rmomentn(fx,Lx,x,0)
             # first assuming a mean generation time of 29
-            ri <- log(R0)/29
+            ri <- log(R0)/T.guess 
             for (i in 1:maxit){ # 10 is more than enough!
                 deltai <- sum(exp(-ri*x)*fx*Lx)-1
                 # the mean generation time self-corrects 

### 2014-09-09 ThanoRepro/R/rMatrixEqual.R

@@ -14,12 +14,20 @@ fa <- with(Data, Fxf[Year == 2000 & Code == "SWE"])
 # 1) use l(a) instead of L(a)
 # further make age strictly integer...
 la2 <- with(Data, lx[Year == 2000 & Code == "SWE"])
+
+L <- rbind(fa, cbind(diag(la2[2:111] / la2[1:110]), 0))
+
 # 2) force d(a) to entail l(a) and vice versa.
 da2 <- -diff(c(la2,0))
 a2      <- 0:110
-(rL2    <- rLotkaCoale(fa,la2,x=a2))
+#(rL2    <- rLotkaCoale(fa,la2,x=a2)
+(rL2 <- log(Re(eigen(L)$values[1])))
 # 5) recalc stable age and births using these 'clean' inputs
 ca2     <- getca(la2,rL2,a2)
+
+ca2 <- Re(eigen(L)$vectors[, 1])
+ca2 <- ca2/sum(ca2)
+
 ba2     <- fa*ca2 
 # 6) and convert these to thanatological equivalents *without*
 # staggering the d(a) vector.
@@ -29,16 +37,64 @@ ba2     <- fa*ca2
 .fy2    <- .by2 / .cy2
 # ok, now make simplified Leslie:
 
-L <- rbind(fa, cbind(diag(la2[2:111] / la2[1:110]), 0))
 Y       <- outer(da2, .fy2, "*") + rbind(cbind(0, diag(110)), 0)
 
-# 
+# wowsa
 log(Re(eigen(L)$values[1]))
 log(Re(eigen(Y)$values[1]))
 
-# 5th decimal. Can it be the close-out?
-
-# same first eigenvalue(roughly), but different first eigenvectors...
-# these are the stable age structures
+## 5th decimal. Can it be the close-out?
+#
+## same first eigenvalue(roughly), but different first eigenvectors...
+## these are the stable age structures
 plot(Re(eigen(L)$vectors[, 1]),type='l')
-lines(abs(Re(eigen(Y)$vectors[, 1])),col="green
+lines(abs(Re(eigen(Y)$vectors[, 1])),col="green")
+
+ca2L <- Re(eigen(L)$vectors[, 1])
+ca2L <- ca2/sum(ca2)
+#ca2L-ca2
+#
[…]

### 2014-09-10 ThanoRepro/ThanoRepro.tex
exploring further matrix properties, and starting rewrite of the matrix section of the paper

@@
 c^\star(y) =& b \int _{a=0}^\infty e^{-ra} d(a+y) \dd a 
-\intertext{or, given the stable chronological age structure and a lifetable:}
-           =& \int _{a=0}^\infty c(a)d(y|a) \dd a\\ \notag \\
+\intertext{or, given the stable chronological age structure, $c(a)$, and a
+lifetable,} =& \int _{a=0}^\infty c(a)d(y|a) \dd a \label{eq:cyfromca}\\ \notag
+\\
            =& \frac{\int _{a=0}^\infty d(a+y)e^{-ra}\dd a}{\int
@@
 and the relationship is asymptotically exact as $r$ approaches zero, but the
-exact relationship is to be found in equation~\eqref{} Further
+exact relationship is to be found in equation~\eqref{eq:cyfromca}. Further
 stable population quantities may be estimated by similarly translating the various common definitions (e.g.,
@@
 focus on the main model rather than on these.
-
-The thanatological renewal model is valid, but this perspective may take some
-time to gestate before accepting that it is also a sound description of how
-populations renew. What should one imagine under the model of thanatological
-population renewal? A useful mnemonic bases itself on Figures~\ref{fig:USdecomp}
-and \ref{fig:USrecomp}. In the chronological age-structured model, new
-generations appear at the bottom of the pyramid, and move up one rung per year.
-All age-classes are subject to attrition, which is spread out over ages and not
-readily visible in the pyramid. In the thanatological \textit{leaf}, each birth cohort increments to the population over the whole range of thanatological age
-according to $d(a)$, as seen in \eqref{eq:thanoren}, becoming the
-shaded layers seen in Figure~\ref{fig:USrecomp}. Each horizontal step is a death
-cohort, and these move one step down the pyramid each year without any decrement
-(indeed incrementing due to births) until reaching the very bottom. In short,
-the locations of increment and decrement, and the direction of movement (when
-so visualized) are all switched. The chronological and thanatological renewal
-models are almost perfectly opposite descriptions of the same process.
 
@@
 matrix, $\textbf{Y}$, is square and of dimension $y \times y$, where $y$ is the number
-of remaining-years classifications on which the population is structured. 
+of thanatological age classifications on which the population is structured. 
 The matrix contains elements for survival and elements for fertility. Unlike
 Leslie matrices, $\textbf{Y}$ is not sparse, but is instead populated primarily
-with positive entries greater than zero.
+with positive entries greater than zero. In the
+following, I illustrate using a 6$\times$6 matrix.
 
-Of interest is that mortality occurs in only the population class with zero
-remaining years of life. Thanatological age 1 in year $t$ moves to 0 in year
-$t + 1$ and dies in the course of that year. Thus, rather than in the
-subdiagonal, we place survival in the superdiagonal. All survival values are 1, since there is no decrement until
-after living through age 0, and the upper-left corner of this superdiagonal
-contains a 0, for full decrement. In the following, I illustrate
-using a 6$\times$6 matrix. The survival component of $\textbf{Y}$ is organized as in
+Mortality occurs in only the population class with zero
+remaining years of life. The population of thanatological age 1 in year $t$
+moves to 0 in year $t + 1$ and dies in the course of that year. Thus, rather than in the
+subdiagonal, survival elements are located in the superdiagonal. All survival
+values are 1, since there is no decrement until after living through age 0. The survival component of $\textbf{Y}$ is organized as in
 Matrix~\ref{matrix:Ysurv}. To imagine a lifecycle graph, each age decrements by
 one through each lower age in successive order until being absorbed by death
-after completing thanatological age zero. This chain unfolds in the opposite
-index-direction of survival in the chronological age system.
+after completing thanatological age zero.
 
@@
  Fertility inputs to the matrix are derived from single-sex thanatological
- fertility and the lifetable $d(a)$ distribution, where $a$ indexes
+ fertility and the lifetable $d_a$ distribution, where $a$ indexes
  chronological age and is equated to $y$ remaining years of life for members
- just born. Fertility in a thanatologically structured human population occurs
- in all but the very highest remaining years classes, which can be thought of as
- structural zeros for simplicity, as they only contain pre-menarchical females
+ just born. It is simpler to imagine $d_a$ in this case as indexing the
+ future lifespan distribution of newborns. Fertility in a thanatologically
+ structured human population occurs in all but the very highest remaining years classes, which can be thought of as structural zeros for simplicity, as they only contain pre-menarchical females
  that will have very long lives.
- For our $6\times 6$ example, say that fertility is observed in classes 0-4, while the
- final class has no fertility, where $f^\star(y)$ indicates the fertility
- probability for class $y$ in the year $t$ entering population (in the matrix
- columns). Each $f^\star(y)$ is then distributed out according to $d(a)$ with no
- further translation, since the $d(a)$ lifetable column already refers to
- the lifespan distribution of newborns.
- Thus, the fertility entry in row $m$ and column $n$ of $\textbf{Y}$ will be
+ For our $6\times 6$ example, say that fertility is observed in
+ thanatological classes 0-4, while the final class has no fertility, where
+ $f^\star_y$ indicates the fertility probability for class $y$ in the year $t$
+ entering population (in the matrix columns). Each $f^\star_y$ is then
+ distributed out over rows according to $d_a$.
+ The fertility entry in row $m$ and column $n$ of $\textbf{Y}$ will therefore be
  $f^\star_n \cdot d_m$. One can assume that those dying over the course of year
@@
  in order to account for the fact that infant mortality is higher in the lower Lexis 
- triangle than in the upper: of those infants who die in the first year of life, a proportion equal to
- $\lambda$ do not make it to December \nth{31} of the calendar year in which
+ triangle than in the upper. Of those infants who die in the first year of life,
+ a proportion equal to $\lambda$ do not make it to December \nth{31} of the calendar year in which
  they were born.\footnote{$\lambda$ can be derived directly from death counts
- data classified by Lexis triangles. In the US, $\lambda$ has behaved similarly
- for males and females, falling steadily from around $0.9$ in 1969 to $0.86$
- around 1990, since which time it has steadily risen to around $0.87$. That is
- to say, $\lambda$ has varied, but not drastically. These numbers are just meant to give a
-feel for the ranges that $\lambda$ can be expected to receive. If the demographer
+ data classified by Lexis triangles. If the demographer
  does not have information to derive $\lambda$ directly, ad hoc or semidirect
- methods may be used to assign a reasonable proportion. } The
- fertility component of $\textbf{Y}$ is then composed as in Matrix~\ref{matrix:Yfert}.
+ methods may be used to assign a reasonable proportion, such as 0.9 for
+ contemporary low mortality populations.} The fertility component of $\textbf{Y}$ is then composed as in Matrix~\ref{matrix:Yfert}.
 
@@
 Figure~\ref{fig:graph} depicts the projection matrix $\textbf{Y}$ in the form of
-a lifecycle graph. In this graph, each thanatological age is a node. Population
-members are born into any of the thanatological age classes, and then follow
+a lifecycle graph.  In this graph, each thanatological age is a node, and death
+is the absorbing state, denoted with an ``X''.
+Population members are born into any of the thanatological age classes, and then follow
 clockwise around the graph on the thick blue paths in order until finally being
@@
 (y)$, and these births distribute out to all other age classes according to the
-lifetable deaths distribution, $d(a)$. Hence, the rate instensity of a fertility
-path is determined by the fertility rate of the origin age, and the probability of dying at the destination age. A fraction of the fertility destined for infant mortality equal
-to $lambda$ leads straight to the absorbing state of death. Fertility paths from
-the \nth{4} thanatological age class have been highlighted. These correspond
-to the fifth column of the example matrix $\textbf{Y}$, plus an additional path
+lifespan distribution from the lifetable, $d_a$. Hence, the rate instensity of a
+fertility path is determined by the fertility rate of the origin age and
+probability of having a lifespan equal to the order of the destination node.  A fraction of
+the fertility destined for infant mortality equal to $\lambda$ leads straight to the absorbing state of death.
+Fertility paths from the \nth{4} thanatological age class have been highlighted.
+These correspond to the fifth column of the example matrix $\textbf{Y}$, plus an additional path
 direct to death for the fraction of infant mortality prior to December \nth{31}
@@
 
+The
+thanatological lifecycle graph is directed and strongly connected. The matrix is
+also redudcible because each thanatological age class is reachable from every
+other thanatological age class, either directly via fertility, or else by ageing toward lower ages. The thanatological matrix is also
+easy to identify as primitive, since it contains self-loops (fertility elements
+on the matrix diagonal). The matrix is said to be regular, since we can
+imagine some value $k$ where each thanatological age class is $k$ steps away
+from every other thanatological age class. In the fertility block of the matrix
+(all rows, but not necessarily all columns), $k$ is 0, due to direct connections
+between all fertile thanatological ages, but in order to make this condition hold for pre-menarcical females with very long lives, $k$ must be equal to the maximum lifespan.
+This can also be demonstrated simply by
+setting $\textbf{Y}^k$ to some large value such that all elements of $\textbf{Y}
+> 0$.
+
 \section*{Some preliminary empirical findings}
@@
 will be necessary to shed more light on this finding.
+
+\section{Discussion}
+The thanatological renewal model is valid, but this perspective may take some
+time to gestate before accepting that it is also a sound description of how
+populations renew. What should one imagine under the model of thanatological
+population renewal? A useful mnemonic bases itself on Figures~\ref{fig:USdecomp}
+and \ref{fig:USrecomp}. In the chronological age-structured model, new
+generations appear at the bottom of the pyramid, and move up one rung per year.
+All age-classes are subject to attrition, which is spread out over ages and not
+readily visible in the pyramid. In the thanatological \textit{leaf}, each birth cohort increments to the population over the whole range of thanatological age
+according to $d(a)$, as seen in \eqref{eq:thanoren}, becoming the
+shaded layers seen in Figure~\ref{fig:USrecomp}. Each horizontal step is a death
+cohort, and these move one step down the pyramid each year without any decrement
+(indeed incrementing due to births) until reaching the very bottom. In short,
+the locations of increment and decrement, and the direction of movement (when
+so visualized) are all switched. The chronological and thanatological renewal
+models are almost perfectly opposite descriptions of the same process.
 

### 2014-09-10 ThanoRepro/R/rMatrixEqual.R

@@ -43,12 +43,42 @@ Y       <- outer(da2, .fy2, "*") + rbind(cbind(0, diag(110)), 0)
 log(Re(eigen(L)$values[1]))
 log(Re(eigen(Y)$values[1]))
 
-## 5th decimal. Can it be the close-out?
+install.packages("matrixcalc")
+library(matrixcalc)
+is.singular.matrix(Y,tol=1e-8)
+is.singular.matrix(Y,tol=1e-9)
+
+is.singular.matrix(L,tol=1e-12)
+is.definite(L)
+
+
+install.packages("expm", repos="http://R-Forge.R-project.org")
+library(expm)
+all(Y%^%100 > 0) # yes, thano is regular
+all(L%^%100 > 0)
+sum(L)
+
+# mixing time
+
+# http://stanford.edu/class/ee363/lectures/pf.pdf
+1/log(1/rev(sort(abs(Re(eigen(Y)$values))))[2])
+1/log(1/rev(sort(abs(Re(eigen(L)$values))))[2])
+
+# decreasing magnitude (from JHJ 2006 standford workshop notes)
+Ldamping <- Mod(eigen(L)$values[2])/Mod(eigen(L)$values[1])  
+Ydamping <- Mod(eigen(Y)$values[2])/Mod(eigen(Y)$values[1])  
+Yperiod       <- 2*pi/ Arg(eigen(Y)$values[2])
+Lperiod       <- 2*pi/ Arg(eigen(L)$values[2])
+
+# approx: number of steps over which deviation from equilibrium
+#distribution decreases by factor e
+#
+# 5th decimal. Can it be the close-out?
 #
 ## same first eigenvalue(roughly), but different first eigenvectors...
 ## these are the stable age structures
-plot(Re(eigen(L)$vectors[, 1]),type='l')
-lines(abs(Re(eigen(Y)$vectors[, 1])),col="green")
+#plot(Re(eigen(L)$vectors[, 1]),type='l')
+#lines(abs(Re(eigen(Y)$vectors[, 1])),col="green")
 
 ca2L <- Re(eigen(L)$vectors[, 1])
 ca2L <- ca2/sum(ca2)
@@ -79,22 +109,71 @@ ca2L <- ca2/sum(ca2)
 #log(Re(eigen(L3)$values[1]))
 #log(Re(eigen(Y3)$values[1]))
 
-da2
-M <- diag(da2)
-col(M)
[…]

### 2014-09-11 ThanoRepro/ThanoRepro.tex
changes throughout text

@@
 mortality is assumed invariant over time, which means that the same survival
-pattern underlies each chronological age class, and we therefore draw from a
-single lifetable. Under these conditions, data classified by chronological age
-can be decomposed and redistributed into thanatological age classes with a
-single mortality pattern. The basic operation is to multiply the probability of
-surviving from exact chronological age $a$ to exact age $a+y$,
-$\frac{l(a+y)}{l(a)}$, and then dying at exact age $a+y$, $\mu(a+y)$. The
-index $y$ in this case means $y$ years in the future. For age 0 this probability
-is simply the $d(a)$ column of the lifetable, standardized to sum to 1.
-For higher ages, simply condition on survival:\footnote{In the reliability literature, our concept of $d(y|a)$ is denoted by \begin{equation} X - t | X > t
+pattern underlies each birth cohort. This is the case we explore, and so the
+following formulas assume a single lifetable.
+
+The basic transformation to thanatological age proceeds by defining a
+deaths density function, similar the lifetable $d(a)$, but over two
+dimensions: chronological age attained, $a$, and remaining years of life, $y$.
+This function we denote $d(y|a)$ --- The density distribution of thanatological
+age depends on chronological age. I use the same function name $d()$, because
+it is essentially the same notion as the familiar $d(a)$, but rescaled as a
+function of chronological age. For ages higher than 0, simply condition
+on survival:\footnote{In the reliability literature, our concept of $d(y|a)$ is denoted by \begin{equation} X - t | X > t
 \end{equation}
@@
 \begin{equation}
+\label{eq:dyadiscret}
 d_{y, a} =  \frac{d_{a+y}+d_{a+y+1}}{L_a + L_{a+1}} \quad\quad \text{,}
@@
 
-\noindent where $\omega$ is the highest age. One
-produces in this way a probability density function over future death times
-for each chronological age, $a$, and the $d(y|a)$- weighted average of $y$ for
-a given $a$ returns the familiar remaining life expectancy column of the lifetable,
-$e(a)$. If $P(a)$ is an chronological age-classified population count,
-we can derive $P(y)$, a thanatological age-classified population count, as
-follows:
+\noindent where $\omega$ is the highest
+age.\footnote{Equation~\eqref{eq:dyadiscret} ill incur some error depending on
+the method used to derive $L_a$. To remove this error, rescale over the $a$ margin so that each distribution conditioned on $a$
+sums to 1. Working with $l_a$ rather than $L_a$ eliminates error and yields
+similar results, but is not necessarily better.} One produces in this way a
+probability density function over future death times for each chronological age,
+$a$. The $d(y|a)$- weighted average of $y$ for a given $a$ returns the familiar
+remaining life expectancy column of the lifetable, $e(a)$. If $P(a)$ is a
+chronological age-classified population count, we can derive $P(y)$, a
+thanatological age-classified population count using ~\eqref{eq:transform}.
 
@@
 population.
-Figure~\ref{fig:USrecomp} shows this \textit{re}composition for the same 2010
+Figure~\ref{fig:USrecomp} shows this recomposition for the same 2010
 US data.
@@
 glance at population attrition given the population stock and lifetable of a
-particular moment, whereas Figure~\ref{fig:USdecomp} is reflective in nature, since a population's chronological age-structure is mostly the fruit of past fertility, but also migration and to a lesser degree attrition.
-Note that any count classified by chronological age can be
-reclassified in this way, as long as you have a lifetable that plausibly represents the population whose data you wish to reclassify.
-This means we can derive thanatological fertility rates, $F_y$, by applying \eqref{eq:transform}
-to age-classified birth counts, $B_a$, to get $B_y$ and again to
-exposure-to-risk, $E_a$ (here we take exposures from all ages), to get $E_y$ and then dividing:
+particular moment, whereas Figure~\ref{fig:USdecomp} is reflective in nature,
+since a population's chronological age-structure is mostly the fruit of past
+fertility, but also migration and to a lesser degree attrition. Any count
+classified by chronological age can be reclassified in this way given a
+corresponding lifetable. Exploiting this, we derive thanatological fertility
+rates, $f^\star(y)$, by applying \eqref{eq:transform} to age-classified birth
+counts, $B(a)$, to get $B(y)$ and again to exposure-to-risk, $E(a)$ (here we
+take exposures from all ages), to get $E(y)$ and then dividing:
 
 \begin{equation}
-f^\star_y = \frac{B_y}{E_y}
+f^\star(y) = \frac{B(y}{E(y)}
 \end{equation}
 
-We are interested to use thanatological fertility rates, $F_y$ in
+We are interested to use thanatological fertility rates, $f^\star(y)$ in
 models of population renewal,\footnote{This is probably premature, but helps to
@@
 populations and years as a more familiar reference, but note that both scales
-are different! With no indication of particular populations or time series, one
+are different. With no indication of particular populations or time series, one
 already concludes that period thanatological fertility rates have a
@@
 rightward shift in both the mean and modal mother's thanatological age at birth
-(TAB) over time (that's unambiguously a good thing). Several populations now
-show modal TABs of over 60 years. Thanatological total fertility rates (TTFR)
+(TAB) over time. Several
+populations now show modal TABs of over 60 years. Thanatological total fertility rates (TTFR)
 track standard period total fertility rates (PTFR) rather closely, but tend to
@@
 
-In general, such fertility rates will not be palatable for purposes of
+In general, thanatological fertility rates will not be useful for purposes of
 projection, unless the demographer believes that their empirical regularity is
-somehow stronger than typical ASFR --- an argument we do not make. There may be
-alternative ways of defining thanatological fertility rates, for instance by
-transforming from a stable population, which would make the resulting model more
-conformable with the Lotka renewal model. First let us define the basic model.
+somehow stronger than typical ASFR. This is a question we do not seek to answer
+in the present treatment. There are alternative ways of defining
+thanatological fertility rates, for instance by transforming the birth and
+population vector from a stable population. As proven in appendix~\ref{app:B},
+thanatological fertility rates derived from a stable population imply a certain
+symmetry with stable populations from the chronological model. First let us
+define the basic model.
 
@@
 of fertility rates shown in Figure~\ref{fig:FySpaghetti}, and we refer to a
-unisex population based on female vital rates. In the following, I
-explain describe the model in four different ways: as a continuous
-model, as a matrix model, with a lifecylcle graph, and intuitively by
-appealing to how population pyramids and leaves (described earlier) change over
-time. Assuming constant vital rates, the births for the present year are given by
+closed, unisex population based on female vital rates. Assuming constant vital rates, the
+births for the present year are given by 
 \begin{align}
@@
 =& \int_{y=0}^\infty \int_{a=0}^\infty f^\star(y) B(t-a)d(a+y)\dd a \dd y
-\quad\text{.} \intertext{Eventually-- usually a bit quicker than is the case for
-chronological age-- strong ergodicity will assert itself, and $B(t)$ will be related to
+\quad\text{.} \intertext{Eventually strong ergodicity will assert itself, and $B(t)$ will be related to
 $B(t-a)$ according to a constant factor $e^{ra}$, where $r$ is the familiar
@@
   =& \int_{a=0}^\infty \int_{y=0}^\infty f(a)d(a+y)e^{-ra}\dd y \dd a
+  \label{eq:lotka2}
 \end{align}
-and note that these are really quite similar, since $\int _{y=0}^\infty
-d(a+y)\dd a = l(a)$. For intuition, notice that $l(a)$ is here split
+and note that \eqref{eq:lotka2} and \eqref{eq:thanoren} are really quite
+similar, since $\int _{y=0}^\infty d(a+y)\dd a = l(a)$. For intuition, notice that $l(a)$ is here split
 up into pieces of $d(a)$, and imagine a 2D surface of these, where one axis is chronological age and the
 other axis is thanatological age. For the chronological case \eqref{eq:lotka},
-we multiply chronological age-specific fertility rates, $f(a)$ over one margin,
-and for the thanatological case we multiply $f^\star(y)$ over the other margin.
-There is a strong parallel here with the case of two-sex age-specific fertility rates. As in the case of
-divergence between male and female single-sex models, there will always be divergence between
-single-sex models under chronological versus thanatological age, unless the
-starting population is stable. A proof of equivalence in the case of
-$f^\star(y)$ being calculated from a stable starting population is provided
-in Appendix~\ref{app:B}. Even though the modeled population stocks are in a way
-commensurable, the rates used here, $f(a)$ versus $f^\star(y)$ are calculated on the basis of differently distributed denominators $E(a)$ versus $E(y)$.
+we multiply chronological age-specific fertility rates, $f(a)$, over the
+chronological age margin, and for the thanatological case we multiply
+$f^\star(y)$ over the thanatological age margin.
 
@@
 although it is rather intuitive, given the much greater density of connections
-within the model; persons from nearly any thanatologcal age\footnote{I say
+within the model; persons from nearly any thanatologcal age can produce offspring that can have any other thanatological age.\footnote{I say
 \textit{nearly} because the very highest thanatological ages tend to be
@@
 transform to thanatological closes out at age 110+, as is the case with HMD data, then fertility rates for thantological ages 101-110+ would be structural
-zeros.} can produce offspring that can have any other thanatological age. In
+zeros.} In
 this sense, the smoothing mechanism at play as the population passes through
 time must be much stronger than that for chronological age, at least in most
-cases.
-A further conjecture is that the jump will also hold for the case of weak ergodicity.
-The mechanisms at play unfold in the same way as those so intuitively described by \citet{arthur1982ergodic}, and said proof may apply here without further modification.
-A proof of the uniqueness of the solution to \eqref{eq:thanoren} is given in Appendix~\ref{app:A}.
+cases.  The mechanisms at play unfold in the same way as those so intuitively
+described by \citet{arthur1982ergodic}, and said proof may apply here without
+further modification. Later in this paper I note how strong ergodicity is
+guaranteed given the properties of the thanatolgical projection matrix. A proof
+of the uniqueness of the solution to \eqref{eq:thanoren} is given in
+Appendix~\ref{app:A}. A proof that the chronological and thanatological renewal
+models must yield the same $r$ if the starting population is already stable is
+given in Appendix~\ref{app:B}. A fast-converging iterative method for finding
+$r$ from equation~\eqref{eq:thanoren} is given in Appendix~\ref{app:C}
+
+If the thanatolgical fertility rates do not derive from a stable
+population, as will typicallybe the case, then the two age perspectives
+will, unless by coincidence, yield divergent stable models. There is a strong
+parallel here with the case of two-sex age-specific fertility rates. As in the
+case of divergence between male and female single-sex models, there will
+virtually always be divergence between single-sex models under chronological
+versus thanatological age, unless the starting population is stable.
+Even though the modeled population stocks are in a way commensurable, the rates used here, $f(a)$ versus $f^\star(y)$ are calculated on the basis of differently distributed denominators $E(a)$ versus $E(y)$.
 
@@
 
-As with Leslie matrices, the above projection matrix may be manipulated using
-standard matrix techniques. Where $\textbf{p}$ is our population vector, we
-project by multiplying $\textbf{Y}$ from the left: $\textbf{p}(t + 1) =
-\textbf{Y}\textbf{p}(t)$. From $\textbf{Y}$, we can extract such information as 
-the intrinsic growth rate, $r$ (natural log of the largest
-real eigenvalue), the stable thanatological age-structure (the real part of the
-eigenvector that corresponds to the largest real eigenvalue), or the pace of
-convergence to stability (the ratio of the \nth{1} to the \nth{2} eigenvalues).
+Thanatological projection matrices may be manipulated using
+standard matrix techniques applicable to the Leslie matrix. Where $\textbf{p}$
+is a population vector classified by thanatological age, projection proceeds by
+multiplying $\textbf{Y}$ from the left:
+$\textbf{p}(t + 1) = \textbf{Y}\textbf{p}(t)$.
+
+For rate schedules typical of human populations, thanatological projection
+matrices are irreducible and primitive, and so will always have a
+dominant positive real eigenvalue, the natural log of which is $r$. This
+observation is sufficient to demonstrate the strong ergodic
+theorem:\footnote{This step was taken for granted in jumping from
+equation~\eqref{eq:ergostep1} to ~\eqref{eq:ergostep2}} i.e., given a fixed
+matrix and enough time, the population will conform to some stable thanatological age distribution.
+The eigenvector corresponding to the dominant eigenvalue gives this stable thanatological age structure.
+The square
+of the, or the pace of convergence to stability (the ratio of the \nth{1} to the \nth{2} eigenvalues).
 \footnote{See \citet[p.86-87]{caswell2001matrix}.}
-
-The
-thanatological lifecycle graph is directed and strongly connected. The matrix is
-also redudcible because each thanatological age class is reachable from every
-other thanatological age class, either directly via fertility, or else by ageing toward lower ages. The thanatological matrix is also
-easy to identify as primitive, since it contains self-loops (fertility elements
-on the matrix diagonal). The matrix is said to be regular, since we can
-imagine some value $k$ where each thanatological age class is $k$ steps away
-from every other thanatological age class. In the fertility block of the matrix
-(all rows, but not necessarily all columns), $k$ is 0, due to direct connections
-between all fertile thanatological ages, but in order to make this condition hold for pre-menarcical females with very long lives, $k$ must be equal to the maximum lifespan.
-This can also be demonstrated simply by
-setting $\textbf{Y}^k$ to some large value such that all elements of $\textbf{Y}
-> 0$.
 
@@
 \section{An iterative method to find the thanatological $r$}
+\label{app:C}
 \citet{coale1957new} proposed a fast-converging iterative approach to estimate

### 2014-09-11 ThanoRepro/R/rMatrixEqual.R

@@ -43,20 +43,37 @@ Y       <- outer(da2, .fy2, "*") + rbind(cbind(0, diag(110)), 0)
 log(Re(eigen(L)$values[1]))
 log(Re(eigen(Y)$values[1]))
 
-install.packages("matrixcalc")
+#install.packages("matrixcalc")
 library(matrixcalc)
-is.singular.matrix(Y,tol=1e-8)
-is.singular.matrix(Y,tol=1e-9)
-
+is.singular.matrix(Y,tol=1e-8) # precision
+is.singular.matrix(Y,tol=1e-9) # here it woks
 is.singular.matrix(L,tol=1e-12)
-is.definite(L)
+
+Leig <- eigen(L)$values
+Yeig <- eigen(Y)$values
+cbind(Leig,Yeig)
+rads <- seq(0,2*pi,length=1000)
+
+pdf("LYeigen.pdf",height=4.5,width=9)
+par(mfrow=c(1,2))
+plot(NULL, xlim=c(-1,1),ylim=c(-1,1),asp=1,axes=FALSE, xlab = "Real", ylab = "Complex",
+        main = "eigenvalues on the unit circle")
+lines(cos(rads),sin(rads))
+points(Re(Leig),Im(Leig),col = "#0000FF60",pch=19,cex=.5)
+points(Re(Yeig),Im(Yeig),col = "#FF000060",pch=19,cex=.5)
+
+plot(0:110, sqrt(Re(Yeig)^2+Im(Yeig)^2), type='l',col="red",ylim=c(0,1),
+        main = "radius")
+lines(0:110, sqrt(Re(Leig)^2+Im(Leig)^2), col="blue")
+dev.off()
+getwd()
 
 
-install.packages("expm", repos="http://R-Forge.R-project.org")
+#install.packages("expm", repos="http://R-Forge.R-project.org")
 library(expm)
 all(Y%^%100 > 0) # yes, thano is regular
-all(L%^%100 > 0)
-sum(L)
+all(L%^%100 > 0) # Leslie not regular?
+
 
 # mixing time
 

### 2014-09-11 ThanoRepro/ThanoRepro.tex
changes to perron frob paragraph

@@
 For rate schedules typical of human populations, thanatological projection
-matrices are irreducible and primitive, and so will always have a
-dominant positive real eigenvalue, the natural log of which is $r$. This
-observation is sufficient to demonstrate the strong ergodic
-theorem:\footnote{This step was taken for granted in jumping from
-equation~\eqref{eq:ergostep1} to ~\eqref{eq:ergostep2}} i.e., given a fixed
-matrix and enough time, the population will conform to some stable thanatological age distribution.
-The eigenvector corresponding to the dominant eigenvalue gives this stable thanatological age structure.
-The square
-of the, or the pace of convergence to stability (the ratio of the \nth{1} to the \nth{2} eigenvalues).
+matrices are filled with non-negative values, mostly greater than zero. Raising
+the matrix to some not-very-large power, $k$, will make all entries greater than zero. This observation, or else by noting that the
+lifecylce graph is strongly connected, lets us conclude that the matrix is
+irreducible and primitive. By the Perron-Frobenius theorem, the matrix will
+ always have a unique, dominant, positive real eigenvalue, the natural
+log of which is the intrinsic growth rate, $r$, and strong ergodicity is
+assured.
+In other words, given a fixed matrix and enough time, the population will conform to some stable thanatological age distribution. Strong
+ergodicity was taken for granted in jumping from equation~\eqref{eq:ergostep1}
+to ~\eqref{eq:ergostep2}, but the presently described properties of the matrix
+model should, albeit unrigorously, satisfy lingering uncertainty on that step.
+The eigenvector corresponding to the dominant eigenvalue gives this stable
+thanatological age structure. If the fertility rates placed in the matrix are
+from the stable population, then the growth rate from the thanatological
+projection matrix is equal to that of the standard Leslie
+matrix.\footnote{Some care needs to be taken in order to demonstrate this
+point, since common approximations and adjustments used in matrix
+construction can throw things off. \texttt{R} code is available from the
+author to demonstrate this point. See also Appendix~\ref{app:B} for the
+corresponding proof from the continuous model.} The square of the, or the pace
+of convergence to stability (the ratio of the \nth{1} to the \nth{2} eigenvalues).
 \footnote{See \citet[p.86-87]{caswell2001matrix}.}

### 2014-09-11 ThanoRepro/ThanoRepro.tex
intro and eq punctuation (not done)

@@
 \newcommand{\dd}{\; \mathrm{d}}
+\newcommand{\ec}{\quad\quad\text{,}}
+\newcommand{\ep}{\quad\quad\text{.}}
 
@@
 All demographic forces vary over age in known and regular ways. Information on
-such forces, along with the size and age structure of a population allows the
+such forces, along with the size and age structure of a population, allows the
 demographer to make predictions about the future size and structure of a population. In this
@@
 until death\footnote{This quantity is referred to as \textit{residual} lifetime
-in the reliability literaure.}. Individuals in this case move in the same
-direction along imaginary life lines, but the reference point for age is now at
-the end of the life line instead of at the beginning. For individuals, the
-timing of events in life is mirrored exactly between the two definitions of age,
+in the reliability literaure.}. The first kind of age is called
+chronological age and the kind of age explored in this paper is
+called thanatological age. Individuals in this case move in the same direction
+along imaginary life lines, but the reference point for age is now at the end of the life line instead of at the beginning.
+For individuals, the timing of events in life is mirrored exactly between the two definitions of age,
 but aggregate patterns differ.
@@
 reproduction are plausibly a function of both age perspectives, and the question
-of how to partition such forces and events over time is a new conundrum that I
-do not tackle here. The model presented here should therefore not be interpreted
-as a suggestion for a better way to model or project population, but as an
-alternative way to tell the story of population renewal.
+of how to best partition such forces and events over time is a new conundrum
+that I do not tackle here. Instead I describe some of the formal consequences
+of the Lotka-Euler renewal model when viewed from the thanatological perspective
+of age. The renewal model presented here is not necessarily a suggestion for a
+better way to model or project population, but in the minimal sense it
+represents an alternative way to tell the story of population renewal.
 
-Thanatological (counting down towards death) and chronological (counting up
-from birth) renewal models, such as the Lotka-Euler renewal model, both
-approximate the movement of population stocks over time, but from opposite
-vantage points. Specifying a model of thanatological renewal expands the toolbox
-available to approach complex temporal population structure. I present a set of
-preliminary findings to help understand the implications of temporal structure.
+The chronological-age classified Lotka-Euler renewal model and the
+thanatological renewal model both approximate the movement of population stocks
+over time, but from opposite vantage points. Specifying a model of thanatological renewal expands the toolbox
+available to approach complex temporal population structure. The next section
+of this article describes the lifetable method of transforming age-classified
+data into thanatological age classes. I then derive the continuous version of
+the thanatological renewal model, and some other stable properties. This is
+followed by the matrix version of the model, which is explained alongside the
+lifecycle graph. I present a set of preliminary empirical findings to help
+understand the implications of temporal structure.
 
@@
 factor in future changes in mortality, and to allow the trajectory of changes
-to vary between ages/cohorts. In the case of a stable population model,
+to vary over ages and cohorts. In the case of a stable population model,
 mortality is assumed invariant over time, which means that the same survival
 pattern underlies each birth cohort. This is the case we explore, and so the
-following formulas assume a single lifetable.
+following formulas assume a single lifetable pattern.
 
@@
 This function we denote $d(y|a)$ --- The density distribution of thanatological
-age depends on chronological age. I use the same function name $d()$, because
-it is essentially the same notion as the familiar $d(a)$, but rescaled as a
+age conditional on chronological age. We use the same function name $d()$,
+because it is essentially the same notion as the familiar $d(a)$, but rescaled as a
 function of chronological age. For ages higher than 0, simply condition
@@
 \label{eq:vaupel1}
-d(y | a) = \mu(a+y)\frac{l(a+y)}{l(a)}
+d(y | a) = \mu(a+y)\frac{l(a+y)}{l(a)} \ep
 \end{equation}
@@
 \label{eq:dyadiscret}
-d_{y, a} =  \frac{d_{a+y}+d_{a+y+1}}{L_a + L_{a+1}} \quad\quad \text{,}
+d_{y, a} =  \frac{d_{a+y}+d_{a+y+1}}{L_a + L_{a+1}} \ec
 \end{equation}
@@
 chronological age-classified population count, we can derive $P(y)$, a
-thanatological age-classified population count using ~\eqref{eq:transform}.
+thanatological age-classified population count using
 
@@
 \label{eq:transform}
-P(y) =& \int_0^\infty P(a) d(y | a) \dd a
+P(y) =& \int_0^\infty P(a) d(y | a) \dd a \ep
 \end{align}
@@
 counts, $B(a)$, to get $B(y)$ and again to exposure-to-risk, $E(a)$ (here we
-take exposures from all ages), to get $E(y)$ and then dividing:
+take exposures from all ages), to get $E(y)$ and then dividing
 
 \begin{equation}
-f^\star(y) = \frac{B(y}{E(y)}
+f^\star(y) = \frac{B(y}{E(y)} \ep
 \end{equation}
@@
 B(t) =& \int _0^\infty P(y,t)f^\star(y) \dd y = \int _0^\infty P(a,t)f(a) \dd a
-\quad\text{,} \intertext{where $a$ indexes chronological age, $y$ indexes
+\ec \intertext{where $a$ indexes chronological age, $y$ indexes
 thanatological age, $P(a), P(y)$ are population counts, and $f^\star(y), f(a)$ are
@@
 be broken down back in terms of chronological age:} =& \int_{y=0}^\infty
-\int_{a=0}^\infty f^\star(y)\frac{P(a,t)d(a+y)}{l(a)} \dd a \dd y \quad\text{,}
+\int_{a=0}^\infty f^\star(y)\frac{P(a,t)d(a+y)}{l(a)} \dd a \dd y \ec
 \intertext{where $d(a)$ is the continuous lifetable death distribution with
@@
 \label{eq:ergostep1}
-=& \int_{y=0}^\infty \int_{a=0}^\infty f^\star(y) B(t-a)d(a+y)\dd a \dd y
-\quad\text{.} \intertext{Eventually strong ergodicity will assert itself, and $B(t)$ will be related to
-$B(t-a)$ according to a constant factor $e^{ra}$, where $r$ is the familiar
+=& \int_{y=0}^\infty \int_{a=0}^\infty f^\star(y) B(t-a)d(a+y)\dd a \dd y \ep 
+\intertext{Eventually strong ergodicity will assert itself, and $B(t)$ will
+be related to $B(t-a)$ according to a constant factor $e^{ra}$, where $r$ is the familiar
 intrinsic rate of growth:}
@@
 =& \int_{y=0}^\infty \int_{a=0}^\infty f^\star(y) B(t)e^{-ra}d(a+y)\dd a \dd y
-\quad\text{.}
+\ep
 \intertext{Divide out $B(t)$ to get back a familiar-looking renewal equation:}
@@
 1 =& \int_{y=0}^\infty \int_{a=0}^\infty f^\star(y) d(a+y)e^{-ra}\dd a \dd y
-\quad\text{.}
+\ep
 \end{align}
@@
 \label{eq:lotka}
-1 =& \int_{a=0}^\infty f(a)l(a)e^{-ra}\dd a \quad\text{,}
-\intertext{which can be expressed as well as:}
-  =& \int_{a=0}^\infty \int_{y=0}^\infty f(a)d(a+y)e^{-ra}\dd y \dd a
+1 =& \int_{a=0}^\infty f(a)l(a)e^{-ra}\dd a \ec
+\intertext{which can be expressed as well as}
+  =& \int_{a=0}^\infty \int_{y=0}^\infty f(a)d(a+y)e^{-ra}\dd y \dd a \ec
   \label{eq:lotka2}
@@
  f^\star(y) \dd a \dd y}{\int _{y=0}^\infty \int _{a=0}^\infty e^{-ra} d(a+y)
- f^\star(y) \dd a \dd y} \quad \text{.}
+ f^\star(y) \dd a \dd y} \ep
 \end{equation}
@@
 b =& \frac{1}{\int _{y=0}^\infty \int _{a=0}^\infty e^{-ra} d(a+y) \dd a \dd
-y}\quad\text{,}
+y}\ec
 \intertext{the denominator of which is the inside of
 equation~\eqref{eq:thanoren} less $f^\star(y)$. This simplifies to} =&
-\frac{1}{ \int _{a=0}^\infty e^{-ra} l(a) \dd a } \quad \text{,}
+\frac{1}{ \int _{a=0}^\infty e^{-ra} l(a) \dd a } \ec
 \intertext{which is exactly the same as that valid for
@@
 known, then we may take the sum of hypothetical births over thanatological age:}
-=& \int _{y=0}^\infty f^\star(y) c^\star(y) \dd y \quad{.}
+=& \int _{y=0}^\infty f^\star(y) c^\star(y) \dd y \ep
 \end{align}
 
-The stable thanatological age structure, $c^\star(y)$, is the proportion of the stable population with remaining years to live $y$:
+The stable thanatological age structure, $c^\star(y)$, is the proportion of the stable population with remaining years to live $y$
 \begin{align}
 \label{eq:cy}
-c^\star(y) =& b \int _{a=0}^\infty e^{-ra} d(a+y) \dd a 
+c^\star(y) =& b \int _{a=0}^\infty e^{-ra} d(a+y) \dd a \ec
 \intertext{or, given the stable chronological age structure, $c(a)$, and a
@@
            =& \frac{\int _{a=0}^\infty d(a+y)e^{-ra}\dd a}{\int
-           _{a=0}^\infty l(a)e^{-ra}\dd a}  \quad\text{.}
+           _{a=0}^\infty l(a)e^{-ra}\dd a}  \ep
 \end{align}

### 2014-09-11 ThanoRepro/ThanoRepro.tex
added caveat to r proof

@@
 the same intrinsic growth rate, $r$, as the Lotka-Euler model if
-the starting population is stable. This is identical to claiming that the right
-side of the rewal equation is equal given the same $r$. Since there is a unique
+the starting population is stable, given a particular set of thanatological
+fertility rates.
+This is identical to claiming that the right side of the rewal equation is equal given the same $r$. Since there is a unique
 real solution, demonstrated in the previous appendix, then it is sufficient to
@@
 \end{align}
+This finding essentially proves that in the stable population there is at least
+one thanatological fertility rate schedule that will satisfy the constraints of
+the lifespan distribution and $r$. However, it is not necessary to derive the
+thanatological fertility distribution in the way prescribed here, and indeed an
+infinite number of such fertility distributions would satisfy the same
+constraints. Imagine these fertility rates cross-classified in
+both age dimensions, where two marginal sums give the thanatological and
+chronological rates, as they would be calculated over the
+corresponding cross-classified exposures. The shape of this rate surface can be
+shifted around in infinitely many ways that would preserve a given growth rate
+and lifespan distribution, and the above proof shows one such mapping.
 

### 2014-09-12 ThanoRepro/ThanoRepro.tex
started to peck at empirical findings. Maybe it'd be cooler to get into eigenvalues more...

@@
 The thanatological renewal equation and the discrete projection matrix can be
-put to work with data. At this time we are only ready to report some early
-results, and we do not yet have explanations some of the patterns that we
-report. Of our 1834 population-years, we optimize $r$ from both
-\eqref{eq:thanoren} and \eqref{eq:lotka}\footnote{It is possible to optimize $r$ using a variety of approximations, or by using a generic optimizer.
-We have used the method proposed by \citet{coale1957new} for Lotka's $r$ and a modified version of the same for
-the thanatological $r$. Details and / or implementation code available on
-request.}. In this sample, both versions of $r$ are only plausibly equal in a
-single instance. Usually thantological $r$ is greater than Lotka's $r$ (1373
-cases). When Lotka's $r$ is positive (693 cases), thanatological $r$ is greater
-just over of 50\% of the time (356), but when Lotka's $r$ is negative,
+put to work with data. In this case a few options are available for determining
+what kind of thanatological fertility rates to use. The two obvious choices are
+to derive rates from the stable population implied by the chronological renewal
+model, or to derive rates such as those from Figure~\ref{fig:FySpaghetti},
+based on a particular stock of population.
+
+Of the 1834 population-years of data on hand, we optimize $r$ from both
+\eqref{eq:thanoren} and \eqref{eq:lotka}. In this sample, both versions of $r$
+are only plausibly equal in a single instance. Usually thanatological $r$ is
+greater than Lotka's $r$ (1373 cases). When Lotka's $r$ is positive (693 cases), thanatological $r$ is greater just over of 50\% of the time (356), but when Lotka's $r$ is negative,
 thanatological $r$ is the greater of the two around 90\% of the time (1017
@@
 than for chronological $r$, with the ratio of variances (thano./chrono.) of
-about 0.75. Usually the two theoretical values of $r$ move in the same direction, but thanatological $r$
-is over time the less erratic of the two, and it usually paints a less dire picture when both are negative.
+about 0.75. The two theoretical values of $r$ covary strongly, but
+thanatological $r$ is the less erratic of the two, and it usually paints a less dire picture when both are negative.
 
@@
 The differences described here are not due primarily to rounding errors in the
-data, but rather to the way in which $f^\star$ has been calculated. If the
-births and exposure used to calculate $f^\star$ come from a population that is not stable
-then $f^\star$ may be subject to some compositional ruses. Specifically, in this
-case we have transformed fertility rates from empirical data with a particular
-chronological age structure. If the chronological age structure was far from its
-stable distribution, then it is hard to imagine that $f^\star$ would remain
-fixed over time. Some simple contrived examples will be presented in a future
-version of this paper in order to eliminate this problem.
+data, but rather to the way in which $f^\star(y)$ has been calculated. 
+Specifically, in this case we have transformed fertility rates from empirical data with a particular
+chronological age structure. If the chronological age structure is far from its
+stable distribution, then it may be hard to imagine that $f^\star(y)$ will
+remain fixed over time. When holding chronological-age classified fertility rates
+fixed, the assumption is just the opposite.
 
@@
 eigenvalue, the damping ratio, is an indicator of the \textit{springiness} of
-the stable population structure, $c_y$ or $c_a$, with respect to disturbances from the stable state
-as determined by vital rate distributions. A higher damping ratio means that the
-population structure oscillates back to its stable state faster, i.e., that
-oscillations decrease in size more rapidly. The thanatological damping ratio was
+the stable population structure, $c^\star(y)$ or $c(a)$, with respect to
+disturbances from the stable state as determined by vital rate distributions. A
+higher damping ratio means that the population structure oscillates back to
+its stable state faster, i.e., that oscillations decrease in size more rapidly. The thanatological damping ratio was
 greater than the Leslie damping ratio for all 1834 population-years included in

### 2014-09-12 ThanoRepro/ThanoRepro.tex
added justifications, per josh's suggestion

@@
 of the Lotka-Euler renewal model when viewed from the thanatological perspective
-of age. The renewal model presented here is not necessarily a suggestion for a
-better way to model or project population, but in the minimal sense it
-represents an alternative way to tell the story of population renewal.
+of age.
+
+The renewal model is the core of modern demography, as it elegantly relates
+fertility, mortality, and population structure, revealing the long
+term consequences of a particular set of conditions. Some might think that
+stable population theory is not relevant to everyday demography, since
+populations seem to always be changing. Still, most projections seem to trail
+off into a stable state, and stability or even stationarity are indeed
+imaginable for the world population of humans one day. I present a renewal model
+based on thanatological age not necessarily as a suggestion for a
+better way to model or project population (or certain aspects of it), although I
+plant this is an open and unexplored empirical question.
+In a minimal sense the thanatological renewal model represents an alternative
+and coherent way to describe the process of population renewal. In a maximal sense, the model may
+actually prove useful, as a model relating renewal to a death flow rather than
+to a birth flow. This enriches our understanding of lifespan and vitality, but also
+of the side effects of our species' renewal, such as the human biomass cycle.
+
+Further, there is likely a range of lifecourse transitions that covary
+more with remaining lifetime that with life lived, but these are mostly unstudied at this
+time,\footnote{I suppose that the prime candidates are late-life transitions.}
+and it is beneficial to expand the formal demographic base by incorporating such information into models. In an even more general sense, by
+thinking about remaining lifetime as a particular class of time-to-transition, one may generalize the lifetable methods presented here into new area
+applications, such as education, partering, the economic lifecycle, or migration.
+The present inquiry is justified possibly for its own sake, and definitely as an
+attempt at accounting for remaining time structure in general in modeling that may prove useful.
 
@@
 thanatological renewal model both approximate the movement of population stocks
-over time, but from opposite vantage points. Specifying a model of thanatological renewal expands the toolbox
-available to approach complex temporal population structure. The next section
-of this article describes the lifetable method of transforming age-classified
+over time, but from opposite vantage points. Specifying a model of
+thanatological renewal expands the toolbox available to approach complex
+temporal population structure. The next section of this article describes the lifetable method of transforming age-classified
 data into thanatological age classes. I then derive the continuous version of

### 2014-09-12 ThanoRepro/ThanoRepro.tex
corrected some spelling

@@
 populations seem to always be changing. Still, most projections seem to trail
-off into a stable state, and stability or even stationarity are indeed
+off into a stable state, and stability or even stationarity are
 imaginable for the world population of humans one day. I present a renewal model
@@
 thinking about remaining lifetime as a particular class of time-to-transition, one may generalize the lifetable methods presented here into new area
-applications, such as education, partering, the economic lifecycle, or migration.
+applications, such as education, partnering, the economic lifecycle, or
+migration.
 The present inquiry is justified possibly for its own sake, and definitely as an
@@
 
-If the thanatolgical fertility rates do not derive from a stable
-population, as will typicallybe the case, then the two age perspectives
+If the thanatological fertility rates do not derive from a stable
+population, as will typically be the case, then the two age perspectives
 will, unless by coincidence, yield divergent stable models. There is a strong
@@
 (y)$, and these births distribute out to all other age classes according to the
-lifespan distribution from the lifetable, $d_a$. Hence, the rate instensity of a
+lifespan distribution from the lifetable, $d_a$. Hence, the rate intensity of a
 fertility path is determined by the fertility rate of the origin age and
@@
 fertility rates.
-This is identical to claiming that the right side of the rewal equation is equal given the same $r$. Since there is a unique
-real solution, demonstrated in the previous appendix, then it is sufficient to
+This is identical to claiming that the right side of the renewal equation is
+equal given the same $r$. Since there is a unique real solution, demonstrated in the previous appendix, then it is sufficient to
 show that the two integral equations are equal under these conditions. The basic
@@
 optimizer. Alternatively, one could use the method of cumulants to estimate $r$,
-bbut this would entail less precision than the above (unless one goes beyond,
+but this would entail less precision than the above (unless one goes beyond,
 say, five cumulants), and this has not yet been worked out at this time.

### 2014-10-17 ThanoRepro/R/DataPrep.R
country selection now different. should be slightly faster now, as we only pull the necessary countries

@@ -16,6 +16,8 @@ library(reshape2)
 
 # IRL and ESP are preliminary countries in HFD at time of this writing
 HFDcountries <- unique(c(getHFDcountries(),c("IRL","ESP")))
+HMDcountries <- getHMDcountries()
+Allcountries <- intersect(HFDcountries, HMDcountries)
 
 # This will take a LONG time to generate. Inefficient web parsing. FYI
 # 1)
@@ -32,7 +34,7 @@ if (!"us" %in% ls()){
 Sys.Date()
 # save it to a little file so you don't forget.
 cat("HFD data downloaded on", as.character(Sys.Date()), file = "Data/HFDdate.txt")
-Bx <- do.call(rbind, lapply(HFDcountries, function(XXX, us, pw){
+Bx <- do.call(rbind, lapply(Allcountries, function(XXX, us, pw){
       Dat <- readHFDweb(CNTRY = XXX, item = "birthsRR", username = us, password = pw)
       data.frame(Code = XXX, Dat, stringsAsFactors = FALSE)
     }, us = us, pw = pw))
@@ -40,7 +42,7 @@ Bx <- do.call(rbind, lapply(HFDcountries, function(XXX, us, pw){
 colnames(Bx)[colnames(Bx) == "Total"] <-"Births"
 
 # same thing for HFD exposures (can be slightly different from HMD exposures)
-Ex <- do.call(rbind, lapply(HFDcountries, function(XXX, us, pw){
+Ex <- do.call(rbind, lapply(Allcountries, function(XXX, us, pw){
       Dat <- readHFDweb(CNTRY = XXX, item = "exposRR", username = us, password = pw)
       data.frame(Code = XXX, Dat, stringsAsFactors = FALSE)
     }, us = us, pw = pw))
@@ -81,12 +83,6 @@ HFD         <- rbind(HFD, m)
 
 # -------------------------------------------------
 # now prep HMD data
-
-#list.files(HMDpath)
-HMDcountries <- getHMDcountries()
-
-Allcountries <- intersect(HFDcountries, HMDcountries)
-
 
 # make similar long format for HMD data, Deaths, Exp, Pop Counts all smacked together. 
 # Later we'll put on Fert data to make a single awesome object.
@@ -191,7 +187,8 @@ save(Data,file = "Data/DataAll.Rdata")
 
 # ------------------------------------------------------------
 # now we need to read in deaths by Lexis triangle to calculate age 0 lambda for
-# thanatological projection matrices.
[…]

### 2014-10-20 ThanoRepro/R/PlotFunctions.R
added plotting functions to paper package

@@ -1,154 +1,0 @@
-
-makeVectorAgeGroups <- function(vec, ages = 0:110, N = 2){
-    x.new  <- ages - ages %% N
-    c(tapply(vec, x.new, sum))
-}
-# draw rect based on x,y of cetroid, plus width and height
-makeRect <- function(x, y, w, h, ...){
-    w2 <- w / 2
-    h2 <- h / 2
-    rect(x - w2, y - h2, x + w2, y + h2, ...)
-}
-
-PyramidOutline2 <- function(males, females, 
-        scale = sum(c(males, females)), 
-        x = 0, y = 0, ...){
-    N       <- length(males)
-    Total   <- sum(c(males, females), na.rm = TRUE)
-    widths  <- rep(1, N)
-    age     <- c(0,cumsum(widths)[-N])
-    u.age   <- age[N] + widths[N]
-    
-    males   <- scale * males / Total
-    females <- scale * females / Total
-    
-    polygon(x = c(0, rep(females, each = 2) + 0,0, rev(c(-0, rep(-males, each = 2) - 0, -0))) + x, 
-            y =  c(rep(c(age, u.age), each = 2), rev(c(rep(c(age, u.age), each = 2)))) + y, ...)
-}
-
-# Males, Females should be the full thano x chrono population matrix, as returned by Thano()
-# Pyramid for age-structure. FALSE for thanatological age structure instead
-# N interval width for remaining-years heterogeneity (age pyramid) or age heterogeneity (thano leaf)
-# BrewerPal the colo ramp from color brewer package.
-# revcol you can decide direction of ramp explicitly
-# x,y for positioning of pyramid
-# ycex, for vertical scaling.
-# ---------------------
-# *only does pyramids / leaves scaled to N  <- 1
-PyramidOrLeafWithHeterogeneity <- function(Males, Females, Pyramid = TRUE,
-        N = 10, BrewerPal = "YlGnBu", revcol = FALSE,x=0,y=0){
-    # define color functions
-    if (revcol){
-        colR        <- grDevices::colorRampPalette(rev(RColorBrewer::brewer.pal(9, BrewerPal)), space = "Lab")
-    } else {
-        colR        <- grDevices::colorRampPalette(RColorBrewer::brewer.pal(9, BrewerPal), space = "Lab")
-    }
-    # if we're plotting a thanatological leaf
-    if (!Pyramid){
[…]

### 2014-10-20 ThanoRepro/R/RiffeFunctions/R/PlotFunctions.R

@@ -1,0 +1,209 @@
+
+# this script contains auximilary functions used in plotting in this paper. These are peculiar and not necessarily useful beyond their application in the paper scripts. Since I am lame and do all plotting in base, in invariably end up with myriad such custom figure functions...
+
+#'
+#' @title makeVectorAgeGroups move single age data into age groups
+#' 
+#' @param vec a vector of data to be grouped
+#' @param ages vector of single ages. \code{0:110} by default to match HMD.
+#' @param N the age interval for the new groups.
+#' 
+#' @return a vector of grouped counts
+#' 
+#' @export
+#' 
+#' 
+
+makeVectorAgeGroups <- function(vec, ages = 0:110, N = 2){
+    x.new  <- ages - ages %% N
+    c(tapply(vec, x.new, sum))
+}
+# draw rect based on x,y of cetroid, plus width and height
+#'
+#' @title makeRect auxiliary function
+#' 
+#' @description draws a rectangle centered on the given coordinates and of specified width and height
+#' 
+#' @param x center x coord.
+#' @param y center y coord.
+#' @param w width.
+#' @param h height.
+#' @param ... optional arguments passed to \code{rect()}.
+#' 
+#' @return NULL function called for its plotting side effects
+#' 
+#' @export
+
+makeRect <- function(x, y, w, h, ...){
+    w2 <- w / 2
+    h2 <- h / 2
+    rect(x - w2, y - h2, x + w2, y + h2, ...)
+}
+#'
+#' @title PyramidOrLeafWithHeterogeneity plots either a leaf or a pyramid
+#' 
+#' @description This function takes cross-classified age matrices (thanatological age in rows and chronological age in columns, and will plot it either as a pyramid with death-cohorts highlighted with a color ramp or a thanatological leaf, with birth cohorts highlighted with a color ramp. I might not be very useful outside of this paper. The figure is added to an already-open device of the appropriate dimension. Scaled to sum to 1, as used in the paper. That's fixed here, so you'l need to re-define the function in order to make it more flexible...
+#' 
+#' @param Males a matrix of counts cross-classified by single chronological and thanatological ages, as returned by \code{Thano()}. These plot on the left.
[…]

### 2014-10-20 ThanoRepro/R/RiffeFunctions/raw/build.R
building now increments version

@@ -16,4 +16,5 @@ library(devtools)
 devtools::document("R/RiffeFunctions")
 #devtools::load_all("R/RiffeFunctions")
 
-
+library("TimUtils")
+IncrementVersion("R/RiffeFunctions","1",package.origin="2014-01-01")

### 2014-10-20 ThanoRepro/R/HFDspaghetti.R
made robust

@@ -9,11 +9,15 @@ if (system("hostname",intern=TRUE)=="triffe-N80Vm"){
 
 
 source("R/Functions.R")
+library(data.table)
 # read in data and select females
 Data <- local(get(load("Data/DataAll.Rdata")))
 Data <- Data[Data$Sex == "f", ]
+DT <- as.data.table(Data)
 
 
+playWithData <- FALSE
+if(playWithData){
 
 # wow, the max age-specific rate is .29! WTF?
 #max(Data$Fx)
@@ -77,10 +81,8 @@ lines(with(Data, Age[Year >= 2000]), with(Data, Fx[Year >= 2000]), col = "#00000
 plot(NULL, type="n",xlim = c(0,95),ylim=c(0,.15))
 lines(with(Data, Age[Year >= 2000]), with(Data, Fy[Year >= 2000]), col = "#00000008")
 
-# try with assumDifferent x and y
+
 # 
-library(data.table)
-DT <- as.data.table(Data)
 
 st <- function(x){
     x/sum(x,na.rm=TRUE)
@@ -103,12 +105,12 @@ lines(Data$Age, Data$FySt, col = "#00000005")
 
 plot(NULL, type="n",xlim = c(0,95),ylim=c(0,.11))
 lines(Data$Age, Data$FxSt, col = "#00000005")
-
+}
 
 graphics.off()
 # Fy 
 png("Figures/FySpaghettiDraft.png")
-#pdf("/home/tim/git/ThanoRepro/ThanoRepro/Figures/FySpaghetti.pdf") # high res vector
+#pdf("Figures/FySpaghetti.pdf") # high res vector
 par(xaxs = "i", yaxs = "i", mai=c(.5,.5,.5,.5))
 plot(NULL, type = "n",xlim = c(0,95),ylim=c(0,.13), axes = FALSE, xlab = "", ylab = "",
         panel.first = list(rect(0,0,95,.15,col = gray(.95),border = NA),
@@ -124,7 +126,7 @@ dev.off()
 
 # Fx 
 png("Figures/FxSpaghettiDraft.png")
-#pdf("/home/tim/git/ThanoRepro/ThanoRepro/Figures/FxSpaghetti.pdf") # high res vector
+#pdf("Figures/FxSpaghetti.pdf") # high res vector
 par(xaxs = "i", yaxs = "i", mai=c(.5,.5,.5,.5))
[…]

### 2014-10-20 ThanoRepro/R/RiffeFunctions/R/PlotFunctions.R
added actually used pyramid functions

@@ -42,7 +42,7 @@ makeRect <- function(x, y, w, h, ...){
 #'
 #' @title PyramidOrLeafWithHeterogeneity plots either a leaf or a pyramid
 #' 
-#' @description This function takes cross-classified age matrices (thanatological age in rows and chronological age in columns, and will plot it either as a pyramid with death-cohorts highlighted with a color ramp or a thanatological leaf, with birth cohorts highlighted with a color ramp. I might not be very useful outside of this paper. The figure is added to an already-open device of the appropriate dimension. Scaled to sum to 1, as used in the paper. That's fixed here, so you'l need to re-define the function in order to make it more flexible...
+#' @description Deprecated. Switched code to use \code{plotPyr()}. This function takes cross-classified age matrices (thanatological age in rows and chronological age in columns, and will plot it either as a pyramid with death-cohorts highlighted with a color ramp or a thanatological leaf, with birth cohorts highlighted with a color ramp. I might not be very useful outside of this paper. The figure is added to an already-open device of the appropriate dimension. Scaled to sum to 1, as used in the paper. That's fixed here, so you'l need to re-define the function in order to make it more flexible...
 #' 
 #' @param Males a matrix of counts cross-classified by single chronological and thanatological ages, as returned by \code{Thano()}. These plot on the left.
 #' @param Females a matrix of counts cross-classified by single chronological and thanatological ages, as returned by \code{Thano()}. These plot on the right.
@@ -207,3 +207,68 @@ colorStrip <- function(x = .0105,y=5,w=.002,h = 100,BrewerPal,revcol = FALSE,ris
     
 }
 
+
+
+#' @title plotPyr plots heterogenous pyramid using output from \code{prepare4plot()}
+#' @description This is an internal function, very specific and rigid.
+#' 
+#' @param XXX list returned by \code{GetThanoMatrices()} with male and female matrix elements
+#' @param x coord of center of figure
+#' @param y corrd of bottom of figure
[…]

### 2014-10-20 ThanoRepro/R/RiffeFunctions/R/PlotFunctions.R
added aggN, replaces makeVectorAgeGroups...

@@ -2,6 +2,22 @@
 # this script contains auximilary functions used in plotting in this paper. These are peculiar and not necessarily useful beyond their application in the paper scripts. Since I am lame and do all plotting in base, in invariably end up with myriad such custom figure functions...
 
 #'
+#' @title aggN aggregate vector of single age (year) data to N-year groups
+#' 
+#' @description This only makes sense if N is 5 or 10 (or 2 I guess), since it uses modulo to find groups
+#'    Ages are assumed to start at 0 and count up in single ages.
+#' 
+#' @param x the vector of single-age-classified data
+#' @param N desired width of interval, e.g., 5 or 10
+#' 
+#' @export 
+#' 
+aggN <- function(x,N){
+  age <- 0:(length(x) - 1)
+  tapply(x,age - age %% N, sum)
+}
+
+#'
 #' @title makeVectorAgeGroups move single age data into age groups
 #' 
 #' @param vec a vector of data to be grouped

### 2014-10-20 ThanoRepro/R/Pyramids.R
script now uses paper package

@@ -7,58 +7,20 @@ if (system("hostname",intern=TRUE)=="triffe-N80Vm"){
   setwd(paste0("/data/commons/",system("whoami",intern=TRUE),"/git/ThanoRepro/ThanoRepro"))
 }
 source("R/Functions.R")
-source("R/PlotFunctions.R")
+#source("R/PlotFunctions.R")
+#
+#
+#devtools::load_all("/home/triffe/git/Leaves/PlosOne/R/RiffeetalFunctions")
 
 
-devtools::load_all("/home/triffe/git/Leaves/PlosOne/R/RiffeetalFunctions")
 
-plotPyr <- function(XXX,x,y,Thano=TRUE){
-    N <- ncol(XXX[["Cm"]])
-    mgrab <- ifelse(Thano,"Tm","Cm")
-    fgrab <- ifelse(Thano, "Tf", "Cf")
-    cgrab <- ifelse(Thano, "Tcol", "Ccol")
-    for (i in N:1){
-        m <- rep(XXX[[mgrab]][, i], each = 2)
-        f <- rep(XXX[[fgrab]][, i], each = 2)
-        polygon(c(m, rev(f)) + x, XXX[["y"]] + y, border = NA, col = XXX[[cgrab]][i])
-    }
-    
-    PyramidOutline(abs(XXX[[mgrab]][, N]), XXX[[fgrab]][, N], 
-            x = x, y = y, scale = 1, border = gray(.1), lwd = .2)
-}
-
-# OK data object in order
-# transform to plot-specific data
-prepare4plot <- function(XXX, colsFun1, colsFun2){
-    Mf          <- XXX$Females
-    Mm          <- XXX$Males
-    TOT         <- sum(Mf) + sum(Mm)
-    
-    Het1f       <- apply(Mf ,2, aggN, N = 10) / TOT             # age with ey het
-    Het1m       <- apply(Mm ,2, aggN, N = 10) / TOT             # age with ey het
-    Het2f       <- apply(Mf ,1, aggN, N = 10) / TOT             # age with ey het
-    Het2m       <- apply(Mm ,1, aggN, N = 10) / TOT 
-    
-# cumulative sum for cleaner plotting
-    cumHet1f    <- t(apply(Het1f, 2, cumsum))
-    cumHet2f    <- t(apply(Het2f, 2, cumsum))
[…]

### 2014-10-20 ThanoRepro/ThanoRepro.tex
tinkering with tikz, won't compile on refugee server due to old ubuntu version...

@@
 \usepackage{subcaption}
+%\usepackage{pgf}
 \usepackage{tikz}
+%\usetikzlibrary{arrows}
 %\usepackage[active,tightpage]{preview}
+%\PreviewEnvironment{tikzpicture}
+%\setlength\PreviewBorder{5pt}
 \usepackage{natbib}
@@
 %-------------------------------------------------------
-
 \begin{document}
@@
 
+% both figures created in R/Pyramids.R
 \begin{figure}[ht!]
@@
 populations now show modal TABs of over 60 years. Thanatological total fertility rates (TTFR)
-track standard period total fertility rates (PTFR) rather closely, but tend to
+track standard period total ferti
+lity rates (PTFR) rather closely, but tend to
 be somewhat higher (not shown).
-
+% both figures created in R/HFDspaghetti.R
 \begin{figure}[h!]
@@
 \label{fig:graph}
-\begin{tikzpicture}[->,shorten >=1pt,auto,node distance=4.3cm,
-  thick,main node/.style={circle,fill=blue!20,draw,font=\sffamily\Large\bfseries}]
+\begin{tikzpicture}[->,shorten>=1pt,auto,node distance=4.3cm,thick,main
+node/.style={circle,fill=blue!20,draw,font=\sffamily\Large\bfseries}]
   \definecolor{gray}{rgb}{0.8,0.8,0.8}
   \definecolor{lightgreen}{rgb}{0,0.8,0.8}
+
   \node[main node] (0) {0};
@@
 
-    \path[every node/.style={font=\sffamily\small}, ultra thick, color =
-    lightgreen ] (5) edge node {} (4)
+    \path[every node/.style={font=\sffamily\small}, ultra thick, color=lightgreen ] 
+    (5) edge node {} (4)
     (4) edge node {} (3)
@@
     (0) edge node {} (X);    
-
 \end{tikzpicture}

### 2014-11-20 ThanoRepro/R/DataPrep.R
update headers for new laptop

@@ -1,6 +1,6 @@
 
 # for Tim, this will choke
-if (system("hostname",intern=TRUE)=="triffe-N80Vm"){
+if (system("hostname",intern=TRUE) %in% c("triffe-N80Vm", "tim-ThinkPad-L440")){
   # if I'm on the laptop
   setwd("/home/tim/git/ThanoRepro/ThanoRepro")
 } else {

### 2014-11-20 ThanoRepro/R/MatrixStuff.R

@@ -1,6 +1,6 @@
 library(popbio)
 # for Tim, this will choke
-if (system("hostname",intern=TRUE)=="triffe-N80Vm"){
+if (system("hostname",intern=TRUE) %in% c("triffe-N80Vm", "tim-ThinkPad-L440")){
   # if I'm on the laptop
   setwd("/home/tim/git/ThanoRepro/ThanoRepro")
 } else {

### 2014-11-20 ThanoRepro/R/rCalc.R

@@ -1,5 +1,5 @@
 # for Tim, this will choke
-if (system("hostname",intern=TRUE)=="triffe-N80Vm"){
+if (system("hostname",intern=TRUE) %in% c("triffe-N80Vm", "tim-ThinkPad-L440")){
   # if I'm on the laptop
   setwd("/home/tim/git/ThanoRepro/ThanoRepro")
 } else {

### 2014-11-22 ThanoRepro/R/DataPrep.R
.

@@ -12,6 +12,7 @@ source("R/Functions.R")
 library(compiler)
 library(data.table)
 library(reshape2)
+library(DemogBerkeley)
 #list.files(HFDpath)
 
 # IRL and ESP are preliminary countries in HFD at time of this writing

### 2014-11-22 ThanoRepro/ThanoRepro.tex
maybe something is broken in tikz. will need to handle it on laptop...
generated figs, fixed tikz to current version, slight change in sentence in damping. Need to think more about how far to go in results based on unstable populations, introduce the concept clearer

@@
 \usepackage{tikz}
-%\usetikzlibrary{arrows}
+\usetikzlibrary{arrows}
 %\usepackage[active,tightpage]{preview}
@@
 \label{fig:graph}
-\begin{tikzpicture}[->,shorten>=1pt,auto,node distance=4.3cm,thick,main
+\begin{tikzpicture}[->,auto,node distance=4.3cm,thick,main
 node/.style={circle,fill=blue!20,draw,font=\sffamily\Large\bfseries}]
@@
 eigenvalue, the damping ratio, is an indicator of the \textit{springiness} of
-the stable population structure, $c^\star(y)$ or $c(a)$, with respect to
-disturbances from the stable state as determined by vital rate distributions. A
-higher damping ratio means that the population structure oscillates back to
+the stable population structure, $c^\star(y)$ or $c(a)$, if perturbed from the
+state of stability.
+A higher damping ratio means that the population structure oscillates back to
 its stable state faster, i.e., that oscillations decrease in size more rapidly. The thanatological damping ratio was

### 2015-10-12 ThanoRepro/R/AgeStructure.R
1) getting started on chernoff pyramids. I can't locate that old work! unbelievable! 2) replacing DemogBerkeley with HMDHFDplus, regenerating results, as new data has since been created in both databases.

@@ -1,0 +1,25 @@
+
+# Author: tim
+###############################################################################
+
+# for Tim, this will choke
+if (system("hostname",intern=TRUE) %in% c("triffe-N80Vm", "tim-ThinkPad-L440")){
+	# if I'm on the laptop
+	setwd("/home/tim/git/ThanoRepro/ThanoRepro")
+} else {
+	# in that case I'm on Berkeley system, and other people in the dept can run this too
+	setwd(paste0("/data/commons/",system("whoami",intern=TRUE),"/git/ThanoRepro/ThanoRepro"))
+}
+
+library(HMDHFDplus)
+source("R/Functions.R")
+
+# 1) read in example survival curves. How about Sweden 1970, just because?
+
+flt <- readHMDweb("SWE","fltper_1x1",password=pw,username=us)
+mlt <- readHMDweb("SWE","mltper_1x1",password=pw,username=us)
+
+
+
+
+

### 2015-10-12 ThanoRepro/R/DataPrep.R

@@ -12,7 +12,7 @@ source("R/Functions.R")
 library(compiler)
 library(data.table)
 library(reshape2)
-library(DemogBerkeley)
+library(HMDHFDplus)
 #list.files(HFDpath)
 
 # IRL and ESP are preliminary countries in HFD at time of this writing

### 2015-10-12 ThanoRepro/R/Functions.R

@@ -14,17 +14,18 @@ PACKAGES <- installed.packages()[,"Package"]
 # see https://github.com/UCBdemography/DemogBerkeley for installation tips.
 # requires some external programs be installed to run properly. If you try and fail, then email 
 # the author. This is only used for reading HMD/HFD data into R.
-if (!"DemogBerkeley" %in% PACKAGES){
-  A <- try(devtools::install_github("UCBdemography/DemogBerkeley", subdir = "DemogBerkeley"))
-  if (class(A) == "try-error"){
-    cat("\nDemogBerkeley didn't install. \nsee https://github.com/UCBdemography/DemogBerkeley \n for installation tips")
-  }
-}
-require(DemogBerkeley)
+#if (!"DemogBerkeley" %in% PACKAGES){
+#  A <- try(devtools::install_github("UCBdemography/DemogBerkeley", subdir = "DemogBerkeley"))
+#  if (class(A) == "try-error"){
+#    cat("\nDemogBerkeley didn't install. \nsee https://github.com/UCBdemography/DemogBerkeley \n for installation tips")
+#  }
+#}
+
+# TR: swapped in HMDHFDplus, much more stable!
+require(HMDHFDplus)
 
 
 devtools::load_all("R/RiffeFunctions")
-
 
 
 

### 2015-10-12 ThanoRepro/R/DataPrep.R
still fixing data prep to regenerate. Saved snapshot of prior paper version. Will now edit the running version. This will cnange character I suspose

@@ -22,14 +22,16 @@ Allcountries <- intersect(HFDcountries, HMDcountries)
 
 # This will take a LONG time to generate. Inefficient web parsing. FYI
 # 1)
-if (!"pw" %in% ls()){
+if (!exists("pw")){
   cat("enter HFD password into console (no quotes) and press enter\n")
   pw <- userInput()
 }
-if (!"us" %in% ls()){
+
+if (!exists("us")){
   cat("enter HFD username into console (no quotes) and press enter\n")
   us <- userInput()
 }
+
 
 # HFD data downloaded on:
 Sys.Date()
@@ -90,7 +92,7 @@ HFD         <- rbind(HFD, m)
 
 # note when we download the data:
 cat("HMD data downloaded on", as.character(Sys.Date()), file = "Data/HMDdate.txt")
-Data <- do.call(rbind, lapply(Allcountries, function(XXX,pw,us){
+HMD <- do.call(rbind, lapply(Allcountries, function(XXX,pw,us){
       
       flt <- readHMDweb(CNTRY = XXX, item = "fltper_1x1", username = us, password = pw)
       mlt <- readHMDweb(CNTRY = XXX, item = "mltper_1x1", username = us, password = pw)
@@ -112,25 +114,26 @@ Data <- do.call(rbind, lapply(Allcountries, function(XXX,pw,us){
       rbind(flt, mlt)
     }, us = us, pw = pw))
 
+HMD <- as.data.frame(HMD)
 HFD <- as.data.frame(HFD)
 
 # ------------------------------------
 # work to combine
 # 1) find years/populations in common:
-HMDxxyr <- paste(Data$Code, Data$Year)
+HMDxxyr <- paste(HMD$Code, HMD$Year)
 HFDxxyr <- paste(HFD$Code, HFD$Year)
 XXXyr   <- intersect(unique(HMDxxyr), unique(HFDxxyr))
-
+#Bx[Bx$Year == 1956 & Bx$Code == "AUT" & Bx$Age > 10, ]
 # cut down to only those pops with both fert and mort:
 HFD     <- HFD[HFDxxyr %in% XXXyr, ]
-Data    <- Data[HMDxxyr %in% XXXyr, ]
+HMD     <- HMD[HMDxxyr %in% XXXyr, ]
 
 # reorder rows
 HFD     <- HFD[with(HFD, order(Code, Year, Sex, Age)),]
[…]

### 2015-10-12 ThanoRepro/R/AgeStructure.R
data recompiled. age structure table in the works. Was quite a pain to get different mx scalers, but finally ofund a simple solution

@@ -12,13 +12,118 @@ if (system("hostname",intern=TRUE) %in% c("triffe-N80Vm", "tim-ThinkPad-L440")){
 }
 
 library(HMDHFDplus)
-source("R/Functions.R")
+devtools::load_all("R/RiffeFunctions")
 
 # 1) read in example survival curves. How about Sweden 1970, just because?
 
-flt <- readHMDweb("SWE","fltper_1x1",password=pw,username=us)
-mlt <- readHMDweb("SWE","mltper_1x1",password=pw,username=us)
+Data <- local(get(load("Data/DataAll.Rdata")))
 
+yr   <- 2010
+Name <- "USA"
+
+# friendly tools
+# mx2dxHMD()
+# da2fya()
+head(Data)
+mxm <- with(Data, mx[Sex == "m" & Code == Name & Year == yr])
+mxf <- with(Data, mx[Sex == "f" & Code == Name & Year == yr])
+
+# closeout:
+malejump <- diff(mxm[c(110,111)])
+mxm[111] + 1:10 * malejump
+mxme <- c(mxm, mxm[111] + 1:10 * malejump)
+femalejump <- diff(mxf[c(110,111)])
+mxf[111] + 1:10 * femalejump
+mxfe <- c(mxf, mxf[111] + 1:10 * femalejump)
+#e0jf <- with(Data, ex[Sex == "f" & Code == "JPN" & Year == 2012 & Age == 0])
+#plot(0:110,with(Data, mx[Sex == "f" & Code == "JPN" & Year == 2012 ]),log="y")
+#lines(0:110,mxf * coefs[4])
+
+
+
+# Note, mxm and mxf were smoothed by HMD into Kannisto. We need a higher closeout
+# age for these plots, like 120. Just assume a straight line to asymptote of 1 from last two points.
+
+
+
+mx2e0 <- function(mx,sex="m"){
+	mx                  <- Mna0(as.numeric(mx))
+	# mean proportion of interval passed at death
+	ax                  <- mx * 0 + .5                      # ax = .5, pg 38 MPv5
+	
+	ax[1]               <- AKm02a0(mx[1], sex)              # v6 a0 protocol
+	
+	ages.mids <- 1:length(mx) - 1 + ax
+	
+	
[…]

### 2015-10-12 ThanoRepro/R/DataPrep.R

@@ -181,9 +181,6 @@ Data[,Fy := FyFun(Exposure, Births, dx), by = list(Code,Sex,Year)]
 Data[,Fyf := FyFun(Exposure, Bxf, dx), by = list(Code,Sex,Year)]
 Data <- as.data.frame(Data)
 
-head(Data[Data$Sex == "f" & Data$Age < 12,])
-head(Data[Data$Sex == "f" & is.na(Data$Fyf),])
-head(Data[Data$Sex == "f" & Data$Year == 1956 & Data$Code == "AUT",],15)
 #-----------------------------------------
 rownames(Data) <- NULL
 #-----------------------------------------
@@ -241,4 +238,9 @@ lambdaLeslie <- do.call(rbind,lapply(Allcountries, function(XXX, us, pw){
                     data.frame(Code = XXX,Year = yrs, lambda.m,lambda.f,stringsAsFactors = FALSE)
                 }, pw = pw, us = us))
 save(lambdaLeslie, file = "Data/lambdaLeslie.Rdata")
+
+
+SRB <- median(PFvec)
+#dput(SRB) # 0.486540388060525
+#(1-SRB)/SRB
 

### 2015-10-12 ThanoRepro/R/rCalc.R

@@ -97,17 +97,17 @@ dev.off()
 var(rT$V1)/var(rL$V1)
 
 par(mfrow=c(1,2))
-matplot(1891:2011, rTM, type = 'l', lty = 1, col = "#0000FF40")
+matplot(1891:2013, rTM, type = 'l', lty = 1, col = "#0000FF40")
 abline(h=0)
-matplot(1891:2011, rLM, type = 'l', lty = 1, col = "#0000FF40")
+matplot(1891:2013, rLM, type = 'l', lty = 1, col = "#0000FF40")
 abline(h=0)
 
 graphics.off()
 for (i in 1:ncol(rTM)){
 plot(NULL, type = "n",xlim=c(1891,2011),ylim=c(-.04,.02),main=colnames(rLM)[i])
 abline(h=0)
-lines(1891:2011, rLM[,i], col = "blue")
-lines(1891:2011, rTM[,i], col = "red")
+lines(1891:2013, rLM[,i], col = "blue")
+lines(1891:2013, rTM[,i], col = "red")
 legend("bottomleft",lty=1,col=c("red","blue"),legend=c("Thano","Lotka"))
 Sys.sleep(1)
 }
@@ -115,6 +115,7 @@ Sys.sleep(1)
 matplot(1891:2011, rLM - rTM, type = 'l', lty = 1, col = "#0000FF40")
 abline(h=0)
 
+sum(!is.na(rDiff))
 sum(abs(rDiff) < 1e-7, na.rm = TRUE) # only one case of virtually identical r values (tol is like 1e-12)
 sum(rLM < rTM, na.rm = TRUE)
 sum(rLM > 0, na.rm = TRUE)
@@ -129,7 +130,7 @@ rLl0 <- sum(rLM < 0, na.rm = TRUE)
 sum(rLM < 0 & rLM > rTM & abs(rDiff) > 1e-7, na.rm = TRUE) / rLl0
 sum(rLM < 0 & rLM < rTM & abs(rDiff) > 1e-7, na.rm = TRUE) / rLl0
 
-sum(sign(rLM) != sign(rTM), na.rm = TRUE) # 138 times where sign was different
+sum(sign(rLM) != sign(rTM), na.rm = TRUE) # 157 times where sign was different
 rLMd <- apply(rLM,2,diff)
 rLTd <- apply(rTM,2,diff)
 

### 2015-10-13 ThanoRepro/ThanoRepro.tex
1) regenerated spaghetti plots 2) generated pyramids and leaves for tables, though section still not written, 3) started to rework intro 4) rename functions in formulas. Still incomplete. Need to decide.

@@
 \AtBeginDocument{\numberwithin{matrix}{section}}
+
+\newcommand{\Py}[2]{\includegraphics[scale=.5]{Figures/MiniPyramids/{Pyr-r#1-e0#2}.pdf}}
+\newcommand{\Lf}[2]{\includegraphics[scale=.5]{Figures/MiniPyramids/{Leaf-r#1-e0#2}.pdf}}
 % end preamble
@@
 \onehalfspacing
-All demographic forces vary over age in known and regular ways. Information on
-such forces, along with the size and age structure of a population, allows the
-demographer to make predictions about the future size and structure of a population. In this
+All demographic phenomena vary over age in regular ways. Information
+on such phenomena, along with the size and age structure of a population, allows
+the demographer to make predictions about the future size and structure of a population. In this
 paper, I explore some formal demographic consequences of a particular
-redefinition of age. Instead of counting age as the time passed between birth and the present (or
-some other moment), consider age as the amount of time left from the present
+redefinition of age. Instead of counting age as the time passed since birth, consider age as the amount of time left
 until death\footnote{This quantity is referred to as \textit{residual} lifetime
@@
 called thanatological age. Individuals in this case move in the same direction
-along imaginary life lines, but the reference point for age is now at the end of the life line instead of at the beginning.
-For individuals, the timing of events in life is mirrored exactly between the two definitions of age,
-but aggregate patterns differ.
+along imaginary life lines, but the reference point for age is now at the end of
+the life line instead of at the beginning. For individuals, the timing of events
+in life is identified exactly using either of the two definitions of age, but
+aggregate patterns differ. Some demographic phenomena are best
+described as a function of time since birth, and others of time until death,
+while still others can be a function of both to some degree \citep{riffe2015ttd}. This observation ought to motivate the formal demographic base by incorporating this aspect of time in into
+models.
+
+%Temporal perspectives and rescalings of the lifecourse are abundant in
+% demography already. Such work has enriched the discipline by producing new insights, questions, and conundrums.
 
 Direct information on the timing of demographic transitions with respect to
-death is only available from long-running panel or linked historical data that are few and far between. Some surveys also ask respondents in various ways about
-their perceived future lifetime, which may underly those
+death is available retrospectively from linked population registers or
+panel surveys with mortality follow-ups.
+Some surveys also ask respondents in various ways about their perceived future lifetime, which may underly those
 demographic transitions governed to some degree by volition. The present inquiry
 draws only from lifetable-projected remaining lifetime, which
-is intended to approximate actual remaining lifetime.
-The lifetable methods applied here follow those of \citet{brouard1986structure}
-or \citet{miller2001increasing}, which redistribute age-classified counts over
-future death times according to a fixed mortality schedule. This allows us to make statements about the remaining lifelines within a current stock of population, under certain assumptions.
+is intended to approximate actual remaining lifetime under the strong
+assumption of homogenous populations.
+The present inquiry is based on lifetable methods, which redistribute
+chronological age-classified counts over future death times according to a fixed
+mortality schedule. This allows us to make statements about the remaining lifelines within a stock of population, under certain assumptions.
 
-Temporal perspectives and rescalings of the lifecourse are abundant in
-demography already. It is my view that these efforts have enriched the
-discipline by producing new insights and questions much more than they have obscured it by
-producing conundrums. Likely some demographic phenomena are best
-described as a function of time since birth, and others of time until death,
-while still others can be a function of both to some degree. Fertility and
-reproduction are plausibly a function of both age perspectives, and the question
-of how to best partition such forces and events over time is a new conundrum
-that I do not tackle here. Instead I describe some of the formal consequences
-of the Lotka-Euler renewal model when viewed from the thanatological perspective
-of age.
+% Fertility and reproduction are plausibly a function of both age perspectives,
+ % and the question of how to best partition them over time is not investigated here.
 
-The renewal model is the core of modern demography, as it elegantly relates
-fertility, mortality, and population structure, revealing the long
-term consequences of a particular set of conditions. Some might think that
-stable population theory is not relevant to everyday demography, since
-populations seem to always be changing. Still, most projections seem to trail
-off into a stable state, and stability or even stationarity are
-imaginable for the world population of humans one day. I present a renewal model
-based on thanatological age not necessarily as a suggestion for a
-better way to model or project population (or certain aspects of it), although I
-plant this is an open and unexplored empirical question.
-In a minimal sense the thanatological renewal model represents an alternative
-and coherent way to describe the process of population renewal. In a maximal sense, the model may
-actually prove useful, as a model relating renewal to a death flow rather than
-to a birth flow. This enriches our understanding of lifespan and vitality, but also
-of the side effects of our species' renewal, such as the human biomass cycle.
+In this paper I describe some formal consequences
+of the Lotka-Euler renewal model when structured by thanatological age. The
+renewal model relates fertility, mortality, and population structure in a parsimonious form, revealing the long term consequences of a particular set of conditions. All diverse lines
+of contemporary demographic research are relatable to the renewal model.
+Some might think that stable population theory is not relevant to everyday
+demography, since populations are always changing. Even if change undercuts the
+notion of stability as a prediction, models of long-run consequnces still give a
+useful index of the demographic force implied by a given set of rates.
+That population projections lead into a stable state means that some
+form of stability is implied by the routine practice of demography.
+Stability or even stationarity are imaginable for the world population of humans
+one day. Models of stationary populations are easily applied by analogy to
+closed or nearly-closed birth cohorts. To present the renewal model based on thanatological
+age not necessarily a suggestion for a better way to model or project
+population. Rather, the thanatological renewal model represents an alternative
+and coherent way to describe the process of population renewal. A model relating
+renewal to a death flow rather than to a birth flow may have applications, but
+I do not speculate on these here.
 
-Further, there is likely a range of lifecourse transitions that covary
-more with remaining lifetime that with life lived, but these are mostly unstudied at this
-time,\footnote{I suppose that the prime candidates are late-life transitions.}
-and it is beneficial to expand the formal demographic base by incorporating such information into models. In an even more general sense, by
-thinking about remaining lifetime as a particular class of time-to-transition, one may generalize the lifetable methods presented here into new area
-applications, such as education, partnering, the economic lifecycle, or
-migration.
-The present inquiry is justified possibly for its own sake, and definitely as an
-attempt at accounting for remaining time structure in general in modeling that may prove useful.
+We have stated that many lifecourse transitions covary more with remaining
+lifetime than with life lived. In an even more general sense, by thinking about remaining lifetime as a
+particular class of time-to-transition, one may generalize the lifetable methods presented here into new area applications, such as education, partnering, the economic lifecycle, or migration.
 
@@
 dimensions: chronological age attained, $a$, and remaining years of life, $y$.
-This function we denote $d(y|a)$ --- The density distribution of thanatological
-age conditional on chronological age. We use the same function name $d()$,
-because it is essentially the same notion as the familiar $d(a)$, but rescaled as a
-function of chronological age. For ages higher than 0, simply condition
+This function we denote $f(y|a)$ --- The density distribution of thanatological
+age conditional on chronological age. This is essentially the same notion as the
+familiar $d(a)$, but rescaled as a function of chronological age. For ages higher than 0, simply condition
 on survival:\footnote{In the reliability literature, our concept of $d(y|a)$ is denoted by \begin{equation} X - t | X > t
@@
 attained.}
-
 \begin{equation}
 \label{eq:vaupel1}
-d(y | a) = \mu(a+y)\frac{l(a+y)}{l(a)} \ep
+f(y | a) = \mu(a+y)\frac{\ell(a+y)}{\ell(a)} \ep
 \end{equation}
@@
 $d_a$ column of the lifetable (here with single ages):
-
 \begin{equation}
 \label{eq:dyadiscret}
-d_{y, a} =  \frac{d_{a+y}+d_{a+y+1}}{L_a + L_{a+1}} \ec
+f_{y, a} =  \frac{d_{a+y}+d_{a+y+1}}{L_a + L_{a+1}} \ec
 \end{equation}
-
-\noindent where $\omega$ is the highest
-age.\footnote{Equation~\eqref{eq:dyadiscret} ill incur some error depending on
-the method used to derive $L_a$. To remove this error, rescale over the $a$ margin so that each distribution conditioned on $a$
-sums to 1. Working with $l_a$ rather than $L_a$ eliminates error and yields
-similar results, but is not necessarily better.} One produces in this way a
-probability density function over future death times for each chronological age,
-$a$. The $d(y|a)$- weighted average of $y$ for a given $a$ returns the familiar
+where $L_a$ is lifetable exposure.\footnote{Equation~\eqref{eq:dyadiscret}
+will incur some error depending on the method used to derive $L_a$. To remove
+this error, rescale over the $a$ margin so that each distribution conditioned on
+$a$ sums to 1. Working with $\ell_a$ rather than $L_a$ eliminates error and
+yields similar results, but is not necessarily better.} One produces in this way a probability density function over future death times for each chronological age,
+$a$. The $f(y|a)$- weighted average of $y$ for a given $a$ returns the familiar
 remaining life expectancy column of the lifetable, $e(a)$. If $P(a)$ is a
@@
 \label{eq:transform}
-P(y) =& \int_0^\infty P(a) d(y | a) \dd a \ep
+P(y) =& \int_0^\infty P(a) f(y | a) \dd a \ep
 \end{align}
 
-Since $1 = \int_0^\infty d(y|a) \dd y$ for each $a$,
+Since $1 = \int_0^\infty f(y|a) \dd y$ for each $a$,
 this density is useful for decomposing data classified by chronological age. For
@@
 corresponding lifetable. Exploiting this, we derive thanatological fertility
-rates, $f^\star(y)$, by applying \eqref{eq:transform} to age-classified birth
+rates, $\gamma(y)$, by applying \eqref{eq:transform} to age-classified birth
 counts, $B(a)$, to get $B(y)$ and again to exposure-to-risk, $E(a)$ (here we
@@
 \begin{equation}
-f^\star(y) = \frac{B(y}{E(y)} \ep
+\gamma(y) = \frac{B(y}{E(y)} \ep
 \end{equation}
 
-We are interested to use thanatological fertility rates, $f^\star(y)$ in
+We are interested to use thanatological fertility rates, $\gamma(y)$ in
 models of population renewal,\footnote{This is probably premature, but helps to
@@
 of data that overlap in the HFD and HMD.\footnote{There are as of this writing
-1834 population-years of overlap between the HMD and HFD, including a wide
+1933 population-years of overlap between the HMD and HFD, including a wide
 variety of fertility and mortality combinations.} Figure~\ref{fig:FxSpaghetti} gives ASFR for the same
@@
 population is) has tended to fall over time. All populations have shown a
-rightward shift in both the mean and modal mother's thanatological age at birth
-(TAB) over time. Several
-populations now show modal TABs of over 60 years. Thanatological total fertility rates (TTFR)
-track standard period total ferti
-lity rates (PTFR) rather closely, but tend to
+rightward shift in both the mean and modal mother's thanatological age at
+childbearing (TAC) over time. Several
+populations now show modal TACs of over 60 years. Thanatological total fertility rates (TTFR)
+track standard period total fertility rates (PTFR) rather closely, but tend to
 be somewhat higher (not shown).
@@
 	\begin{tiny}
-	*AUT, $1951-2010$; BGR, $1947-2009$; BLR, $1964-2009$; CAN, $1921-2007$; 
-	CHE, $1932-2011$; CZE, $1950-2011$; DEUTE, $1956-2010$; DEUTNP, $1990-2010$; 
-	DEUTW, $1956-2010$; ESP, $1922-2006$; EST, $1959-2010$; FIN, $1939-2009$; 
-	FRATNP, $1946-2010$; GBR\_NIR, $1974-2009$; GBR\_NP, $1974-2009$; GBR\_SCO,
-	$1945-2009$; GBRTENW, $1938-2009$; HUN, $1950-2009$; IRL, $1955-2009$; JPN, $1947-2009$; 
-	LTU, $1959-2010$; NLD, $1950-2009$; NOR, $1967-2009$; PRT, $1940-2009$; 
-	RUS, $1959-2010$; SVK, $1950-2009$; SVN, $1983-2009$; SWE, $1891-2010$; 
-	TWN, $1976-2010$; UKR, $1959-2006$; USA, $1933-2010$
+	*AUT, $1951-2010$; BGR, $1947-2009$; BLR, $1964-2012$; CAN, $1921-2009$; 
+	CHE, $1932-2011$; CHL, $1992-2005$; CZE, $1950-2011$; DEUTE, $1956-2011$; 
+	DEUTNP, $1990-2011$; DEUTW, $1956-2011$; ESP, $1922-2012$; EST, $1959-2011$; 
+	FIN, $1939-2009$; FRATNP, $1946-2013$; GBR\_NIR, $1974-2011$; GBR\_NP,
+	$1974-2011$; GBR\_SCO, $1945-2011$; GBRTENW, $1938-2011$; HUN, $1950-2009$;
+	IRL, $1955-2009$; ISL, $1963-2010$; JPN, $1947-2012$; LTU, $1959-2011$; NLD, $1950-2009$; 
+	NOR, $1967-2009$; PRT, $1940-2012$; RUS, $1959-2010$; SVK, $1950-2009$; 
+	SVN, $1983-2009$; SWE, $1891-2011$; TWN, $1976-2010$; UKR, $1959-2009$; 
+	USA, $1933-2010$
 	\end{tiny}
@@
 \begin{align}
-B(t) =& \int _0^\infty P(y,t)f^\star(y) \dd y = \int _0^\infty P(a,t)f(a) \dd a
-\ec \intertext{where $a$ indexes chronological age, $y$ indexes
-thanatological age, $P(a), P(y)$ are population counts, and $f^\star(y), f(a)$ are
-exact specific fertility probabilities (rates), which here are best thought of
+B(t) =& \int _0^\infty P(y,t)\gamma(y) \dd y = \int _0^\infty
+P(a,t)m(a) \dd a \ec \intertext{where $a$ indexes chronological age, $y$ indexes
+thanatological age, $P(a), P(y)$ are population counts, and $\gamma(y), m(a)$ are
+exact specific fertility probabilities (rates), best imagined
 as the fertility rates of females born to females. The thanatological integral can
 be broken down back in terms of chronological age:} =& \int_{y=0}^\infty
-\int_{a=0}^\infty f^\star(y)\frac{P(a,t)d(a+y)}{l(a)} \dd a \dd y \ec
+\int_{a=0}^\infty \gamma(y)\frac{P(a,t)d(a+y)}{\ell(a)} \dd a \dd y \ec
 \intertext{where $d(a)$ is the continuous lifetable death distribution with
-radix of 1, so $d(a)/l(0)$. We can relate the present population to past births
-with $P(a,t) = B(t-a)l(a)$:}
+radix of 1, so $d(a)/\ell(0)$. We can relate the present population to past births
+with $P(a,t) = B(t-a)\ell(a)$:}
 \label{eq:ergostep1}
-=& \int_{y=0}^\infty \int_{a=0}^\infty f^\star(y) B(t-a)d(a+y)\dd a \dd y \ep 
+=& \int_{y=0}^\infty \int_{a=0}^\infty \gamma(y) B(t-a)d(a+y)\dd a \dd y \ep 
 \intertext{Eventually strong ergodicity will assert itself, and $B(t)$ will
@@
 \label{eq:ergostep2}
-=& \int_{y=0}^\infty \int_{a=0}^\infty f^\star(y) B(t)e^{-ra}d(a+y)\dd a \dd y
+=& \int_{y=0}^\infty \int_{a=0}^\infty \gamma(y) B(t)e^{-ra}d(a+y)\dd a \dd y
 \ep
@@
 \label{eq:thanoren}
-1 =& \int_{y=0}^\infty \int_{a=0}^\infty f^\star(y) d(a+y)e^{-ra}\dd a \dd y
+1 =& \int_{y=0}^\infty \int_{a=0}^\infty \gamma(y) d(a+y)e^{-ra}\dd a \dd y
 \ep
@@
 \label{eq:lotka}
-1 =& \int_{a=0}^\infty f(a)l(a)e^{-ra}\dd a \ec
+1 =& \int_{a=0}^\infty m(a)l(a)e^{-ra}\dd a \ec
 \intertext{which can be expressed as well as}
-  =& \int_{a=0}^\infty \int_{y=0}^\infty f(a)d(a+y)e^{-ra}\dd y \dd a \ec
+  =& \int_{a=0}^\infty \int_{y=0}^\infty m(a)d(a+y)e^{-ra}\dd y \dd a \ec
   \label{eq:lotka2}
@@
 
+\section*{Stable age structure}
+
+\begin{figure}[]
+\centering
+\caption{Stable age structures by life expectancy, $e(0)$, and growth rate,
+$r$.*}
+\begin{subfigure}[b]{\textwidth}
+\centering
+\caption{Chronological age structure, $c(a)$.}
+\label{tab:chronostr}
+\begin{tabular}{rccccc}
+$e(0)\rightarrow$ \\
+$\downarrow r $  & 55  & 65 & 75 & 85 & 95 \\
+0.02  &\Py{0.02}{55}  &\Py{0.02}{65} &\Py{0.02}{75}&\Py{0.02}{85}  &\Py{0.02}{95}\\
+0.01  &\Py{0.01}{55}  &\Py{0.01}{65} &\Py{0.01}{75} &\Py{0.01}{85} &\Py{0.01}{95}\\
+0     &\Py{0}{55}     &\Py{0}{65}    &\Py{0}{75}    &\Py{0}{85}    &\Py{0}{95}\\
+-0.01 &\Py{-0.01}{55} &\Py{-0.01}{65}&\Py{-0.01}{75}&\Py{-0.01}{85}&\Py{-0.01}{95}\\
+-0.02 &\Py{-0.02}{55} &\Py{-0.02}{65}&\Py{-0.02}{75}&\Py{-0.02}{85}&\Py{-0.02}{95}
+\end{tabular}
+\end{subfigure}
+\\ \vspace{2em}
+\begin{subfigure}[b]{\textwidth}
+\centering
+\caption{Thanatological age structure, $c^\star(y)$.}
+\label{tab:thanostr}
+\begin{tabular}{rccccc}
+$e(0)\rightarrow$ \\
+$\downarrow r $  & 55  & 65 & 75 & 85 & 95 \\
+0.02  &\Lf{0.02}{55}  &\Lf{0.02}{65} &\Lf{0.02}{75}&\Lf{0.02}{85}  &\Lf{0.02}{95}\\
+0.01  &\Lf{0.01}{55}  &\Lf{0.01}{65} &\Lf{0.01}{75} &\Lf{0.01}{85} &\Lf{0.01}{95}\\
+0     &\Lf{0}{55}     &\Lf{0}{65}    &\Lf{0}{75}    &\Lf{0}{85}    &\Lf{0}{95}\\
+-0.01 &\Lf{-0.01}{55} &\Lf{-0.01}{65}&\Lf{-0.01}{75}&\Lf{-0.01}{85}&\Lf{-0.01}{95}\\
+-0.02 &\Lf{-0.02}{55} &\Lf{-0.02}{65}&\Lf{-0.02}{75}&\Lf{-0.02}{85}&\Lf{-0.02}{95}
+\end{tabular}
+\end{subfigure}
+\vspace{2em}
+*Underlying mortality pattern based on 2010 USA (HMD)
+\end{figure}
+
 \section*{Some preliminary empirical findings}

### 2015-10-13 ThanoRepro/R/AgeStructure.R

@@ -105,26 +105,41 @@ for (i in 1:length(e0vec)){
 			maximum = FALSE, 
 			tol = 1e-7)$minimum
 		}
-e0avg(-10.7133251,mxm,mxf,PF)
-plot(0:120, mx2lx(mxme * getcoef(coefs[1], mxme, "m"),"m"), type = 'l', ylim = c(0,1))
-lines(0:120, mx2lx(mxme * getcoef(coefs[2], mxme, "m"),"m"))
-lines(0:120, mx2lx(mxme * getcoef(coefs[3], mxme, "m"),"m"))
-lines(0:120, mx2lx(mxme * getcoef(coefs[4], mxme, "m"),"m"))
-lines(0:120, mx2lx(mxme * getcoef(coefs[5], mxme, "m"),"m"))
-# so, coefs are what we need to use to derive the right avg e0, just multiply in mxm, mxf
-# to get back corresponding lx, rescale using PF and make pyramids.
-age <- 0:120
+
 rs <- c(-.02,-.01,0,.01,.02)
 par(mai = c(0,0,0,0))
 
-# ready!!!
-plot(NULL, type = "n", axes = FALSE, xlab = "", ylab = "",xlim=c(-2,2),ylim=c(0,115))
-PyramidOutline(males = mx2lx(mxme*getcoef(coefs[1], mxme, "f"),"m") * (1 - PF) * exp(-age*.02),
-		females = mx2lx(mxfe*getcoef(coefs[1], mxfe, "f"),"f") * PF * exp(-age*.02),
-		scale = 100,
-		col = "gray",
-		border =  NA)
+for (r in rs){
+	for (e in 1:length(e0vec)){
+		pdf(paste0("Figures/MiniPyramids/","Pyr-r",r,"-e0",e0vec[e],".pdf"), width=1,height=1)
+		par(mai=c(0,0,0,0), xaxs="i", yaxs = "i")
+		plot(NULL, type = "n", axes = FALSE, xlab = "", ylab = "",xlim=c(-1.6,1.6),ylim=c(0,115))
+		PyramidOutline(males = mx2lx(mxme*getcoef(coefs[e], mxme, "m"),"m") * (1 - PF) * exp(-age*r),
+				females = mx2lx(mxfe*getcoef(coefs[e], mxfe, "f"),"f") * PF * exp(-age*r),
+				scale = 100,
+				col = "gray",
[…]

### 2015-10-13 ThanoRepro/R/HFDspaghetti.R

@@ -108,6 +108,9 @@ lines(Data$Age, Data$FxSt, col = "#00000005")
 }
 
 graphics.off()
+
+# just for a cleaner figure
+Data$Fy[Data$Age > 100] <- NA
 # Fy 
 png("Figures/FySpaghettiDraft.png")
 #pdf("Figures/FySpaghetti.pdf") # high res vector
@@ -124,6 +127,8 @@ text(mean(c(0,95)),-.01,"Thanatological age (years left)",xpd=TRUE)
 #text(c(25,25,80,80),c(.03,.12,.03,.12),"DRAFT",cex=4,srt=45,col="#BBBBBB95",xpd=TRUE)
 dev.off()
 
+# just for a cleaner figure
+Data$Fx[Data$Fx==0] <- NA
 # Fx 
 png("Figures/FxSpaghettiDraft.png")
 #pdf("Figures/FxSpaghetti.pdf") # high res vector
@@ -140,6 +145,7 @@ text(mean(c(12,55)),-.02,"Chronological age (years lived)",xpd=TRUE)
 #text(c(25,25,45,45),c(.03,.25,.03,.25),"DRAFT",cex=4,srt=45,col="#BBBBBB95",xpd=TRUE)
 dev.off()
 
+length(unique(paste(Data$Code,Data$Year)))
 #
 ## Fyst
 #par(xaxs = "i", yaxs = "i", mai=c(1,1,1,1))
@@ -173,8 +179,8 @@ dev.off()
 #
 #
 ## for caption
-#ranges <- tapply(Data$Year, Data$Code, function(yrs){
-#            paste0("$",paste(min(yrs),max(yrs),sep = "-"),"$")
-#        })
-#paste(paste(names(ranges),ranges, sep = ", "), collapse = "; ")
-#length(ranges)
+ranges <- tapply(Data$Year, Data$Code, function(yrs){
+            paste0("$",paste(min(yrs),max(yrs),sep = "-"),"$")
+        })
+paste(paste(names(ranges),ranges, sep = ", "), collapse = "; ")
+length(ranges)

### 2015-10-13 ThanoRepro/R/RiffeFunctions/R/lifetablefunctions.R

@@ -9,7 +9,7 @@
 #' 
 #' @return q0 the estimate of q0 according to the identity between a0, m0, q0
 #' 
-#' @author Tim Riffe \email{triffe@@demog.berkeley.edu}
+#' @author Tim Riffe \email{tim.riffe@@gmail.com}
 #' 
 #' @export
 #' 
@@ -28,7 +28,7 @@ AKm02q0 <- function(m0,constant,slope){
 #' 
 #' @return a0, the estimated average age at death of those dying in the first year of life, either a single value or a vector of a_0 values.
 #' 
-#' @author Tim Riffe \email{triffe@@demog.berkeley.edu}
+#' @author Tim Riffe \email{tim.riffe@@gmail.com}
 #' 
 #' @export
 
@@ -52,7 +52,7 @@ AKm02a0 <- function(m0, sex = "m"){
 #' 
 #' @return dx a vector the same length as mx, radix 1.
 #' 
-#' @author Tim Riffe \email{triffe@@demog.berkeley.edu}
+#' @author Tim Riffe \email{tim.riffe@@gmail.com}
 #' 
 #' @export
 #' 
@@ -85,7 +85,37 @@ mx2dxHMD <- cmpfun(function(mx, sex = "m"){
   })
 
 #'
-#' @title mx2dxHMD derive dx from mx using the HMD lifetable protocol
+#' @title derive DYA from d(a) 
+#' 
+#' @description This function makes a triangle out of d(a)
+#' 
+#' @param da the lifetable deaths distribution classified by chronological age
+#' 
+#' @return d(Y,A) the lifetable remaining lifetime distribution, not conditioned on surviva;
+#' 
+#' @author Tim Riffe \email{tim.riffe@@gmail.com}
+#' 
+#' @export
+#' 
+
+da2DYA <- function(da){
+	N       <- length(da)
+	ay      <- 1:N - 1
+	
+	da      <- Mna0(da)   # remove NAs if any       
+	da      <- c(da, da * 0) / sum(da) # pad out with 0s
+	fya     <- matrix(da[col(matrix(nrow = N, 
+									ncol = N)) + ay], 
+			nrow = N, 
+			ncol = N, 
+			dimnames = list(Ex = ay, 
+					Age = ay)
+	)
+	fya
+}
+
+#'
+#' @title derive f(y,a) from d(a) 
 #' 
[…]

### 2015-10-14 ThanoRepro/ThanoRepro.tex
further massaging of intro

@@
 \onehalfspacing
-All demographic phenomena vary over age in regular ways. Information
-on such phenomena, along with the size and age structure of a population, allows
-the demographer to make predictions about the future size and structure of a population. In this
-paper, I explore some formal demographic consequences of a particular
-redefinition of age. Instead of counting age as the time passed since birth, consider age as the amount of time left
-until death\footnote{This quantity is referred to as \textit{residual} lifetime
-in the reliability literaure.}. The first kind of age is called
-chronological age and the kind of age explored in this paper is
+In this paper, I explore some formal demographic consequences of a particular redefinition of age.
+Instead of counting age as the time passed since birth, consider age as the amount of time left until death\footnote{This quantity is referred to as \textit{residual} lifetime
+in the reliability literaure, or just time-to-death (TTD) elsewhere.}. The first
+kind of age is called chronological age and the kind of age explored in this paper is
 called thanatological age. Individuals in this case move in the same direction
@@
 in life is identified exactly using either of the two definitions of age, but
-aggregate patterns differ. Some demographic phenomena are best
-described as a function of time since birth, and others of time until death,
-while still others can be a function of both to some degree \citep{riffe2015ttd}. This observation ought to motivate the formal demographic base by incorporating this aspect of time in into
+aggregate patterns differ. The models of demography are prefaced by the observation
+that demographic phenomena vary over age and time in empirically regular ways.
+Some such phenomena are best described in the aggregatate as a function
+of time since birth, and others of time until death, while still others can be a function of both to some degree \citep{riffe2015ttd}.
+This observation ought to motivate the formal demographic base by incorporating this aspect of time in into
 models.
@@
 % demography already. Such work has enriched the discipline by producing new insights, questions, and conundrums.
+% Fertility and reproduction are plausibly a function of both age perspectives,
+ % and the question of how to best partition them over time is not investigated here.
 
+I describe the Lotka-Euler-Leslie renewal model when structured by
+thanatological age.
+The renewal model relates fertility, mortality, and population structure in a parsimonious form, revealing the long term consequences of a particular set of conditions.
+All diverse lines
+of contemporary demographic research are relatable to the renewal model.
+Some might say that stable population theory is not relevant to everyday
+demography, since populations are always changing. Even if change precludes the
+notion of stability as a prediction, models of long-run consequnces still give a
+useful measure of the demographic force implied by a given set of rates.
+In this and other ways, stability is never far away. That population projections
+lead into a stable state means that some form of stability is implied by the routine practice of demography.
+Stability or even stationarity are imaginable for the world population of humans
+one day. Models of stationary populations are easily applied by analogy to
+closed or nearly-closed birth cohorts. Finally, stable population theory is
+relavant to other non-human structured populations, such as stars.
+
+%We have stated that many lifecourse transitions covary more with remaining
+%lifetime than with life lived. In an even more general sense, by thinking about
+% remaining lifetime as a particular class of time-to-transition, one may generalize the lifetable methods presented here into new area applications, such as education, partnering, the economic lifecycle, or migration.
+
+To present the renewal model based on thanatological age is not necessarily a
+suggestion for a better way to model or project population. Rather, the
+thanatological renewal model represents an alternative and coherent way to
+describe the process of population renewal by relating renewal to a death flow
+rather than to a birth flow. The chronological-age classified renewal model and the thanatological renewal model both approximate the movement of population stocks
+over time, but from opposite vantage points. Specifying a model of
+thanatological renewal expands the toolbox available to approach complex
+temporal population structure. The next section of this article describes the lifetable method of transforming age-classified
+data into thanatological age classes. I then derive the continuous version of
+the thanatological renewal model, and some other stable properties. This is
+followed by the matrix version of the model, which is explained alongside the
+lifecycle graph.
+
+\section*{A lifetable approximation of thanatological age}
 Direct information on the timing of demographic transitions with respect to
@@
 is intended to approximate actual remaining lifetime under the strong
-assumption of homogenous populations.
-The present inquiry is based on lifetable methods, which redistribute
-chronological age-classified counts over future death times according to a fixed
-mortality schedule. This allows us to make statements about the remaining lifelines within a stock of population, under certain assumptions.
-
-% Fertility and reproduction are plausibly a function of both age perspectives,
- % and the question of how to best partition them over time is not investigated here.
-
-In this paper I describe some formal consequences
-of the Lotka-Euler renewal model when structured by thanatological age. The
-renewal model relates fertility, mortality, and population structure in a parsimonious form, revealing the long term consequences of a particular set of conditions. All diverse lines
-of contemporary demographic research are relatable to the renewal model.
-Some might think that stable population theory is not relevant to everyday
-demography, since populations are always changing. Even if change undercuts the
-notion of stability as a prediction, models of long-run consequnces still give a
-useful index of the demographic force implied by a given set of rates.
-That population projections lead into a stable state means that some
-form of stability is implied by the routine practice of demography.
-Stability or even stationarity are imaginable for the world population of humans
-one day. Models of stationary populations are easily applied by analogy to
-closed or nearly-closed birth cohorts. To present the renewal model based on thanatological
-age not necessarily a suggestion for a better way to model or project
-population. Rather, the thanatological renewal model represents an alternative
-and coherent way to describe the process of population renewal. A model relating
-renewal to a death flow rather than to a birth flow may have applications, but
-I do not speculate on these here.
-
-We have stated that many lifecourse transitions covary more with remaining
-lifetime than with life lived. In an even more general sense, by thinking about remaining lifetime as a
-particular class of time-to-transition, one may generalize the lifetable methods presented here into new area applications, such as education, partnering, the economic lifecycle, or migration.
-
-The chronological-age classified Lotka-Euler renewal model and the
-thanatological renewal model both approximate the movement of population stocks
-over time, but from opposite vantage points. Specifying a model of
-thanatological renewal expands the toolbox available to approach complex
-temporal population structure. The next section of this article describes the lifetable method of transforming age-classified
-data into thanatological age classes. I then derive the continuous version of
-the thanatological renewal model, and some other stable properties. This is
-followed by the matrix version of the model, which is explained alongside the
-lifecycle graph. I present a set of preliminary empirical findings to help
-understand the implications of temporal structure.
-
-\section*{A lifetable approximation of thanatological age}
+assumption of homogenous populations. Chronological age-classified counts are
+redistributed over future death times according to a fixed mortality schedule.
+This allows us to make statements about the remaining lifetime distribution
+within a stock of population, under certain assumptions.
 

### 2015-10-14 ThanoRepro/R/DataPrep.R
changes taken. Ugly conflicts because I forgot to fold in PC changes. Now merged finally. Sucks you can't overwrite remote when using merge tool...

@@ -1,14 +1,22 @@
 
-# for Tim, this will choke
-if (system("hostname",intern=TRUE) %in% c("triffe-N80Vm", "tim-ThinkPad-L440")){
-  # if I'm on the laptop
-  setwd("/home/tim/git/ThanoRepro/ThanoRepro")
-} else {
-  # in that case I'm on Berkeley system, and other people in the dept can run this too
-  setwd(paste0("/data/commons/",system("whoami",intern=TRUE),"/git/ThanoRepro/ThanoRepro"))
-}
 
-source("R/Functions.R")
+if (system("hostname",intern=TRUE) %in% c("triffe-N80Vm", "tim-ThinkPad-L440")){
+	# if I'm on the laptop
+	setwd("/home/tim/git/ThanoRepro/ThanoRepro")
+} else {
+	if (system("hostname",intern=TRUE) == "PC-403478"){
+		# on MPIDR PC
+		setwd("U://git//ThanoRepro//ThanoRepro")
+	} else {
+		# in that case I'm on Berkeley system, and other people in the dept can run this too
+		setwd(paste0("/data/commons/",system("whoami",intern=TRUE),"/git/ThanoRepro/ThanoRepro"))
+	}
+}
+getwd()
+
+
+#source("R/Functions.R")
+devtools::load_all(file.path("R","RiffeFunctions"))
 library(compiler)
 library(data.table)
 library(reshape2)

### 2015-10-14 ThanoRepro/R/DepRatio.R

@@ -1,85 +1,92 @@
-# note, Robert's neat dependency ratio parlor trick:
-
-# P65+ / P18-64 = alpha
-# how does alpha change with r?
-
-# change in alpha with r = (xbar18-64 - xbar65+) / 100
-# (or something like that)
+## code deprecated. It was incomplete, and I decided I didn't care for the time being #
+#######################################################################################
 
 
-# example data, HMD, I forget which country, year, sex.... legit tho
-Lx <- c(0.99273, 0.99194, 0.99147, 0.99114, 0.99089, 0.99068, 0.99048, 
-        0.99031, 0.99015, 0.98999, 0.98982, 0.98963, 0.98942, 0.98918, 
-        0.98888, 0.98846, 0.98786, 0.98705, 0.98601, 0.98478, 0.98347, 
-        0.9821, 0.98074, 0.97941, 0.97809, 0.9768, 0.97556, 0.97431, 
-        0.97304, 0.97176, 0.97047, 0.96916, 0.96779, 0.96634, 0.96476, 
-        0.9631, 0.96135, 0.95946, 0.9574, 0.9552, 0.95283, 0.95027, 0.94753, 
-        0.94453, 0.94127, 0.93777, 0.93401, 0.92996, 0.92553, 0.9208, 
-        0.91584, 0.91063, 0.9051, 0.899, 0.89265, 0.88585, 0.87828, 0.87005, 
-        0.86113, 0.85155, 0.8412, 0.83022, 0.8183, 0.80528, 0.79125, 
-        0.77627, 0.76047, 0.74349, 0.72533, 0.70625, 0.68594, 0.66435, 
-        0.64136, 0.61696, 0.59165, 0.56534, 0.53797, 0.50965, 0.4803, 
-        0.44982, 0.41786, 0.38513, 0.35184, 0.31817, 0.2848, 0.2517, 
-        0.21943, 0.18854, 0.1596, 0.13276, 0.10839, 0.08691, 0.06812, 
-        0.05206, 0.03879, 0.02823, 0.02006, 0.01387, 0.00932, 0.00608, 
-        0.00385, 0.00237, 0.00141, 0.00081, 0.00046, 0.00025, 0.00013, 
-        7e-05, 3e-05, 2e-05, 1e-05)
 
-Plotka <- function(Lx, r, a){
-    sum(Lx * exp(-r*a))
-}
-xbarLotka <- function(Lx, r, a){
-    sum(a * Lx * exp(-r*a)) /  sum(Lx * exp(-r*a))
-}
 
[…]

### 2015-10-14 ThanoRepro/R/Functions.R

@@ -1,12 +1,17 @@
 
-# for Tim, this will choke
 if (system("hostname",intern=TRUE) %in% c("triffe-N80Vm", "tim-ThinkPad-L440")){
-  # if I'm on the laptop
-  setwd("/home/tim/git/ThanoRepro/ThanoRepro")
+	# if I'm on the laptop
+	setwd("/home/tim/git/ThanoRepro/ThanoRepro")
 } else {
-  # in that case I'm on Berkeley system, and other people in the dept can run this too
-  setwd(paste0("/data/commons/",system("whoami",intern=TRUE),"/git/ThanoRepro/ThanoRepro"))
+	if (system("hostname",intern=TRUE) == "PC-403478"){
+		# on MPIDR PC
+		setwd("U://git//ThanoRepro//ThanoRepro")
+	} else {
+		# in that case I'm on Berkeley system, and other people in the dept can run this too
+		setwd(paste0("/data/commons/",system("whoami",intern=TRUE),"/git/ThanoRepro/ThanoRepro"))
+	}
 }
+getwd()
 
 PACKAGES <- installed.packages()[,"Package"]
 

### 2015-10-14 ThanoRepro/R/HFDspaghetti.R

@@ -1,19 +1,24 @@
-# for Tim, this will choke
+
 if (system("hostname",intern=TRUE) %in% c("triffe-N80Vm", "tim-ThinkPad-L440")){
-  # if I'm on the laptop
-  setwd("/home/tim/git/ThanoRepro/ThanoRepro")
+	# if I'm on the laptop
+	setwd("/home/tim/git/ThanoRepro/ThanoRepro")
 } else {
-  # in that case I'm on Berkeley system, and other people in the dept can run this too
-  setwd(paste0("/data/commons/",system("whoami",intern=TRUE),"/git/ThanoRepro/ThanoRepro"))
+	if (system("hostname",intern=TRUE) == "PC-403478"){
+		# on MPIDR PC
+		setwd("U://git//ThanoRepro//ThanoRepro")
+	} else {
+		# in that case I'm on Berkeley system, and other people in the dept can run this too
+		setwd(paste0("/data/commons/",system("whoami",intern=TRUE),"/git/ThanoRepro/ThanoRepro"))
+	}
 }
+getwd()
 
-
-source("R/Functions.R")
+devtools::load_all(file.path("R","RiffeFunctions"))
 library(data.table)
 # read in data and select females
 Data <- local(get(load("Data/DataAll.Rdata")))
 Data <- Data[Data$Sex == "f", ]
-DT <- as.data.table(Data)
+DT   <- as.data.table(Data)
 
 
 playWithData <- FALSE
@@ -112,6 +117,9 @@ graphics.off()
 # just for a cleaner figure
 Data$Fy[Data$Age > 100] <- NA
 # Fy 
+Data$Fy[Data$Fy==0] <- NA
+Data$Fx[Data$Fx==0] <- NA
+
 png("Figures/FySpaghettiDraft.png")
 #pdf("Figures/FySpaghetti.pdf") # high res vector
 par(xaxs = "i", yaxs = "i", mai=c(.5,.5,.5,.5))

### 2015-10-14 ThanoRepro/R/MatrixStuff.R

@@ -1,36 +1,50 @@
-library(popbio)
-# for Tim, this will choke
+
 if (system("hostname",intern=TRUE) %in% c("triffe-N80Vm", "tim-ThinkPad-L440")){
-  # if I'm on the laptop
-  setwd("/home/tim/git/ThanoRepro/ThanoRepro")
+	# if I'm on the laptop
+	setwd("/home/tim/git/ThanoRepro/ThanoRepro")
 } else {
-  # in that case I'm on Berkeley system, and other people in the dept can run this too
-  setwd(paste0("/data/commons/",system("whoami",intern=TRUE),"/git/ThanoRepro/ThanoRepro"))
+	if (system("hostname",intern=TRUE) == "PC-403478"){
+		# on MPIDR PC
+		setwd("U://git//ThanoRepro//ThanoRepro")
+	} else {
+		# in that case I'm on Berkeley system, and other people in the dept can run this too
+		setwd(paste0("/data/commons/",system("whoami",intern=TRUE),"/git/ThanoRepro/ThanoRepro"))
+	}
 }
+getwd()
 # 
-source("R/Functions.R")
+# install.packages("popbio")
+library(popbio)
+devtools::load_all(file.path("R","RiffeFunctions"))
 
 Data <- local(get(load("Data/DataAll.Rdata")))
 Data <- Data[Data$Sex == "f", ]
 
 lambdaT <- local(get(load("Data/lambda.Rdata")))
 lambdaL <- local(get(load("Data/lambdaLeslie.Rdata")))
+
+makeMatrices <- FALSE
+if (makeMatrices){
 # made in commented-out code below:
 #AllMatrices <- local(get(load("Data/AllMatrices.Rdata")))
 #print(object.size(AllMatrices),units = "Mb")
-#AllMatrices <- lapply(split(Data, with(Data, paste(Code,Year))), function(Dat, lambdaL, lambdaT){
-#           yr <- unique(Dat$Year)
-#           Cd <- unique(Dat$Code)
-#           N  <- nrow(Dat)
-#           px <- Mna0(Dat$lx[2:N] / Dat$lx[1:(N-1)])
-#           lL <- lambdaL$lambda.f[lambdaL$Code == Cd & lambdaL$Year == yr]
-#           lT <- lambdaT$lambda.f[lambdaL$Code == Cd & lambdaL$Year == yr]
-#         
[…]

### 2015-10-14 ThanoRepro/R/Oscillation.R

@@ -1,11 +1,18 @@
-# for Tim, this will choke
+
 if (system("hostname",intern=TRUE) %in% c("triffe-N80Vm", "tim-ThinkPad-L440")){
-  # if I'm on the laptop
-  setwd("/home/tim/git/ThanoRepro/ThanoRepro")
+	# if I'm on the laptop
+	setwd("/home/tim/git/ThanoRepro/ThanoRepro")
 } else {
-  # in that case I'm on Berkeley system, and other people in the dept can run this too
-  setwd(paste0("/data/commons/",system("whoami",intern=TRUE),"/git/ThanoRepro/ThanoRepro"))
+	if (system("hostname",intern=TRUE) == "PC-403478"){
+		# on MPIDR PC
+		setwd("U://git//ThanoRepro//ThanoRepro")
+	} else {
+		# in that case I'm on Berkeley system, and other people in the dept can run this too
+		setwd(paste0("/data/commons/",system("whoami",intern=TRUE),"/git/ThanoRepro/ThanoRepro"))
+	}
 }
+getwd()
+# 
 # seems like oscillation comparisons should be done on standard populations.
 
 # could start with Px, get Py, and then derive Cohen's measure, but this adds an initial smoothing
@@ -30,9 +37,9 @@ library(parallel)
 P0mat <- matrix(runif(111*1e3),ncol=1e3)
 
 
-apply(P0mat,2,function(P0, AY){
-       CohenA(P0, AY[["L"]]) / CohenA(P0, AY[["Y"]])
-        }, AY = AllMatrices[[1]])
+#apply(P0mat,2,function(P0, AY){
+#       CohenA(P0, AY[["L"]]) / CohenA(P0, AY[["Y"]])
+#        }, AY = AllMatrices[[1]])
 
 extremeCase1 <- c(1,rep(0,110))
 CohenA(extremeCase1, AllMatrices[[1]][["L"]]) / CohenA(extremeCase1, AllMatrices[[1]][["Y"]])

### 2015-10-14 ThanoRepro/R/Pyramids.R

@@ -1,12 +1,19 @@
-# for Tim, this will choke
+
 if (system("hostname",intern=TRUE) %in% c("triffe-N80Vm", "tim-ThinkPad-L440")){
-  # if I'm on the laptop
-  setwd("/home/tim/git/ThanoRepro/ThanoRepro")
+	# if I'm on the laptop
+	setwd("/home/tim/git/ThanoRepro/ThanoRepro")
 } else {
-  # in that case I'm on Berkeley system, and other people in the dept can run this too
-  setwd(paste0("/data/commons/",system("whoami",intern=TRUE),"/git/ThanoRepro/ThanoRepro"))
+	if (system("hostname",intern=TRUE) == "PC-403478"){
+		# on MPIDR PC
+		setwd("U://git//ThanoRepro//ThanoRepro")
+	} else {
+		# in that case I'm on Berkeley system, and other people in the dept can run this too
+		setwd(paste0("/data/commons/",system("whoami",intern=TRUE),"/git/ThanoRepro/ThanoRepro"))
+	}
 }
-source("R/Functions.R")
+getwd()
+# 
+devtools::load_all(file.path("R","RiffeFunctions"))
 #source("R/PlotFunctions.R")
 #
 #

### 2015-10-14 ThanoRepro/R/RiffeFunctions/R/stablefunctions.R

@@ -138,7 +138,7 @@ rLotkaCoale <- cmpfun(function(fa,La,a=.5:110.5, T.guess = 29, maxit = 1e2, tol 
 
 Leslie <- function(fa, Sa, lambda){
   fa <- fa * lambda 
-  cbind(rbind(fa[1:length(pa)], diag(Sa)), 0)
+  cbind(rbind(fa[1:length(Sa)], diag(Sa)), 0)
 }
 
 #' @title ThanoMin residual functin for optimizing thanatological r

### 2015-10-14 ThanoRepro/R/UniqueProofTests.R

@@ -1,13 +1,19 @@
-# for Tim, this will choke
-if (system("hostname",intern=TRUE) %in% c("triffe-N80Vm", "tim-ThinkPad-L440")){
-  # if I'm on the laptop
-  setwd("/home/tim/git/ThanoRepro/ThanoRepro")
-} else {
-  # in that case I'm on Berkeley system, and other people in the dept can run this too
-  setwd(paste0("/data/commons/",system("whoami",intern=TRUE),"/git/ThanoRepro/ThanoRepro"))
-}
 
-source("R/Functions.R")
+if (system("hostname",intern=TRUE) %in% c("triffe-N80Vm", "tim-ThinkPad-L440")){
+	# if I'm on the laptop
+	setwd("/home/tim/git/ThanoRepro/ThanoRepro")
+} else {
+	if (system("hostname",intern=TRUE) == "PC-403478"){
+		# on MPIDR PC
+		setwd("U://git//ThanoRepro//ThanoRepro")
+	} else {
+		# in that case I'm on Berkeley system, and other people in the dept can run this too
+		setwd(paste0("/data/commons/",system("whoami",intern=TRUE),"/git/ThanoRepro/ThanoRepro"))
+	}
+}
+getwd()
+# 
+
 Data <- local(get(load("Data/DataAll.Rdata")))
 Data <- Data[Data$Sex == "f", ]
 

### 2015-10-14 ThanoRepro/R/rCalc.R

@@ -1,13 +1,20 @@
 # for Tim, this will choke
 if (system("hostname",intern=TRUE) %in% c("triffe-N80Vm", "tim-ThinkPad-L440")){
-  # if I'm on the laptop
-  setwd("/home/tim/git/ThanoRepro/ThanoRepro")
+	# if I'm on the laptop
+	setwd("/home/tim/git/ThanoRepro/ThanoRepro")
 } else {
-  # in that case I'm on Berkeley system, and other people in the dept can run this too
-  setwd(paste0("/data/commons/",system("whoami",intern=TRUE),"/git/ThanoRepro/ThanoRepro"))
+	if (system("hostname",intern=TRUE) == "PC-403478"){
+		# on MPIDR PC
+		setwd("U://git//ThanoRepro//ThanoRepro")
+	} else {
+		# in that case I'm on Berkeley system, and other people in the dept can run this too
+		setwd(paste0("/data/commons/",system("whoami",intern=TRUE),"/git/ThanoRepro/ThanoRepro"))
+	}
 }
+getwd()
+# 
 
-source("R/Functions.R")
+devtools::load_all(file.path("R","RiffeFunctions"))
 #install.packages("data.table")
 library(data.table)
 # read in data and select females
@@ -160,21 +167,21 @@ ncol(rDiff)
 
 
 # ---------------------------------------
-ay      <- 0:110
-dx2      <- c(dx, dx * 0) / sum(dx) # pad out with 0s
-dM     <- matrix(dx2[col(matrix(nrow = 111, ncol = 111)) + ay], 
-        nrow = 111, ncol = 111, dimnames = list(Ex = ay, Age = ay))
-
-Fyx <- Minf0(Mna0(Thano(Dat$Bxf, dx) / Thano(Dat$Exposure, dx)))
-r <- 0
-sum(dM*exp(-r*.5:110.5)*Fyx)
-sum(t(dM)*exp(-r*.5:110.5)*t(Fyx))
-# i.e. if fertility is not collapsed to Fy only, then we get same results from age and thano
-
-sum(dM*exp(-r*.5:110.5)*Fyx)
-sum(t(dM)*exp(-r*.5:110.5)*t(Fyx))
-
-
+#ay      <- 0:110
+#dx2      <- c(dx, dx * 0) / sum(dx) # pad out with 0s
[…]

### 2015-10-14 ThanoRepro/R/rEqual.R

@@ -1,13 +1,20 @@
-# for Tim, this will choke
-if (system("hostname",intern=TRUE) %in% c("triffe-N80Vm", "tim-ThinkPad-L440")){
-  # if I'm on the laptop
-  setwd("/home/tim/git/ThanoRepro/ThanoRepro")
-} else {
-  # in that case I'm on Berkeley system, and other people in the dept can run this too
-  setwd(paste0("/data/commons/",system("whoami",intern=TRUE),"/git/ThanoRepro/ThanoRepro"))
-}
 
-source("R/Functions.R")
+if (system("hostname",intern=TRUE) %in% c("triffe-N80Vm", "tim-ThinkPad-L440")){
+	# if I'm on the laptop
+	setwd("/home/tim/git/ThanoRepro/ThanoRepro")
+} else {
+	if (system("hostname",intern=TRUE) == "PC-403478"){
+		# on MPIDR PC
+		setwd("U://git//ThanoRepro//ThanoRepro")
+	} else {
+		# in that case I'm on Berkeley system, and other people in the dept can run this too
+		setwd(paste0("/data/commons/",system("whoami",intern=TRUE),"/git/ThanoRepro/ThanoRepro"))
+	}
+}
+getwd()
+# 
+
+devtools::load_all(file.path("R","RiffeFunctions"))
 #  thanatological age 
 
 # took me a while to get to this...

### 2015-10-14 ThanoRepro/R/rMatrixEqual.R

@@ -1,13 +1,19 @@
-# for Tim, this will choke
-if (system("hostname",intern=TRUE) %in% c("triffe-N80Vm", "tim-ThinkPad-L440")){
-  # if I'm on the laptop
-  setwd("/home/tim/git/ThanoRepro/ThanoRepro")
-} else {
-  # in that case I'm on Berkeley system, and other people in the dept can run this too
-  setwd(paste0("/data/commons/",system("whoami",intern=TRUE),"/git/ThanoRepro/ThanoRepro"))
-}
 
-source("R/Functions.R")
+if (system("hostname",intern=TRUE) %in% c("triffe-N80Vm", "tim-ThinkPad-L440")){
+	# if I'm on the laptop
+	setwd("/home/tim/git/ThanoRepro/ThanoRepro")
+} else {
+	if (system("hostname",intern=TRUE) == "PC-403478"){
+		# on MPIDR PC
+		setwd("U://git//ThanoRepro//ThanoRepro")
+	} else {
+		# in that case I'm on Berkeley system, and other people in the dept can run this too
+		setwd(paste0("/data/commons/",system("whoami",intern=TRUE),"/git/ThanoRepro/ThanoRepro"))
+	}
+}
+getwd()
+# 
+devtools::load_all(file.path("R","RiffeFunctions"))
 
 # I finally got the discrete (pseudo continuous) equation to match
 # perfectly, so now it seems relevant to try again with the matrix.
@@ -80,8 +86,8 @@ getwd()
 
 #install.packages("expm", repos="http://R-Forge.R-project.org")
 library(expm)
-all(Y%^%100 > 0) # yes, thano is regular
-all(L%^%100 > 0) # Leslie not regular?
+all(Y%^%110 > 0) # yes, thano is regular
+all(L%^%110 > 0) # Leslie not regular?
 
 
 # mixing time
@@ -91,10 +97,10 @@ all(L%^%100 > 0) # Leslie not regular?
 1/log(1/rev(sort(abs(Re(eigen(L)$values))))[2])
 
 # decreasing magnitude (from JHJ 2006 standford workshop notes)
-Ldamping <- Mod(eigen(L)$values[2])/Mod(eigen(L)$values[1])  
-Ydamping <- Mod(eigen(Y)$values[2])/Mod(eigen(Y)$values[1])  
[…]

### 2015-10-15 ThanoRepro/ThanoRepro.tex
changes taken. Ugly conflicts because I forgot to fold in PC changes. Now merged finally. Sucks you can't overwrite remote when using merge tool...
change affil
added funding disclaimer, since work was mostly done in Berkeley

@@
 \usepackage{tikz}
-\usetikzlibrary{arrows}
+%\usetikzlibrary{arrows}
 %\usepackage[active,tightpage]{preview}
@@
 \usepackage{setspace}
+\newcommand\ackn[1]{%
+  \begingroup
+  \renewcommand\thefootnote{}\footnote{#1}%
+  \addtocounter{footnote}{-1}%
+  \endgroup
+}
+
 % matrix numbering
@@
 years of life}
-\author{Tim Riffe \\ University of California, Berkeley}
+\author{Tim Riffe \\ Max Planck Institute for Demographic Research}
 \maketitle
@@
 by chronological versus thanatological age. Results from the two age perspectives are conformable if the starting
-population is stable, but are otherwise divergent.
+population is stable, but are otherwise divergent.\ackn{Research reported in this manuscript
+was supported by the U.S.
+National Institute On Aging of the National Institutes of Health under award numbers R01-AG011552 and R01-AG040245 and by the Spanish Ministry of Science and Innovation under award
+number ECO2013-48326-C2-1-P. The content is solely the responsibility of the
+author and does not necessarily represent the official views of the funding
+agencies.}
 \end{abstract}

### 2015-10-16 ThanoRepro/ThanoRepro.tex
more tinkering and making things consistent. still need to change f(y|a)

@@
 \onehalfspacing
+\section*{Introduction}
 In this paper, I explore some formal demographic consequences of a particular redefinition of age.
@@
 in the reliability literaure, or just time-to-death (TTD) elsewhere.}. The first
-kind of age is called chronological age and the kind of age explored in this paper is
-called thanatological age. Individuals in this case move in the same direction
-along imaginary life lines, but the reference point for age is now at the end of
-the life line instead of at the beginning. For individuals, the timing of events
-in life is identified exactly using either of the two definitions of age, but
-aggregate patterns differ. The models of demography are prefaced by the observation
-that demographic phenomena vary over age and time in empirically regular ways.
-Some such phenomena are best described in the aggregatate as a function
-of time since birth, and others of time until death, while still others can be a function of both to some degree \citep{riffe2015ttd}.
-This observation ought to motivate the formal demographic base by incorporating this aspect of time in into
-models.
-
+kind of age is called chronological age, and the second kind of age is called
+thanatological age. Individuals in this case move in the same direction along
+imaginary life lines, but the reference point for thanatological age is at
+the end of the life line instead of at the beginning. For individuals, the timing of events in life is exactly identified using either of the two definitions of age, but
+aggregate patterns differ. Demography is built upon a small set of empirical
+regularities with respect to chronological age patterns. The models and methods
+of demography are in varying degrees prefaced on such regularity. There is evidence that many demographic phenomena are best described in the
+aggregatate as a function of time since birth, and others of time until death,
+while still others can be a function of both to some degree
+\citep{riffe2015ttd}. This observation ought to motivate the formal demographic
+base to incorporate this aspect of time in into models. 
 %Temporal perspectives and rescalings of the lifecourse are abundant in
@@
 All diverse lines
-of contemporary demographic research are relatable to the renewal model.
-Some might say that stable population theory is not relevant to everyday
-demography, since populations are always changing. Even if change precludes the
+of contemporary demographic research are relatable to the renewal model, and by
+extention to stable population theory. Some might say that stable population
+theory is not relevant to everyday demography, since populations are always changing. Even if change precludes the
 notion of stability as a prediction, models of long-run consequnces still give a
@@
 closed or nearly-closed birth cohorts. Finally, stable population theory is
-relavant to other non-human structured populations, such as stars.
+relavant for modelling some non-human populations.
 
@@
 describe the process of population renewal by relating renewal to a death flow
-rather than to a birth flow. The chronological-age classified renewal model and the thanatological renewal model both approximate the movement of population stocks
-over time, but from opposite vantage points. Specifying a model of
-thanatological renewal expands the toolbox available to approach complex
+rather than to a birth flow. The chronological-age classified renewal model and
+the thanatological renewal model both describe the movement of population stocks
+over time, but from opposite vantage points. Specifying a model of thanatological renewal expands the toolbox available to approach complex
 temporal population structure. The next section of this article describes the lifetable method of transforming age-classified
@@
 chronological age margin, and for the thanatological case we multiply
-$f^\star(y)$ over the thanatological age margin.
+$\gamma(y)$ over the thanatological age margin.
 
@@
 versus thanatological age, unless the starting population is stable.
-Even though the modeled population stocks are in a way commensurable, the rates used here, $f(a)$ versus $f^\star(y)$ are calculated on the basis of differently distributed denominators $E(a)$ versus $E(y)$.
+Even though the modeled population stocks are in a way commensurable, the rates used here, $f(a)$ versus $\gamma(y)$ are calculated on the basis of differently distributed denominators $E(a)$ versus $E(y)$.
 
@@
  T^\star =  \frac{\int _{y=0}^\infty \int _{a=0}^\infty y e^{-ra} d(a+y)
- f^\star(y) \dd a \dd y}{\int _{y=0}^\infty \int _{a=0}^\infty e^{-ra} d(a+y)
- f^\star(y) \dd a \dd y} \ep
+ \gamma(y) \dd a \dd y}{\int _{y=0}^\infty \int _{a=0}^\infty e^{-ra} d(a+y)
+ \gamma(y) \dd a \dd y} \ep
 \end{equation}
@@
 \intertext{the denominator of which is the inside of
-equation~\eqref{eq:thanoren} less $f^\star(y)$. This simplifies to} =&
+equation~\eqref{eq:thanoren} less $\gamma(y)$. This simplifies to} =&
 \frac{1}{ \int _{a=0}^\infty e^{-ra} l(a) \dd a } \ec
@@
 known, then we may take the sum of hypothetical births over thanatological age:}
-=& \int _{y=0}^\infty f^\star(y) c^\star(y) \dd y \ep
+=& \int _{y=0}^\infty \gamma(y) c^\star(y) \dd y \ep
 \end{align}
@@
 stable population quantities may be estimated by similarly translating the various common definitions (e.g.,
-in the glossary of \citet{coale1972growth}) to the present perspective. We will
-focus on the main model rather than on these.
+in the glossary of \citet{coale1972growth}) to the present perspective. I focus on the main model rather than on these.
 
@@
  thanatological classes 0-4, while the final class has no fertility, where
- $f^\star_y$ indicates the fertility probability for class $y$ in the year $t$
- entering population (in the matrix columns). Each $f^\star_y$ is then
+ $\gamma_y$ indicates the fertility probability for class $y$ in the year $t$
+ entering population (in the matrix columns). Each $\gamma_y$ is then
  distributed out over rows according to $d_a$.
  The fertility entry in row $m$ and column $n$ of $\textbf{Y}$ will therefore be
- $f^\star_n \cdot d_m$. One can assume that those dying over the course of year
+ $\gamma_n \cdot d_m$. One can assume that those dying over the course of year
  $t$ (the first column) are exposed to fertility for half of the year,\footnote{One might be tempted to not allow for fertility at all for females dying in year $t$, but recall that fertility is measured in the moment of
@@
  accordingly.
- Further, infant mortality, $f^\star_y \cdot d_0$, located in the first row, must also be discounted, since part
+ Further, infant mortality, $\gamma_y \cdot d_0$, located in the first row, must also be discounted, since part
  of the mortality will occur in the same year $t$ and the rest in year $t + 1$. 
@@
   {e_y } \vspace{.6em}&                0_t  & 1_t  & 2_t  & 3_t  & 4_t  & 5_t\cr 
-   0_{t+1} \vspace{.6em}& (1-\lambda) \tfrac{f^\star_0d_0}{2} & (1-\lambda) f^\star_1d_0 & (1-\lambda)
-   f^\star_2d_0 & (1-\lambda) f^\star_3d_0 & (1-\lambda) f^\star_4d_0 & 0 \cr 
-   1_{t+1} \vspace{.6em}& \tfrac{f^\star_0d_1}{2} & f^\star_1d_1 & f^\star_2d_1 & f^\star_3d_1 & f^\star_4d_1
-   & 0   \cr 2_{t+1} \vspace{.6em}& \tfrac{f^\star_0d_2}{2} & f^\star_1d_2 & f^\star_2d_2 & f^\star_3d_2 & f^\star_4d_2
-   & 0   \cr 3_{t+1} \vspace{.6em}& \tfrac{f^\star_0d_3}{2} & f^\star_1d_3 & f^\star_2d_3 & f^\star_3d_3 & f^\star_4d_3
-   & 0   \cr 4_{t+1} \vspace{.6em}& \tfrac{f^\star_0d_4}{2} & f^\star_1d_4 & f^\star_2d_4 & f^\star_3d_4 & f^\star_4d_4
-   & 0   \cr 5_{t+1} \vspace{.6em}& \tfrac{f^\star_0d_5}{2} & f^\star_1d_5 & f^\star_2d_5 & f^\star_3d_5 & f^\star_4d_5
+   0_{t+1} \vspace{.6em}& (1-\lambda) \tfrac{\gamma_0d_0}{2} & (1-\lambda) \gamma_1d_0 & (1-\lambda)
+   \gamma_2d_0 & (1-\lambda) \gamma_3d_0 & (1-\lambda) \gamma_4d_0 & 0 \cr 
+   1_{t+1} \vspace{.6em}& \tfrac{\gamma_0d_1}{2} & \gamma_1d_1 & \gamma_2d_1 & \gamma_3d_1 & \gamma_4d_1
+   & 0   \cr 2_{t+1} \vspace{.6em}& \tfrac{\gamma_0d_2}{2} & \gamma_1d_2 & \gamma_2d_2 & \gamma_3d_2 & \gamma_4d_2
+   & 0   \cr 3_{t+1} \vspace{.6em}& \tfrac{\gamma_0d_3}{2} & \gamma_1d_3 & \gamma_2d_3 & \gamma_3d_3 & \gamma_4d_3
+   & 0   \cr 4_{t+1} \vspace{.6em}& \tfrac{\gamma_0d_4}{2} & \gamma_1d_4 & \gamma_2d_4 & \gamma_3d_4 & \gamma_4d_4
+   & 0   \cr 5_{t+1} \vspace{.6em}& \tfrac{\gamma_0d_5}{2} & \gamma_1d_5 & \gamma_2d_5 & \gamma_3d_5 & \gamma_4d_5
    & 0   }$
@@
   {e_y } \vspace{.6em} & 0_t  & 1_t  & 2_t  & 3_t  & 4_t  & 5_t\cr 
-  0_{t+1} \vspace{.6em}&  (1-\lambda) \tfrac{f^\star_0d_0}{2} & (1-\lambda) f^\star_1d_0 + 1 &
-  (1-\lambda) f^\star_2d_0 & (1-\lambda) f^\star_3d_0 & (1-\lambda) f^\star_4d_0 & 0 \cr 
-    1_{t+1} \vspace{.6em}& \tfrac{f^\star_0d_1}{2} & f^\star_1d_1 & f^\star_2d_1 + 1 & f^\star_3d_1 & f^\star_4d_1 & 0 \cr 
-    2_{t+1} \vspace{.6em}& \tfrac{f^\star_0d_2}{2} & f^\star_1d_2 & f^\star_2d_2 & f^\star_3d_2 + 1 & f^\star_4d_2 & 0 \cr 
-   3_{t+1} \vspace{.6em}& \tfrac{f^\star_0d_3}{2} & f^\star_1d_3 & f^\star_2d_3 & f^\star_3d_3 & f^\star_4d_3 + 1 & 0 \cr 
-   4_{t+1} \vspace{.6em}& \tfrac{f^\star_0d_4}{2} & f^\star_1d_4 & f^\star_2d_4 & f^\star_3d_4 & f^\star_4d_4 & 1 \cr 
-   5_{t+1} \vspace{.6em}& \tfrac{f^\star_0d_5}{2} & f^\star_1d_5 & f^\star_2d_5 & f^\star_3d_5 & f^\star_4d_5 & 0 }$
+  0_{t+1} \vspace{.6em}&  (1-\lambda) \tfrac{\gamma_0d_0}{2} & (1-\lambda) \gamma_1d_0 + 1 &
+  (1-\lambda) \gamma_2d_0 & (1-\lambda) \gamma_3d_0 & (1-\lambda) \gamma_4d_0 & 0 \cr 
+    1_{t+1} \vspace{.6em}& \tfrac{\gamma_0d_1}{2} & \gamma_1d_1 & \gamma_2d_1 + 1 & \gamma_3d_1 & \gamma_4d_1 & 0 \cr 
+    2_{t+1} \vspace{.6em}& \tfrac{\gamma_0d_2}{2} & \gamma_1d_2 & \gamma_2d_2 & \gamma_3d_2 + 1 & \gamma_4d_2 & 0 \cr 
+   3_{t+1} \vspace{.6em}& \tfrac{\gamma_0d_3}{2} & \gamma_1d_3 & \gamma_2d_3 & \gamma_3d_3 & \gamma_4d_3 + 1 & 0 \cr 
+   4_{t+1} \vspace{.6em}& \tfrac{\gamma_0d_4}{2} & \gamma_1d_4 & \gamma_2d_4 & \gamma_3d_4 & \gamma_4d_4 & 1 \cr 
+   5_{t+1} \vspace{.6em}& \tfrac{\gamma_0d_5}{2} & \gamma_1d_5 & \gamma_2d_5 & \gamma_3d_5 & \gamma_4d_5 & 0 }$
 \end{matrix}
@@
 than zero, which for this toy matrix are all age classes except the \nth{5}
-(highest). The y$^{th}$ age class is subject to the fertility rate $f^\star
+(highest). The y$^{th}$ age class is subject to the fertility rate $\gamma
 (y)$, and these births distribute out to all other age classes according to the
@@
     \path[every node/.style={font=\sffamily\small}]
-    (4) edge [bend right] node {$f^\star (4)d(5)$} (5)
-        edge [bend left] node {$f^\star (4)d(3)$} (3)
-        edge [bend left] node {$f^\star (4)d(2)$} (2)
-        edge [bend left] node {$f^\star (4)d(1)$} (1)
-        edge [bend left] node {$(1-\lambda)f^\star (4)d(0)$} (0)
-        edge [loop above] node {$f^\star (4)d(4)$} (4)
-        edge [bend left] node {$\lambda f^\star (4)d(0)$} (X);
+    (4) edge [bend right] node {$\gamma(4)d(5)$} (5)
+        edge [bend left] node {$\gamma(4)d(3)$} (3)
+        edge [bend left] node {$\gamma(4)d(2)$} (2)
+        edge [bend left] node {$\gamma(4)d(1)$} (1)
+        edge [bend left] node {$(1-\lambda)\gamma (4)d(0)$} (0)
+        edge [loop above] node {$\gamma(4)d(4)$} (4)
+        edge [bend left] node {$\lambda \gamma(4)d(0)$} (X);
 
@@
 The differences described here are not due primarily to rounding errors in the
-data, but rather to the way in which $f^\star(y)$ has been calculated. 
+data, but rather to the way in which $\gamma(y)$ has been calculated. 
 Specifically, in this case we have transformed fertility rates from empirical data with a particular
 chronological age structure. If the chronological age structure is far from its
-stable distribution, then it may be hard to imagine that $f^\star(y)$ will
+stable distribution, then it may be hard to imagine that $\gamma(y)$ will
 remain fixed over time. When holding chronological-age classified fertility rates
@@
 \citet{pressat1973analyse}. Define a convenience function, $I(r)$, for the
-integrand of \eqref{eq:thanoren} for a given $r$ and fixed $f^\star(y)$ and $d(a)$:
+integrand of \eqref{eq:thanoren} for a given $r$ and fixed $\gamma(y)$ and $d(a)$:
 
 \begin{equation}
-I(r) = \int_{y=0}^\infty \int_{a=0}^\infty f^\star(y) d(a+y)e^{-ra}\dd a \dd y
+I(r) = \int_{y=0}^\infty \int_{a=0}^\infty \gamma(y) d(a+y)e^{-ra}\dd a \dd y
 \end{equation}
 Since the death distribution function, $d(a)$, and fertility
-function, $f^\star(y)$, are continuous and non-negative, $\lim_{r \to +\infty} I(r)
+function, $\gamma(y)$, are continuous and non-negative, $\lim_{r \to +\infty} I(r)
 = 0$ and $\lim_{r \to -\infty} I(r)= \infty$. If $r_2 > r_1$, then $I(r_1) >
@@
 \int _{a=0}^\infty l(a)e^{-ra}f(a) \dd a &= \int _{y=0}^\infty \int
-_{a=0}^\infty d(a+y)e^{-ra}f^\star (y) \dd a \dd y
+_{a=0}^\infty d(a+y)e^{-ra}\gamma (y) \dd a \dd y
 \intertext{when}
-f^\star(y) &= \frac{b^\star(y)}{c^\star(y)} \\
+\gamma(y) &= \frac{b^\star(y)}{c^\star(y)} \\
 b^\star (y) &= \int_{a=0}^\infty b(a) \mu(a+y)\frac{l(a+y)}{l(a)} \dd a \\
@@
 b(a) &= f(a)c(a)
-\end{align}Replacing $f^\star (y)$ with the full expanded
+\end{align}Replacing $\gamma(y)$ with the full expanded
 expression and plugging into the thanatological renewal equation gives:
@@
 \widehat{T^\star} = \frac{\int _{y=0}^\infty \int _{a=0}^\infty y d (a+y)
-f^\star(y) \dd a \dd y}{\int _{y=0}^\infty \int _{a=0}^\infty d(a+y) f^\star(y)
+\gamma(y) \dd a \dd y}{\int _{y=0}^\infty \int _{a=0}^\infty d(a+y) \gamma(y)
 \dd a \dd y}
@@
  \begin{equation}
-  R_0 = \int _{y=0}^\infty \int _{a=0}^\infty d(a+y) f^\star(y) \dd a
+  R_0 = \int _{y=0}^\infty \int _{a=0}^\infty d(a+y) \gamma(y) \dd a
 \dd y

### 2015-10-19 ThanoRepro/ThanoRepro.tex
added force of increment paper

@@
 aggregatate as a function of time since birth, and others of time until death,
-while still others can be a function of both to some degree
+while others can be a function of both to some degree
 \citep{riffe2015ttd}. This observation ought to motivate the formal demographic
@@
 I describe the Lotka-Euler-Leslie renewal model when structured by
-thanatological age.
-The renewal model relates fertility, mortality, and population structure in a parsimonious form, revealing the long term consequences of a particular set of conditions.
-All diverse lines
-of contemporary demographic research are relatable to the renewal model, and by
-extention to stable population theory. Some might say that stable population
-theory is not relevant to everyday demography, since populations are always changing. Even if change precludes the
-notion of stability as a prediction, models of long-run consequnces still give a
-useful measure of the demographic force implied by a given set of rates.
-In this and other ways, stability is never far away. That population projections
-lead into a stable state means that some form of stability is implied by the routine practice of demography.
-Stability or even stationarity are imaginable for the world population of humans
-one day. Models of stationary populations are easily applied by analogy to
-closed or nearly-closed birth cohorts. Finally, stable population theory is
-relavant for modelling some non-human populations.
-
+thanatological age. The renewal model relates
+fertility, mortality, and population structure in a parsimonious form, revealing the long term consequences of a particular set of conditions.
+All diverse lines of contemporary demographic research are relatable to the renewal model, and by extention to stable population theory.
+%Some might say that stable population
+%theory is not relevant to everyday demography, since populations are always
+% changing. Even if change precludes the notion of stability as a prediction, models of long-run consequnces still give a
+%useful measure of the demographic force implied by a given set of rates.
+%In this and other ways, stability is never far away. That population
+% projections lead into a stable state means that some form of stability is implied by the routine practice of demography.
+%Stability or even stationarity are imaginable for the world population of
+% humans one day. Models of stationary populations are easily applied by analogy to
+%closed or nearly-closed birth cohorts. Finally, stable population theory is
+%relavant for modelling some non-human populations. 
 %We have stated that many lifecourse transitions covary more with remaining
@@
 % remaining lifetime as a particular class of time-to-transition, one may generalize the lifetable methods presented here into new area applications, such as education, partnering, the economic lifecycle, or migration.
-
 To present the renewal model based on thanatological age is not necessarily a
@@
 describe the process of population renewal by relating renewal to a death flow
-rather than to a birth flow. The chronological-age classified renewal model and
-the thanatological renewal model both describe the movement of population stocks
-over time, but from opposite vantage points. Specifying a model of thanatological renewal expands the toolbox available to approach complex
+rather than to a birth flow. I have described a limited case of this
+perspective on renewal for the case of stationary populations
+\citep{riffe2015force}, and here aim to provide a broader treatment allowing for
+growth in the population.
+The chronological-age classified renewal model and the thanatological renewal model both describe the movement of population stocks over time, but from opposite vantage points.
+Specifying a model of thanatological renewal expands the toolbox available to approach complex
 temporal population structure. The next section of this article describes the lifetable method of transforming age-classified
@@
 \section*{A lifetable approximation of thanatological age}
-Direct information on the timing of demographic transitions with respect to
-death is available retrospectively from linked population registers or
-panel surveys with mortality follow-ups.
-Some surveys also ask respondents in various ways about their perceived future lifetime, which may underly those
-demographic transitions governed to some degree by volition. The present inquiry
-draws only from lifetable-projected remaining lifetime, which
-is intended to approximate actual remaining lifetime under the strong
-assumption of homogenous populations. Chronological age-classified counts are
-redistributed over future death times according to a fixed mortality schedule.
-This allows us to make statements about the remaining lifetime distribution
-within a stock of population, under certain assumptions.
-
-A lifetable approach to transforming chronologically classified data into
-thanatologically classified data is valid when the given lifetable accurately
-describes the future mortality of age-classified population. To make
-a projective statement about a current stock of population, one does well to
-factor in future changes in mortality, and to allow the trajectory of changes
-to vary over ages and cohorts. In the case of a stable population model,
-mortality is assumed invariant over time, which means that the same survival
-pattern underlies each birth cohort. This is the case we explore, and so the
-following formulas assume a single lifetable pattern.
+%Direct information on the timing of demographic transitions with respect to
+%death is available retrospectively from linked population registers or
+%panel surveys with mortality follow-ups.
+%Some surveys also ask respondents in various ways about their perceived future
+% lifetime, which may underly those demographic transitions governed to some degree by volition. 
+In general, to make a lifetable-based projective statement about the remaining
+time distribution of a current stock of population, it is best to to account for population subgroups
+with different mortality and to account for uncertainty about future changes in
+mortality.
+The present inquiry approximates thanatological age using lifetable-projected remaining lifetime,
+which is intended to approximate remaining lifetime under the strong
+assumptions of homogenous populations und fixed vital rates. 
+%Chronological age-classified counts are redistributed over future death times
+% according to a fixed mortality schedule.
+%In general, a lifetable approach to transforming chronologically classified
+%data into thanatologically classified data is sound if the given lifetable
+%accurately describes the future mortality of the age-classified population, and
+%if the chronological and thantological age patterns are independent.
+In this case, each birth cohort undergoes the same
+pattern of attrition. In a period setting, this means that we can distribute the
+remaining lifespans of the population in each age class using the
+age-conditioned lifetable.
+%This is the case we explore, and so the following formulas assume a single lifetable pattern.
 
 The basic transformation to thanatological age proceeds by defining a
-deaths density function, similar the lifetable $d(a)$, but over two
-dimensions: chronological age attained, $a$, and remaining years of life, $y$.
-This function we denote $f(y|a)$ --- The density distribution of thanatological
-age conditional on chronological age. This is essentially the same notion as the
-familiar $d(a)$, but rescaled as a function of chronological age. For ages higher than 0, simply condition
-on survival:\footnote{In the reliability literature, our concept of $d(y|a)$ is denoted by \begin{equation} X - t | X > t
-\end{equation}
-where $X$ is the total lifespan and $t$ is the current chronological age
-attained.}
+conditional deaths density function, $f(y|a)$. For chronological
+age zero, $f(y|0)$ is identical to $d(a)$. For ages higher than 0, simply
+condition on survival:\footnote{In the reliability literature, $f(y|a)$ is
+denoted by $ X - t | X > t $, where $X$ is the total lifespan and $t$ is the
+current chronological age attained.}
 \begin{equation}
@@
 chronological age-classified population count, we can derive $P(y)$, a
-thanatological age-classified population count using
+thanatological age-classified population count using\footnote{An identical
+definition is found in \citet{brouard1989mouvements}.}
 
@@
 with historical stocks. \citet{brouard1989mouvements}, and later
-\citet{vaupel2009life}, showed that in stationary populations the chronological
-and thanatological age structures are identical, and this provides a
-limited basis for comparison between the two perspectives. \citet{wachter2014essential} hints at, and
-\citet{vaupel2014stable} elaborates on how to make chronological and
-thanatological age structure commensurate under a constant growth rate, $r$.
+\citet{vaupel2009life}, showed that in stationary populations the
+marginal chronological and thanatological age structures are identical, and this
+provides a limited basis for comparison between the two perspectives. 
+%\citet{wachter2014essential} hints at, and
+%\citet{vaupel2014stable} elaborates on how to make chronological and
+%thanatological age structure commensurate under a constant growth rate, $r$.
 These results will also follow from the following specification of a
@@
   reproduction, $\widehat{T^\star}$, akin to Lotka's mean generation time, $T$.
-  To start, a good-enough guess is to just assume $r=0$:
+  To start, a good-enough guess is to just assume $r=0$:\footnote{Using the
+  Brouard-Carey equality, the denominator could also be: $\int _{y=0}^\infty
+  \ell(y) \gamma(y) \dd y$}
 \begin{equation}
@@
 \end{equation}
-  \item A first rough guess at the net reproduction rate, $R_0$ is given by
+  \item A first rough guess at the net reproduction rate, $R_0^\star$ is given
+  by
  \begin{equation}
-  R_0 = \int _{y=0}^\infty \int _{a=0}^\infty d(a+y) \gamma(y) \dd a
+  R_0^\star = \int _{y=0}^\infty \int _{a=0}^\infty d(a+y) \gamma(y) \dd a
 \dd y
@@
    \begin{equation}
-   r^0 = \frac{ln(R_0)}{\widehat{T^\star}}
+   r^0 = \frac{ln(R_0^\star)}{\widehat{T^\star}}
    \end{equation}

### 2015-10-19 ThanoRepro/R/rEqual.R

@@ -27,6 +27,22 @@ Data$Fy[is.na(Data$Fy)]     <- 0
 Data$Fyf[is.nan(Data$Fyf)]  <- 0
 DATA                        <- data.table(Data)
 
+#lx1 <- Data$lx[1:111]
+#fx1 <- Data$Fxf[1:111]
+#fy1 <- Data$Fyf[1:111]
+#
+#dx1 <- Data$dx[1:111]
+#lx1 <- rev(cumsum(rev(dx1)))
+#
+#by1 <- fx1 * lx1
+#
+#by1 <- rowSums(Thano(by1, dx1, stagger=FALSE))
+#fy1 <- by1 / lx1
+#fy1[is.nan(fy1)] <- 0
+#
+#sum(lx1*fx1)
+#sum(lx1*fy1)
+
 # head(Data)
 # f(a) is the fertility function, not a density function!
 fa <- with(Data, Fxf[Year == 2000 & Code == "SWE"])
@@ -216,3 +232,38 @@ sum(fa * dxM * growth)
 # bam!
 sum(fa * la2 * growth)
 # this should serve to figure it out on paper
+
+
+################################################3
+# this is a side test.
+# what about homogenous stationary vs homogenous stationary?
+#dx2lx <- function(dx){
+#	rev(cumsum(rev(dx)))
+#}
+#dx1 <- Data$dx[1:111]
+#sum(.5:110.5*dx1)
+#
+#
+#dx2 <- with(Data,dx[Year==2000 & Sex == "f" & Code == "AUT"])
+#
+#p <- .5
+#
+#plot(0:110,dx1,type='l',col="red")
+#lines(0:110,dx2,col="blue")
+#
+#lines(0:110,(dx1 * p +dx2 * (1-p)),col="green")
+#
+#lx1 <- dx2lx(dx1)
+#lx2 <- dx2lx(dx2)
+#
+#plot(0:110, (lx1*p+lx2*(1-p)), type = 'l', col = "green")
+#lines(0:110, lx1, col = "red")
+#lines(0:110, lx2, col = "blue")
+
+
+
+
+
+
+
+

### 2015-10-19 ThanoRepro/ThanoRepro.tex
.

@@
 perspective on renewal for the case of stationary populations
-\citep{riffe2015force}, and here aim to provide a broader treatment allowing for
+\citep{riffe2015force}, and here provide a broader treatment allowing for
 growth in the population.
@@
 pattern of attrition. In a period setting, this means that we can distribute the
-remaining lifespans of the population in each age class using the
-age-conditioned lifetable.
+remaining lifespans of the population in each age class using the same
+lifetable.
 %This is the case we explore, and so the following formulas assume a single lifetable pattern.
@@
 chronological age-classified population count, we can derive $P(y)$, a
-thanatological age-classified population count using\footnote{An identical
-definition is found in \citet{brouard1989mouvements}.}
+thanatological age-classified population count using\footnote{An
+different-looking but identical definition is found in
+\citet{brouard1989mouvements}.}
 
@@
 this density is useful for decomposing data classified by chronological age. For
-example, we can break down a population pyramid into thanatological age classes.
+example, under these assumptions one can break down a population pyramid into
+thanatological age classes.
 Figure~\ref{fig:USdecomp} shows this exercise for 2010 US data, assuming the

### 2015-10-20 ThanoRepro/ThanoRepro.tex
simplified references file

@@
 Since $1 = \int_0^\infty f(y|a) \dd y$ for each $a$,
-this density is useful for decomposing data classified by chronological age. For
-example, under these assumptions one can break down a population pyramid into
-thanatological age classes.
-Figure~\ref{fig:USdecomp} shows this exercise for 2010 US data, assuming the
-2010 period mortality schedule remains constant. When looking from the
-chronological age perspective, this decomposition reveals thanatological
+this density is useful for decomposing data classified by chronological age,
+such as the population pyramid, into thanatological age classes.
+Figure~\ref{fig:USdecomp} shows this exercise for 2010 US data \citep{HMD},
+assuming the 2010 period mortality schedule remains constant. When looking from the chronological age perspective, this decomposition reveals thanatological
 heterogeneity, but when viewed from the thanatological perspective

### 2015-10-20 ThanoRepro/ThanoRepro.tex
working through the proof again

@@
 
-It bears repeating that lifetable-derived thanatological age structure is probabilistic until after death, since it refers to the future according to an assumed mortality pattern.
-As such, Figure~\ref{fig:USrecomp} is projective in nature, a forward-looking
-glance at population attrition given the population stock and lifetable of a
-particular moment, whereas Figure~\ref{fig:USdecomp} is reflective in nature,
-since a population's chronological age-structure is mostly the fruit of past
-fertility, but also migration and to a lesser degree attrition. Any count
+It bears repeating that lifetable-derived thanatological age structure is
+probabilistic until after death, since it refers to the future according to an
+assumed mortality pattern; in this case a period mortality pattern is used.
+As such, Figure~\ref{fig:USrecomp} is \textit{projective} in nature, a
+forward-looking glance at population attrition given the population stock and lifetable of a
+particular moment, whereas Figure~\ref{fig:USdecomp} is \textit{reflective} in
+nature, since a population's chronological age-structure is mostly the fruit of past
+fertility, but also migration, and to a lesser degree attrition. Any count
 classified by chronological age can be reclassified in this way given a
-corresponding lifetable. Exploiting this, we derive thanatological fertility
-rates, $\gamma(y)$, by applying \eqref{eq:transform} to age-classified birth
-counts, $B(a)$, to get $B(y)$ and again to exposure-to-risk, $E(a)$ (here we
-take exposures from all ages), to get $E(y)$ and then dividing
+corresponding lifetable. To derive thanatological fertility
+rates under these assumptions, $\gamma(y)$, apply \eqref{eq:transform} to
+chronological age-classified birth counts, $B(a)$, to get $B(y)$ and again to
+exposure-to-risk, $E(a)$, to calcualte $E(y)$ and then divide in the usual way:
 
@@
 
-We are interested to use thanatological fertility rates, $\gamma(y)$ in
-models of population renewal,\footnote{This is probably premature, but helps to
-understand the nature and provenance of model components.} rather than to
-explicitly study the nature of these rates. Nonetheless it is best not to throw such rates into a model blindly, so we offer a schematic overview of some of their characteristics.
-Figure~\ref{fig:FySpaghetti} represents the full variety in
-female thanatological period fertility rates that can be found for all years
+Thanatological fertility rates, $\gamma(y)$, are briefly shown because they
+are used in the renewal model here studied. Figure~\ref{fig:FySpaghetti}
+represents the full variety in female thanatological period fertility rates that can be found for all years
 of data that overlap in the HFD and HMD.\footnote{There are as of this writing
@@
 populations and years as a more familiar reference, but note that both scales
-are different. With no indication of particular populations or time series, one
-already concludes that period thanatological fertility rates have a
-characteristic shape and are not random, erratic or informationless: There is a
-pattern to fertility rates by remaining years of life, and it varies over time
-and between populations within some range of normality. Thanatological fertility
-rates have a wider distribution than do chronological rates. Note that this
-spaghetti plot includes both fertility booms and busts, as well as some
-mortality crises (1918, WWII). Some patterns to note, but which we do not
-separate graphically here, are that the left tail (a gauge of how orphan-prone a
-population is) has tended to fall over time. All populations have shown a
-rightward shift in both the mean and modal mother's thanatological age at
-childbearing (TAC) over time. Several
-populations now show modal TACs of over 60 years. Thanatological total fertility rates (TTFR)
-track standard period total fertility rates (PTFR) rather closely, but tend to
-be somewhat higher (not shown).
+are different. The shape of thanatological fertility curves is speculative and
+due to the above-mentioned strong assumptions. Nonetheless it is best not to
+include such rates into a model blindly, so we offer this schematic overview of
+some of their characteristics.
+
+Period thanatological fertility rates estimated in this way have a
+characteristic shape. The distribution is wider than that of chronological
+fertility rates. The range of post-childbearing lifespans stretches from 
+zero, due e.g., to maternal mortality, to at least 100.\footnote{For instance,
+Jeanne Calment lived for another 100 years after giving birth to her daughter}
+This spaghetti plot includes both fertility booms and busts, as well as some
+mortality crises (1918, WWII). Some patterns in the data (not
+shown) include that the left tail (a gauge of how orphan-prone a population is)
+has tended to fall over time. All populations have shown an increase in both the
+mean and mode of thanatological age at childbearing over time. Several
+contemporary populations now show modes of over 60 years. Period thanatological
+total fertility rates track standard period total fertility rates closely, but tend to be somewhat higher (not shown).
 % both figures created in R/HFDspaghetti.R
 \begin{figure}[h!]
-	\caption{Chronological and thantological fertility rates, all 1600
+	\caption{Chronological and thantological fertility rates, all 1933
 	country-year combinations present in both the HFD and HMD.* Note different x
@@
 In general, thanatological fertility rates will not be useful for purposes of
-projection, unless the demographer believes that their empirical regularity is
-somehow stronger than typical ASFR. This is a question we do not seek to answer
-in the present treatment. There are alternative ways of defining
-thanatological fertility rates, for instance by transforming the birth and
+projection, unless their empirical regularity is
+stronger than typical ASFR. This is a question I do not seek to answer
+in the present treatment, but in any case such rates are not susceptible to
+direct measurement.
+There are alternative ways of defining thanatological fertility rates, for instance by transforming the birth and
 population vector from a stable population. As proven in appendix~\ref{app:B},
 thanatological fertility rates derived from a stable population imply a certain
-symmetry with stable populations from the chronological model. First let us
+symmetry with stable populations from the chronological model. First I
 define the basic model.
@@
 \section*{The thanatological renewal model}
-For the renewal model that follows, we are most interested in using the kind
-of fertility rates shown in Figure~\ref{fig:FySpaghetti}, and we refer to a
-closed, unisex population based on female vital rates. Assuming constant vital rates, the
-births for the present year are given by 
+The renewal model that follows uses thanatological fertility rates, $\gamma(y)$,
+the form of which is suggested in Figure~\ref{fig:FySpaghetti}, and we refer to
+a homogenous, closed, unisex population based on female vital rates. Assuming constant vital rates, the births for the present year are given by 
 \begin{align}
@@
 thanatological age, $P(a), P(y)$ are population counts, and $\gamma(y), m(a)$ are
-exact specific fertility probabilities (rates), best imagined
-as the fertility rates of females born to females. The thanatological integral can
-be broken down back in terms of chronological age:} =& \int_{y=0}^\infty
-\int_{a=0}^\infty \gamma(y)\frac{P(a,t)d(a+y)}{\ell(a)} \dd a \dd y \ec
-\intertext{where $d(a)$ is the continuous lifetable death distribution with
-radix of 1, so $d(a)/\ell(0)$. We can relate the present population to past births
-with $P(a,t) = B(t-a)\ell(a)$:}
+exact fertility probabilities (rates), best imagined
+as the fertility rates of females born to females. The thanatological
+(first) integral can be broken down back in terms of chronological age:} =&
+\int_{y=0}^\infty \int_{a=0}^\infty \gamma(y)P(a,t)f(y|a) \dd a \dd y \ec
+\intertext{where $f(y|a)$ is defined per equation~\eqref{eq:vaupel1}, so
+$d(a)/\ell(0)$.
+We can relate the present population to past births with $P(a,t) = B(t-a)\ell(a)$:}
 \label{eq:ergostep1}
@@
 and note that \eqref{eq:lotka2} and \eqref{eq:thanoren} are really quite
-similar, since $\int _{y=0}^\infty d(a+y)\dd a = l(a)$. For intuition, notice that $l(a)$ is here split
-up into pieces of $d(a)$, and imagine a 2D surface of these, where one axis is chronological age and the
-other axis is thanatological age. For the chronological case \eqref{eq:lotka},
-we multiply chronological age-specific fertility rates, $f(a)$, over the
-chronological age margin, and for the thanatological case we multiply
+similar, since $\int _{y=0}^\infty d(a+y)\dd a = l(a)$. For intuition, notice
+that $l(a)$ is here decomposed into $d(a)$, and imagine a matrix of these,
+where one margin is chronological age and the other margin is thanatological age.
+For the chronological case \eqref{eq:lotka}, we multiply chronological age-specific fertility rates, $f(a)$, over the chronological age margin, and for the thanatological case we multiply
 $\gamma(y)$ over the thanatological age margin.
@@
 There are gaps in the above line of development, since the jump from
-\eqref{eq:ergostep1} to \eqref{eq:ergostep2} (strong ergodicity) is unproven,
-although it is rather intuitive, given the much greater density of connections
-within the model; persons from nearly any thanatologcal age can produce offspring that can have any other thanatological age.\footnote{I say
-\textit{nearly} because the very highest thanatological ages tend to be
-composed of pre-menarchical females, so these can be thought of as structural
-zeros. I am uncertain if it is a record, but it is certainly an extreme case
-useful for illustration: Jeanne Calment, who lived to 122, had a daughter at age
-22, which means she was about thanatological age 100
-(\url{http://en.wikipedia.org/wiki/Jeanne_Calment}). If the lifetable used to
-transform to thanatological closes out at age 110+, as is the case with HMD data, then fertility rates for thantological ages 101-110+ would be structural
+\eqref{eq:ergostep1} to \eqref{eq:ergostep2} (strong ergodicity) is assumed for
+the moment. The step has some intuition, due to
+the much greater density of connections within the model; persons from nearly any
+thanatologcal age can produce offspring that can have any other
+thanatological age.\footnote{I say \textit{nearly} because the very highest thanatological ages tend to be composed of pre-menarchical females, so these can be thought of as structural
 zeros.} In
@@
 \end{figure}
+\FloatBarrier
 
@@
 
-\section{Discussion}
+\section*{Discussion}
 The thanatological renewal model is valid, but this perspective may take some
@@
 \begin{align}
-\int _{a=0}^\infty l(a)e^{-ra}f(a) \dd a &= \int _{y=0}^\infty \int
+\int _{a=0}^\infty \ell(a)e^{-ra}f(a) \dd a &= \int _{y=0}^\infty \int
 _{a=0}^\infty d(a+y)e^{-ra}\gamma (y) \dd a \dd y
@@
 \gamma(y) &= \frac{b^\star(y)}{c^\star(y)} \\
-b^\star (y) &= \int_{a=0}^\infty b(a) \mu(a+y)\frac{l(a+y)}{l(a)} \dd a \\
-c^\star (y) &= \int_{a=0}^\infty c(a) \mu(a+y)\frac{l(a+y)}{l(a)} \dd a
+b^\star (y) &= \int_{a=0}^\infty b(a) \mu(a+y)\frac{\ell(a+y)}{\ell(a)} \dd a \\
+c^\star (y) &= \int_{a=0}^\infty c(a) \mu(a+y)\frac{\ell(a+y)}{\ell(a)} \dd a
 \intertext{Note that:}
-c(a) &= \frac{l(a)e^{-ra}}{\int l(a) e^{-ra} \dd a} \\
+c(a) &= \frac{\ell(a)e^{-ra}}{\int \ell(a) e^{-ra} \dd a} \\
 b(a) &= f(a)c(a)
@@
 _{a=0}^\infty d(a+y)e^{-ra} \frac{\frac{\int_{t=0}^\infty
-f(t)l(t)e^{-rt}\mu(a+y)\frac{l(t+y)}{l(t)} \dd t}{\int l(t) e^{-rt} \dd
-t}}{\frac{\int_{t=0}^\infty l(t)e^{-rt} \mu(t+y)\frac{l(t+y)}{l(t)} \dd t}{\int
-l(t) e^{-rt} \dd t}} \dd a \dd y
+f(t)\ell(t)e^{-rt}\mu(a+y)\frac{\ell(t+y)}{\ell(t)} \dd t}{\int \ell(t) e^{-rt} \dd
+t}}{\frac{\int_{t=0}^\infty \ell(t)e^{-rt} \mu(t+y)\frac{\ell(t+y)}{\ell(t)} \dd t}{\int
+\ell(t) e^{-rt} \dd t}} \dd a \dd y
 \intertext{The denominator of $c(a)$ cancels out:}
@@
 \int_{t=0}^\infty
-f(t)l(t)e^{-rt}\mu(t+y)\frac{l(t+y)}{l(t)} \dd t}{
-\int_{t=0}^\infty l(t)e^{-rt} \mu(t+y)\frac{l(t+y)}{l(t)} \dd t} \dd a \dd y
+f(t)\ell(t)e^{-rt}\mu(t+y)\frac{\ell(t+y)}{\ell(t)} \dd t}{
+\int_{t=0}^\infty \ell(t)e^{-rt} \mu(t+y)\frac{\ell(t+y)}{\ell(t)} \dd t} \dd a \dd y
 \intertext{Some $l(a)$'s also cancel out:}
@@
 \int_{t=0}^\infty
-f(t)e^{-rt}\mu(t+y)l(t+y) \dd t}{
-\int_{t=0}^\infty e^{-rt} \mu(t+y)l(t+y) \dd t} \dd a \dd y
-\intertext{Now $d(a+y) = \mu (a+y)l(a+y)$, so we can cancel the inner integral
+f(t)e^{-rt}\mu(t+y)\ell(t+y) \dd t}{
+\int_{t=0}^\infty e^{-rt} \mu(t+y)\ell(t+y) \dd t} \dd a \dd y
+\intertext{Now $d(a+y) = \mu (a+y)\ell(a+y)$, so we can cancel the inner integral
 with the denominator:}
@@
 \int_{a=0}^\infty
-f(a)e^{-ra}\mu(a+y)l(a+y) \dd a \dd y
+f(a)e^{-ra}\mu(a+y)\ell(a+y) \dd a \dd y
 \intertext{Note also that $l(a) = \int _{=a}^\infty d(x) \dd x$ (present
 livings are future deaths), which brings us back to the chronological
-formulation:} =& \int_{a=0}^\infty e^{-ra}l(a)f(a)
+formulation:} =& \int_{a=0}^\infty e^{-ra}\ell(a)f(a)
 \dd a

### 2015-11-06 ThanoRepro/ThanoRepro.tex
minor edits to full manuscript. Removed empirical section.

@@
 A unisex model of population renewal for populations structured by
-thanatological age is presented in both continuous and discrete form. Various
-stable and transient properties populations are compared and related when viewed
+remaining years of life is presented in both continuous and discrete form.
+Various stable and transient properties populations are compared and related when viewed
 by chronological versus thanatological age. Results from the two age perspectives are conformable if the starting
@@
 while others can be a function of both to some degree
-\citep{riffe2015ttd}. This observation ought to motivate the formal demographic
-base to incorporate this aspect of time in into models. 
+\citep{riffe2015ttd,wolf2015disability}. This observation ought to motivate the
+formal demographic base to incorporate this aspect of time in into models. 
 %Temporal perspectives and rescalings of the lifecourse are abundant in
-% demography already. Such work has enriched the discipline by producing new insights, questions, and conundrums.
+% demography already. Such work has enriched the discipline by producing new
+% insights, questions, and conundrums.
 % Fertility and reproduction are plausibly a function of both age perspectives,
@@
 Specifying a model of thanatological renewal expands the toolbox available to approach complex
-temporal population structure. The next section of this article describes the lifetable method of transforming age-classified
+temporal population structure.
+
+The next section of this article describes the lifetable method of transforming age-classified
 data into thanatological age classes. I then derive the continuous version of
@@
 %if the chronological and thantological age patterns are independent.
-In this case, each birth cohort undergoes the same
+Under these assumptions, each birth cohort undergoes the same
 pattern of attrition. In a period setting, this means that we can distribute the
@@
 conditional deaths density function, $f(y|a)$. For chronological
-age zero, $f(y|0)$ is identical to $d(a)$. For ages higher than 0, simply
-condition on survival:\footnote{In the reliability literature, $f(y|a)$ is
+age zero, $f(y|0)$ is identical to $d(a)$, the lifetable deaths distribution
+(based on a unity radix).
+For ages higher than 0, simply condition on survival:\footnote{In the reliability literature, $f(y|a)$ is
 denoted by $ X - t | X > t $, where $X$ is the total lifespan and $t$ is the
@@
 Figure~\ref{fig:USrecomp} shows this recomposition for the same 2010
-US data.
+US data. The outer profile of Figure~\ref{fig:USrecomp} is $P(y)$, under our assumptions.
 
@@
 differences in French age structure, at times combining this projective method
-with historical stocks. \citet{brouard1989mouvements}, and later
-\citet{vaupel2009life}, showed that in stationary populations the
+with historical stocks. \citet{brouard1989mouvements}, and \citet{vaupel2009life}, showed that in stationary populations the
 marginal chronological and thanatological age structures are identical, and this
@@
 %thanatological age structure commensurate under a constant growth rate, $r$.
-These results will also follow from the following specification of a
+These results also follow from the following specification of a
 thanatological renewal model.
 
-It bears repeating that lifetable-derived thanatological age structure is
+Lifetable-derived thanatological age structure is
 probabilistic until after death, since it refers to the future according to an
@@
 nature, since a population's chronological age-structure is mostly the fruit of past
-fertility, but also migration, and to a lesser degree attrition. Any count
+fertility, but also to a lesser degree migration and attrition. Any count
 classified by chronological age can be reclassified in this way given a
-corresponding lifetable. To derive thanatological fertility
-rates under these assumptions, $\gamma(y)$, apply \eqref{eq:transform} to
+pertinent lifetable. To
+derive thanatological fertility rates under these assumptions, $\gamma(y)$, apply \eqref{eq:transform} to
 chronological age-classified birth counts, $B(a)$, to get $B(y)$ and again to
@@
 
-Thanatological fertility rates, $\gamma(y)$, are briefly shown because they
-are used in the renewal model here studied. Figure~\ref{fig:FySpaghetti}
+Figure~\ref{fig:FySpaghetti}
 represents the full variety in female thanatological period fertility rates that can be found for all years
@@
 1933 population-years of overlap between the HMD and HFD, including a wide
-variety of fertility and mortality combinations.} Figure~\ref{fig:FxSpaghetti} gives ASFR for the same
-populations and years as a more familiar reference, but note that both scales
-are different. The shape of thanatological fertility curves is speculative and
-due to the above-mentioned strong assumptions. Nonetheless it is best not to
+variety of fertility and mortality combinations.} I show $\gamma(y)$ calculated in this way because
+it enters into the renewal model here studied. Figure~\ref{fig:FxSpaghetti}
+gives ASFR for the same populations and years as a more familiar reference, but
+note that both the x and y axes are different. The shape of thanatological
+fertility curves is speculative and due to the above-mentioned strong assumptions. Nonetheless it is best not to
 include such rates into a model blindly, so we offer this schematic overview of
@@
 zero, due e.g., to maternal mortality, to at least 100.\footnote{For instance,
-Jeanne Calment lived for another 100 years after giving birth to her daughter}
+Jeanne Calment lived for another 100 years after giving birth to her daughter.}
 This spaghetti plot includes both fertility booms and busts, as well as some
 mortality crises (1918, WWII). Some patterns in the data (not
-shown) include that the left tail (a gauge of how orphan-prone a population is)
-has tended to fall over time. All populations have shown an increase in both the
+highlighted in the figure) include that the left tail (a gauge of how
+orphan-prone a population is) has tended to fall over time. All populations have shown an increase in both the
 mean and mode of thanatological age at childbearing over time. Several
-contemporary populations now show modes of over 60 years. Period thanatological
-total fertility rates track standard period total fertility rates closely, but tend to be somewhat higher (not shown).
+contemporary populations now show modes of over 60 years. Period thanatological total
+fertility rates track standard period total fertility rates closely, but tend to be somewhat higher (not shown).
 % both figures created in R/HFDspaghetti.R
@@
 \end{align}
-Now compare this to Lotka's chronological formulation
+Now compare this to Lotka's chronological formulation:
 \begin{align}
@@
 similar, since $\int _{y=0}^\infty d(a+y)\dd a = l(a)$. For intuition, notice
-that $l(a)$ is here decomposed into $d(a)$, and imagine a matrix of these,
-where one margin is chronological age and the other margin is thanatological age.
-For the chronological case \eqref{eq:lotka}, we multiply chronological age-specific fertility rates, $f(a)$, over the chronological age margin, and for the thanatological case we multiply
-$\gamma(y)$ over the thanatological age margin.
+that $l(a)$ is here decomposed into $d(a)$. Imagine a matrix of this
+decomposition, where one margin is chronological age and the other margin is
+thanatological age.
+For the chronological case \eqref{eq:lotka}, we multiply chronological
+age-specific fertility rates, $m(a)$, over the chronological age margin, and for the thanatological case we multiply $\gamma(y)$ over the thanatological age margin.
 
@@
 \eqref{eq:ergostep1} to \eqref{eq:ergostep2} (strong ergodicity) is assumed for
-the moment. The step has some intuition, due to
+the moment.\footnote{The step to exponential stability is proven
+for the discrete version of the model.} The step has some intuition, due to
 the much greater density of connections within the model; persons from nearly any
@@
 time must be much stronger than that for chronological age, at least in most
-cases.  The mechanisms at play unfold in the same way as those so intuitively
-described by \citet{arthur1982ergodic}, and said proof may apply here without
+cases. The mechanisms at play unfold in the same way
+as those so intuitively described by \citet{arthur1982ergodic}, and said proof may apply here without
 further modification. Later in this paper I note how strong ergodicity is
@@
 Literally, this is the mean of the remaining lifespans of new mothers in the
-stable population. The net reproduction rate, $R_0$ is related by, e.g.,
+stable population. The thanatological net reproduction rate, $R_0^\star$ is
+related by, e.g.,
 \begin{equation}
 \label{eq:R0fromTy}
-R_0 = e^{r T^\star}
+R_0^\star = e^{r T^\star}
 \end{equation}
@@
 \intertext{or, given the stable chronological age structure, $c(a)$, and a
-lifetable,} =& \int _{a=0}^\infty c(a)d(y|a) \dd a \label{eq:cyfromca}\\ \notag
+lifetable,} =& \int _{a=0}^\infty c(a)f(y|a) \dd a \label{eq:cyfromca}\\ \notag
 \\
@@
 matrix, $\textbf{Y}$, is square and of dimension $y \times y$, where $y$ is the number
-of thanatological age classifications on which the population is structured. 
+of equal-interval thanatological age classes into which the population is
+structured.
 The matrix contains elements for survival and elements for fertility. Unlike
-Leslie matrices, $\textbf{Y}$ is not sparse, but is instead populated primarily
+Leslie matrices, $\textbf{Y}$ is not sparse, but is populated primarily
 with positive entries greater than zero. In the
@@
  just born. It is simpler to imagine $d_a$ in this case as indexing the
- future lifespan distribution of newborns. Fertility in a thanatologically
- structured human population occurs in all but the very highest remaining years classes, which can be thought of as structural zeros for simplicity, as they only contain pre-menarchical females
- that will have very long lives.
+ future lifespan distribution of newborns \citep{riffe2015force}. Fertility in a
+ thanatologically structured human population occurs in all but the very highest
+ thanatological age classes, which can be thought of as structural zeros for
+ simplicity, as they only contain pre-menarchical females that will have very long lives.
+
  For our $6\times 6$ example, say that fertility is observed in
@@
  entering population (in the matrix columns). Each $\gamma_y$ is then
- distributed out over rows according to $d_a$.
+ distributed over rows according to $d_a$.
  The fertility entry in row $m$ and column $n$ of $\textbf{Y}$ will therefore be
@@
  accordingly.
+
  Further, infant mortality, $\gamma_y \cdot d_0$, located in the first row, must also be discounted, since part
@@
  triangle than in the upper. Of those infants who die in the first year of life,
- a proportion equal to $\lambda$ do not make it to December \nth{31} of the calendar year in which
- they were born.\footnote{$\lambda$ can be derived directly from death counts
+ a proportion equal to $\lambda$ do not survive to December \nth{31} of the
+ calendar year in which they were born.\footnote{$\lambda$ can be derived directly from death counts
  data classified by Lexis triangles. If the demographer
@@
 clockwise around the graph on the thick blue paths in order until finally being
-absorbed into death (the node labeled ``X'', not included in the projection matrix). The light gray paths represent fertility. Thse originate from any age class that has a fertility rate greater
-than zero, which for this toy matrix are all age classes except the \nth{5}
+absorbed into death (the node labeled ``X'', not included in the projection
+matrix). The light gray paths represent fertility. These originate from any age class that has a fertility rate greater than zero, which for this toy matrix are all age classes except the \nth{5}
 (highest). The y$^{th}$ age class is subject to the fertility rate $\gamma
@@
 the fertility destined for infant mortality equal to $\lambda$ leads straight to the absorbing state of death. 
-Fertility paths from the \nth{4} thanatological age class have been highlighted.
-These correspond to the fifth column of the example matrix $\textbf{Y}$, plus an additional path
-direct to death for the fraction of infant mortality prior to December \nth{31}
+Fertility paths from the \nth{4} thanatological age
+class are highlighted.
+These correspond to the fifth column of the example matrix $\textbf{Y}$, plus an additional path direct to death for the fraction of infant mortality prior to December \nth{31}
 in year $t$. Paths that loop back into the same age class simply indicate those
@@
 author to demonstrate this point. See also Appendix~\ref{app:B} for the
-corresponding proof from the continuous model.} The square of the, or the pace
-of convergence to stability (the ratio of the \nth{1} to the \nth{2} eigenvalues).
-\footnote{See \citet[p.86-87]{caswell2001matrix}.}
+corresponding proof from the continuous model.} The square of the ratio of
+the \nth{1} to the \nth{2} eigenvalues is the damping ratio, an indicator of the
+\textit{springiness} of the stable population structure, $c^\star(y)$, if perturbed from the
+state of stability.\footnote{See \citet[p.86-87]{caswell2001matrix}.} A higher
+damping ratio means that the population structure oscillates back to its stable
+state faster, i.e., that oscillations decrease in size more rapidly.
+Thanatological damping ratios are usually higher than chronological damping
+ratios in the data used elsewhere in this paper.
 
 \section*{Stable age structure}
+stable age structure is
 
@@
 
-\section*{Some preliminary empirical findings}
-The thanatological renewal equation and the discrete projection matrix can be
-put to work with data. In this case a few options are available for determining
-what kind of thanatological fertility rates to use. The two obvious choices are
-to derive rates from the stable population implied by the chronological renewal
-model, or to derive rates such as those from Figure~\ref{fig:FySpaghetti},
-based on a particular stock of population.
+%\section*{Some preliminary empirical findings}
+%The thanatological renewal equation and the discrete projection matrix can be
+%put to work with data. In this case a few options are available for determining
+%what kind of thanatological fertility rates to use. The two obvious choices are
+%to derive rates from the stable population implied by the chronological renewal
+%model, or to derive rates such as those from Figure~\ref{fig:FySpaghetti},
+%based on a particular stock of population.
 
-Of the 1834 population-years of data on hand, we optimize $r$ from both
-\eqref{eq:thanoren} and \eqref{eq:lotka}. In this sample, both versions of $r$
-are only plausibly equal in a single instance. Usually thanatological $r$ is
-greater than Lotka's $r$ (1373 cases). When Lotka's $r$ is positive (693 cases), thanatological $r$ is greater just over of 50\% of the time (356), but when Lotka's $r$ is negative,
-thanatological $r$ is the greater of the two around 90\% of the time (1017
-cases). These two approximations of $r$ are of opposite sign 138 times. We
-provide a comparison of the $r$ distributions in Figure~\ref{fig:rDist}. Mean
-locations for each distribution are indicated with vertical dashed lines; thano.
--0.0010; chrono. -0.0027. The distribution of thanatological $r$ is more compact
-than for chronological $r$, with the ratio of variances (thano./chrono.) of
-about 0.75. The two theoretical values of $r$ covary strongly, but
-thanatological $r$ is the less erratic of the two, and it usually paints a less dire picture when both are negative.
+%Of the 1834 population-years of data on hand, we optimize $r$ from both
+%\eqref{eq:thanoren} and \eqref{eq:lotka}. In this sample, both versions of $r$
+%are only plausibly equal in a single instance. Usually thanatological $r$ is
+%greater than Lotka's $r$ (1373 cases). When Lotka's $r$ is positive (693
+% cases), thanatological $r$ is greater just over of 50\% of the time (356), but when Lotka's $r$ is negative, thanatological $r$ is the greater of the two around 90\% of the time (1017
+%cases). These two approximations of $r$ are of opposite sign 138 times. We
+%provide a comparison of the $r$ distributions in Figure~\ref{fig:rDist}. Mean
+%locations for each distribution are indicated with vertical dashed lines;
+% thano.
+%-0.0010; chrono. -0.0027. The distribution of thanatological $r$ is more
+% compact than for chronological $r$, with the ratio of variances (thano./chrono.) of
+%about 0.75. The two theoretical values of $r$ covary strongly, but
+%thanatological $r$ is the less erratic of the two, and it usually paints a less
+% dire picture when both are negative.
 
-\begin{figure}[h!]
-	\caption{Distribution of $r$, chronological (Lotka) and thanatological*.}
-	\begin{center}
-		\label{fig:rDist}
-		\includegraphics[scale=.7]{Figures/rDist.pdf}
-	\end{center}
-	\begin{tiny}
-     * Data from HMD and HFD. Countries and years listed in
-     Figure~\ref{fig:Fxcompare}.
-	\end{tiny}
-\end{figure}
+%\begin{figure}[h!]
+%	\caption{Distribution of $r$, chronological (Lotka) and thanatological*.}
+%	\begin{center}
+%		\label{fig:rDist}
+%		\includegraphics[scale=.7]{Figures/rDist.pdf}
+%	\end{center}
+%	\begin{tiny}
+ %    * Data from HMD and HFD. Countries and years listed in
+  %   Figure~\ref{fig:Fxcompare}.
+	%\end{tiny}
+%\end{figure}
 
-The differences described here are not due primarily to rounding errors in the
-data, but rather to the way in which $\gamma(y)$ has been calculated. 
-Specifically, in this case we have transformed fertility rates from empirical data with a particular
-chronological age structure. If the chronological age structure is far from its
-stable distribution, then it may be hard to imagine that $\gamma(y)$ will
-remain fixed over time. When holding chronological-age classified fertility rates
-fixed, the assumption is just the opposite.
+%The differences described here are not due primarily to rounding errors in the
+%data, but rather to the way in which $\gamma(y)$ has been calculated. 
+%Specifically, in this case we have transformed fertility rates from empirical
+% data with a particular chronological age structure. If the chronological age structure is far from its
+%stable distribution, then it may be hard to imagine that $\gamma(y)$ will
+%remain fixed over time. When holding chronological-age classified fertility
+% rates fixed, the assumption is just the opposite.
 
-From the projection matrix, the ratio of the largest to the second largest
-eigenvalue, the damping ratio, is an indicator of the \textit{springiness} of
-the stable population structure, $c^\star(y)$ or $c(a)$, if perturbed from the
-state of stability.
-A higher damping ratio means that the population structure oscillates back to
-its stable state faster, i.e., that oscillations decrease in size more rapidly. The thanatological damping ratio was
-greater than the Leslie damping ratio for all 1834 population-years included in
-our empirical analysis. Leslie damping ratios ranged from 1.01258 to 1.0518,
-while thanatological damping ratios ranged from 1.0455 to 1.0868. Again, this
-was for the raw data, and I am not sure whether this consistent finding is a
-natural property of the model or an artifact of the unstable starting state of
-the data used to derive thanatological fertility rates. Intuitively, the
-thanatological smoothing process should be an order of magnitude stronger than
-the chronological smoothing process because each thanatological age class
-(node) -- except for the oldest-old -- is connected directly to each other
-thanatological age class in both directions. Again, I think contrived examples
-will be necessary to shed more light on this finding.
+%From the projection matrix, the ratio of the largest to the second largest
+%eigenvalue, the damping ratio, is an indicator of the \textit{springiness} of
+%the stable population structure, $c^\star(y)$ or $c(a)$, if perturbed from the
+%state of stability.
+%A higher damping ratio means that the population structure oscillates back to
+%its stable state faster, i.e., that oscillations decrease in size more rapidly.
+% The thanatological damping ratio was greater than the Leslie damping ratio for all 1834 population-years included in
+%our empirical analysis. Leslie damping ratios ranged from 1.01258 to 1.0518,
+%while thanatological damping ratios ranged from 1.0455 to 1.0868. Again, this
+%was for the raw data, and I am not sure whether this consistent finding is a
+%natural property of the model or an artifact of the unstable starting state of
+%the data used to derive thanatological fertility rates. Intuitively, the
+%thanatological smoothing process should be an order of magnitude stronger than
+%the chronological smoothing process because each thanatological age class
+%(node) -- except for the oldest-old -- is connected directly to each other
+%thanatological age class in both directions. Again, I think contrived examples
+%will be necessary to shed more light on this finding.
 

### 2015-11-06 ThanoRepro/ThanoRepro.tex
edits to age structure section

@@
 \begin{abstract}
-A unisex model of population renewal for populations structured by
-remaining years of life is presented in both continuous and discrete form.
-Various stable and transient properties populations are compared and related when viewed
-by chronological versus thanatological age. Results from the two age perspectives are conformable if the starting
-population is stable, but are otherwise divergent.\ackn{Research reported in this manuscript
+I present a unisex model of population renewal for populations structured by
+remaining years of life in both continuous and discrete form.
+Various stable properties of this model are compared and related when viewed
+by chronological versus thanatological age.\ackn{Research reported in this manuscript
 was supported by the U.S.
@@
 \end{abstract}
-
-*This is a work in progress. All findings are preliminary at this time, so
-please don't cite without permission of the author.
-\vspace{2em}
 
@@
 while others can be a function of both to some degree
-\citep{riffe2015ttd,wolf2015disability}. This observation ought to motivate the
-formal demographic base to incorporate this aspect of time in into models. 
+\citep{riffe2015ttd,wolf2015disability}. This observation motivates
+incorporating this aspect of time into formal demographic methods and models.
 %Temporal perspectives and rescalings of the lifecourse are abundant in
@@
 \section*{Stable age structure}
-stable age structure is
+The thanatological stable age structure, as given in equation~\eqref{eq:cy},
+displays some regular patterns worthy of mention. In Figuree~\ref{fig:str} the
+same underlying mortality pattern, based on US males and females from 2010 (HMD) is rescaled to different mortality levels and growth
+rates. Figure~\ref{tab:chronostr} shows stable age structures by chronological
+age, while Figure~\ref{tab:thanostr} shows stable age structure of the same
+population by thanatological age. Rows indicate growth rates, descending from
+positive to negative, while columns indicate the both-sex mean life expectancy,
+$e(0)$, to which 2010 US mortality is scaled. The rows and columns of
+Figures~\ref{tab:chronostr} and \ref{tab:thanostr} are organized in identical
+fashion, each element referring to the same intrinsic growth rate and average
+$e(0)$.
+
+The Brouard-Carey equality is visible by noting that the middle rows from the
+chronological and thanatological matrices are identical. In comparing any other
+rows, however, it is as if the thanatological matrix were flipped: a growingy is population in the chronological perspective looks like a shrinking
+population in the thanatological perspective, and vice versa. While the bottom row of
+Figure~\ref{tab:chronostr} looks like the top row of Figure~\ref{tab:thanostr},
+the profiles ar enot rigorously identical, and the symmetry is only noteworthy
+as rule of thumb. In thinking of the birth flow and the death flow, the
+approximate reflection of age structure over the $r$-axis does make sense: in a
+shrinking population new birth cohorts are smaller than old birth cohorts, which
+means that relatively fewer people are born, and relatively more are in older
+ages, closer to dying. In this case, the base of the chronological pyramid is
+relatively thin, while the base of the thanatological leaf is relatively thick.
+The same holds in reverse for growing populations.
+
+While this heuristic applies to stable populations, but it does not necessarily
+hold for real observed populations under changing mortality. Further, the observation is only
+valid for the period perspective in stability. In the cohort perspective, i.e.,
+within closed birth cohorts, the analogy to stationary populations is spot on--- That is, within birth cohorts
+the Brouard-Carey equality is exactly true when applied to the lifespan
+distribution.
 
@@
 \caption{Stable age structures by life expectancy, $e(0)$, and growth rate,
-$r$.*}
+$r$.*} \label{fig:str}
 \begin{subfigure}[b]{\textwidth}
@@
 models are almost perfectly opposite descriptions of the same process.
+
+This symmetry is visible in comparing the
+chronological and thanatological renewal equations~\eqref{eq:lotka2}
+and \eqref{eq:thanoren}, and it is also seen in the approximate symmetry between
+chronological and thanatological age structure profiles under intrinsic growth
+rates of equal magnitude but opposite sign. The thanatological projection matrix may be thought of as a
+sort of dual to the Leslie matrix--- a twin construct. The thanatological perspective on renewal I present is
+essentially a hitherto undescribed property of the Lotka-Leslie renewal model. Possible applications of
+this perspective on renewal are mentioned in \citet{riffe2015force}, the only
+difference in the present inquiry is that we now account for constant growth rates in the
+population. The primary contribution of this inquiry is to broaden our
+understanding of the renewal process.
 

### 2015-11-09 ThanoRepro/ThanoRepro.tex
restructuring lead-in

@@
 
-\title{Renewal and stability in populations structured by remaining
-years of life}
+\title{Renewal and stability in populations structured by remaining years of
+life}
 \author{Tim Riffe \\ Max Planck Institute for Demographic Research}
@@
 \begin{abstract}
-I present a unisex model of population renewal for populations structured by
-remaining years of life in both continuous and discrete form.
-Various stable properties of this model are compared and related when viewed
-by chronological versus thanatological age.\ackn{Research reported in this manuscript
-was supported by the U.S.
-National Institute On Aging of the National Institutes of Health under award numbers R01-AG011552 and R01-AG040245 and by the Spanish Ministry of Science and Innovation under award
-number ECO2013-48326-C2-1-P. The content is solely the responsibility of the
+Models of age structured population renewal are the core of formal demography.
+I derive a specification of the classic renewal equation that is structured by
+thanatological age rather than by chronological age. I give both continuous
+and discrete variants of the derived model, and relate these to the well-known
+age structured model. In stability, the two kinds of age
+structure are commensurable, and imply identical intrinsic growth rates.
+I demonstrate approximate symmetry between chronological and
+thanatological age structure in stability when subject to
+intrinsic growth rates equal magnitude and opposite sign. This paper offers a
+new perspective on the Lotka-Leslie renewal process.\ackn{Research reported in
+this manuscript was supported by the U.S.
+National Institute On Aging of the National Institutes of Health under award numbers R01-AG011552 and R01-AG040245. The content is solely the responsibility of the
 author and does not necessarily represent the official views of the funding
@@
 \section*{Introduction}
-In this paper, I explore some formal demographic consequences of a particular redefinition of age.
-Instead of counting age as the time passed since birth, consider age as the amount of time left until death\footnote{This quantity is referred to as \textit{residual} lifetime
-in the reliability literaure, or just time-to-death (TTD) elsewhere.}. The first
-kind of age is called chronological age, and the second kind of age is called
-thanatological age. Individuals in this case move in the same direction along
-imaginary life lines, but the reference point for thanatological age is at
-the end of the life line instead of at the beginning. For individuals, the timing of events in life is exactly identified using either of the two definitions of age, but
-aggregate patterns differ. Demography is built upon a small set of empirical
+
+Demography is built upon a small set of empirical
 regularities with respect to chronological age patterns. The models and methods
@@
 \citep{riffe2015ttd,wolf2015disability}. This observation motivates
-incorporating this aspect of time into formal demographic methods and models.
+incorporating this aspect time-to-death into formal demographic methods and
+models.
+
+In this paper, I explore some formal demographic consequences of a particular redefinition of age.
+Instead of counting age as the time passed since birth, consider age as the amount of time left until death.
+The first
+kind of age is called chronological age, and the second kind of age is called
+thanatological age.\footnote{This quantity is referred to as \textit{residual} lifetime
+in the reliability literaure, or just time-to-death (TTD) elsewhere.} Individuals in this case move in the same direction along
+imaginary life lines, but the reference point for thanatological age is at
+the end of the life line instead of at the beginning. For individuals, the timing of events in life is exactly identified using either of the two definitions of age, but
+aggregate patterns differ. 
 %Temporal perspectives and rescalings of the lifecourse are abundant in
@@
 
-I describe the Lotka-Euler-Leslie renewal model when structured by
-thanatological age. The renewal model relates
-fertility, mortality, and population structure in a parsimonious form, revealing the long term consequences of a particular set of conditions.
-All diverse lines of contemporary demographic research are relatable to the renewal model, and by extention to stable population theory.
+The renewal model relates
+fertility, mortality, and population structure in a parsimonious form, revealing
+the long term consequences of a particular set of conditions. All diverse lines
+of contemporary demographic research are relatable to the renewal model, and by
+extention to stable population theory. The purpose of this paper is to describe
+the Lotka-Leslie renewal model when structured by thanatological age.
 %Some might say that stable population
@@
 growth in the population.
-The chronological-age classified renewal model and the thanatological renewal model both describe the movement of population stocks over time, but from opposite vantage points.
-Specifying a model of thanatological renewal expands the toolbox available to approach complex
-temporal population structure.
+The chronological-age classified renewal model and the thanatological renewal
+model both describe the movement of population stocks over time, but from opposite vantage points.
+%Specifying a model of thanatological renewal expands the toolbox available to approach complex temporal population structure.
 
@@
 followed by the matrix version of the model, which is explained alongside the
-lifecycle graph.
+lifecycle graph. Finally, I explore patterns in stable age structure compared
+between the two models. I conclude that thanatological renewal offers
+a new perspective on the familiar Lotka-Leslie model.
 
-\section*{A lifetable approximation of thanatological age}
+\section*{Requisites of a thanatological renewal model}
+The Lotka-Leslie renewal model \citep{leslie1945use,sharpe1911problem} is
+situated in a unisex homogenous population framework. The model expressed in continous mathematics contains a
+component for survival, $\ell(a)$, a scaling factor for growth, $e^{-ra}$, and a
+component for fertility, $m(a)$:
+
+\begin{equation}
+\label{eq:lotka}
+1 = \int_{a=0}^\infty l(a)e^{-ra}m(a)\dd a \ep
+\end{equation}
+
+Each of these elements is structured by
+chronological age, $a$. The model can be approximated by observed data because
+chronological age is typically known. The survival function, $\ell(a)$, derives
+from a lifetable, usually based on age-specific event-exposure mortality rates,
+which requires knowing the age of decedentsand an estimate of the population
+exposed to risk in the same age classes. Female fertility rates, $m(a)$, are
+classified by age-of-mother
+
+\subsection*{A lifetable approximation of thanatological age}
 %Direct information on the timing of demographic transitions with respect to
@@
 % lifetime, which may underly those demographic transitions governed to some degree by volition.
+
 In general, to make a lifetable-based projective statement about the remaining
@@
 chronological age-classified population count, we can derive $P(y)$, a
-thanatological age-classified population count using\footnote{An
+thanatological age-classified population count using\footnote{A
 different-looking but identical definition is found in
@@
 nature, since a population's chronological age-structure is mostly the fruit of past
-fertility, but also to a lesser degree migration and attrition. Any count
+fertility, but also to a lesser degree migration and attrition.
+
+\subsection{Thanatological fertility rates.}
+Any count
 classified by chronological age can be reclassified in this way given a
@@
 \end{align}
-Now compare this to Lotka's chronological formulation:
-\begin{align}
-\label{eq:lotka}
-1 =& \int_{a=0}^\infty m(a)l(a)e^{-ra}\dd a \ec
-\intertext{which can be expressed as well as}
-  =& \int_{a=0}^\infty \int_{y=0}^\infty m(a)d(a+y)e^{-ra}\dd y \dd a \ec
+Now compare this to Lotka's chronological formulation in
+Equation~\eqref{eq:lotka},which can be expressed as well as:
+\begin{equation}
+ 1 = \int_{a=0}^\infty \int_{y=0}^\infty m(a)d(a+y)e^{-ra}\dd y \dd a \ec
   \label{eq:lotka2}
-\end{align}
+\end{equation}
 and note that \eqref{eq:lotka2} and \eqref{eq:thanoren} are really quite
@@
 and \ref{fig:USrecomp}. In the chronological age-structured model, new
-generations appear at the bottom of the pyramid, and move up one rung per year.
+generations appear at the bottom of the pyramid, and move up one class per year.
 All age-classes are subject to attrition, which is spread out over ages and not
@@
 c(a) &= \frac{\ell(a)e^{-ra}}{\int \ell(a) e^{-ra} \dd a} \\
-b(a) &= f(a)c(a)
+b(a) &= m(a)c(a)
 \end{align}Replacing $\gamma(y)$ with the full expanded
@@
 _{a=0}^\infty d(a+y)e^{-ra} \frac{\frac{\int_{t=0}^\infty
-f(t)\ell(t)e^{-rt}\mu(a+y)\frac{\ell(t+y)}{\ell(t)} \dd t}{\int \ell(t) e^{-rt} \dd
+m(t)\ell(t)e^{-rt}\mu(a+y)\frac{\ell(t+y)}{\ell(t)} \dd t}{\int \ell(t) e^{-rt} \dd
 t}}{\frac{\int_{t=0}^\infty \ell(t)e^{-rt} \mu(t+y)\frac{\ell(t+y)}{\ell(t)} \dd t}{\int
@@
 \int_{t=0}^\infty
-f(t)\ell(t)e^{-rt}\mu(t+y)\frac{\ell(t+y)}{\ell(t)} \dd t}{
+m(t)\ell(t)e^{-rt}\mu(t+y)\frac{\ell(t+y)}{\ell(t)} \dd t}{
 \int_{t=0}^\infty \ell(t)e^{-rt} \mu(t+y)\frac{\ell(t+y)}{\ell(t)} \dd t} \dd a \dd y
-\intertext{Some $l(a)$'s also cancel out:}
+\intertext{Some $\ell(a)$'s also cancel out:}
 =& \int _{y=0}^\infty \int
@@
 \int_{t=0}^\infty
-f(t)e^{-rt}\mu(t+y)\ell(t+y) \dd t}{
+m(t)e^{-rt}\mu(t+y)\ell(t+y) \dd t}{
 \int_{t=0}^\infty e^{-rt} \mu(t+y)\ell(t+y) \dd t} \dd a \dd y
@@
 \int_{a=0}^\infty
-f(a)e^{-ra}\mu(a+y)\ell(a+y) \dd a \dd y
+m(a)e^{-ra}\mu(a+y)\ell(a+y) \dd a \dd y
 \intertext{Note also that $l(a) = \int _{=a}^\infty d(x) \dd x$ (present
 livings are future deaths), which brings us back to the chronological
-formulation:} =& \int_{a=0}^\infty e^{-ra}\ell(a)f(a)
+formulation:} =& \int_{a=0}^\infty e^{-ra}\ell(a)m(a)
 \dd a

### 2015-11-09 ThanoRepro/ThanoRepro.tex
edits throughout, esp intro and conclusions

@@
 
+I couch the present inquiry in terms of aggregate renewal of human populations,
+but the renewal model is more generally a description of the birth-death
+process \citep{cox1962,feller1941integral}. As such, the thanatological
+renewal model may be applicable in myriad other areas, such as queueing,
+infectious disease processes, or aspects of renewable aggregates in general.
+Given the breadth of birth-death models and the speculative applicability of the thanatological renewal model, I am
+content for the present inquiry to establish the validity of the model. Perhaps
+there exists a birth-death process for which modelling as a death-birth process
+makes the most sense. In order to establish the \textit{soundness} of this model as a description of the human renewal process
+would require considerable empirical groundwork--- It would imply demonstrating
+that the thanatological age pattern of fertility is empirically regular, and
+that the time-to-death structure of the population is empirically regular.
+Such an endeavor would be misplaced, as the objective of this inquiry is not to
+improve the practice of empirical demography. Instead the objective is to deepen
+our understanding of the Lotka-Leslie renewal model. The chronological and
+thanatological renewal models imply one another in the state of stability. We
+therefore present this model as a property implied by the Lotka-Leslie renewal
+model, as two sides of the same coin. The result is however more general than
+that of literal birth and death in human populations.
+
 The next section of this article describes the lifetable method of transforming age-classified
@@
 from a lifetable, usually based on age-specific event-exposure mortality rates,
-which requires knowing the age of decedentsand an estimate of the population
-exposed to risk in the same age classes. Female fertility rates, $m(a)$, are
-classified by age-of-mother
+which require counts of decedents and an estimate of the
+population exposed to risk in the same chronological age classes.
+Chronological age in the growth factor, $e^{-ra}$ pertains to the passage of
+time more than it does to chronological age (these are synchronous for birth
+cohorts). The product of this growth factor and $\ell(a)$ gives a
+function proportional to the stable age structure, $c(a)$, and it is often
+easier to imagine these two elements together as one for this reason. Female
+fertility rates, $m(a)$, are based on birth counts classified by age-of-mother, and the same variety of exposure to risk.
+For the two essential rate componenents underlying the renewal equation~\eqref{eq:lotka}, classification by chronological age is essentially given--- it is deterministic in the sense that time passed since birth is typically known by individuals.
+The kinds of errors in such classification are mostly due to innumeracy, rounding, exaggeration, and
+similar kinds of biases, but even in such cases, chronological age is
+fundamentally knowable.
+
+Thanatological age is different in this respect because we do not know when we
+will die--- it is only knowable retrospectively or probabilistically. Under a set of simple assumptions, the same lifetable used to
+estimate $\ell(a)$ can be used to
+partition and recombine counts observed by chronological age into estimates
+classified by thanatological age. Our objective is then to estimate an age
+structure component, similar to $\ell(a)e^{-ra}$, and a fertility component
+simialr to $m(a)$, but each for the thanatological case. Taken together, these
+componenents will yield a thanatological renewal model.
 
@@
 \label{eq:vaupel1}
-f(y | a) = \mu(a+y)\frac{\ell(a+y)}{\ell(a)} \ep
-\end{equation}
+f(y | a) = \mu(a+y)\frac{\ell(a+y)}{\ell(a)} \ec
+\end{equation}
+where $\mu(a)$ is the force of mortality.
 
@@
 $a$. The $f(y|a)$- weighted average of $y$ for a given $a$ returns the familiar
-remaining life expectancy column of the lifetable, $e(a)$. If $P(a)$ is a
+remaining life expectancy column of the lifetable, $e(a)$. Since $1 = \int_0^\infty f(y|a) \dd y$ for each $a$,
+this density is useful for decomposing data classified by chronological age,
+such as the population pyramid, into thanatological age classes. If $P(a)$ is a
 chronological age-classified population count, we can derive $P(y)$, a
@@
 
-Since $1 = \int_0^\infty f(y|a) \dd y$ for each $a$,
-this density is useful for decomposing data classified by chronological age,
-such as the population pyramid, into thanatological age classes.
 Figure~\ref{fig:USdecomp} shows this exercise for 2010 US data \citep{HMD},
@@
 %thanatological age structure commensurate under a constant growth rate, $r$.
-These results also follow from the following specification of a
-thanatological renewal model.
+%These results also follow from the following specification of a
+%thanatological renewal model.
 
 Lifetable-derived thanatological age structure is
-probabilistic until after death, since it refers to the future according to an
-assumed mortality pattern; in this case a period mortality pattern is used.
+probabilistic until after death in several senses. In this case, note that
+counts from a given age class are distributed over a range of higher ages
+according to a given mortality rate schedule; in this
+case a period mortality pattern is used.
 As such, Figure~\ref{fig:USrecomp} is \textit{projective} in nature, a
@@
 particular moment, whereas Figure~\ref{fig:USdecomp} is \textit{reflective} in
-nature, since a population's chronological age-structure is mostly the fruit of past
-fertility, but also to a lesser degree migration and attrition.
+nature, since a population's chronological age-structure is the fruit of past
+demographic forces.
 
@@
 Any count
-classified by chronological age can be reclassified in this way given a
-pertinent lifetable. To
+classified by chronological age can be reclassified as in
+equation~\eqref{eq:transform} given a pertinent lifetable. To
 derive thanatological fertility rates under these assumptions, $\gamma(y)$, apply \eqref{eq:transform} to
 chronological age-classified birth counts, $B(a)$, to get $B(y)$ and again to
-exposure-to-risk, $E(a)$, to calcualte $E(y)$ and then divide in the usual way:
+exposure-to-risk, $E(a)$, to get $E(y)$ and then divide in the usual way:
 
@@
 thanatological fertility rates derived from a stable population imply a certain
-symmetry with stable populations from the chronological model. First I
-define the basic model.
+symmetry with stable populations from the chronological model.
 
@@
 \section*{Discussion}
-The thanatological renewal model is valid, but this perspective may take some
-time to gestate before accepting that it is also a sound description of how
-populations renew. What should one imagine under the model of thanatological
-population renewal? A useful mnemonic bases itself on Figures~\ref{fig:USdecomp}
-and \ref{fig:USrecomp}. In the chronological age-structured model, new
-generations appear at the bottom of the pyramid, and move up one class per year.
+The thanatological renewal model is valid, but I do not expect the reader to
+accept that it is also a sound description of the fundamental forces of
+population renewal. Specifically, I expect that the fertility rates are best
+described in the margin by chronological age and not by thanatological age. The
+main contribution of this inquiry is to demonstrate a transformation of the
+Lotka-Leslie renewal model, and to highlight a parallel aspect of renewal that
+occurs irrespective of its empirical regularity. That is, it is the case
+that in the moment of birth, both progenitor and baby have a certain number of
+remaining years of life, and that the range of potential values for both is
+great. The thanatological renewal
+model therefore is an abstraction of a real process, and not merely an adornment of the chronological renewal model.
+
+What should one imagine under the model of thanatological population renewal?
+A useful mnemonic bases itself on Figures~\ref{fig:USdecomp} and \ref{fig:USrecomp}.
+In the chronological age-structured model, new generations appear at the bottom of the pyramid, and move up one class per year.
 All age-classes are subject to attrition, which is spread out over ages and not
@@
 the locations of increment and decrement, and the direction of movement (when
-so visualized) are all switched. The chronological and thanatological renewal
+so visualized) are all swapped. The chronological and thanatological renewal
 models are almost perfectly opposite descriptions of the same process.
@@
 rates of equal magnitude but opposite sign. The thanatological projection matrix may be thought of as a
-sort of dual to the Leslie matrix--- a twin construct. The thanatological perspective on renewal I present is
-essentially a hitherto undescribed property of the Lotka-Leslie renewal model. Possible applications of
+sort of dual to the Leslie matrix--- a twin construct.
+The thanatological perspective on renewal I present is a hitherto undescribed property of the Lotka-Leslie renewal model.
+Possible applications of
 this perspective on renewal are mentioned in \citet{riffe2015force}, the only
 difference in the present inquiry is that we now account for constant growth rates in the
-population. The primary contribution of this inquiry is to broaden our
-understanding of the renewal process.
+population. The present treatment is broader, and potentially applicable to a
+wider range of birth-death processes. In the first place, we offer a deeper understanding
+of the Lotka-Leslie renewal models.
 
@@
 \end{align}
-This finding essentially proves that in the stable population there is at least
+This proof establishes that in the stable population there is at least
 one thanatological fertility rate schedule that will satisfy the constraints of
@@
 but this would entail less precision than the above (unless one goes beyond,
-say, five cumulants), and this has not yet been worked out at this time.
+say, five cumulants), and this is not explored.
 

### 2015-11-10 ThanoRepro/ThanoRepro.tex
version submitted to DR

@@
-\documentclass{article}
+\documentclass[12pt,oneside,a4paper,leqno]{article}
 \usepackage{appendix}
@@
 \usepackage{tikz}
-%\usetikzlibrary{arrows}
-%\usepackage[active,tightpage]{preview}
-%\PreviewEnvironment{tikzpicture}
+
 %\setlength\PreviewBorder{5pt}
@@
 \usepackage{setspace}
+\doublespacing
+
 \newcommand\ackn[1]{%
@@
 life}
-\author{Tim Riffe \\ Max Planck Institute for Demographic Research}
+\author{(author redacted)}
 \maketitle
 
-\begin{abstract}
-Models of age structured population renewal are the core of formal demography.
+%\ackn{Research
+%reported in this manuscript was supported by the U.S.
+%National Institute On Aging of the National Institutes of Health under award
+% numbers R01-AG011552 and R01-AG040245. The content is solely the responsibility of the author and does not necessarily represent the official views of the funding
+%agency.}
+
+\section*{Structured abstract}
+
+\subsection*{Background}
+The Lotka-Leslie renewal model is the core of formal demography. This model is
+structured by chronological age, and it does not account for thanatological age.
+
+\subsection*{Objective}
 I derive a specification of the classic renewal equation that is structured by
 thanatological age rather than by chronological age. I give both continuous
-and discrete variants of the derived model, and relate these to the well-known
-age structured model. In stability, the two kinds of age
-structure are commensurable, and imply identical intrinsic growth rates.
+and discrete variants of the derived model, and relate these to the Lotka-Leslie
+renewal model.
+
+\subsection*{Results}
+In stability, the thanatological and chronological renewal models are
+commensurable, implying identical intrinsic growth rates.
 I demonstrate approximate symmetry between chronological and
 thanatological age structure in stability when subject to
-intrinsic growth rates equal magnitude and opposite sign. This paper offers a
-new perspective on the Lotka-Leslie renewal process.\ackn{Research reported in
-this manuscript was supported by the U.S.
-National Institute On Aging of the National Institutes of Health under award numbers R01-AG011552 and R01-AG040245. The content is solely the responsibility of the
-author and does not necessarily represent the official views of the funding
-agencies.}
-\end{abstract}
-
-\onehalfspacing
+intrinsic growth rates equal magnitude and opposite sign.
+
+\subsection*{Conclusions}
+Birth-death renewal processes can be expressed as death-birth processes, and
+vice versa.
+
+\subsection*{Contribution}
+The thanatological renewal model offers a new perspective on population renewal,
+and it is valid more generally as an aspect of birth-death processes.
+\vspace{2cm}
+
 \section*{Introduction}
 
-Demography is built upon a small set of empirical
+Contemporary demography is built upon a small set of empirical
 regularities with respect to chronological age patterns. The models and methods

### 2015-11-10 ThanoRepro/ThanoReproMPIDR.tex
MPIDR version committed. This has been submitted to post as a working paper while under review at DR

--- 2015-11-10 ThanoRepro/ThanoRepro.tex
@@
 life}
-\author{(author redacted)}
+\author{Tim Riffe \\ Max Planck Institute for Demographic Research}
 \maketitle
 
-%\ackn{Research
-%reported in this manuscript was supported by the U.S.
-%National Institute On Aging of the National Institutes of Health under award
-% numbers R01-AG011552 and R01-AG040245. The content is solely the responsibility of the author and does not necessarily represent the official views of the funding
-%agency.}
-
+\vspace{8cm}
+\ackn{Research
+reported in this manuscript was supported by the U.S.
+National Institute On Aging of the National Institutes of Health under award
+ numbers R01-AG011552 and R01-AG040245. The content is solely the responsibility of the author and does not necessarily represent the official views of the funding
+agency.}
+
+\pagebreak
 \section*{Structured abstract}

### 2016-01-22 ThanoRepro/ReplyReviewers.tex
starting to make changes for first round of review. started draft of response to reviewers

\documentclass{article}

\usepackage{amsmath}
\usepackage{natbib}
\bibpunct{(}{)}{,}{a}{}{;} 
\newcommand{\dd}{\; \mathrm{d}}
\usepackage{url}

\begin{document}
\title{Reply to Reviewers of submission 3043: ``Renewal and stability in populations structured by remaining years of
life''}
\author{Author Redacted}
\maketitle

I thank reviewers A and B for taking the time to engage with and comment on this
submission, and the assistant editor for his suggestions and guidance, and to
all involved for a swift first round of review. I've made an attempt to increase
appeal by reducing unneeded sentences and statements, and by changing
language, for instance swapping out thanatological age for years left in most
cases. Reviewers comments are then treated individually.

However, I first wish to politely state that I think DR is a good home for
purely theoretical or mathematical investigations because (in general and in no
order) a) DR has no space limitations, b) readers of DR come in almost
exclusively via content-based internet searches rather than periodical
readership (ergo there is no need to profile readership), c) the alternatives
(e.g., Theoretical Population Biology and Mathematical Population Studies) are not open access and so I have ethical objections to
them, and d) DR already sponsors a formal relationships series (ergo formal demography is demonstrably in-universe). I
should like to think that DR covers the full breadth of researchers that
demographers do. This article is standard fare for formal demography, and formal
demography is the foundation of the discipline. Browsing through the titles of
recently publishes articles in DR, I'm lead to believe that appealing to a broad
audience is not itself a criterion for publication, and I do not believe that it
should be either.

That said, I've followed advice to make the article more palatable and
approachable, and these offset somewhat the space saved by following through on
reviewer A's wish for the article to be more compact.

Comments in order:

Editor:

\end{document}

### 2016-01-22 ThanoRepro/ThanoRepro.tex
starting to make changes for first round of review. started draft of response to reviewers

@@
 \newcommand{\ep}{\quad\quad\text{.}}
-
+\defcitealias{HMD}{HMD}
 \usepackage[top=2in, bottom=1.5in, left=1in, right=1in]{geometry}
@@
 The Lotka-Leslie renewal model is the core of formal demography. This model is
-structured by chronological age, and it does not account for thanatological age.
+structured by age, and it does not account for time until death.
 
@@
 I derive a specification of the classic renewal equation that is structured by
-thanatological age rather than by chronological age. I give both continuous
+years left rather than by years lived. I give both continuous
 and discrete variants of the derived model, and relate these to the Lotka-Leslie
@@
 \subsection*{Results}
-In stability, the thanatological and chronological renewal models are
+In stability, the years lived and left renewal models are
 commensurable, implying identical intrinsic growth rates.
-I demonstrate approximate symmetry between chronological and
-thanatological age structure in stability when subject to
-intrinsic growth rates equal magnitude and opposite sign.
+I demonstrate approximate symmetry between years lived and left age structure in
+stability when subject to intrinsic growth rates equal magnitude and opposite sign.
 
@@
 \subsection*{Contribution}
-The thanatological renewal model offers a new perspective on population renewal,
+The years left renewal model offers a new perspective on population renewal,
 and it is valid more generally as an aspect of birth-death processes.
@@
 \section*{Introduction}
-
-Contemporary demography is built upon a small set of empirical
-regularities with respect to chronological age patterns. The models and methods
+Contemporary demography is built upon a small set of empirically regular age
+patterns. The models and methods
 of demography are in varying degrees prefaced on such regularity. There is evidence that many demographic phenomena are best described in the
@@
 while others can be a function of both to some degree
-\citep{riffe2015ttd,wolf2015disability}. This observation motivates
-incorporating this aspect time-to-death into formal demographic methods and
-models.
-
-In this paper, I explore some formal demographic consequences of a particular redefinition of age.
-Instead of counting age as the time passed since birth, consider age as the amount of time left until death.
-The first
-kind of age is called chronological age, and the second kind of age is called
-thanatological age.\footnote{This quantity is referred to as \textit{residual} lifetime
-in the reliability literaure, or just time-to-death (TTD) elsewhere.} Individuals in this case move in the same direction along
-imaginary life lines, but the reference point for thanatological age is at
-the end of the life line instead of at the beginning. For individuals, the timing of events in life is exactly identified using either of the two definitions of age, but
-aggregate patterns differ. 
+\citep{riffe2015ttd,wolf2015disability}. Further, aggregate indices
+motivated by measures of remaining lifetime have also become widely-used
+indicators of ageing \citep{sanderson2007new}. These observations motivate
+incorporating time-to-death into formal demographic methods and models.
+
+In this paper, I explore some formal demographic consequences of a particular
+redefinition of age. Instead of counting age as the time passed since birth,
+consider the amount of time left until death.\footnote{Elsewhere this quantity
+is referred to as \textit{residual} lifetime, time-to-death (TTD), prospective
+age \citep{sanderson2007new}, or thanatological age \citep{riffe2015force}.}
+Individuals in this case move in the same direction along imaginary life lines,
+but the reference point for measurement is switched from birth to death. For
+individuals, the timing of events in life is exactly identified using either of
+years lived or left, but aggregate patterns differ.
 %Temporal perspectives and rescalings of the lifecourse are abundant in
@@
 fertility, mortality, and population structure in a parsimonious form, revealing
-the long term consequences of a particular set of conditions. All diverse lines
+the long term consequences of a set of vital rates. All lines
 of contemporary demographic research are relatable to the renewal model, and by
 extention to stable population theory. The purpose of this paper is to describe
-the Lotka-Leslie renewal model when structured by thanatological age.
+the Lotka-Leslie renewal model when structured by years of life left.
 %Some might say that stable population
@@
 % remaining lifetime as a particular class of time-to-transition, one may generalize the lifetable methods presented here into new area applications, such as education, partnering, the economic lifecycle, or migration.
-To present the renewal model based on thanatological age is not necessarily a
-suggestion for a better way to model or project population. Rather, the
-thanatological renewal model represents an alternative and coherent way to
+%To present the renewal model based on thanatological age is not necessarily a
+%suggestion for a better way to model or project population. Rather, 
+The years left renewal model represents an alternative and coherent way to
 describe the process of population renewal by relating renewal to a death flow
@@
 growth in the population.
-The chronological-age classified renewal model and the thanatological renewal
+The years lived renewal model and the years left renewal
 model both describe the movement of population stocks over time, but from opposite vantage points.
@@
 but the renewal model is more generally a description of the birth-death
-process \citep{cox1962,feller1941integral}. As such, the thanatological
-renewal model may be applicable in myriad other areas, such as queueing,
-infectious disease processes, or aspects of renewable aggregates in general.
-Given the breadth of birth-death models and the speculative applicability of the thanatological renewal model, I am
-content for the present inquiry to establish the validity of the model. Perhaps
-there exists a birth-death process for which modelling as a death-birth process
-makes the most sense. In order to establish the \textit{soundness} of this model as a description of the human renewal process
-would require considerable empirical groundwork--- It would imply demonstrating
-that the thanatological age pattern of fertility is empirically regular, and
-that the time-to-death structure of the population is empirically regular.
-Such an endeavor would be misplaced, as the objective of this inquiry is not to
-improve the practice of empirical demography. Instead the objective is to deepen
-our understanding of the Lotka-Leslie renewal model. The chronological and
-thanatological renewal models imply one another in the state of stability. We
-therefore present this model as a property implied by the Lotka-Leslie renewal
-model, as two sides of the same coin. The result is however more general than
-that of literal birth and death in human populations.
-
-The next section of this article describes the lifetable method of transforming age-classified
-data into thanatological age classes. I then derive the continuous version of
-the thanatological renewal model, and some other stable properties. This is
+process \citep{cox1962,feller1941integral}, and so may be valid for other kinds
+of renewable aggregates.
+The present inquiry establishes the validity of the model, but does not argue
+for its empirical practicality or soundness.
+%Perhaps there exists a birth-death process for which modelling as a death-birth
+% process makes the most sense. In order to establish the \textit{soundness} of this model as a description of the human renewal process
+%would require considerable empirical groundwork--- It would imply demonstrating
+%that the thanatological age pattern of fertility is empirically regular, and
+%that the time-to-death structure of the population is empirically regular.
+%Such an endeavor would be misplaced, as the objective of this inquiry is not to
+%improve the practice of empirical demography. 
+The years lived and left renewal models imply one another in the state of stability. We
+therefore present the years left model as a property implied by the
+standard Lotka-Leslie renewal model, as two sides of the same coin. The objective in this paper is to deepen
+our understanding of the Lotka-Leslie renewal model, and to offer an extention
+of stable population theory.
+
+The next section of this article reviews the lifetable method of transforming
+age-classified data into remaining years classes. I then derive the continuous version of
+the years left renewal model, and some other stable properties. This is
 followed by the matrix version of the model, which is explained alongside the
 lifecycle graph. Finally, I explore patterns in stable age structure compared
-between the two models. I conclude that thanatological renewal offers
+between the two models. I conclude that years left renewal offers
 a new perspective on the familiar Lotka-Leslie model.
 
-\section*{Requisites of a thanatological renewal model}
+\section*{Requisites of a years left renewal model}
 The Lotka-Leslie renewal model \citep{leslie1945use,sharpe1911problem} is
-situated in a unisex homogenous population framework. The model expressed in continous mathematics contains a
-component for survival, $\ell(a)$, a scaling factor for growth, $e^{-ra}$, and a
-component for fertility, $m(a)$:
+situated in a closed, unisex, homogenous population framework. With
+parameters equated, the continuous form of the model contains a component for
+survival, $\ell(a)$, a scaling factor for growth, $e^{-ra}$, and a component for fertility, $m(a)$:
 
@@
 Each of these elements is structured by
-chronological age, $a$. The model can be approximated by observed data because
-chronological age is typically known. The survival function, $\ell(a)$, derives
-from a lifetable, usually based on age-specific event-exposure mortality rates,
-which require counts of decedents and an estimate of the
-population exposed to risk in the same chronological age classes.
-Chronological age in the growth factor, $e^{-ra}$ pertains to the passage of
+age, $a$. The model can be approximated by observed data because
+age is typically known.% The survival function, $\ell(a)$, derives
+%from a lifetable, usually based on age-specific event-exposure mortality rates,
+%which require counts of decedents and an estimate of the
+%population exposed to risk in the same chronological age classes.
+Age in the growth factor, $e^{-ra}$, pertains to the passage of
 time more than it does to chronological age (these are synchronous for birth
@@
 function proportional to the stable age structure, $c(a)$, and it is often
-easier to imagine these two elements together as one for this reason. Female
-fertility rates, $m(a)$, are based on birth counts classified by age-of-mother, and the same variety of exposure to risk.
-For the two essential rate componenents underlying the renewal equation~\eqref{eq:lotka}, classification by chronological age is essentially given--- it is deterministic in the sense that time passed since birth is typically known by individuals.
-The kinds of errors in such classification are mostly due to innumeracy, rounding, exaggeration, and
-similar kinds of biases, but even in such cases, chronological age is
-fundamentally knowable.
-
-Thanatological age is different in this respect because we do not know when we
+intuitive to imagine these two elements together as one for this reason. Female
+fertility rates, $m(a)$, are based on daughters born to mothers in this limited
+single-sex model. This model lends itself to commonly collected data because age
+is typically known, and can be asked.
+% For the two essential rate componenents underlying the renewal
+ % equation~\eqref{eq:lotka}, classification by chronological age is essentially given--- it is deterministic in the sense that time passed since birth is typically known by individuals.
+The kinds of errors in such classification are mostly due to innumeracy, rounding, exaggeration, and similar kinds of biases, but even in such cases, chronological age is fundamentally knowable.
+
+Remaining lifetime is different in this respect because we do not know when we
 will die--- it is only knowable retrospectively or probabilistically. Under a set of simple assumptions, the same lifetable used to
 estimate $\ell(a)$ can be used to
-partition and recombine counts observed by chronological age into estimates
-classified by thanatological age. Our objective is then to estimate an age
+partition and recombine counts observed by age into estimates
+classified by years of life left. Our objective is then to estimate a population
 structure component, similar to $\ell(a)e^{-ra}$, and a fertility component
-simialr to $m(a)$, but each for the thanatological case. Taken together, these
-componenents will yield a thanatological renewal model.
-
-\subsection*{A lifetable approximation of thanatological age}
+similar to $m(a)$, but each for the years left case. Taken together, these
+componenents yield the years left renewal model.
+
+\subsection*{A lifetable approximation of life left}
 %Direct information on the timing of demographic transitions with respect to
@@
 
-In general, to make a lifetable-based projective statement about the remaining
-time distribution of a current stock of population, it is best to to account for population subgroups
-with different mortality and to account for uncertainty about future changes in
-mortality.
-The present inquiry approximates thanatological age using lifetable-projected remaining lifetime,
-which is intended to approximate remaining lifetime under the strong
+%In general, to make a lifetable-based projective statement about the remaining
+%time distribution of a current stock of population, it is best to to account
+% for population subgroups with different mortality and to account for uncertainty about future changes in
+%mortality.
+The present inquiry approximates years left in the aggregate using
+lifetable-projected remaining lifetime, which is intended to approximate remaining lifetime under the strong
 assumptions of homogenous populations und fixed vital rates. 
@@
 Under these assumptions, each birth cohort undergoes the same
-pattern of attrition. In a period setting, this means that we can distribute the
-remaining lifespans of the population in each age class using the same
-lifetable.
+pattern of attrition. %In a period setting, this means that we can distribute
+% the remaining lifespans of the population in each age class using the same
+%lifetable.
 %This is the case we explore, and so the following formulas assume a single lifetable pattern.
 
-The basic transformation to thanatological age proceeds by defining a
-conditional deaths density function, $f(y|a)$. For chronological
-age zero, $f(y|0)$ is identical to $d(a)$, the lifetable deaths distribution
-(based on a unity radix).
-For ages higher than 0, simply condition on survival:\footnote{In the reliability literature, $f(y|a)$ is
-denoted by $ X - t | X > t $, where $X$ is the total lifespan and $t$ is the
-current chronological age attained.}
+The basic transformation to years left starts with the well-known remaining
+years density function, $f(y|a)$.
+%For age zero, $f(y|0)$ is identical to $d(a)$, the lifetable deaths
+% distribution (based on a unity radix).
+%For ages higher than 0, simply condition on survival:
+%\footnote{In the
+% reliability literature, $f(y|a)$ is denoted by $ X - t | X > t $, where $X$ is the total lifespan and $t$ is the
+%current chronological age attained.}
 \begin{equation}
@@
 where $\mu(a)$ is the force of mortality.
-
-A convenient discretization of \eqref{eq:vaupel1} is to simply work with the
-$d_a$ column of the lifetable (here with single ages):
-\begin{equation}
-\label{eq:dyadiscret}
-f_{y, a} =  \frac{d_{a+y}+d_{a+y+1}}{L_a + L_{a+1}} \ec
-\end{equation}
-where $L_a$ is lifetable exposure.\footnote{Equation~\eqref{eq:dyadiscret}
-will incur some error depending on the method used to derive $L_a$. To remove
-this error, rescale over the $a$ margin so that each distribution conditioned on
-$a$ sums to 1. Working with $\ell_a$ rather than $L_a$ eliminates error and
-yields similar results, but is not necessarily better.} One produces in this way a probability density function over future death times for each chronological age,
-$a$. The $f(y|a)$- weighted average of $y$ for a given $a$ returns the familiar
-remaining life expectancy column of the lifetable, $e(a)$. Since $1 = \int_0^\infty f(y|a) \dd y$ for each $a$,
-this density is useful for decomposing data classified by chronological age,
-such as the population pyramid, into thanatological age classes. If $P(a)$ is a
-chronological age-classified population count, we can derive $P(y)$, a
-thanatological age-classified population count using\footnote{A
-different-looking but identical definition is found in
-\citet{brouard1989mouvements}.}
+%A convenient discretization of \eqref{eq:vaupel1} is to simply work with the
+%$d_a$ column of the lifetable (here with single ages):
+%\begin{equation}
+%\label{eq:dyadiscret}
+%f_{y, a} =  \frac{d_{a+y}+d_{a+y+1}}{L_a + L_{a+1}} \ec
+%\end{equation}
+%where $L_a$ is lifetable exposure.\footnote{Equation~\eqref{eq:dyadiscret}
+%will incur some error depending on the method used to derive $L_a$. To remove
+%this error, rescale over the $a$ margin so that each distribution conditioned
+% on $a$ sums to 1. Working with $\ell_a$ rather than $L_a$ eliminates error and
+%yields similar results, but is not necessarily better.} One produces in this
+% way a probability density function over future death times for each chronological age, $a$. 
+The $f(y|a)$- weighted average of $y$ for a given $a$ returns the remaining life
+expectancy column of the lifetable, $e(a)$. $f(y|a)$ is also useful for
+decomposing age-classified data, such as the
+population pyramid, into remaining years classes. If $P(a)$ is an
+age-classified population count, we can derive the years left version $P(y)$,
+using\footnote{An identical definition is found in
+\citet{brouard1989mouvements} and illustrated in \citet{brouard1986structure}.}
 
@@
 
-Figure~\ref{fig:USdecomp} shows this exercise for 2010 US data \citep{HMD},
-assuming the 2010 period mortality schedule remains constant. When looking from the chronological age perspective, this decomposition reveals thanatological
-heterogeneity, but when viewed from the thanatological perspective
-(Figure~\ref{fig:USdecomp}) it reveals chronological age heterogeneity and the
-overall thanatological age profile, the future death flow of the current
-population.
-Figure~\ref{fig:USrecomp} shows this recomposition for the same 2010
-US data. The outer profile of Figure~\ref{fig:USrecomp} is $P(y)$, under our assumptions.
+Figure~\ref{fig:pop} shows this exercise for 2010 US data \citep{HMD},
+assuming homogeneity and that the 2010 period mortality schedule is held fixed.
+When looking from the age perspective (Figure~\ref{fig:USdecomp}), this
+decomposition reveals projected years left heterogeneity within a stock of known age, but when viewed from the
+years left perspective (Figure~\ref{fig:USrecomp}) it reveals age heterogeneity
+and the projected overall years left profile--- the future death flow of the
+current population, $P(y)$ under our assumptions.
 
@@
 	\caption{2010 US population structure}
+	\label{fig:pop}
 	\begin{center}
 	\begin{subfigure}{.45\textwidth}
-		\caption{Chronological age-structure (years lived) decomposed by
-		thanatological age (years left).}
+		\caption{Age structure (years lived) decomposed by years left.}
 		\label{fig:USdecomp}
@@
 	\begin{subfigure}{.45\textwidth}
-		\caption{Thanatological age-structure (years left) decomposed by chronological
-		age (years lived).}
+		\caption{Years left structure decomposed by years lived.}
 		\label{fig:USrecomp}
@@
 chronological age-structured Leslie matrix, the thanatological projection
-matrix, $\textbf{Y}$, is square and of dimension $y \times y$, where $y$ is the number
-of equal-interval thanatological age classes into which the population is
+matrix, $\textbf{Y}$, is square and of dimension $\omega \times \omega$, where $\omega$ is
+the number of equal-interval thanatological age classes into which the population is
 structured.
@@
 Leslie matrices, $\textbf{Y}$ is not sparse, but is populated primarily
-with positive entries greater than zero. In the
+with non-zero positive entries. In the
 following, I illustrate using a 6$\times$6 matrix.
@@
 subdiagonal, survival elements are located in the superdiagonal. All survival
-values are 1, since there is no decrement until after living through age 0. The survival component of $\textbf{Y}$ is organized as in
-Matrix~\ref{matrix:Ysurv}. To imagine a lifecycle graph, each age decrements by
-one through each lower age in successive order until being absorbed by death
+values are 1, since there is no decrement until after living through age 0. The
+survival component of $\textbf{Y}$, \textbf{S}, is organized as in
+Matrix~\ref{matrix:Ysurv}. To imagine a lifecycle graph, each age decrements by one through each lower age in successive order until being absorbed by death
 after completing thanatological age zero.
@@
 \centering
-\caption{Survival component of unisex thanatological projection matrix,
-$\textbf{Y}$}
+\caption{Survival component,  \textbf{S}, of unisex thanatological projection
+matrix.}
 \label{matrix:Ysurv}
@@
  thanatologically structured human population occurs in all but the very highest
- thanatological age classes, which can be thought of as structural zeros for
- simplicity, as they only contain pre-menarchical females that will have very long lives.
+ thanatological age classes, which can be thought of as structural zeros, as they only contain pre-menarchical females that will have very long lives.
 
@@
  methods may be used to assign a reasonable proportion, such as 0.9 for
- contemporary low mortality populations.} The fertility component of $\textbf{Y}$ is then composed as in Matrix~\ref{matrix:Yfert}.
+ contemporary low mortality populations.} The fertility component of
+ $\textbf{Y}$, \textbf{$\Gamma$}, is then composed as in
+ Matrix~\ref{matrix:Yfert}.
 
@@
 \centering
-\caption{Fertility component of unisex thanatological projection matrix,
-$\textbf{Y}$}
+\caption{Fertility component, \textbf{$\Gamma$}, of unisex thanatological
+projection matrix.}
 \label{matrix:Yfert}
@@
 
-The survival and fertility components of $\textbf{Y}$ add together elementwise,
-thus the full 6$\times$6 matrix is composed as in Matrix~\ref{matrix:Y}.
-
-\begin{matrix}[h!]
-\centering
-\caption{A full unisex thanatological projection matrix, $\textbf{Y}$} 
-\label{matrix:Y}
-$\textbf{Y} = \bordermatrix{
-  {e_y } \vspace{.6em} & 0_t  & 1_t  & 2_t  & 3_t  & 4_t  & 5_t\cr 
-  0_{t+1} \vspace{.6em}&  (1-\lambda) \tfrac{\gamma_0d_0}{2} & (1-\lambda) \gamma_1d_0 + 1 &
-  (1-\lambda) \gamma_2d_0 & (1-\lambda) \gamma_3d_0 & (1-\lambda) \gamma_4d_0 & 0 \cr 
-    1_{t+1} \vspace{.6em}& \tfrac{\gamma_0d_1}{2} & \gamma_1d_1 & \gamma_2d_1 + 1 & \gamma_3d_1 & \gamma_4d_1 & 0 \cr 
-    2_{t+1} \vspace{.6em}& \tfrac{\gamma_0d_2}{2} & \gamma_1d_2 & \gamma_2d_2 & \gamma_3d_2 + 1 & \gamma_4d_2 & 0 \cr 
-   3_{t+1} \vspace{.6em}& \tfrac{\gamma_0d_3}{2} & \gamma_1d_3 & \gamma_2d_3 & \gamma_3d_3 & \gamma_4d_3 + 1 & 0 \cr 
-   4_{t+1} \vspace{.6em}& \tfrac{\gamma_0d_4}{2} & \gamma_1d_4 & \gamma_2d_4 & \gamma_3d_4 & \gamma_4d_4 & 1 \cr 
-   5_{t+1} \vspace{.6em}& \tfrac{\gamma_0d_5}{2} & \gamma_1d_5 & \gamma_2d_5 & \gamma_3d_5 & \gamma_4d_5 & 0 }$
-\end{matrix}
+The survival matrix, \textbf{S}, and the fertility matrix, \textbf{$\Gamma$} sum
+to create the thanatological projection matrix, \textbf{Y}:
+
+\begin{equation}
+\textbf{Y} = \textbf{S} + \textbf{$\Gamma$}
+\end{equation}
+
+%\begin{matrix}[h!]
+%\centering
+%\caption{A full unisex thanatological projection matrix, $\textbf{Y}$} 
+%\label{matrix:Y}
+%$\textbf{Y} = \bordermatrix{
+%  {e_y } \vspace{.6em} & 0_t  & 1_t  & 2_t  & 3_t  & 4_t  & 5_t\cr 
+%  0_{t+1} \vspace{.6em}&  (1-\lambda) \tfrac{\gamma_0d_0}{2} & (1-\lambda)
+  % \gamma_1d_0 + 1 & (1-\lambda) \gamma_2d_0 & (1-\lambda) \gamma_3d_0 & (1-\lambda) \gamma_4d_0 & 0 \cr 
+%    1_{t+1} \vspace{.6em}& \tfrac{\gamma_0d_1}{2} & \gamma_1d_1 & \gamma_2d_1 +
+    % 1 & \gamma_3d_1 & \gamma_4d_1 & 0 \cr 2_{t+1} \vspace{.6em}& \tfrac{\gamma_0d_2}{2} & \gamma_1d_2 & \gamma_2d_2 & \gamma_3d_2 + 1 & \gamma_4d_2 & 0 \cr 
+%   3_{t+1} \vspace{.6em}& \tfrac{\gamma_0d_3}{2} & \gamma_1d_3 & \gamma_2d_3 &
+   % \gamma_3d_3 & \gamma_4d_3 + 1 & 0 \cr 4_{t+1} \vspace{.6em}& \tfrac{\gamma_0d_4}{2} & \gamma_1d_4 & \gamma_2d_4 & \gamma_3d_4 & \gamma_4d_4 & 1 \cr 
+%   5_{t+1} \vspace{.6em}& \tfrac{\gamma_0d_5}{2} & \gamma_1d_5 & \gamma_2d_5 &
+   % \gamma_3d_5 & \gamma_4d_5 & 0 }$
+%\end{matrix}
 
@@
 clockwise around the graph on the thick blue paths in order until finally being
-absorbed into death (the node labeled ``X'', not included in the projection
-matrix). The light gray paths represent fertility. These originate from any age class that has a fertility rate greater than zero, which for this toy matrix are all age classes except the \nth{5}
+absorbed into death.
+The light gray paths represent fertility.
+These originate from any age class that has a fertility rate greater than zero, which for this toy matrix are all age classes except the \nth{5}
 (highest). The y$^{th}$ age class is subject to the fertility rate $\gamma
@@
 
-Thanatological age-classes will ideally terminate at the highest value permitted
-by data. For the data used in this paper, there are 111 total age classes, which
-translate to 111 total remaining-years classes (0-110+). In practice $\textbf{Y}$ becomes
-a 111$\times$111 matrix, with most entries non-zero (nearly complete
-connectivity, such as in Figure~\ref{fig:graph}). Construction may appear tedious for
-this reason.
+%Thanatological age-classes will ideally terminate at the highest value
+% permitted by data. For the data used in this paper, there are 111 total age classes, which
+%translate to 111 total remaining-years classes (0-110+). In practice
+% $\textbf{Y}$ becomes a 111$\times$111 matrix, with most entries non-zero (nearly complete
+%connectivity, such as in Figure~\ref{fig:graph}). Construction may appear
+% tedious for this reason.
 However, note that the bulk of fertility entries can be derived as the outer (tensor) product $d_a \otimes f_y$, leaving only the first row and first column mortality discounting followed by the addition of the
-survival superdiagonal. In most statistical programming languages constructing $\textbf{Y}$ entails only
-a couple more lines of code than constructing a Leslie matrix.
+%survival superdiagonal. In most statistical programming languages constructing
+% $\textbf{Y}$ entails only a couple more lines of code than constructing a Leslie matrix.
 
@@
 author to demonstrate this point. See also Appendix~\ref{app:B} for the
-corresponding proof from the continuous model.} The square of the ratio of
-the \nth{1} to the \nth{2} eigenvalues is the damping ratio, an indicator of the
-\textit{springiness} of the stable population structure, $c^\star(y)$, if perturbed from the
-state of stability.\footnote{See \citet[p.86-87]{caswell2001matrix}.} A higher
-damping ratio means that the population structure oscillates back to its stable
-state faster, i.e., that oscillations decrease in size more rapidly.
-Thanatological damping ratios are usually higher than chronological damping
-ratios in the data used elsewhere in this paper.
+corresponding proof from the continuous model.}
+
+%The square of the ratio of
+%the \nth{1} to the \nth{2} eigenvalues is the damping ratio, an indicator of
+% the \textit{springiness} of the stable population structure, $c^\star(y)$, if perturbed from the
+%state of stability.\footnote{See \citet[p.86-87]{caswell2001matrix}.} A higher
+%damping ratio means that the population structure oscillates back to its stable
+%state faster, i.e., that oscillations decrease in size more rapidly.
+%Thanatological damping ratios are usually higher than chronological damping
+%ratios in the data used elsewhere in this paper.
 

### 2016-02-01 ThanoRepro/ThanoRepro.tex
.

@@
 consider the amount of time left until death.\footnote{Elsewhere this quantity
-is referred to as \textit{residual} lifetime, time-to-death (TTD), prospective
+is referred to as \textit{residual} life, time-to-death (TTD), prospective
 age \citep{sanderson2007new}, or thanatological age \citep{riffe2015force}.}

### 2016-05-20 ThanoRepro/ThanoRepro.tex
minor edits from some weeks ago

@@
 I demonstrate approximate symmetry between years lived and left age structure in
-stability when subject to intrinsic growth rates equal magnitude and opposite sign.
+stability when subject to intrinsic growth rates of equal magnitude and opposite
+sign.
 
@@
 \subsection*{Contribution}
-The years left renewal model offers a new perspective on population renewal,
-and it is valid more generally as an aspect of birth-death processes.
+The years-left renewal model offers a new perspective on population renewal.
 \vspace{2cm}

### 2016-10-28 ThanoRepro/ThanoRepro.tex
minor edits to get the ball rolling in anticipation of having some empirical fertility rates by years lived and left

@@
 %suggestion for a better way to model or project population. Rather, 
-The years left renewal model represents an alternative and coherent way to
+The years-left renewal model represents an alternative and coherent way to
 describe the process of population renewal by relating renewal to a death flow
@@
 growth in the population.
-The years lived renewal model and the years left renewal
+The years-lived renewal model and the years-left renewal
 model both describe the movement of population stocks over time, but from opposite vantage points.
@@
 
-\section*{Requisites of a years left renewal model}
-The Lotka-Leslie renewal model \citep{leslie1945use,sharpe1911problem} is
+\section*{Requisites of a years-left renewal model}
+The Lotka-Leslie renewal model \citep{sharpe1911problem, leslie1945use} is
 situated in a closed, unisex, homogenous population framework. With
@@
 age, $a$. The model can be approximated by observed data because
-age is typically known.% The survival function, $\ell(a)$, derives
+age is typically known. % The survival function, $\ell(a)$, derives
 %from a lifetable, usually based on age-specific event-exposure mortality rates,
@@
 Age in the growth factor, $e^{-ra}$, pertains to the passage of
-time more than it does to chronological age (these are synchronous for birth
+time more than it does to chronological age (this is synchronous with birth
 cohorts). The product of this growth factor and $\ell(a)$ gives a
@@
 fertility rates, $m(a)$, are based on daughters born to mothers in this limited
-single-sex model. This model lends itself to commonly collected data because age
+single-sex model. The functions $m(a)$ and $\ell(a)$
+together intuitively make for net female fertility, and their age-wise product
+is sometimes written as $\phi(a)$. This model lends itself to commonly collected data because age
 is typically known, and can be asked.

### 2016-11-29 ThanoRepro/ThanoRepro.tex
starting to erase former lifetable transformation work. It distracts from the renewal aspect of the paper.

@@
 Each of these elements is structured by
-age, $a$. The model can be approximated by observed data because
-age is typically known. % The survival function, $\ell(a)$, derives
+age, $a$. %The model can be approximated by observed data because
+%age is typically known. % The survival function, $\ell(a)$, derives
 %from a lifetable, usually based on age-specific event-exposure mortality rates,
@@
 decomposing age-classified data, such as the
-population pyramid, into remaining years classes. If $P(a)$ is an
-age-classified population count, we can derive the years left version $P(y)$,
-using\footnote{An identical definition is found in
+population pyramid, into years-left classes. If $P(a)$ is an
+age-classified population count, under fixed mortality we can infer the
+years-left structure, $P(y)$, using\footnote{An identical definition is found in
 \citet{brouard1989mouvements} and illustrated in \citet{brouard1986structure}.}
@@
 
-Figure~\ref{fig:pop} shows this exercise for 2010 US data \citep{HMD},
-assuming homogeneity and that the 2010 period mortality schedule is held fixed.
-When looking from the age perspective (Figure~\ref{fig:USdecomp}), this
-decomposition reveals projected years left heterogeneity within a stock of known age, but when viewed from the
-years left perspective (Figure~\ref{fig:USrecomp}) it reveals age heterogeneity
-and the projected overall years left profile--- the future death flow of the
-current population, $P(y)$ under our assumptions.
+%Figure~\ref{fig:pop} shows this exercise for 2010 US data \citep{HMD},
+%assuming homogeneity and that the 2010 period mortality schedule is held fixed.
+%When looking from the age perspective (Figure~\ref{fig:USdecomp}), this
+%decomposition reveals projected years left heterogeneity within a stock of known age, but when viewed from the
+%years left perspective (Figure~\ref{fig:USrecomp}) it reveals age heterogeneity
+%and the projected overall years left profile--- the future death flow of the
+%current population, $P(y)$ under our assumptions.
 
 % both figures created in R/Pyramids.R
-\begin{figure}[ht!]
-	\caption{2010 US population structure}
-	\label{fig:pop}
-	\begin{center}
-	\begin{subfigure}{.45\textwidth}
-		\caption{Age structure (years lived) decomposed by years left.}
-		\label{fig:USdecomp}
-		\includegraphics[scale=.6]{Figures/US2010Age.pdf}
-	\end{subfigure}
-	~
-	\begin{subfigure}{.45\textwidth}
-		\caption{Years left structure decomposed by years lived.}
-		\label{fig:USrecomp}
-		\includegraphics[scale=.6]{Figures/US2010Thano.pdf}
-	\end{subfigure}
-	\\
-	\small{*Population and mortality data from the HMD.}
-	\end{center}
-\end{figure}
+%\begin{figure}[ht!]
+%	\caption{2010 US population structure}
+%	\label{fig:pop}
+%	\begin{center}
+%	\begin{subfigure}{.45\textwidth}
+%		\caption{Age structure (years lived) decomposed by years left.}
+%		\label{fig:USdecomp}
+%		\includegraphics[scale=.6]{Figures/US2010Age.pdf}
+%	\end{subfigure}
+%	~
+%	\begin{subfigure}{.45\textwidth}
+%		\caption{Years left structure decomposed by years lived.}
+%		\label{fig:USrecomp}
+%		\includegraphics[scale=.6]{Figures/US2010Thano.pdf}
+%	\end{subfigure}
+%	\\
+%	\small{*Population and mortality data from the HMD.}
+%	\end{center}
+%\end{figure}
 
@@
 with historical stocks. \citet{brouard1989mouvements}, and \citet{vaupel2009life}, showed that in stationary populations the
-marginal chronological and thanatological age structures are identical, and this
-provides a limited basis for comparison between the two perspectives. 
+marginal years-lived and years-left distributions are identical,
+and this provides a limited basis for comparison between the two perspectives. 
 %\citet{wachter2014essential} hints at, and
@@
 
-Lifetable-derived thanatological age structure is
+Lifetable-derived years-left structure is
 probabilistic until after death in several senses. In this case, note that
@@
 case a period mortality pattern is used.
-As such, Figure~\ref{fig:USrecomp} is \textit{projective} in nature, a
+Redistribution as in \eqref{eq:transform} is \textit{projective}, a
 forward-looking glance at population attrition given the population stock and lifetable of a
-particular moment, whereas Figure~\ref{fig:USdecomp} is \textit{reflective} in
-nature, since a population's chronological age-structure is the fruit of past
-demographic forces.
-
-\subsection{Thanatological fertility rates.}
-Any count
-classified by chronological age can be reclassified as in
-equation~\eqref{eq:transform} given a pertinent lifetable. To
-derive thanatological fertility rates under these assumptions, $\gamma(y)$, apply \eqref{eq:transform} to
-chronological age-classified birth counts, $B(a)$, to get $B(y)$ and again to
-exposure-to-risk, $E(a)$, to get $E(y)$ and then divide in the usual way:
-
-\begin{equation}
-\gamma(y) = \frac{B(y}{E(y)} \ep
-\end{equation}
-
-Figure~\ref{fig:FySpaghetti}
-represents the full variety in female thanatological period fertility rates that can be found for all years
-of data that overlap in the HFD and HMD.\footnote{There are as of this writing
-1933 population-years of overlap between the HMD and HFD, including a wide
-variety of fertility and mortality combinations.} I show $\gamma(y)$ calculated in this way because
-it enters into the renewal model here studied. Figure~\ref{fig:FxSpaghetti}
-gives ASFR for the same populations and years as a more familiar reference, but
-note that both the x and y axes are different. The shape of thanatological
-fertility curves is speculative and due to the above-mentioned strong assumptions. Nonetheless it is best not to
-include such rates into a model blindly, so we offer this schematic overview of
-some of their characteristics.
-
-Period thanatological fertility rates estimated in this way have a
-characteristic shape. The distribution is wider than that of chronological
-fertility rates. The range of post-childbearing lifespans stretches from 
-zero, due e.g., to maternal mortality, to at least 100.\footnote{For instance,
-Jeanne Calment lived for another 100 years after giving birth to her daughter.}
-This spaghetti plot includes both fertility booms and busts, as well as some
-mortality crises (1918, WWII). Some patterns in the data (not
-highlighted in the figure) include that the left tail (a gauge of how
-orphan-prone a population is) has tended to fall over time. All populations have shown an increase in both the
-mean and mode of thanatological age at childbearing over time. Several
-contemporary populations now show modes of over 60 years. Period thanatological total
-fertility rates track standard period total fertility rates closely, but tend to be somewhat higher (not shown).
-% both figures created in R/HFDspaghetti.R
-\begin{figure}[h!]
-	\caption{Chronological and thantological fertility rates, all 1933
-	country-year combinations present in both the HFD and HMD.* Note different x
-	and y scales.}
-	\label{fig:Fxcompare}
-	\begin{center}
-	\makebox[\textwidth]{
-	\begin{subfigure}{.45\textwidth}
-		\caption{Chronological period fertility rates.}
-		\label{fig:FxSpaghetti}
-		\includegraphics[scale=.45]{Figures/FxSpaghettiDraft.png}
-	\end{subfigure}
-	~
-	\begin{subfigure}{.45\textwidth}
-		\caption{Thanatological period fertility rates.}
-		\label{fig:FySpaghetti}
-		\includegraphics[scale=.45]{Figures/FySpaghettiDraft.png}
-	\end{subfigure}
-	}
-	\\
-	\end{center}
-	\begin{tiny}
-	*AUT, $1951-2010$; BGR, $1947-2009$; BLR, $1964-2012$; CAN, $1921-2009$; 
-	CHE, $1932-2011$; CHL, $1992-2005$; CZE, $1950-2011$; DEUTE, $1956-2011$; 
-	DEUTNP, $1990-2011$; DEUTW, $1956-2011$; ESP, $1922-2012$; EST, $1959-2011$; 
-	FIN, $1939-2009$; FRATNP, $1946-2013$; GBR\_NIR, $1974-2011$; GBR\_NP,
-	$1974-2011$; GBR\_SCO, $1945-2011$; GBRTENW, $1938-2011$; HUN, $1950-2009$;
-	IRL, $1955-2009$; ISL, $1963-2010$; JPN, $1947-2012$; LTU, $1959-2011$; NLD, $1950-2009$; 
-	NOR, $1967-2009$; PRT, $1940-2012$; RUS, $1959-2010$; SVK, $1950-2009$; 
-	SVN, $1983-2009$; SWE, $1891-2011$; TWN, $1976-2010$; UKR, $1959-2009$; 
-	USA, $1933-2010$
-	\end{tiny}
-\end{figure}
-
-In general, thanatological fertility rates will not be useful for purposes of
-projection, unless their empirical regularity is
-stronger than typical ASFR. This is a question I do not seek to answer
-in the present treatment, but in any case such rates are not susceptible to
-direct measurement.
-There are alternative ways of defining thanatological fertility rates, for instance by transforming the birth and
-population vector from a stable population. As proven in appendix~\ref{app:B},
-thanatological fertility rates derived from a stable population imply a certain
-symmetry with stable populations from the chronological model.
-
+particular moment, whereas the classic pyramid is \textit{reflective}, since a
+population's age-structure owes to past demographic forces.
+
+Other age-structured quantities can be reclassified under the same assumptions.
+If birth counts and population exposures are reclassified as in
+\eqref{eq:transform}, their ratio yields like-classified fertility rates,
+$\gamma{y}$.
+
+%\subsection{Years-left fertility rates.}
+%Any age-classified can be reclassified as in
+%equation~\eqref{eq:transform} given a pertinent lifetable. To
+%derive thanatological fertility rates under these assumptions, $\gamma(y)$, apply \eqref{eq:transform} to
+%chronological age-classified birth counts, $B(a)$, to get $B(y)$ and again to
+%exposure-to-risk, $E(a)$, to get $E(y)$ and then divide in the usual way:
+%
+%\begin{equation}
+%\gamma(y) = \frac{B(y}{E(y)} \ep
+%\end{equation}
+%
+%Figure~\ref{fig:FySpaghetti}
+%represents the full variety in female thanatological period fertility rates that can be found for all years
+%of data that overlap in the HFD and HMD.\footnote{There are as of this writing
+%1933 population-years of overlap between the HMD and HFD, including a wide
+%variety of fertility and mortality combinations.} I show $\gamma(y)$ calculated in this way because
+%it enters into the renewal model here studied. Figure~\ref{fig:FxSpaghetti}
+%gives ASFR for the same populations and years as a more familiar reference, but
+%note that both the x and y axes are different. The shape of thanatological
+%fertility curves is speculative and due to the above-mentioned strong assumptions. Nonetheless it is best not to
+%include such rates into a model blindly, so we offer this schematic overview of
+%%some of their characteristics.
+
+%Period thanatological fertility rates estimated in this way have a
+%characteristic shape. The distribution is wider than that of chronological
+%fertility rates. The range of post-childbearing lifespans stretches from 
+%zero, due e.g., to maternal mortality, to at least 100.\footnote{For instance,
+%Jeanne Calment lived for another 100 years after giving birth to her daughter.}
+%This spaghetti plot includes both fertility booms and busts, as well as some
+%mortality crises (1918, WWII). Some patterns in the data (not
+%highlighted in the figure) include that the left tail (a gauge of how
+%orphan-prone a population is) has tended to fall over time. All populations have shown an increase in both the
+%mean and mode of thanatological age at childbearing over time. Several
+%contemporary populations now show modes of over 60 years. Period thanatological total
+%fertility rates track standard period total fertility rates closely, but tend to be somewhat higher (not shown).
+%% both figures created in R/HFDspaghetti.R
+%\begin{figure}[h!]
+%	\caption{Chronological and thantological fertility rates, all 1933
+%	country-year combinations present in both the HFD and HMD.* Note different x
+%	and y scales.}
+%	\label{fig:Fxcompare}
+%	\begin{center}
+%	\makebox[\textwidth]{
+%	\begin{subfigure}{.45\textwidth}
+%		\caption{Chronological period fertility rates.}
+%		\label{fig:FxSpaghetti}
+%		\includegraphics[scale=.45]{Figures/FxSpaghettiDraft.png}
+%	\end{subfigure}
+%	~
+%	\begin{subfigure}{.45\textwidth}
+%		\caption{Thanatological period fertility rates.}
+%		\label{fig:FySpaghetti}
+%		\includegraphics[scale=.45]{Figures/FySpaghettiDraft.png}
+%	\end{subfigure}
+%	}
+%	\\
+%	\end{center}
+%	\begin{tiny}
+%	*AUT, $1951-2010$; BGR, $1947-2009$; BLR, $1964-2012$; CAN, $1921-2009$; 
+%	CHE, $1932-2011$; CHL, $1992-2005$; CZE, $1950-2011$; DEUTE, $1956-2011$; 
+%	DEUTNP, $1990-2011$; DEUTW, $1956-2011$; ESP, $1922-2012$; EST, $1959-2011$; 
+%	FIN, $1939-2009$; FRATNP, $1946-2013$; GBR\_NIR, $1974-2011$; GBR\_NP,
+%	$1974-2011$; GBR\_SCO, $1945-2011$; GBRTENW, $1938-2011$; HUN, $1950-2009$;
+%	IRL, $1955-2009$; ISL, $1963-2010$; JPN, $1947-2012$; LTU, $1959-2011$; NLD, $1950-2009$; 
+%	NOR, $1967-2009$; PRT, $1940-2012$; RUS, $1959-2010$; SVK, $1950-2009$; 
+%	SVN, $1983-2009$; SWE, $1891-2011$; TWN, $1976-2010$; UKR, $1959-2009$; 
+%	USA, $1933-2010$
+%	\end{tiny}
+%\end{figure}
+%
+%In general, thanatological fertility rates will not be useful for purposes of
+%projection, unless their empirical regularity is
+%stronger than typical ASFR. This is a question I do not seek to answer
+%in the present treatment, but in any case such rates are not susceptible to
+%direct measurement.
+%There are alternative ways of defining thanatological fertility rates, for instance by transforming the birth and
+%population vector from a stable population. As proven in appendix~\ref{app:B},
+%thanatological fertility rates derived from a stable population imply a certain
+%symmetry with stable populations from the chronological model. 
+%
 \FloatBarrier
-\section*{The thanatological renewal model}
+\section*{The years-left renewal model}
 The renewal model that follows uses thanatological fertility rates, $\gamma(y)$,

### 2016-11-29 ThanoRepro/ThanoRepro.tex
further tinkering with text around lotka renewal eq

@@
 The Lotka-Leslie renewal model \citep{sharpe1911problem, leslie1945use} is
-situated in a closed, unisex, homogenous population framework. With
-parameters equated, the continuous form of the model contains a component for
-survival, $\ell(a)$, a scaling factor for growth, $e^{-ra}$, and a component for fertility, $m(a)$:
-
-\begin{equation}
+situated in a closed, unisex, homogenous population framework. The continuous form of the model contains a component for
+survival, $\ell(a)$, a scaling factor for growth, $e^{-ra}$, and a component for
+female fertility, $m(a)$:
+
+\begin{align}
 \label{eq:lotka}
-1 = \int_{a=0}^\infty l(a)e^{-ra}m(a)\dd a \ep
-\end{equation}
+B(t) &= \int_{0}^\infty B(t)\ell(a)e^{-ra}m(a)\dd a \ec \\
+\intertext{or when parameters are equated}
+1 &= \int_{0}^\infty \ell(a)e^{-ra}m(a)\dd a \ep
+\end{align}
 
@@
 fertility rates, $m(a)$, are based on daughters born to mothers in this limited
-single-sex model. The functions $m(a)$ and $\ell(a)$
-together intuitively make for net female fertility, and their age-wise product
-is sometimes written as $\phi(a)$. This model lends itself to commonly collected data because age
-is typically known, and can be asked.
+single-sex model.  The age-wise product of $m(a)$ and $\ell(a)$
+can be interpreted as net female fertility, and this
+is sometimes written as $\phi(a)$. As with most demographic models and
+measures, this abstraction of renewal lends itself to
+commonly collected data because age is typically known.
 % For the two essential rate componenents underlying the renewal

### 2016-11-29 ThanoRepro/ThanoRepro.tex
mostly tinkering around with text. Big structural changes to come!

@@
 is sometimes written as $\phi(a)$. As with most demographic models and
-measures, this abstraction of renewal lends itself to
-commonly collected data because age is typically known.
+measures, this model of renewal is tractable with
+commonly collected population and vital data because age is typically known.
 % For the two essential rate componenents underlying the renewal

### 2016-11-30 ThanoRepro/R/Quebec.R
change Quebec script to remive immigrants (only 1k or so of them), and to also return lx

@@ -54,8 +54,9 @@ Q         <- Q[!is.na(Q$BD),]
 # Q into two, M, for mothers, and Ch, for children.
 M         <- Ch      <- Q
 
-# filter potential mothers down to females with non-NA L:
-M         <- M[!is.na(M$L) & M$sexe == "f", ]
+# filter potential mothers down to females with non-NA L and non-immigrants,
+# for the sake of an ubiased denominator (child mort erased otherwise).
+M         <- M[!is.na(M$L) & M$sexe == "f" & M$immigrant == 0, ]
 
 # children need to be filtered down to only those
 # whose mother is still in M:
@@ -151,13 +152,17 @@ get.birth.rates    <- function(
 	age <- seq(0, omega, by = ageN)
 	
 	# calculate Lx  (denominator)
-	denom               <- rep(0, length(age))
+	Exposure               <- rep(0, length(age))
+	# also get lx-proportionate
+	ExpTrunc               <- rep(0, length(age))
 	for (i in 1:length(age))  {
 		# ceiling() will work for single, but need fancy otherwise
-		ind             <- (cohort$L - (cohort$L %% ageN) + ageN) == age[i]
+		ind                <- (cohort$L - (cohort$L %% ageN) + ageN) == age[i]
 		# one more step with %% to get right age group.
-		dec             <- sum(ageN - (age[i] - cohort$L[ind]))
-		denom[i]        <- sum(cohort$L >= age[i]) + dec
+		dec                <- sum(ageN - (age[i] - cohort$L[ind]))
+		ExpTrunc[i]        <- sum(cohort$L >= age[i])
+		Exposure[i]        <- ExpTrunc[i] + dec
+		
 	} 
 	
 	# optionally, just select female births:
@@ -178,12 +183,13 @@ get.birth.rates    <- function(
 	
 	out                 <- data.frame(Cohort = startyear,
 			                          Age = age, 
-			                          Exposure = denom, 
+			                          Exposure = Exposure, 
 									  BA = NA, 
 									  BT = NA, 
[…]

### 2016-11-30 ThanoRepro/R/rEqual.R
rename fstar to gamma, as in paper

@@ -122,7 +122,7 @@ rThanoCoale(.fy2,da2,.a=a2)
 # easier to prove on paper if we know this is true...
 
 # one thing to verify:
-#f^\star(y) is the c(a)d(y|a)-weighted average of f(a)...
+#gamma(y) is the c(a)d(y|a)-weighted average of f(a)...
 
 sum(.fy2)
 sum(fa)
@@ -199,7 +199,7 @@ sum(
                 colSums(dxM * growth / sum(dxM * growth))
 )
 
-# f^\star(y)
+# gamma(y)
 colSums(fa * dxM * growth / sum(dxM * growth)) / 
         colSums(dxM * growth / sum(dxM * growth))
 

### 2016-11-30 ThanoRepro/ThanoRepro.tex
still in midst of edits pre thano renewal

@@
 \usepackage{setspace}
-\doublespacing
+%\doublespacing
 
@@
 
-Remaining lifetime is different in this respect because we do not know when we
-will die--- it is only knowable retrospectively or probabilistically. Under a set of simple assumptions, the same lifetable used to
-estimate $\ell(a)$ can be used to
-partition and recombine counts observed by age into estimates
-classified by years of life left. Our objective is then to estimate a population
-structure component, similar to $\ell(a)e^{-ra}$, and a fertility component
+Fertility rates by remaining lifetime is different in this respect because death
+dates can not be registered on birth certificates in the way that
+maternal age or birth year can; Death timing is only knowable retrospectively or probabilistically. Since
+it is difficult to imagine what a fertility schedule might look like when
+classified by remaining years of life, I give two examples, first of empirically
+tabulated years-left female fertiltiy rates, and second rates derived
+probabilistically using a lifetable. Empirical years-left fertility rates refer
+to the
+
+ Under
+a set of simple assumptions, the same lifetable used to estimate $\ell(a)$ can be used to partition and recombine counts observed by age into estimates classified by years of life left.
+Our objective is then to estimate a population structure component, similar to $\ell(a)e^{-ra}$, and a fertility component
 similar to $m(a)$, but each for the years left case. Taken together, these

### 2016-11-30 ThanoRepro/R/Quebec.R
add lx calc for Quebec. Can now actually calculate r, etc for Quebec

@@ -239,73 +239,78 @@ Rates1x5 <- do.call(rbind,
 
 save(Rates1x5,file = "/home/tim/git/ThanoRepro/ThanoRepro/Data/QuebecRates1x5.Rdata")
 
-Rates2x5 <- do.call(rbind,
-		lapply(coh5, function(cohi, Ch, M){
-					get.birth.rates(Ch = Ch,
-							M = M, 
-							startyear = cohi, 
-							cohN = 5,
-							ageN = 2,
-							omega = 110,
-							sex = "f")
-				},Ch = Ch , M = M))
+# ---------------------------------------#
+head(Rates1x5)
 
 
-save(Rates2x5,file = "/home/tim/git/ThanoRepro/ThanoRepro/Data/QuebecRates2x5.Rdata")
-
-Rates3x5 <- do.call(rbind,
-		lapply(coh5, function(cohi, Ch, M){
-					get.birth.rates(Ch = Ch,
-							M = M, 
-							startyear = cohi, 
-							cohN = 5,
-							ageN = 3,
-							omega = 110,
-							sex = "f")
-				},Ch = Ch , M = M))
-
-
-save(Rates3x5,file = "/home/tim/git/ThanoRepro/ThanoRepro/Data/QuebecRates3x5.Rdata")
-
-Rates5x5 <- do.call(rbind,
-		lapply(coh5, function(cohi, Ch, M){
-					get.birth.rates(Ch = Ch,
-							M = M, 
-							startyear = cohi, 
-							cohN = 5,
-							ageN = 5,
-							omega = 110,
-							sex = "f")
-				},Ch = Ch , M = M))
-
-
-save(Rates5x5,file = "/home/tim/git/ThanoRepro/ThanoRepro/Data/QuebecRates5x5.Rdata")
-coh10 <- seq(from = 1660, to = 1740, by = 10) 
-Rates1x10 <- do.call(rbind,
-		lapply(coh10, function(cohi, Ch, M){
-					get.birth.rates(Ch = Ch,
-							M = M, 
-							startyear = cohi, 
-							cohN = 10,
-							omega = 110,
-							sex = "f")
-				},Ch = Ch , M = M))
-
-
-save(Rates1x10,file = "/home/tim/git/ThanoRepro/ThanoRepro/Data/QuebecRates1x10.Rdata")
-Rates2x10 <- do.call(rbind,
-		lapply(coh10, function(cohi, Ch, M){
-					get.birth.rates(Ch = Ch,
[…]

### 2016-11-30 ThanoRepro/R/RiffeFunctions/R/stablefunctions.R
remove T.guess arg from r function, also fix age arg

@@ -37,7 +37,7 @@ getca <- function(La,r,a=.5:110.5){
 #' 
 #' @importFrom compiler cmpfun
 
-rThanoCoale <- cmpfun(function(fy, da, .a = .5:110.5, maxit = 1e2, tol = 1e-15){  
+rThanoCoale <- cmpfun(function(fy, da, a = .5:110.5, maxit = 1e2, tol = 1e-15){  
     
     # Based on Coale (1957), modified.
     N    <- length(fy)
@@ -48,13 +48,13 @@ rThanoCoale <- cmpfun(function(fy, da, .a = .5:110.5, maxit = 1e2, tol = 1e-15){
       dxi <- dxi[2:length(dxi) ]
     }     
     R0      <- sum(dxM * fy)
-    T.guess <- sum(.a * dxM * fy) / R0 # assuming r = 0
+    T.guess <- sum(a * dxM * fy) / R0 # assuming r = 0
     r.i     <- log(R0) / T.guess
     # be careful to discount Fy by SRB appropriately for males / females
     # prior to specification
     for (i in 1:maxit){ # 15 is more than enough!
       #cat(r2,i,"\n")
-      deltai <- 1 - sum(colSums(dxM * exp(-r.i * .a)) * fy)
+      deltai <- 1 - sum(colSums(dxM * exp(-r.i * a)) * fy)
       # the mean generation time self-corrects 
       # according to the error produced by the Lotka equation
       r.i <- r.i - (deltai / (T.guess - (deltai / r.i))) 
@@ -93,7 +93,6 @@ Rmomentn <- cmpfun(function(fa,La,a,n=0){
 #' @param fa ASFR, all ages, including zeros.
 #' @param La lifetable exposure (or survival, if you're lazy).
 #' @param a age (midpoints, usually).
-#' @param T.guess a guess at mean generation length. default 29.
 #' @param maxit maximum number of iterations to converge, default 100.
 #' @param tol convergence tolerance, default 1e-15.
 #' 
@@ -104,10 +103,11 @@ Rmomentn <- cmpfun(function(fa,La,a,n=0){
 #' @export
 #' 
 #' @importFrom compiler cmpfun
-rLotkaCoale <- cmpfun(function(fa,La,a=.5:110.5, T.guess = 29, maxit = 1e2, tol = 1e-15){
+rLotkaCoale <- cmpfun(function(fa,La,a=.5:110.5, maxit = 1e2, tol = 1e-15){
[…]
